### 2022-05-15 Scripts/Evaluations/identifying_viral_contigs_20220515.Rmd
First viral sorting rules pipeline iteration.

---
title: "Identifying Viral Contigs"
output: html_notebook
---

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(treemapify)
library(tidyr)
library(pheatmap)
library(dplyr)
library(gridExtra)
library(readr)
```

# All Viral Contigs

## checkV
```{r}
checkV <- read_tsv("../../ToolOutput/quality_summary.tsv")
checkV$contig <- sub("\\.", "_", checkV$contig_id)
```

## VIBRANT
```{r}
vb_c <- read_tsv("../../ToolOutput/VIBRANT_genome_quality_combined_virus_refseq.tsv")
vb_c$contig <- sub("\\.", "_", vb_c$scaffold)
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
vb_c <- vb_c[!duplicated(vb_c$contig),]
```

## DeepVirFinder
```{r}
dvf_c <- read_tsv("../../ToolOutput/combined_virus_refseq_2000000.fasta_gt1bp_dvfpred.txt")
dvf_c$contig <- sub("\\.", "_", dvf_c$name)
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
```

## VirSorter2 
```{r}
vs2_c <- read_tsv("../../ToolOutput/final-viral-score.tsv")
vs2_c <- separate(vs2_c, col = seqname, into = c("Contig", "Type"), "\\|\\|")
vs2_c$contig <- sub("\\.", "_", vs2_c$Contig)
vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
```

## Merging
```{r}
viruses <- full_join(x=checkV, y=dvf_c, by = c("contig"))
viruses <- full_join(x=viruses, y=vb_c, by = c("contig"))
viruses <- full_join(x=viruses, y=vs2_c, by = c("contig"))
```

```{r}
head(viruses)
```

```{r}
viruses$percent_host <- viruses$host_genes/viruses$gene_count*100
viruses$percent_viral <- viruses$viral_genes/viruses$gene_count*100
viruses$percent_unknown <- 100-(viruses$host_genes+viruses$viral_genes)/viruses$gene_count*100

viruses$Quality[is.na(viruses$Quality)] <- 0
viruses$score[is.na(viruses$score)] <- 0
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
```

######################################################################
# New rules for keeping

## Getting Score for Deciding Whether to Keep
```{r}
viruses$keep_score_vibrant <- 0
viruses$keep_score_deepvirfinder <- 0
viruses$keep_score_virsorter2 <- 0
viruses$keep_score_checkV <- 0

#vibrant
viruses$keep_score_vibrant[viruses$Quality=="high quality draft"] <- 1
viruses$keep_score_vibrant[viruses$Quality=="medium quality draft"] <- 0.52
viruses$keep_score_vibrant[viruses$Quality=="low quality draft"] <- 0.333
viruses$keep_score_vibrant[viruses$viral_genes==0 & viruses$host_genes>=1] <- 0
viruses$keep_score_vibrant[(viruses$viral_genes*3) <= viruses$host_genes] <- 0
viruses$keep_score_vibrant[viruses$contig_length>50000 & viruses$hallmark==0] <- 0

#deepvirfinder
viruses$keep_score_deepvirfinder[viruses$bh_pvalue<=0.05] <- 1
viruses$keep_score_deepvirfinder[viruses$viral_genes==0 & viruses$host_genes>=1] <- 0
viruses$keep_score_deepvirfinder[(viruses$viral_genes*3) <= viruses$host_genes] <- 0
viruses$keep_score_deepvirfinder[viruses$contig_length>50000 & viruses$hallmark==0] <- 0

#virsorter2
viruses$keep_score_virsorter2[viruses$viral>=50] <- 0.333
viruses$keep_score_virsorter2[viruses$hallmark>2] <- 1
viruses$keep_score_virsorter2[viruses$viral>=95] <- viruses$keep_score_virsorter2[viruses$viral>=95] + 1
viruses$keep_score_virsorter2[viruses$viral_genes==0 & viruses$host_genes>=1] <- 0
viruses$keep_score_virsorter2[(viruses$viral_genes*3) <= viruses$host_genes] <- 0
viruses$keep_score_virsorter2[viruses$contig_length>50000 & viruses$hallmark==0] <- 0

#checkV
viruses$keep_score_checkV[(viruses$viral>=50 & viruses$host_genes==0) & viruses$viral_genes==0] <- viruses$keep_score_checkV[(viruses$viral>=50 & viruses$host_genes==0) & viruses$viral_genes==0] + 0.52
viruses$keep_score_checkV[viruses$percent_viral>=50] <- viruses$keep_score_checkV[viruses$percent_viral>=50] + 1
viruses$keep_score_checkV[viruses$viral_genes==0 & viruses$host_genes>=1] <- 0
viruses$keep_score_checkV[(viruses$viral_genes*3) <= viruses$host_genes] <- 0
viruses$keep_score_checkV[viruses$contig_length>50000 & viruses$hallmark==0] <- 0
```

### keep high quality by any method (minus discard rules at bottom), medium quality in at least 3, low if consensus with others 
```{r}
viruses$keep_score <- viruses$keep_score_checkV + viruses$keep_score_virsorter2 +
  viruses$keep_score_deepvirfinder + viruses$keep_score_vibrant

viruses$keep_score_high <- 0
viruses$keep_score_high[viruses$keep_score>=1] <- 1
```

```{r}
table(viruses$keep_score_high)
```

```{r}
viruses_high <- viruses[viruses$keep_score_high==1,]
```

```{r}
head(viruses_high)
```

# Plot of how many methods identify a contig
```{r}
viruses_high_mod <- viruses_high %>% select("keep_score_vibrant":"keep_score_checkV")
viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x > 0) {x <- 1} else {x <- 0}})
viruses_high_mod <- as_tibble(viruses_high_mod)

sm <- as.data.frame(rbind(c(0, table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vibrant>0,]))),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_deepvirfinder>0,])),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_checkV>0,])),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_virsorter2>0,]))
                          )
)

sm$method <- c("vibrant", "deepvirfinder", "checkV_polishing", "virsorter2")

colnames(sm) <- c("one","two","three","four", "method")

sm
```

```{r}
sm_m <- melt(sm)
colnames(sm_m) <- c("primary_method", "number_methods", "number")
```

```{r}
ggplot(sm_m, aes(x=primary_method, y=number,
                   fill=number_methods)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Number of Methods',
                     values = alpha(c(viridis(7)), 1)) +
  xlab("Primary Method") +
  ylab("Count of Viral Contigs")
```

```{r}
hist(viruses$keep_score[viruses$keep_score_high==1])
```

```{r}

ggplot(viruses_high, aes(x=percent_host, y=percent_viral,
                   color=keep_score, fill=keep_score)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_viridis(option = "viridis") +
  scale_color_viridis(option = "viridis") +
  xlab("Percent Genes Host") +
  ylab("Percent Genes Viral") +
  coord_flip()

ggplot(viruses_high, aes(x=percent_host, y=percent_viral,
                   color=score, fill=score)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_viridis(option = "magma") +
  scale_color_viridis(option = "magma") +
  xlab("Percent Genes Host") +
  ylab("Percent Genes Viral") +
  coord_flip()

ggplot(viruses_high, aes(x=percent_host, y=percent_viral,
                   color=log10(0.1+hallmark), 
                   fill=log10(0.1+hallmark))) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_viridis(option = "viridis") +
  scale_color_viridis(option = "viridis") +
  xlab("Percent Genes Host") +
  ylab("Percent Genes Viral") +
  coord_flip()

ggplot(viruses_high, aes(x=percent_host, y=percent_viral,
                   color=log10(contig_length), fill=log10(contig_length))) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_viridis(option = "magma") +
  scale_color_viridis(option = "magma") +
  xlab("Percent Genes Host") +
  ylab("Percent Genes Viral") +
  coord_flip()

ggplot(viruses_high, aes(x=percent_viral, y=contig_length,
                   color=log10(0.1+hallmark), 
                   fill=log10(0.1+hallmark))) +
  geom_point(alpha=0.1) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_viridis(option = "viridis") +
  scale_color_viridis(option = "viridis") +
  xlab("Percent Viral") +
  ylab("Contig Length") +
  scale_y_log10()

ggplot(viruses_high, aes(x=percent_viral, y=percent_unknown,
                   color=log10(contig_length), 
                   fill=log10(contig_length))) +
  geom_point(alpha=0.1) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_viridis(option = "magma") +
  scale_color_viridis(option = "magma") +
  xlab("Percent Viral") +
  ylab("Percent Unknown") 
```

### 2022-06-01 Scripts/Evaluations/identifying_viral_contigs_20220521.Rmd
Added updated rule set.

---
title: "Viral Sequence Sorting Evaluation"
output: html_notebook
---

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
```

# All Viral Contigs

## checkV
```{r}
checkV <- read_tsv("../TestingSetOutput/quality_summary_20220521.tsv")
checkV <- separate(checkV, col = contig_id, into = c("seqtype", "contig"), sep="--")
checkV$contig <- sub("\\.", "_", checkV$contig)
```

## VIBRANT
```{r}
vb_c <- read_tsv("../TestingSetOutput/VIBRANT_genome_quality_metagenomic_testing_set_20220521.tsv")
vb_c <- separate(vb_c, col = scaffold, into = c("seqtype", "contig"), sep="--")
vb_c$contig <- sub("\\.", "_", vb_c$contig)
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
vb_c <- vb_c[!duplicated(vb_c$contig),]
```

## DeepVirFinder
```{r}
dvf_c <- read_tsv("../TestingSetOutput/dvf_output_20220521.tsv")
dvf_c <- separate(dvf_c, col = name, into = c("seqtype", "contig"), sep="--")
dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
```

## VirSorter2 
```{r}
vs2_c <- read_tsv("../TestingSetOutput/final-viral-score_20220521.tsv")
vs2_c <- separate(vs2_c, col = seqname, into = c("seqtype", "contig"), sep="--")
vs2_c <- separate(vs2_c, col = contig, into = c("contig", "vs2type"), "\\|\\|")
vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
```

## Kaiju
(parsing when reading the file in "fails" for sequences that weren't classified)
```{r}
kj_c <- read_tsv("../VSTE/ToolOutput/metagenomic_testing_set_nrandeuk.kaiju.names.out", col_names = F)
colnames(kj_c) <- c("Classified", "Contig", "NCBI_taxon", "len", "ID_best", "IDs_all", "Seq", "Name")
kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
kj_c$contig <- sub("\\.", "_", kj_c$contig)
```

```{r}
table(kj_c$seqtype, kj_c$Kaiju_Viral)
```

## VirSorter
```{r}
vs_c <- read_csv("../TestingSetOutput/VIRSorter_global-phage-signal.csv")
vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
vs_c <- separate(vs_c, col = `## Contig_id`, into = c("seqtype", "contig"), sep="--")
vs_c$contig <- sub("-circular", "", vs_c$contig)
vs_c <- vs_c[!duplicated(vs_c$contig),]
```

## Merging

```{r}
viruses <- full_join(x=vs2_c, y=dvf_c, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=checkV, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=vb_c, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs_c, by = c("contig", "seqtype"))
viruses <- full_join(x=viruses, y=kj_c, by = c("contig", "seqtype"))
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$host_genes/viruses$gene_count*100
viruses$percent_viral <- viruses$viral_genes/viruses$gene_count*100
viruses$percent_unknown <- 100-(viruses$host_genes+viruses$viral_genes)/viruses$gene_count*100
```

get rid of NAs for downstream processing
```{r}
viruses$Quality[is.na(viruses$Quality)] <- 0
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
viruses$Category[is.na(viruses$Category)] <- 0
```

##################################################

```{r}
getting_viral_set_1 <- function(input_seqs, include_vibrant=FALSE, include_virsorter2=FALSE,
                              include_deepvirfinder=FALSE, include_checkV=FALSE,
                              include_kaiju=FALSE, include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$Category==1] <- keep_score[input_seqs$Category==1] + 1
    keep_score[input_seqs$Category==2] <- keep_score[input_seqs$Category==2] + 1
    keep_score[input_seqs$Category==3] <- keep_score[input_seqs$Category==3] + 0.5
    keep_score[input_seqs$Category==4] <- keep_score[input_seqs$Category==4] + 1
    keep_score[input_seqs$Category==5] <- keep_score[input_seqs$Category==5] + 1
    keep_score[input_seqs$Category==6] <- keep_score[input_seqs$Category==6] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
  }

  if (include_kaiju) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
  }

  return(keep_score)

}

```

```{r}
assess_performance <- function(seqtype, keep_score) {

  truepositive <- rep("not viral", length(seqtype))
  truepositive[seqtype=="virus"] <- "viral"

  #make confusion matrix
  confusion_matrix <- rep("true negative", length(keep_score))
  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"

  precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])

  recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])

  F1 <- 2*precision*recall/(precision+recall)

  performance <- c(precision, recall, F1)
  names(performance) <- c("precision", "recall", "F1")

  return(performance)
}
```

```{r}
comp_pipelines <- tibble(pipeline="example", precision=0, recall=0, F1=0)
```

## just vibrant
```{r}
viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb)
```

## vibrant + checkV
```{r}
viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb_cv)
```

## vibrant + checkV + kaiju
```{r}
viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj)
```

## vibrant + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj_dvf)
```

## vibrant + virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vb_vs2_cv_kj)
```

#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_all)
```

## just virsorter2
```{r}
viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vs2)
```

## virsorter2 + checkV
```{r}
viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
```

## virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj_dvf)
```

## just checkV
```{r}
viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_cv)
```

## checkV + kaiju
```{r}
viruses$keep_score_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

assess_performance(viruses$seqtype, viruses$keep_score_cv_kj)
```

## just kaiju
```{r}
viruses$keep_score_k <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = F)

assess_performance(viruses$seqtype, viruses$keep_score_k)
```

## just deepvirfinder
```{r}
viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_dvf)
```

## deepvirfinder + checkV
```{r}
viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv)
```

## deepvirfinder + checkV + kaiju
```{r}
viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = F)

assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv_kj)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
```

## just virsorter
```{r}
viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs)
```

## virsorter + checkV
```{r}
viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv)
```

## virsorter + checkV + kaiju
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
```

## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
```

## virsorter + checkV + kaiju + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb)
```

## virsorter + checkV + kaiju + virsorter2 + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb_vs2)
```

## virsorter + checkV + kaiju + virsorter2
```{r}
viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vs2_dvf)
```

# Plot of how many methods identify a contig
```{r}
viruses_high <- viruses[viruses$keep_score_all>=1,]
viruses_high_mod <- viruses_high %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
viruses_high_mod <- as_tibble(viruses_high_mod)

sm <- as.data.frame(rbind(c(0,table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb_cv>0,]))),
                          c(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf_cv>0,]))),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2_cv>0,])),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_cv_kj>0,]))
                          )
)

sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "kaiju")

colnames(sm) <- c("one","two","three","all","method")

sm
```

```{r}
sm_m <- melt(sm)
colnames(sm_m) <- c("primary_method", "number_methods", "number")
```

```{r}
ggplot(sm_m, aes(x=primary_method, y=number,
                   fill=number_methods)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Number of Methods',
                     values = alpha(c(viridis(5)), 1)) +
  xlab("Primary Method") +
  ylab("Count of Viral Contigs")
```

# Visualizing which tools led to a contig being selected
(alluvial plot)
with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/

```{r}
alluvial_data <- viruses %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  apply(c(1,2), function(x) {if (x >= 1) {x <- 1} else if (x==0.5) {x <- 0.5} else {x <- 0}}) %>%
  as_tibble() %>%
  count(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
```

```{r}
alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  rowSums()
```

```{r}
alluvial_data$viral <- "not viral"
alluvial_data$viral[alluvial_data$cum_sum>=1] <- "one method"
alluvial_data$viral[alluvial_data$cum_sum>=2] <- "two methods"
alluvial_data$viral[alluvial_data$cum_sum>=3] <- "three methods"
alluvial_data$viral[alluvial_data$cum_sum>=4] <- "all methods"

alluvial_data$viral <- factor(alluvial_data$viral, levels=c("not viral", "one method",
                                                            "two methods", "three methods",
                                                            "all methods"))
```

```{r}
alluvial_data <- alluvial_data %>% 
  mutate(across(c(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj), as.factor))
```

```{r}
head(alluvial_data)
```

```{r}
library(ggalluvial)
```

```{r}
ggplot(data = alluvial_data,
       aes(axis1 = as.character(keep_score_vb_cv), 
           axis2 = as.character(keep_score_vs2_cv),
           axis3 = as.character(keep_score_dvf_cv),
           axis4 = as.character(keep_score_cv_kj),
           y = n)) +
  geom_alluvium(aes(fill = viral)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "VirSorter2", "DeepVirFinder", "Kaiju"),
                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
  ylab("Number of Sequences") +
  theme_classic()
```

# Ignoring True Negatives and visualizing flows
(alluvial plot)
with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/

```{r}
viruses$confusion_matrix <- "true negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_all<=1] <- "false negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
```

```{r}
table(viruses$confusion_matrix, viruses$keep_score_all)
```

```{r}
alluvial_data <- viruses[viruses$confusion_matrix!="true negative",] %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all, confusion_matrix) %>%
  count(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all, confusion_matrix)
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")

ggplot(data = alluvial_data,
       aes(axis1 = as.character(keep_score_vb_cv), 
           axis2 = as.character(keep_score_dvf_cv),
           axis3 = as.character(keep_score_vs2_cv),
           axis4 = as.character(keep_score_cv_kj),
           y = n)) +
  geom_alluvium(aes(fill = confusion_matrix)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_manual(name="",
                     values = alpha(pal(4)[c(4,3,1)]), 0.7,
                    labels=c("false negative", "false positive", "true positive")) + 
  scale_x_discrete(limits = c("VIBRANT", "DeepVirFinder", "VirSorter2", "Kaiju"),
                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
  ylab("Number of Sequences") +
  theme_classic()
```

visualizing confusion matrix by taxa
```{r}
confusion_by_taxa <- melt(table(viruses$confusion_matrix, viruses$seqtype))
colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","count")
```

```{r}
ggplot(confusion_by_taxa, aes(x=count, y=seqtype,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_bar(stat="identity") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") 
```

# Using NMDS to compare the tools
get matrix want to use for beta diversity
```{r}
virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_cv, keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all)

```

```{r}
virus_score_matrix <- virus_score_matrix[rowSums(virus_score_matrix)>0,]
```

```{r}
library(vegan)
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

example_NMDS <- metaMDS(t(virus_score_matrix), distance = "bray")

ss <- as.data.frame(example_NMDS$points)
ss$tool <- colnames(virus_score_matrix)

p <- ggplot() + 
  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
   theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text=element_text(size=14, colour="black"),
        axis.title=element_text(size=16),
        legend.key=element_blank(),
        legend.position="right",
        legend.background = element_blank()) +
  guides(size=FALSE) + xlab("NMDS 1") + ylab("NMDS 2") + 
  coord_equal() +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(nrow(ss))), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(nrow(ss))), 1)) +
  scale_shape_manual(name="",
                     values = c(20,21,21,22,22,23,23,24,24))

p
```

## Looking at False Positives
```{r}
viruses_false_positive <- viruses[viruses$confusion_matrix=="false positive",]
```

```{r}
table(viruses_false_positive$provirus)
```

```{r}
table(viruses_false_positive$vs2type)
```

```{r}
table(viruses_false_positive$percent_viral>50)
```

```{r}
table(viruses_false_positive$hallmark>5)
```

```{r}
table(viruses_false_positive$len.x<=10000)
```

```{r}
table(viruses_false_positive$seqtype)
```

### 2022-07-01 Scripts/Evaluations/EarlyIterations/identifying_viral_contigs_20220515.Rmd
Reorganizing scripts.

--- 2022-05-15 Scripts/Evaluations/identifying_viral_contigs_20220515.Rmd
@@
 viruses$keep_score_vibrant[viruses$Quality=="medium quality draft"] <- 0.52
-viruses$keep_score_vibrant[viruses$Quality=="low quality draft"] <- 0.333
+viruses$keep_score_vibrant[viruses$Quality=="low quality draft"] <- 0.52
 viruses$keep_score_vibrant[viruses$viral_genes==0 & viruses$host_genes>=1] <- 0
@@
 #virsorter2
-viruses$keep_score_virsorter2[viruses$viral>=50] <- 0.333
+viruses$keep_score_virsorter2[viruses$viral>=50] <- 0.52
 viruses$keep_score_virsorter2[viruses$hallmark>2] <- 1
@@
 ```
+
+# Visualizing which tools led to a contig being selected
+(alluvial plot)
+with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/
+
+```{r}
+alluvial_data <- viruses %>% select(keep_score_vibrant:keep_score_checkV) %>%
+  count(keep_score_vibrant, keep_score_virsorter2, keep_score_deepvirfinder, keep_score_checkV)
+```
+
+```{r}
+alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vibrant:keep_score_checkV) %>% 
+  rowSums()
+```
+
+```{r}
+alluvial_data$viral <- "not viral"
+alluvial_data$viral[alluvial_data$cum_sum>=1] <- "one method"
+alluvial_data$viral[alluvial_data$cum_sum>=2] <- "two methods"
+alluvial_data$viral[alluvial_data$cum_sum>=3] <- "three methods"
+alluvial_data$viral[alluvial_data$cum_sum>=4] <- "all methods"
+
+alluvial_data$viral <- factor(alluvial_data$viral, levels=c("not viral", "one method",
+                                                            "two methods", "three methods",
+                                                            "all methods"))
+```
+
+```{r}
+alluvial_data <- alluvial_data %>% 
+  mutate(across(keep_score_vibrant:keep_score_checkV, as.factor))
+```
+
+```{r}
+head(alluvial_data)
+```
+
+```{r}
+library(ggalluvial)
+```
+
+```{r}
+ggplot(data = alluvial_data,
+       aes(axis1 = as.character(keep_score_vibrant), 
+           axis2 = as.character(keep_score_virsorter2),
+           axis3 = as.character(keep_score_deepvirfinder),
+           axis4 = as.character(keep_score_checkV),
+           y = n)) +
+  geom_alluvium(aes(fill = viral)) +
+  geom_stratum() +
+  geom_text(stat = "stratum",
+            aes(label = after_stat(stratum))) +
+  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "VirSorter2", "DeepVirFinder", "CheckV"),
+                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
+  ylab("Number of Sequences") +
+  theme_classic()
+```
+
+#########################################################
+# Slightly modified rules for visualization purposes
+
+######################################################################
+# New rules for keeping
+
+## Getting Score for Deciding Whether to Keep
+```{r}
+viruses$keep_score_vibrant <- 0
+viruses$keep_score_deepvirfinder <- 0
+viruses$keep_score_virsorter2 <- 0
+viruses$keep_score_checkV <- 0
+
+#vibrant
+viruses$keep_score_vibrant[viruses$Quality=="high quality draft"] <- 1
+viruses$keep_score_vibrant[viruses$Quality=="medium quality draft"] <- 0.5
+viruses$keep_score_vibrant[viruses$Quality=="low quality draft"] <- 0.5
+
+#deepvirfinder
+viruses$keep_score_deepvirfinder[viruses$bh_pvalue<=0.05] <- 1
+
+#virsorter2
+viruses$keep_score_virsorter2[viruses$viral>=50] <- 0.5
+viruses$keep_score_virsorter2[viruses$hallmark>2] <- 1
+viruses$keep_score_virsorter2[viruses$viral>=95] <- viruses$keep_score_virsorter2[viruses$viral>=95] + 1
+
+#checkV
+viruses$keep_score_checkV[(viruses$viral>=50 & viruses$host_genes==0) & viruses$viral_genes==0] <- viruses$keep_score_checkV[(viruses$viral>=50 & viruses$host_genes==0) & viruses$viral_genes==0] + 0.5
+viruses$keep_score_checkV[viruses$percent_viral>=50] <- viruses$keep_score_checkV[viruses$percent_viral>=50] + 1
+viruses$keep_score_checkV[viruses$viral_genes==0 & viruses$host_genes>=1] <- -5
+viruses$keep_score_checkV[(viruses$viral_genes*3) <= viruses$host_genes] <- -5
+viruses$keep_score_checkV[viruses$contig_length>50000 & viruses$hallmark==0] <- -5
+```
+
+# Visualizing which tools led to a contig being selected
+(alluvial plot)
+with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/
+
+```{r}
+alluvial_data <- viruses %>% select(keep_score_vibrant:keep_score_checkV) %>%
+  count(keep_score_vibrant, keep_score_virsorter2, keep_score_deepvirfinder, keep_score_checkV)
+```
+
+```{r}
+alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vibrant:keep_score_checkV) %>% 
+  rowSums()
+```
+
+```{r}
+alluvial_data$viral <- "not viral"
+alluvial_data$viral[alluvial_data$cum_sum>=1] <- "one method"
+alluvial_data$viral[alluvial_data$cum_sum>=2] <- "two methods"
+alluvial_data$viral[alluvial_data$cum_sum>=3] <- "three methods"
+alluvial_data$viral[alluvial_data$cum_sum>=4] <- "all methods"
+
+alluvial_data$viral <- factor(alluvial_data$viral, levels=c("not viral", "one method",
+                                                            "two methods", "three methods",
+                                                            "all methods"))
+```
+
+```{r}
+alluvial_data <- alluvial_data %>% 
+  mutate(across(keep_score_vibrant:keep_score_checkV, as.factor))
+```
+
+```{r}
+head(alluvial_data)
+```
+
+```{r}
+ggplot(data = alluvial_data,
+       aes(axis1 = as.character(keep_score_vibrant), 
+           axis2 = as.character(keep_score_virsorter2),
+           axis3 = as.character(keep_score_deepvirfinder),
+           axis4 = as.character(keep_score_checkV),
+           y = n)) +
+  geom_alluvium(aes(fill = viral)) +
+  geom_stratum() +
+  geom_text(stat = "stratum",
+            aes(label = after_stat(stratum))) +
+  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "VirSorter2", "DeepVirFinder", "CheckV"),
+                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
+  ylab("Number of Sequences") +
+  theme_classic()
+```
+
+=> once have the testing set data, make this figure but split by type of sequence
+
+# Other visualizations of output
 

### 2022-07-01 Scripts/Evaluations/EarlyIterations/viral_contig_sorting_testing_set_20220521_to_share.Rmd
Reorganizing scripts.

--- 2022-06-01 Scripts/Evaluations/identifying_viral_contigs_20220521.Rmd
@@
 ```{r}
-kj_c <- read_tsv("../VSTE/ToolOutput/metagenomic_testing_set_nrandeuk.kaiju.names.out", col_names = F)
+kj_c <- read_tsv("../TestingSetOutput/metagenomic_testing_set_nrandeuk.kaiju.names.out", col_names = F)
 colnames(kj_c) <- c("Classified", "Contig", "NCBI_taxon", "len", "ID_best", "IDs_all", "Seq", "Name")
@@
 
-## VirSorter
-```{r}
-vs_c <- read_csv("../TestingSetOutput/VIRSorter_global-phage-signal.csv")
-vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
-vs_c <- separate(vs_c, col = `## Contig_id`, into = c("seqtype", "contig"), sep="--")
-vs_c$contig <- sub("-circular", "", vs_c$contig)
-vs_c <- vs_c[!duplicated(vs_c$contig),]
-```
-
 ## Merging
@@
 viruses <- full_join(x=viruses, y=vb_c, by = c("contig", "seqtype"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("contig", "seqtype"))
 viruses <- full_join(x=viruses, y=kj_c, by = c("contig", "seqtype"))
@@
 viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
-viruses$Category[is.na(viruses$Category)] <- 0
 ```
@@
 ```{r}
-getting_viral_set_1 <- function(input_seqs, include_vibrant=FALSE, include_virsorter2=FALSE,
-                              include_deepvirfinder=FALSE, include_checkV=FALSE,
-                              include_kaiju=FALSE, include_virsorter=FALSE) {
+getting_viral_set_1 <- function(input_seqs, include_vibrant, include_virsorter2,
+                              include_deepvirfinder, include_checkV,
+                              include_kaiju) {
 
@@
 
-  if (include_virsorter) {
-    keep_score[input_seqs$Category==1] <- keep_score[input_seqs$Category==1] + 1
-    keep_score[input_seqs$Category==2] <- keep_score[input_seqs$Category==2] + 1
-    keep_score[input_seqs$Category==3] <- keep_score[input_seqs$Category==3] + 0.5
-    keep_score[input_seqs$Category==4] <- keep_score[input_seqs$Category==4] + 1
-    keep_score[input_seqs$Category==5] <- keep_score[input_seqs$Category==5] + 1
-    keep_score[input_seqs$Category==6] <- keep_score[input_seqs$Category==6] + 0.5
-  }
-
   if (include_deepvirfinder) {
     keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
-#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
   }
@@
                                             include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = F)
+                                            include_deepvirfinder = F)
 
@@
 
-## deepvirfinder + checkV + kaiju
-```{r}
-viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = F)
-
-assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv_kj)
-```
-
 ## virsorter2 + checkV + kaiju
@@
 assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
-```
-
-## just virsorter
-```{r}
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs)
-```
-
-## virsorter + checkV
-```{r}
-viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs_cv)
-```
-
-## virsorter + checkV + kaiju
-```{r}
-viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
-```
-
-## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
-```{r}
-viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
-```
-
-## virsorter + checkV + kaiju + vibrant
-```{r}
-viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb)
-```
-
-## virsorter + checkV + kaiju + virsorter2 + vibrant
-```{r}
-viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb_vs2)
-```
-
-## virsorter + checkV + kaiju + virsorter2
-```{r}
-viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-
-assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vs2_dvf)
 ```
@@
 ```
-
-# Using NMDS to compare the tools
-get matrix want to use for beta diversity
-```{r}
-virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_cv, keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all)
-
-```
-
-```{r}
-virus_score_matrix <- virus_score_matrix[rowSums(virus_score_matrix)>0,]
-```
-
-```{r}
-library(vegan)
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
-
-example_NMDS <- metaMDS(t(virus_score_matrix), distance = "bray")
-
-ss <- as.data.frame(example_NMDS$points)
-ss$tool <- colnames(virus_score_matrix)
-
-p <- ggplot() + 
-  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
-   theme(panel.background = element_blank(), 
-        axis.line = element_line(colour = "black"),
-        panel.border = element_rect(colour = "black", fill=NA, size=1),
-        axis.text=element_text(size=14, colour="black"),
-        axis.title=element_text(size=16),
-        legend.key=element_blank(),
-        legend.position="right",
-        legend.background = element_blank()) +
-  guides(size=FALSE) + xlab("NMDS 1") + ylab("NMDS 2") + 
-  coord_equal() +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(nrow(ss))), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(nrow(ss))), 1)) +
-  scale_shape_manual(name="",
-                     values = c(20,21,21,22,22,23,23,24,24))
-
-p
-```
-
-## Looking at False Positives
-```{r}
-viruses_false_positive <- viruses[viruses$confusion_matrix=="false positive",]
-```
-
-```{r}
-table(viruses_false_positive$provirus)
-```
-
-```{r}
-table(viruses_false_positive$vs2type)
-```
-
-```{r}
-table(viruses_false_positive$percent_viral>50)
-```
-
-```{r}
-table(viruses_false_positive$hallmark>5)
-```
-
-```{r}
-table(viruses_false_positive$len.x<=10000)
-```
-
-```{r}
-table(viruses_false_positive$seqtype)
-```

### 2022-07-26 identifying_viral_contigs_environmental.Rmd
added most updated rules
Made specifically for running on environmental samples

---
title: "Viral Sequence Sorting Tools Evaluation"
author: Bridget Hegarty, James Riddell
date: 07-22-2022
output: html_notebook
---
This Rmarkdown file takes outputs from viral identification tools and merges
them into one file, "{dataset}_merged_viral_id_outputs.csv", which is the input
for identifying_viral_metrics_environmental.Rmd.

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (email) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

# Inputs
1) For this script's inputs, the user must merge each viral identification 
tool output with the following specified filenames into one tab-separated file. 
Replace the filenames with their paths if needed into the variables in the 
'inputs' chunk.

    General filenames for each tool:
        Checkv: quality_summary.tsv
        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
        Virsorter: VIRSorter_global-phage-signal.csv
        Virsorter2: final-viral-score.tsv
        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
        Kaiju: ${assembly}.kaiju.names.out

2) Make sure all fasta headers (assembly, contig) are consistent within and
across tools. Chunks for each tool do contain some lines for cleaning up these features, but due to their variability it will be the user's responsibility to 
make sure they match across tools.

3) Check each chunk and ensure all columns are accounted for. For example,
if you don't have "assembly" and called it something else, please either
remove it from this script or modify your input files.

4) This script is designed for contigs > 3000 basepairs. It can be modified
to be higher or lower, but going lower will greatly increase the size of the
dataframe and memory usage.

```{r inputs}
# dataset name for file organization and outputs for metrics file
dataset_name <- "lake_michigan"
# checkV
checkV_path <- "merged_lake_michigan_checkv_assemblies.tsv"
# VIBRANT
vibrant_path <- "merged_lake_michigan_vibrant_assemblies.tsv"
# DeepVirFinder
dvf_path <- "merged_lake_michigan_dvf_assemblies.tsv"
# Virsorter
vs_path <- "merged_lake_michigan_virsorter_assemblies.tsv"
# Virsorter2
vs2_path <- "merged_lake_michigan_vs2_assemblies.tsv"
# Kaiju
kj_path <- "merged_lake_michigan_kaiju_assemblies.tsv"
# KB cutoff
KB_CUTOFF <- 3000

```

# All Viral Contigs

## checkV
```{r}
checkV <- fread(checkV_path, 
                sep="\t",
                header = T, 
                select = c(
                    'contig_id',
                    'assembly',
                    'provirus',
                    'completeness',
                    'contamination',
                    'viral_genes',
                    'host_genes',
                    'gene_count',
                    'contig_length',
                    'checkv_quality'
                    )
                ) %>% 
    rename(
        contig = contig_id,
        checkv_provirus = provirus,
        checkv_completeness = completeness,
        checkv_contamination = contamination,
        checkv_viral_genes = viral_genes,
        checkv_host_genes = host_genes,
        checkv_total_genes = gene_count,
        checkv_length = contig_length
        )
checkV$method = 'checkv'
checkV$contig <- gsub('_', '', checkV$contig)
checkV$checkv_uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
```

## VIBRANT
```{r}
vb_c <- fread(vibrant_path,
              header = T,
              sep = "\t",
              select = c(
                  'scaffold',
                  'assembly',
                  'type',
                  'Quality'
                  )
              ) %>%
    rename(
        contig = scaffold,
        vibrant_quality = Quality
        )
vb_c$method <- "vibrant"
vb_c$vibrant_prophage <- "No"
vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c$contig <- gsub('_', '', vb_c$contig)
vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
vb_c <- vb_c[!duplicated(vb_c$uniq_contig),]
```

## DeepVirFinder
```{r}
dvf_c <- fread(dvf_path,
               header = T,
               sep = "\t",
               select = c(
                   'name',
                   'assembly',
                   'score',
                   'pvalue'
                   )
               ) %>% 
    rename(
        contig = name
        )
dvf_c$contig <- gsub('_', '', dvf_c$contig)
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
```

## VirSorter
```{r}
vs_c <- fread(vs_path,
              select = c(
                  '## Contig_id',
                  'Nb genes contigs',
                  'Nb genes',
                  'Category',
                  'Nb phage hallmark genes'
                  )
              ) %>% drop_na(`## Contig_id`)
vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
vs_c$`## Contig_id` <- sub("-circular", "", vs_c$`## Contig_id`)
vs_c$`## Contig_id` <- gsub("_", "", vs_c$`## Contig_id`)
vs_c <- separate(vs_c, col = `## Contig_id`, into = c("assembly", "contig"), sep="--")

vs_c <- vs_c[!duplicated(vs_c$contig),]
vs_c <- vs_c %>% drop_na(contig)
```

## VirSorter2 
```{r}
vs2_c <- fread(vs2_path,
                    header = T,
                    sep = '\t',
                    select = c(
                        'seqname',
                        'assembly',
                        'dsDNAphage',
                        'RNA',
                        'ssDNA',
                        'max_score',
                        'max_score_group',
                        'hallmark',
                        'viral',
                        'cellular'
                    )
               ) %>% 
              separate(
                  col = seqname,
                  into = c("contig", "type"), 
                  sep = "\\|\\|",
                  remove = T
                  )
vs2_c$contig <- gsub("_", "", vs2_c$contig)
vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
vs2_c <- vs2_c[!duplicated(vs2_c$uniq_contig),]
```

## Kaiju
(parsing when reading the file in "fails" for sequences that weren't classified)
```{r}
kj_c <- fread(kj_path)

colnames(kj_c) <-c(
    "Classified",
    "Contig",
    "NCBI_taxon",
    "len",
    "ID_best",
    "IDs_all",
    "Seq",
    "Name"
    )
kj_c$Contig <- gsub('_', '', kj_c$Contig)
kj_c <- setDT(kj_c)[, c("assembly", "contig") := tstrsplit(Contig, "--")]
kj_c <- setDT(kj_c)[, c("Kaiju_Viral", "Kingdom") := tstrsplit(Name, ";", keep = c(1, 2))]
```

```{r}
# Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
table(kj_c$assembly, kj_c$Kaiju_Viral)
```

## Merging

```{r}
viruses <- full_join(x=vs2_c, y=dvf_c, by = c("contig", "assembly"))
viruses <- full_join(x=viruses, y=checkV, by = c("contig", "assembly"))
viruses <- full_join(x=viruses, y=vb_c, by = c("contig", "assembly"))
viruses <- full_join(x=viruses, y=vs_c, by = c("contig", "assembly"))
viruses <- full_join(x=viruses, y=kj_c, by = c("contig", "assembly"))
```

Remove contigs not greater than the basepair length cutoff (3000)

```{r}
viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
```

Use the next code chunk to check if any contigs are missing from checkv.
If yes, then there is a mix up in contig names since checkv should contain
all contig names even if they are identified as non-viral.
```{r}
v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
```

get rid of NAs for downstream processing
```{r}

# checkV
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0

# VIBRANT
viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0

# DeepVirFinder
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0

# VirSorter2
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0

#Virsorter
viruses$Category[is.na(viruses$Category)] <- 0

# Kaiju
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"

```

```{r}
viruses_filename <- 'merged_viral_id_outputs.csv'
write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
```

Data cleaning is done! Export viruses to tsv and load into identifying_viral_metrics.Rmd for metrics on tool outputs. The code for metrics is also pasted below, but it is not recommended running it in the same R session due to memory issues. The construction of the final dataframe essentially doubles the memory being used, and even with garbage collector can make Rstudio unresponsive.

---
title: "identifying_viral_metrics_environmental"
author: Bridget Hegarty, James Riddell
date: 7/23/2022
output: html_document
---

TODO: ADD ENVIRONMENTAL METRICS LIKE SHANNON, ALPHA DIVERSITY.

This section defines a viralness score "keep_score" based on the tool classifications. 
A final keep_score above 1 indicates we will keep that sequence and call it viral.

VIBRANT
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus == TRUE: +0.5

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +1
    Hallmark > 2: +1

Virsorter
    Category <= 5: +1
    Category == 6: +0.5

DeepVirFinder:
    Score >= 0.7: +0.5

Kaiju:
    Kaiju_viral = "cellular organisms": -1
    Kaiju_viral = "Viruses": +1

CheckV
    viral_genes == 0 and host_genes >= 1: keep_score = 0
    If 3*viral_genes <= host_genes: keep_score = 0
    If length > 50,000 and hallmark == 0: keep_score = 0
    If %unknown >= 75: +0.5

This script produces visualizations of these combined viral scorings and
includes ecological metrics like alpha diversity.

Users can decide which combination is appropriate for them and only need to run
the chunks for the appropriate tools.

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_kaiju=FALSE,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$Category==1] <- keep_score[input_seqs$Category==1] + 1
    keep_score[input_seqs$Category==2] <- keep_score[input_seqs$Category==2] + 1
    keep_score[input_seqs$Category==3] <- keep_score[input_seqs$Category==3] + 0.5
    keep_score[input_seqs$Category==4] <- keep_score[input_seqs$Category==4] + 1
    keep_score[input_seqs$Category==5] <- keep_score[input_seqs$Category==5] + 1
    keep_score[input_seqs$Category==6] <- keep_score[input_seqs$Category==6] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
  }

  if (include_kaiju) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
  }

  return(keep_score)

}

```

## just vibrant
```{r}
viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb)
```

## vibrant + checkV
```{r}
viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv)
```

## vibrant + checkV + kaiju
```{r}
viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj)
```

## vibrant + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

# # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj_dvf)
```

## vibrant + virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

# # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_vs2_cv_kj)
```

#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_all)
```

## just virsorter2
```{r}
viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2)
```

## virsorter2 + checkV
```{r}
viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
```

## virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj_dvf)
```

## just checkV
```{r}
viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_cv)
```

## checkV + kaiju
```{r}
viruses$keep_score_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_cv_kj)
```

## just kaiju
```{r}
viruses$keep_score_k <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_k)
```

## just deepvirfinder
```{r}
viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf)
```

## deepvirfinder + checkV
```{r}
viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv)
```

## deepvirfinder + checkV + kaiju
```{r}
viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = F)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv_kj)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
```

## just virsorter
```{r}
viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs)
```

## virsorter + checkV
```{r}
viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv)
```

## virsorter + checkV + kaiju
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
```

## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
```{r}
viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
```

## virsorter + checkV + kaiju + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb)
```

## virsorter + checkV + kaiju + virsorter2 + vibrant
```{r}
viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb_vs2)
```

## virsorter + checkV + kaiju + virsorter2
```{r}
viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T,
                                            include_virsorter = T)

# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vs2_dvf)
```

 TODO WHAT DO I DO ABOUT THE ZEROS????
# Plot of how many methods identify a contig
```{r}
viruses_high <- viruses[viruses$keep_score_all>=1,]
viruses_high_mod <- viruses_high %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_vs_cv, keep_score_cv_kj)
viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
viruses_high_mod <- as_tibble(viruses_high_mod)

virusHighCount <- function(df, tool_combo) {
    tb <- rowSums(df[df[,tool_combo]>0,])
    if (length(tb) == 0) {
        return (0)
    }
    else {
        return (table(tb))
    }
}

sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb_cv>0,])),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf_cv>0,])),
                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_cv_kj>0,]))
                          )
)

sm <- as.data.frame(
    rbind(
        virusHighCount(viruses_high_mod, "keep_score_vb_cv"),
        virusHighCount(viruses_high_mod, "keep_score_dvf_cv"),
        virusHighCount(viruses_high_mod, "keep_score_vs2_cv"),
        virusHighCount(viruses_high_mod, "keep_score_vs_cv"),
        virusHighCount(viruses_high_mod, "keep_score_cv_kj")
    )
)

sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "kaiju")
sm
```

Create list of contigs considered viral
```{r}
c_list <- data.frame(viruses_high$checkv_uniq_contig)
write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
```

Export keep_score_high and list of contigs considered viral by the combined tool.

```{r}
write_tsv(viruses_high, paste(dataset_name, "viruses_high.tsv", sep='_'))
```

```{r}
sm_m <- reshape2::melt(sm)
colnames(sm_m) <- c("primary_method", "number_methods", "number")
```

```{r}
ggplot(sm_m, aes(x=primary_method, y=number,
                   fill=number_methods)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Number of Methods',
                     values = alpha(c(viridis(5)), 1)) +
  xlab("Primary Method") +
  ylab("Count of Viral Contigs")
```

# Visualizing which tools led to a contig being selected
(alluvial plot)
with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/

```{r}
alluvial_data <- viruses %>% select(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  apply(c(1,2), function(x) {if (x >= 1) {x <- 1} else if (x==0.5) {x <- 0.5} else {x <- 0}}) %>%
  as_tibble() %>%
  count(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
```

```{r}
alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv, keep_score_vs2_cv, keep_score_cv_kj) %>% 
  rowSums()
```

```{r}
alluvial_data$keep_score <- "0-0.5"
alluvial_data$keep_score[alluvial_data$cum_sum>=1] <- "1-1.5"
alluvial_data$keep_score[alluvial_data$cum_sum>=2] <- "2-2.5"
alluvial_data$keep_score[alluvial_data$cum_sum>=3] <- "3-3.5"
alluvial_data$keep_score[alluvial_data$cum_sum>=4] <- "4-4.5"
alluvial_data$keep_score[alluvial_data$cum_sum>=5] <- "5+"

alluvial_data$keep_score <- factor(
    alluvial_data$keep_score,
    levels=c(
        "0-0.5",
        "1-1.5",
        "2-2.5",
        "3-3.5",
        "4-4.5",
        "5+"
        )
    )
```

```{r}
alluvial_data <- alluvial_data %>% 
  mutate(
      across(
          c(keep_score_vb_cv,
            keep_score_vs_cv,
            keep_score_dvf_cv,
            keep_score_vs2_cv,
            keep_score_cv_kj
            ),
          as.factor
          )
      )
```

```{r}
head(alluvial_data)
```

```{r}
library(ggalluvial)
```

```{r}
ggplot(data = alluvial_data,
       aes(axis1 = as.character(keep_score_vb_cv), 
           axis2 = as.character(keep_score_vs_cv),
           axis3 = as.character(keep_score_vs2_cv),
           axis4 = as.character(keep_score_dvf_cv),
           axis5 = as.character(keep_score_cv_kj),
           y = n)) +
  geom_alluvium(aes(fill = keep_score)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "Virsorter", "VirSorter2", "DeepVirFinder", "Kaiju"),
                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
  ylab("Number of Sequences") +
  xlab("method")
  theme_classic()
```

Get proportion of contigs from "viruses" classified as virus by combined tool:

```{r}
prop_viral <- paste(nrow(viruses_high) / nrow(viruses) * 100, "%", sep="")
prop_viral
```

Distribution of keep_score across "viruses" dataframe.
```{r}
keep_scores <- reshape2::melt(table(viruses$keep_score_all))

colnames(keep_scores) <- c('keep_score', 'count')
keep_scores[,'keep_score'] <- as.factor(keep_scores[,'keep_score'])

ggplot(data = keep_scores, aes(x = keep_score, y = count)) +
    geom_bar(stat="identity", width = 0.3, fill = 'steelblue')
```

Determine number of viral sequences with kaiju taxonomic classification

```{r}
tax_counts <- viruses_high %>% count(Kaiju_Viral, name = "num")
tax_counts$prop <- tax_counts$num / sum(tax_counts$num)
```

### 2022-08-10 Data/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Created folder for testing pipeline on DW.

--- 2022-07-26 identifying_viral_contigs_environmental.Rmd
@@
 Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
-Bridget Hegarty (email) regarding any issues, or open an issue on github.
+Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.
 
@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "lake_michigan"
+dataset_name <- "drinking_water"
 # checkV
-checkV_path <- "merged_lake_michigan_checkv_assemblies.tsv"
+checkV_path <- "merged_drinking_water_checkv_assemblies.tsv"
 # VIBRANT
-vibrant_path <- "merged_lake_michigan_vibrant_assemblies.tsv"
+vibrant_path <- "../Results/VIBRANT_genome_quality_contigs_3000.tsv"
 # DeepVirFinder

### 2022-08-10 Data/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updated iding viral contigs to work for VB, CV, VS, VS2.

@@
 Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.
+
+TO DO:
+- add kaiju and DVF
 
@@
 # checkV
-checkV_path <- "merged_drinking_water_checkv_assemblies.tsv"
+checkV_path <- "DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
 # VIBRANT
-vibrant_path <- "../Results/VIBRANT_genome_quality_contigs_3000.tsv"
+vibrant_path <- "DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
 # DeepVirFinder
-dvf_path <- "merged_lake_michigan_dvf_assemblies.tsv"
+dvf_path <- "DrinkingWaterToolOutput/merged_lake_michigan_dvf_assemblies.tsv"
 # Virsorter
-vs_path <- "merged_lake_michigan_virsorter_assemblies.tsv"
+vs_path <- "DrinkingWaterToolOutput/vs_merged_all.txt"
 # Virsorter2
-vs2_path <- "merged_lake_michigan_vs2_assemblies.tsv"
+vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
 # Kaiju
-kj_path <- "merged_lake_michigan_kaiju_assemblies.tsv"
+kj_path <- "DrinkingWaterToolOutput/merged_lake_michigan_kaiju_assemblies.tsv"
 # KB cutoff
@@
 checkV$method = 'checkv'
-checkV$contig <- gsub('_', '', checkV$contig)
-checkV$checkv_uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
+checkV$contig <- sub("\\.", "_", checkV$contig)
+checkV$uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
 ```
@@
                   'type',
-                  'Quality'
+                  'quality'
                   )
@@
         contig = scaffold,
-        vibrant_quality = Quality
+        vibrant_quality = quality
         )
@@
 vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
-vb_c$contig <- gsub('_', '', vb_c$contig)
+vb_c$contig <- sub("\\.", "_", vb_c$contig)
 vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
@@
 
-## DeepVirFinder
+## DeepVirFinder - to run
 ```{r}
@@
         )
-dvf_c$contig <- gsub('_', '', dvf_c$contig)
+dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
+
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
@@
               select = c(
-                  '## Contig_id',
-                  'Nb genes contigs',
-                  'Nb genes',
-                  'Category',
-                  'Nb phage hallmark genes'
-                  )
-              ) %>% drop_na(`## Contig_id`)
-vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
-vs_c$`## Contig_id` <- sub("-circular", "", vs_c$`## Contig_id`)
-vs_c$`## Contig_id` <- gsub("_", "", vs_c$`## Contig_id`)
-vs_c <- separate(vs_c, col = `## Contig_id`, into = c("assembly", "contig"), sep="--")
+                  'assembly',
+                  'contig_id',
+                  'category'
+                  ),
+              ) %>% 
+  rename(contig = contig_id) %>%
+  drop_na(contig)
+vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
+vs_c$contig <- sub("-circula", "", vs_c$contig)
+vs_c$contig <- sub("uth", "South", vs_c$contig)
+vs_c$contig <- sub("pannenburg", "Spannenburg", vs_c$contig)
+vs_c$contig <- sub("\\.", "_", vs_c$contig)
+vs_c$uniq_contig <- paste(vs_c$assembly, vs_c$contig, sep = "--")
 
-vs_c <- vs_c[!duplicated(vs_c$contig),]
-vs_c <- vs_c %>% drop_na(contig)
+vs_c <- vs_c[!duplicated(vs_c$uniq_contig),]
+vs_c <- vs_c %>% drop_na(uniq_contig)
 ```
@@
                         'dsDNAphage',
-                        'RNA',
                         'ssDNA',
@@
                   )
-vs2_c$contig <- gsub("_", "", vs2_c$contig)
+vs2_c$assembly <- sub("_3000", "", vs2_c$assembly)
+vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
 vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
@@
 
-## Kaiju
+## Kaiju - to run
 (parsing when reading the file in "fails" for sequences that weren't classified)
@@
 kj_c$Contig <- gsub('_', '', kj_c$Contig)
+kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
+
 kj_c <- setDT(kj_c)[, c("assembly", "contig") := tstrsplit(Contig, "--")]
@@
 ## Merging
-
+to do: this step isn't working properly
 ```{r}
-viruses <- full_join(x=vs2_c, y=dvf_c, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=checkV, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=vb_c, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=kj_c, by = c("contig", "assembly"))
+viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
+viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
+viruses <- full_join(x=viruses, y=vb_c, by = c("uniq_contig", "assembly"))
+viruses <- full_join(x=viruses, y=vs_c, by = c("uniq_contig", "assembly"))
+viruses <- full_join(x=viruses, y=kj_c, by = c("uniq_contig", "assembly"))
 ```
@@
 #Virsorter
-viruses$Category[is.na(viruses$Category)] <- 0
+viruses$category[is.na(viruses$category)] <- 0
 
@@
 Virsorter
-    Category <= 5: +1
-    Category == 6: +0.5
+    category <= 5: +1
+    category == 6: +0.5
 
@@
   if (include_vibrant) {
-    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
-    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
-    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
   }
@@
   if (include_virsorter) {
-    keep_score[input_seqs$Category==1] <- keep_score[input_seqs$Category==1] + 1
-    keep_score[input_seqs$Category==2] <- keep_score[input_seqs$Category==2] + 1
-    keep_score[input_seqs$Category==3] <- keep_score[input_seqs$Category==3] + 0.5
-    keep_score[input_seqs$Category==4] <- keep_score[input_seqs$Category==4] + 1
-    keep_score[input_seqs$Category==5] <- keep_score[input_seqs$Category==5] + 1
-    keep_score[input_seqs$Category==6] <- keep_score[input_seqs$Category==6] + 0.5
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
   }
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj)
 ```
@@
                                             include_deepvirfinder = T)
-
-# # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj_dvf)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_vs2_cv_kj)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_all)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
 ```
@@
                                             include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj_dvf)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_cv)
 ```
@@
                                             include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_cv_kj)
 ```
@@
 ```{r}
-viruses$keep_score_k <- getting_viral_set_1(viruses, include_vibrant = F,
+viruses$keep_score_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                             include_virsorter2 = F,
@@
                                             include_virsorter = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_k)
 ```
@@
                                             include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf)
 ```
@@
                                             include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv)
 ```
@@
                                             include_virsorter = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv_kj)
 ```
@@
                                             include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb_vs2)
 ```
@@
                                             include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vs2_dvf)
 ```
@@
   ylab("Number of Sequences") +
-  xlab("method")
-  theme_classic()
+  xlab("method") +
+  theme_bw()
 ```

### 2022-08-10 identifying_viral_contigs_environmental.Rmd
Fixed bug that lost all vibrant sequences.

@@
 VIBRANT
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus == TRUE: +0.5
+    vibrant_quality == "High Quality Draft": +1
+    vibrant_quality == "Medium Quality Draft": +1
+    vibrant_quality == "Low Quality Draft" & provirus == TRUE: +0.5
 

### 2022-08-10 EnvironmentalData/EarlyIterations/example_envsamples_viralseqsorting_v20220524.Rmd
Updated directory structure.

---
title: "Viral Sequence Sorting A2"
output: html_notebook
---

based on VSTE v 20220524

```{r}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
```

# All Viral Contigs

## checkV
```{r}
checkV <- read_tsv("../Results/contamination.tsv")
checkV$contig <- sub("\\.", "_", checkV$contig_id)
```

## VIBRANT
```{r}
vb_c <- read_tsv("../Results/VIBRANT_genome_quality_contigs_3000.tsv")
vb_c$contig <- sub("\\.", "_", vb_c$scaffold)
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c <- vb_c[!duplicated(vb_c$contig),]
```

## DeepVirFinder
```{r}
dvf_c <- read_tsv("../Results/contigs_3000.fasta_gt1bp_dvfpred.txt")
dvf_c$contig <- sub("\\.", "_", dvf_c$name)
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
```

## VirSorter2 
```{r}
vs2_c <- read_tsv("../Results/merged_final_viral_score_A2.tsv")
vs2_c <- separate(vs2_c, col = seqname, into = c("contig", "vs2type"), "\\|\\|")
vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
```

## Kaiju
(parsing when reading the file in "fails" for sequences that weren't classified)
```{r}
kj_c <- read_tsv("../Results/A2.nr_euk.kaiju.names.out", col_names = F)
colnames(kj_c) <- c("Classified", "Contig", "NCBI_taxon", "len", "ID_best", "IDs_all", "Seq", "Name")
kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
kj_c$contig <- sub("\\.", "_", kj_c$Contig)
```

```{r}
table(kj_c$seqtype, kj_c$Kaiju_Viral)
```

## Merging

```{r}
viruses <- full_join(x=vs2_c, y=dvf_c, by = c("contig"))
viruses <- full_join(x=viruses, y=checkV, by = c("contig"))
viruses <- full_join(x=viruses, y=vb_c, by = c("contig"))
viruses <- full_join(x=viruses, y=kj_c, by = c("contig"))
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$host_genes/viruses$total_genes*100
viruses$percent_viral <- viruses$viral_genes/viruses$total_genes*100
viruses$percent_unknown <- 100-(viruses$host_genes+viruses$viral_genes)/viruses$total_genes*100
```

get rid of NAs for downstream processing
```{r}
viruses$Quality[is.na(viruses$Quality)] <- 0
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
```

##################################################

```{r}
getting_viral_set_1 <- function(input_seqs, include_vibrant=FALSE, include_virsorter2=FALSE,
                              include_deepvirfinder=FALSE, include_checkV=FALSE,
                              include_kaiju=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
  }

  if (include_kaiju) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
  }

  return(keep_score)

}

```

## just vibrant
```{r}
viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## vibrant + checkV
```{r}
viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## vibrant + checkV + kaiju
```{r}
viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## vibrant + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## vibrant + virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## just virsorter2
```{r}
viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## virsorter2 + checkV
```{r}
viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## virsorter2 + checkV + kaiju + deepvirfinder
```{r}
viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = T,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## just checkV
```{r}
viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = F)
```

## just kaiju
```{r}
viruses$keep_score_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## checkV + kaiju
```{r}
viruses$keep_score_kj_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = F)
```

## just deepvirfinder
```{r}
viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = F,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)
```

## deepvirfinder + checkV
```{r}
viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = F,
                                            include_deepvirfinder = T)
```

## deepvirfinder + checkV + kaiju
```{r}
viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

## virsorter2 + checkV + kaiju
```{r}
viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
                                            include_virsorter2 = F,
                                            include_checkV = T,
                                            include_kaiju = T,
                                            include_deepvirfinder = T)
```

# Plot of how many methods identify a contig
```{r}
viruses_high <- viruses[viruses$keep_score_all>=1,]
viruses_high_mod <- viruses_high %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
viruses_high_mod <- as_tibble(viruses_high_mod)

sm <- as.data.frame(rbind(c(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb_cv>0,]))),
                          c(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf_cv>0,]))),
                          c(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2_cv>0,]))),
                          c(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_cv_kj>0,])))
                          )
)

sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "kaiju")

colnames(sm) <- c("one","two","three","all","method")

sm
```

```{r}
sm_m <- melt(sm)
colnames(sm_m) <- c("primary_method", "number_methods", "number")
```

```{r}
ggplot(sm_m, aes(x=primary_method, y=number,
                   fill=number_methods)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name = 'Number of Methods',
                     values = alpha(c(viridis(5)), 1)) +
  xlab("Primary Method") +
  ylab("Count of Viral Contigs")
```

# Visualizing which tools led to a contig being selected
(alluvial plot)
with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/

```{r}
alluvial_data <- viruses %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  apply(c(1,2), function(x) {if (x >= 1) {x <- 1} else if (x==0.5) {x <- 0.5} else {x <- 0}}) %>%
  as_tibble() %>%
  count(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
```

```{r}
alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
  rowSums()
```

```{r}
alluvial_data$viral <- "not viral"
alluvial_data$viral[alluvial_data$cum_sum>=1] <- "one method"
alluvial_data$viral[alluvial_data$cum_sum>=2] <- "two methods"
alluvial_data$viral[alluvial_data$cum_sum>=3] <- "three methods"
alluvial_data$viral[alluvial_data$cum_sum>=4] <- "all methods"

alluvial_data$viral <- factor(alluvial_data$viral, levels=c("not viral", "one method",
                                                            "two methods", "three methods",
                                                            "all methods"))
```

```{r}
alluvial_data <- alluvial_data %>% 
  mutate(across(c(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj), as.factor))
```

```{r}
head(alluvial_data)
```

```{r}
library(ggalluvial)
```

```{r}
ggplot(data = alluvial_data,
       aes(axis1 = as.character(keep_score_vb_cv), 
           axis2 = as.character(keep_score_vs2_cv),
           axis3 = as.character(keep_score_dvf_cv),
           axis4 = as.character(keep_score_cv_kj),
           y = n)) +
  geom_alluvium(aes(fill = viral)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "VirSorter2", "DeepVirFinder", "Kaiju"),
                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
  ylab("Number of Sequences") +
  theme_classic()
```

# Using NMDS to compare the tools
get matrix want to use for beta diversity
```{r}
virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_kj, keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_kj_cv, keep_score_all)

```

```{r}
virus_score_matrix <- virus_score_matrix[rowSums(virus_score_matrix)>0,]
```

```{r}
library(vegan)
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

example_NMDS <- metaMDS(t(virus_score_matrix), distance = "bray")

ss <- as.data.frame(example_NMDS$points)
ss$tool <- colnames(virus_score_matrix)
```

```{r}
p <- ggplot() + 
  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
   theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text=element_text(size=14, colour="black"),
        axis.title=element_text(size=16),
        legend.key=element_blank(),
        legend.position="right",
        legend.background = element_blank()) +
  guides(size="none") + xlab("NMDS 1") + ylab("NMDS 2") + 
  coord_equal() +
  scale_fill_manual(name="",
                     values = alpha(c("black", "#F1CE63", "#B6992D", "#8CD17D", "#59A14F", "#FFBE7D", "#F28E2B", "#A0CBE8", "#4E79A7"), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(c("black","#F1CE63", "#B6992D", "#8CD17D", "#59A14F", "#FFBE7D", "#F28E2B", "#A0CBE8", "#4E79A7"), 1)) +
  scale_shape_manual(name="",
                     values = c(20,21,21,22,22,25,23,23,24,24))

p
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

virus_score_matrix_binary <- apply(virus_score_matrix, c(1,2), function(x) if (x>=1) {x <- 1} else {x <- 0})

example_NMDS <- metaMDS(t(virus_score_matrix_binary), distance = "bray")

ss <- as.data.frame(example_NMDS$points)
ss$tool <- colnames(virus_score_matrix)
```

```{r}
p <- ggplot() + 
  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
   theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text=element_text(size=14, colour="black"),
        axis.title=element_text(size=16),
        legend.key=element_blank(),
        legend.position="right",
        legend.background = element_blank()) +
  guides(size="none") + xlab("NMDS 1") + ylab("NMDS 2") + 
  coord_equal() +
  scale_fill_manual(name="",
                     values = alpha(c("black", "#F1CE63", "#B6992D", "#8CD17D", "#59A14F", "#FFBE7D", "#F28E2B", "#A0CBE8", "#4E79A7"), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(c("black","#F1CE63", "#B6992D", "#8CD17D", "#59A14F", "#FFBE7D", "#F28E2B", "#A0CBE8", "#4E79A7"), 1)) +
  scale_shape_manual(name="",
                     values = c(20,21,21,22,22,25,23,23,24,24))

p
```

```{r}
virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_kj, keep_score_vb_cv, keep_score_dvf_cv, keep_score_vs2_cv, keep_score_kj_cv, keep_score_vb_cv_kj, keep_score_dvf_cv_kj, keep_score_vs2_cv_kj, keep_score_vb_vs2_cv_kj, keep_score_vb_cv_kj_dvf, keep_score_vs2_cv_kj_dvf, keep_score_all)
```

```{r}
virus_score_matrix <- virus_score_matrix[rowSums(virus_score_matrix)>0,]
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

example_NMDS <- metaMDS(t(virus_score_matrix), distance = "bray")

ss <- as.data.frame(example_NMDS$points)
ss$tool <- colnames(virus_score_matrix)
```

```{r}
p <- ggplot() + 
  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
   theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text=element_text(size=14, colour="black"),
        axis.title=element_text(size=16),
        legend.key=element_blank(),
        legend.position="right",
        legend.background = element_blank()) +
  guides(size="none") + xlab("NMDS 1") + ylab("NMDS 2") + 
  coord_equal() +
  scale_fill_manual(name="",
                     values = alpha(c("black", "#F1CE63", "#4E79A7", "#59A14F", "#8CD17D", "#4E79A7", "#FFBE7D", "#4E79A7", "#59A14F", "#F1CE63", "#E15759", "#E15759", "#4E79A7","#59A14F", "#F1CE63"), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(c("black", "#F1CE63", "#4E79A7", "#59A14F", "#8CD17D", "#4E79A7", "#FFBE7D", "#4E79A7", "#59A14F", "#F1CE63", "#E15759", "#E15759", "#4E79A7","#59A14F", "#F1CE63"), 1)) +
  scale_shape_manual(name="",
                     values = c(20,21,21,21,22,22,23,23,23,23,23,24,24,24,24))

p
```

```{r}
virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_kj, keep_score_vb_cv, keep_score_dvf_cv, keep_score_vs2_cv, keep_score_kj_cv, keep_score_vb_cv_kj, keep_score_dvf_cv_kj, keep_score_vs2_cv_kj, keep_score_vb_vs2_cv_kj, keep_score_vb_cv_kj_dvf, keep_score_vs2_cv_kj_dvf, keep_score_all)
```

```{r}
virus_score_matrix <- virus_score_matrix[rowSums(virus_score_matrix)>0,]
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")

virus_score_matrix_binary <- apply(virus_score_matrix, c(1,2), function(x) if (x>=1) {x <- 1} else {x <- 0})

example_NMDS <- metaMDS(t(virus_score_matrix_binary), distance = "bray")

ss <- as.data.frame(example_NMDS$points)
ss$tool <- colnames(virus_score_matrix)
```

```{r}
p <- ggplot() + 
  geom_point(data=ss, aes(x=MDS1,y=MDS2, color=tool, fill=tool, shape=tool, size=4)) +
   theme(panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text=element_text(size=14, colour="black"),
        axis.title=element_text(size=16),
        legend.key=element_blank(),
        legend.position="right",
        legend.background = element_blank()) +
  guides(size="none") + xlab("NMDS 1") + ylab("NMDS 2") + 
  coord_equal() +
  scale_fill_manual(name="",
                     values = alpha(c("black", "#F1CE63", "#4E79A7", "#59A14F", "#8CD17D", "#4E79A7", "#FFBE7D", "#4E79A7", "#59A14F", "#F1CE63", "#E15759", "#E15759", "#4E79A7","#59A14F", "#F1CE63"), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(c("black", "#F1CE63", "#4E79A7", "#59A14F", "#8CD17D", "#4E79A7", "#FFBE7D", "#4E79A7", "#59A14F", "#F1CE63", "#E15759", "#E15759", "#4E79A7","#59A14F", "#F1CE63"), 1)) +
  scale_shape_manual(name="",
                     values = c(20,21,21,21,22,22,23,23,23,23,23,24,24,24,24))

p
```

### 2022-08-10 EnvironmentalData/EarlyIterations/identifying_viral_contigs_20220521.Rmd
Updated directory structure.

--- 2022-06-01 Scripts/Evaluations/identifying_viral_contigs_20220521.Rmd
@@
 }
-```
-
-```{r}
-comp_pipelines <- tibble(pipeline="example", precision=0, recall=0, F1=0)
 ```
@@
 ```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=seqtype,
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+p2 <- ggplot(confusion_by_taxa, aes(x=count, y=seqtype,
                    fill=confusion_matrix,
@@
     axis.ticks.y = element_blank(),
-    legend.position = "bottom"
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
   ) +
@@
   ylab("")
+
+p2
+```
+
+```{r}
+png("testing_set_accuracy_by_taxa.png", width=6.8, height=3.2, units="in", res=400)
+
+p2
+
+dev.off()
+```
+
+# Visualizing confusion matrix by number of tools
+
+```{r}
+viruses$confusion_matrix <- "true positive"
+viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
+viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
+viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_all<1] <- "true negative"
+
+viruses$keep_score_visualize <- viruses$keep_score_all
+viruses$keep_score_visualize[viruses$keep_score_all>1] <- ">1"
+
+viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
+                                       levels=c("-0.5", "-1", "0", "0.5","1", ">1"))
+viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
+                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
+```
+
+```{r}
+levels(factor(viruses$keep_score_visualize))
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
+
+p1 <- ggplot(viruses, aes(x=confusion_matrix,
+                   fill=keep_score_visualize, color=keep_score_visualize)) +
+  geom_bar(stat="count", position="stack") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16)
+  ) +
+  scale_color_manual(name = 'Viral Score',
+                     values = alpha(c(pal(4)), 1)) +
+  scale_fill_manual(name = 'Viral Score',
+                     values = alpha(c(pal(4)), 0.5)) +
+  xlab("") +
+  ylab("Sequence Count")
+
+p1
+```
+
+```{r}
+png("testing_set_accuracy_confusion_matrix.png", width=6.8, height=2.6, units="in", res=400)
+
+p1
+
+dev.off()
 ```
@@
 ```{r}
-virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_cv, keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj, keep_score_all)
+virus_score_matrix <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_cv, keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_kj_cv, keep_score_all)
 
@@
 
+# Correlating Scores and Other parameters
+
+```{r}
+cor(viruses$keep_score_all, viruses$hallmark, method="pearson")
+```
+
+```{r}
+cor(viruses$keep_score_all, viruses$percent_unknown, method="pearson")
+```
+
+```{r}
+cor(viruses$keep_score_all, viruses$percent_viral, method="pearson")
+```
+
 ## Looking at False Positives

### 2022-08-18 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
Changed Virsorter2 filter
Changed virsorter2 from >=0.5 +0.5 and >=0.9 +1 to both being +0.5. This way a viral sequence scored 0.9 will yield a score from vs2 of 1 instead of 1.5

--- 2022-07-26 identifying_viral_contigs_environmental.Rmd
@@
 
-vs_c <- vs_c[!duplicated(vs_c$contig),]
 vs_c <- vs_c %>% drop_na(contig)
@@
   if (include_vibrant) {
-    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
-    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
-    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
+    # keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
   }
@@
     keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
   }

### 2022-08-23 TestingPipeline/Scripts/checkV.sh
Create checkV.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+echo start
+echo $1
+samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
+database="/home/hegartyb/checkv-db-v0.6/"
+threads=4
+
+echo contamination
+checkv contamination $samplename ./ -t $threads -d $database
+echo completeness
+checkv completeness $samplename ./ -t $threads -d $database
+echo complete genomes
+checkv complete_genomes $samplename ./
+echo quality summary
+checkv quality_summary $samplename ./
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/checkv_slurm.sh
Create checkv_slurm.sh

@@ -1,0 +1,45 @@
+#!/bin/bash
+
+###############################
+#                             #
+#  1) Job Submission Options  #
+#                             #
+###############################
+
+# Name
+#SBATCH --job-name=checkv
+
+# Resources
+# For MPI, increase ntasks-per-node
+# For multithreading, increase cpus-per-task
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --cpus-per-task=2
+#SBATCH --mem-per-cpu=80G
+#SBATCH --time=5-00:00:00
+
+# Account
+#SBATCH --account=kwigg1
+#SBATCH --partition=standard
+
+# Logs
+#SBATCH --mail-user=hegartyb@umich.edu
+#SBATCH --mail-type=BEGIN,END,FAIL
+
+# Environment
+##SBATCH --export=ALL
+
+#SBATCH --array=1-10
+
+#  Show list of CPUs you ran on
+echo $SLURM_JOB_NODELIST
+
+echo start
+
+echo "activate conda environment"
+source activate  /home/hegartyb/miniconda3/envs/checkV
+
+echo "run checkV on all testing sets"
+bash checkv.sh ${SLURM_ARRAY_TASK_ID}
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/checkV.sh
Updated tool scripts to auto check num threads.

@@ -4,7 +4,7 @@ echo start
 echo $1
 samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
 database="/home/hegartyb/checkv-db-v0.6/"
-threads=4
+threads=$2
 
 echo contamination
 checkv contamination $samplename ./ -t $threads -d $database

### 2022-08-23 TestingPipeline/Scripts/Tools/checkv_slurm.sh

@@ -40,6 +40,6 @@ echo "activate conda environment"
 source activate  /home/hegartyb/miniconda3/envs/checkV
 
 echo "run checkV on all testing sets"
-bash checkv.sh ${SLURM_ARRAY_TASK_ID}
+bash checkv.sh ${SLURM_ARRAY_TASK_ID} {SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/checkV.sh
Scripts to run tools on testing sets.

@@ -5,6 +5,11 @@ echo $1
 samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
 database="/home/hegartyb/checkv-db-v0.6/"
 threads=$2
+echo $threads
+
+rm -r "Checkv_${1}"
+mkdir "Checkv_${1}"
+cd "Checkv_${1}"
 
 echo contamination
 checkv contamination $samplename ./ -t $threads -d $database

### 2022-08-23 TestingPipeline/Scripts/Tools/dvf.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+echo start
+
+rm -r "DVF_{1}"
+mkdir "DVF_${1}"
+
+samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
+threads=$2
+outfile="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/DVF/DVF_${1}"
+
+python /scratch/kwigg_root/kwigg/hegartyb/DeepVirFinder/dvf.py -i $samplename -o $outfile -c $threads
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/dvf_slurm.sh

@@ -7,15 +7,15 @@
 ###############################
 
 # Name
-#SBATCH --job-name=checkv
+#SBATCH --job-name=dvf
 
 # Resources
 # For MPI, increase ntasks-per-node
 # For multithreading, increase cpus-per-task
 #SBATCH --nodes=1
 #SBATCH --ntasks-per-node=1
-#SBATCH --cpus-per-task=2
-#SBATCH --mem-per-cpu=80G
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=32G
 #SBATCH --time=5-00:00:00
 
 # Account
@@ -37,12 +37,11 @@ echo $SLURM_JOB_NODELIST
 echo start
 
 echo "activate conda environment"
-#source /home/hegartyb/.bashrc
-source activate  /home/hegartyb/miniconda3/envs/checkV
+source activate  /home/hegartyb/miniconda3/envs/dvf
 
-echo "run checkV on all testing sets"
-mkdir ../Checkv
-cd ../Checkv
-bash ../Scripts/checkv.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_PER_TASK}
+echo "run dvf on all testing sets"
+mkdir ../DVF
+cd ../DVF
+bash ../Scripts/dvf.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/kaiju.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+echo start
+
+rm -r "Kaiju_${1}"
+mkdir "Kaiju_${1}"
+cd "Kaiju_${1}"
+pwd
+
+samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
+threads=$2
+input_ndmp="/scratch/duhaimem_root/duhaimem0/shared_data/kaijudb/nr_euk/nodes.dmp"
+input_names="/scratch/duhaimem_root/duhaimem0/shared_data/kaijudb/nr_euk/names.dmp"
+input_vmi="/nfs/turbo/cee-kwigg/hegartyb/Kaiju/Kaijudb-refseq/kaiju_db_refseq.fmi"
+outname="${1}.nreuk.kaiju.out"
+outname_taxa="${1}.nreuk.kaiju.names.out"
+
+/nfs/turbo/lsa-duhaimem/Lake_Michigan_JGI_viral_metaGs/kaiju/bin/kaiju -z $2 -t $input_ndmp -f $input_vmi -i $samplename -o $outname -v
+/nfs/turbo/lsa-duhaimem/Lake_Michigan_JGI_viral_metaGs/kaiju/bin/kaiju-addTaxonNames -t $input_ndmp -n $input_names -i $outname -o $outname_taxa -p
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/kaiju_slurm.sh

@@ -7,15 +7,15 @@
 ###############################
 
 # Name
-#SBATCH --job-name=checkv
+#SBATCH --job-name=kaiju
 
 # Resources
 # For MPI, increase ntasks-per-node
 # For multithreading, increase cpus-per-task
 #SBATCH --nodes=1
 #SBATCH --ntasks-per-node=1
-#SBATCH --cpus-per-task=2
-#SBATCH --mem-per-cpu=80G
+#SBATCH --cpus-per-task=4
+#SBATCH --mem-per-cpu=20G
 #SBATCH --time=5-00:00:00
 
 # Account
@@ -37,12 +37,11 @@ echo $SLURM_JOB_NODELIST
 echo start
 
 echo "activate conda environment"
-#source /home/hegartyb/.bashrc
-source activate  /home/hegartyb/miniconda3/envs/checkV
+source activate  /home/hegartyb/miniconda3/envs/cluster_kaiju
 
-echo "run checkV on all testing sets"
-mkdir ../Checkv
-cd ../Checkv
-bash ../Scripts/checkv.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_PER_TASK}
+echo "run dvf on all testing sets"
+mkdir ../Kaiju
+cd ../Kaiju
+bash ../Scripts/kaiju.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/running_tools_on_testing_sets_slurm.sh

@@ -1,0 +1,44 @@
+#!/bin/bash
+
+###############################
+#                             #
+#  1) Job Submission Options  #
+#                             #
+###############################
+
+# Name
+#SBATCH --job-name=startingruntoolstestingset
+
+# Resources
+# For MPI, increase ntasks-per-node
+# For multithreading, increase cpus-per-task
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=6
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=1000M
+#SBATCH --time=01:00:00
+
+# Account
+#SBATCH --account=kwigg1
+#SBATCH --partition=standard
+
+# Logs
+#SBATCH --mail-user=hegartyb@umich.edu
+#SBATCH --mail-type=BEGIN,END,FAIL
+
+# Environment
+##SBATCH --export=ALL
+
+#  Show list of CPUs you ran on
+echo $SLURM_JOB_NODELIST
+
+echo start
+
+sbatch /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/Scripts/checkv_slurm.sh 
+sbatch /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/Scripts/dvf_slurm.sh 
+sbatch /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/Scripts/kaiju_slurm.sh 
+sbatch /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/Scripts/vibrant_slurm.sh 
+sbatch /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/Scripts/vs_slurm.sh 
+sbatch /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/Scripts/vs2_slurm.sh
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/vibrant.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+echo start
+
+rm "Vibrant_${1}"
+mkdir "Vibrant_${1}"
+cd "Vibrant_${1}"
+pwd
+
+samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
+threads=$2
+database="/home/hegartyb/VIBRANT/databases/"
+vibrantloc="/home/hegartyb/VIBRANT/files/"
+
+python3 /home/hegartyb/VIBRANT/VIBRANT_run.py -i $samplename -t $threads -folder ./ -d $database -m $vibrantloc
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/vibrant_slurm.sh

@@ -7,7 +7,7 @@
 ###############################
 
 # Name
-#SBATCH --job-name=checkv
+#SBATCH --job-name=vibrant
 
 # Resources
 # For MPI, increase ntasks-per-node
@@ -15,7 +15,7 @@
 #SBATCH --nodes=1
 #SBATCH --ntasks-per-node=1
 #SBATCH --cpus-per-task=2
-#SBATCH --mem-per-cpu=80G
+#SBATCH --mem-per-cpu=8G
 #SBATCH --time=5-00:00:00
 
 # Account
@@ -37,12 +37,12 @@ echo $SLURM_JOB_NODELIST
 echo start
 
 echo "activate conda environment"
-#source /home/hegartyb/.bashrc
-source activate  /home/hegartyb/miniconda3/envs/checkV
+source activate  /home/hegartyb/miniconda3/envs/vibrant3.7
 
-echo "run checkV on all testing sets"
-mkdir ../Checkv
-cd ../Checkv
-bash ../Scripts/checkv.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_PER_TASK}
+echo "run dvf on all testing sets"
+mkdir ../Vibrant
+cd ../Vibrant
+pwd
+bash ../Scripts/vibrant.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/vs.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+echo start
+
+rm -r "VS_${1}"
+#mkdir "VS_${1}"
+#cd "VS_${1}"
+pwd
+
+samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
+threads=$2
+path="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/VS/VS_${1}"
+vsdirectory="/nfs/turbo/lsa-duhaimem/software/virsorter-data"
+
+/nfs/turbo/cee-kwigg/hegartyb/VirSorter/wrapper_phage_contigs_sorter_iPlant.pl \
+        -f $samplename --db 1 \
+        --wdir $path \
+        --data-dir $vsdirectory \
+        --ncpu $threads
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/vs2.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+echo start
+
+rm -r "VS2_${1}"
+mkdir "VS2_${1}"
+cd "VS2_${1}"
+pwd
+
+samplename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSet/metagenomic_testing_set_${1}.fna"
+threads=$2
+path="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/VS2/VS2_${1}"
+threads=10
+
+virsorter run -w $path -i $samplename -j $threads all
+
+echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/vs2_slurm.sh

@@ -7,15 +7,15 @@
 ###############################
 
 # Name
-#SBATCH --job-name=checkv
+#SBATCH --job-name=vs2
 
 # Resources
 # For MPI, increase ntasks-per-node
 # For multithreading, increase cpus-per-task
 #SBATCH --nodes=1
 #SBATCH --ntasks-per-node=1
-#SBATCH --cpus-per-task=2
-#SBATCH --mem-per-cpu=80G
+#SBATCH --cpus-per-task=20
+#SBATCH --mem-per-cpu=300mb
 #SBATCH --time=5-00:00:00
 
 # Account
@@ -37,12 +37,11 @@ echo $SLURM_JOB_NODELIST
 echo start
 
 echo "activate conda environment"
-#source /home/hegartyb/.bashrc
-source activate  /home/hegartyb/miniconda3/envs/checkV
+source activate  /home/hegartyb/miniconda3/envs/vs2_2
 
-echo "run checkV on all testing sets"
-mkdir ../Checkv
-cd ../Checkv
-bash ../Scripts/checkv.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_PER_TASK}
+echo "run virsorter on all testing sets"
+mkdir ../VS2
+cd ../VS2
+bash ../Scripts/vs2.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-08-23 TestingPipeline/Scripts/Tools/vs_slurm.sh

@@ -7,15 +7,15 @@
 ###############################
 
 # Name
-#SBATCH --job-name=checkv
+#SBATCH --job-name=vs
 
 # Resources
 # For MPI, increase ntasks-per-node
 # For multithreading, increase cpus-per-task
 #SBATCH --nodes=1
 #SBATCH --ntasks-per-node=1
-#SBATCH --cpus-per-task=2
-#SBATCH --mem-per-cpu=80G
+#SBATCH --cpus-per-task=4
+#SBATCH --mem-per-cpu=4G
 #SBATCH --time=5-00:00:00
 
 # Account
@@ -37,12 +37,13 @@ echo $SLURM_JOB_NODELIST
 echo start
 
 echo "activate conda environment"
-#source /home/hegartyb/.bashrc
-source activate  /home/hegartyb/miniconda3/envs/checkV
+module load Bioinformatics
+source activate  /home/hegartyb/miniconda3/envs/virsorter2
 
-echo "run checkV on all testing sets"
-mkdir ../Checkv
-cd ../Checkv
-bash ../Scripts/checkv.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_PER_TASK}
+echo "run virsorter on all testing sets"
+mkdir ../VS
+cd ../VS
+pwd
+bash ../Scripts/vs.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-09-07 TestingPipeline/Scripts/MergingScripts/merge_checkv.py
Added scripts to merge together multiple runs of the tools.

@@ -1,0 +1,20 @@
+with open("/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/testing_sets_checkv_output.tsv", 'wt') as outfile:
+    j=0
+    for index in range(1,11):
+        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/Testing_Set_%s/quality_summary.tsv" %(index)
+        with open(filename, 'rt') as v:
+            if (j==0):
+                i=0
+                for line in v:
+                    if (i==0):
+                        outfile.write('Index\t%s' %(line))
+                    if (i>0):
+                        outfile.write('%s\t%s' %(index, line))
+                    i+=1
+            if (j>0):
+                i=0
+                for line in v:
+                    if (i>0):
+                        outfile.write('%s\t%s' %(index, line))
+                    i+=1
+        j+=1

### 2022-09-07 TestingPipeline/Scripts/MergingScripts/merge_deepvirfinder.py

@@ -1,0 +1,20 @@
+with open("/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/testing_sets_vibrant_output.tsv", 'wt') as outfile:
+    j=0
+    for index in range(1,11):
+        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/Testing_Set_%s/VIBRANT_genome_quality_metagenomic_testing_set_%s.tsv" %(index,index)
+        with open(filename, 'rt') as v:
+            if (j==0):
+                i=0
+                for line in v:
+                    if (i==0):
+                        outfile.write('Index\t%s' %(line))
+                    if (i>0):
+                        outfile.write('%s\t%s' %(index, line))
+                    i+=1
+            if (j>0):
+                i=0
+                for line in v:
+                    if (i>0):
+                        outfile.write('%s\t%s' %(index, line))
+                    i+=1
+        j+=1

### 2022-09-07 TestingPipeline/Scripts/MergingScripts/merge_kaiju.py

@@ -1,0 +1,7 @@
+with open("/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/merged.nreuk.kaiju.names.out", 'wt') as outfile:
+    outfile.write('Index\tClassified\tContig\tNCBI_taxon\tlen\tID_best\tIDs_all\tSeq\tName\n')
+    for index in range(1,11):
+        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/Testing_Set_%s/%s.nreuk.kaiju.names.out" %(index,index)
+        with open(filename, 'rt') as v:
+            for line in v:
+                outfile.write('%s\t%s' %(index, line))

### 2022-09-07 TestingPipeline/Scripts/MergingScripts/merge_virsorter.py

@@ -1,0 +1,19 @@
+with open("/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/testing_sets_virsorter_output.tsv", 'wt') as outfile:
+    outfile.write("index\tcontig\tcategory\n")
+    for index in range(1,11):
+        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/Testing_Set_%s/VIRSorter_global-phage-signal.csv" %(index)
+        with open(filename, 'rt') as v:
+            for line in v:
+                c1 = line.split(" ")[0]
+                if (c1 == "##"):
+                    if ("(" in line):
+                        c2 = line.split(" ")[1]
+                        category=c2
+                if (c1 != "##"):
+                    contig=line.split(",")[0]
+                    numgenes1=line.split(",")[1]
+                    numgenes2=line.split(",")[3]
+                    numphagegenes=line.split(",")[5]
+                    phagessig=line.split(",")[6]
+                    noncaudphagesig=line.split(",")[7]
+                    outfile.write('%s\t%s\t%s\n' %(index,contig,category))

### 2022-09-07 TestingPipeline/Scripts/MergingScripts/merge_virsorter2.py

@@ -1,0 +1,25 @@
+#!/usr/bin/python
+"""merge_vs2.py (author:BEH)
+create a list of all the vs2 viral contigs
+"""
+
+with open("/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/testing_sets_virsorter2_output.tsv", 'wt') as outfile:
+    j=0
+    for index in range(1,11):
+        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/Testing_Set_%s/final-viral-score.tsv" %(index)
+        with open(filename, 'rt') as v:
+            if (j==0):
+                i=0
+                for line in v:
+                    if (i==0):
+                        outfile.write('Index\t%s' %(line))
+                    if (i>0):
+                        outfile.write('%s\t%s' %(index, line))
+                    i+=1
+            if (j>0):
+                i=0
+                for line in v:
+                    if (i>0):
+                        outfile.write('%s\t%s' %(index, line))
+                    i+=1
+        j+=1

### 2022-09-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Added R notebooks used for rule tuning to develop pipeline.

---
title: "Viral Sequence Sorting Tools Evaluation"
author: Bridget Hegarty, James Riddell
date: 07-22-2022
output: html_notebook
---
This Rmarkdown file assesses the output of CheckV, DeepVirFinder, Kaiju,
VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
primarily from NCBI. Created from fungal, viral, bacterial, archeael, protist,
and plasmid DNA sequences

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

# Inputs
1) For this script's inputs, the user must merge each viral identification 
tool output file into one tab-separated file. 
Replace the filenames with their paths if needed into the variables in the 
'inputs' chunk.

    General filenames for each tool:
        Checkv: quality_summary.tsv
        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
        Virsorter: VIRSorter_global-phage-signal.csv
        Virsorter2: final-viral-score.tsv
        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
        Kaiju: ${assembly}.kaiju.names.out

2) Make sure all fasta headers (assembly, contig) are consistent within and
across tools. Chunks for each tool do contain some lines for cleaning up these 
features, but due to their variability it will be the user's responsibility to 
make sure they match across tools.

3) Check each chunk and ensure all columns are accounted for.

4) This script is designed for contigs > 3000 basepairs. It can be modified
to be higher or lower, but going lower will greatly increase the size of the
dataframe and memory usage.

```{r inputs}
# dataset name for file organization and outputs for metrics file
dataset_name <- "testing_set"
# checkV
checkV_path <- "TestingSetToolOutputs/testing_sets_checkv_output.tsv"
# VIBRANT
vibrant_path <- "TestingSetToolOutputs/testing_sets_vibrant_output.tsv"
# DeepVirFinder
dvf_path <- "TestingSetToolOutputs/testing_sets_deepvirfinder_output.tsv"
# Virsorter
vs_path <- "TestingSetToolOutputs/testing_sets_virsorter_output.tsv"
# Virsorter2
vs2_path <- "TestingSetToolOutputs/testing_sets_virsorter2_output.tsv"
# Kaiju
kj_path <- "TestingSetToolOutputs/merged.nreuk.kaiju.names.out"
# KB cutoff
KB_CUTOFF <- 3000

```

# All Viral Contigs

## checkV
```{r}
checkV <- fread(checkV_path, 
                sep="\t",
                header = T, 
                select = c(
                    'Index',
                    'contig_id',
                    'provirus',
                    'completeness',
                    'contamination',
                    'viral_genes',
                    'host_genes',
                    'gene_count',
                    'contig_length',
                    'checkv_quality'
                    )
                ) %>% 
    rename(
        contig = contig_id,
        checkv_provirus = provirus,
        checkv_completeness = completeness,
        checkv_contamination = contamination,
        checkv_viral_genes = viral_genes,
        checkv_host_genes = host_genes,
        checkv_total_genes = gene_count,
        checkv_length = contig_length
        )
checkV$method = 'checkv'
checkV <- separate(checkV, col = contig, into = c("seqtype", "contig"), sep="--")
checkV$contig <- sub("\\.", "_", checkV$contig)
checkV$contig <- sub("\\|", "_", checkV$contig)
```

## VIBRANT
```{r}
vb_c <- fread(vibrant_path,
              header = T,
              sep = "\t",
              select = c(
                  'Index',
                  'scaffold',
                  'type',
                  'Quality'
                  )
              ) %>%
    rename(
        contig = scaffold,
        vibrant_quality = Quality
        )
vb_c$method <- "vibrant"
vb_c$vibrant_prophage <- "No"
vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
vb_c <- separate(vb_c, col = contig, into = c("seqtype", "contig"), sep="--")
vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
vb_c$contig <- sub("\\.", "_", vb_c$contig)
vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
vb_c <- vb_c[!duplicated(vb_c$contig),]
```

## DeepVirFinder
```{r}
dvf_c <- fread(dvf_path,
               header = T,
               sep = "\t",
               select = c(
                   'Index',
                   'name',
                   'score',
                   'pvalue'
                   )
               ) %>% 
    rename(
        contig = name
        )
dvf_c$contig <- sub("\\.", "_", dvf_c$contig)

dvf_c <- separate(dvf_c, col = contig, into = c("seqtype", "contig"), sep="--")
dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
```

## VirSorter
```{r}
vs_c <- fread(vs_path,
              select = c(
                  'index',
                  'contig',
                  'category'
                  )
              ) %>% 
    rename(
        Index = index
        )

vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
vs_c$contig <- sub("-circular", "", vs_c$contig)
vs_c <- separate(vs_c, col = `contig`, into = c("seqtype", "contig"), sep="--")
vs_c$contig <- sub("\\.", "_", vs_c$contig)

vs_c <- vs_c %>% drop_na(contig)
vs_c <- vs_c[!duplicated(vs_c$contig),]
```

## VirSorter2 
```{r}
vs2_c <- fread(vs2_path,
                    header = T,
                    sep = '\t',
                    select = c(
                        'Index',
                        'seqname',
                        'dsDNAphage',
                        'ssDNA',
                        'max_score',
                        'max_score_group',
                        'hallmark',
                        'viral',
                        'cellular'
                    )
               ) %>% 
              separate(
                  col = seqname,
                  into = c("contig", "vs2type"), 
                  sep = "\\|\\|",
                  remove = T
                  )
vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
vs2_c <- separate(vs2_c, col = contig, into = c("seqtype", "contig"), sep="--")
vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
```

## Kaiju 
```{r}
kj_c <- read_tsv(kj_path, col_names = T)

kj_c$Contig <- sub("\\.", "_", kj_c$Contig)

kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")

kj_c$contig <- sub("\\.", "_", kj_c$contig)
```
Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
```{r}
table(kj_c$seqtype, kj_c$Kaiju_Viral)
```

## Merging
```{r}
viruses <- full_join(x=checkV, y=kj_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=dvf_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vb_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs_c, by = c("Index", "contig", "seqtype"))
viruses <- full_join(x=viruses, y=vs2_c, by = c("Index", "contig", "seqtype"))
```

Remove contigs not greater than the basepair length cutoff (3000)
```{r}
viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
```

Use the next code chunk to check if any contigs are missing from checkv.
If yes, then there is a mix up in contig names since checkv should contain
all contig names even if they are identified as non-viral.
```{r}
v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
```

calculate percent viral/host/unknown
```{r}
viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
```

get rid of NAs for downstream processing
```{r}
# checkV
viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0

# VIBRANT
viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0

# DeepVirFinder
viruses$score[is.na(viruses$score)] <- 0
viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0

# VirSorter2
viruses$viral[is.na(viruses$viral)] <- 0
viruses$hallmark[is.na(viruses$hallmark)] <- 0

#Virsorter
viruses$category[is.na(viruses$category)] <- 0

# Kaiju
viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
```

This section defines a viralness score "keep_score" based on the tool classifications. 
A final keep_score above 1 indicates we will keep that sequence and call it viral.

VIBRANT
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus == TRUE: +0.5

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +1
    Hallmark > 2: +1

Virsorter
    category ==  1,2,4,5: +1
    category == 3,6: +0.5

DeepVirFinder:
    Score >= 0.7: +0.5

Kaiju:
    Kaiju_viral = "cellular organisms": -1
    Kaiju_viral = "Viruses": +1

CheckV
    viral_genes == 0 and host_genes >= 1: keep_score = 0
    If 3*viral_genes <= host_genes: keep_score = 0
    If length > 50,000 and hallmark == 0: keep_score = 0
    If %unknown >= 75: +0.5

This script produces visualizations of these combined viral scorings and
includes ecological metrics like alpha diversity.

You can decide which combination is appropriate for them and only need use the
tools appropriate for your data.

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_checkV=FALSE,
                                include_kaiju=FALSE,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
  }

  if (include_kaiju) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
  }

  if (include_checkV) {
    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
  }

  return(keep_score)

}

```

# Assessing performance against the "truth"
note that this is only as accurate as the annotations of the input sequences

this function calculates the precision, recall, and F1 score for each pipeline
```{r}
assess_performance <- function(seqtype, keep_score) {

  truepositive <- rep("not viral", length(seqtype))
  truepositive[seqtype=="virus"] <- "viral"

  #make confusion matrix
  confusion_matrix <- rep("true negative", length(keep_score))
  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"

  precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])

  recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])

  F1 <- 2*precision*recall/(precision+recall)

  performance <- c(precision, recall, F1)
  names(performance) <- c("precision", "recall", "F1")

  return(performance)
}
```

combination of tools list
```{r}
combos_list <- read_csv("combinations_list.csv", col_names = T)
```

this function builds a list of all of the combinations that the user wants to 
test. 
In this case, we're comparing the performance of all unique combinations of the 
six tools.
```{r}
build_score_list <- function(input_seqs, combos) {
  output <- data.frame(precision=rep(0, nrow(combos)),
                       recall=rep(0, nrow(combos)),
                       F1=rep(0, nrow(combos)))
  for (i in 1:nrow(combos)) {
    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
                                            include_virsorter = combos$VS[i],
                                            include_virsorter2 = combos$VS2[i],
                                            include_checkV = combos$CheckV[i],
                                            include_kaiju = combos$Kaiju[i],
                                            include_deepvirfinder = combos$DVF[i])

    output[i,1:3] <- assess_performance(input_seqs$seqtype, keep_score)

    output$toolcombo[i] <- paste(combos$CheckV[i],combos$DVF[i],
                                 combos$Kaiju[i], combos$VIBRANT[i],
                                 combos$VS[i], combos$VS2[i])
  }

  output[is.na(output)] <- 0

  #return(keep_score)
  return (output)
}
```

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- build_score_list(viruses[viruses$Index==1], combos_list)

accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10))

accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1], combos_list))

for (i in 2:10) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i], combos_list)))
}

```

## Visualize how the precision, recall, and F1 scores change across pipelines.
```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")

p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("F1 Score")

p2

ggplot(accuracy_scores, aes(x=toolcombo, y=precision)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("Precision")

ggplot(accuracy_scores, aes(x=toolcombo, y=recall)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
  ylab("Recall")
```

### 2022-09-09 TestingPipeline/Scripts/comparing_rule_sets.Rmd
Playing around with other rule sets.

--- 2022-09-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "Comparing rule sets"
 author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+date: 09-08-2022
 output: html_notebook
 ---
-This Rmarkdown file assesses the output of CheckV, DeepVirFinder, Kaiju,
-VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
+This Rmarkdown file will compare the viral predictions of multiple pipelines 
+based on the output of CheckV, DeepVirFinder, Kaiju,VIBRANT, VirSorter, and 
+VirSorter2 on multiple training sets of microbial DNA, 
 primarily from NCBI. Created from fungal, viral, bacterial, archeael, protist,
-and plasmid DNA sequences
+and plasmid DNA sequences.
 
@@
 # checkV
-checkV_path <- "TestingSetToolOutputs/testing_sets_checkv_output.tsv"
+checkV_path <- "../ToolOutput/testing_sets_checkv_output.tsv"
 # VIBRANT
-vibrant_path <- "TestingSetToolOutputs/testing_sets_vibrant_output.tsv"
+vibrant_path <- "../ToolOutput/testing_sets_vibrant_output.tsv"
 # DeepVirFinder
-dvf_path <- "TestingSetToolOutputs/testing_sets_deepvirfinder_output.tsv"
+dvf_path <- "../ToolOutput/testing_sets_deepvirfinder_output.tsv"
 # Virsorter
-vs_path <- "TestingSetToolOutputs/testing_sets_virsorter_output.tsv"
+vs_path <- "../ToolOutput/testing_sets_virsorter_output.tsv"
 # Virsorter2
-vs2_path <- "TestingSetToolOutputs/testing_sets_virsorter2_output.tsv"
+vs2_path <- "../ToolOutput/testing_sets_virsorter2_output.tsv"
 # Kaiju
-kj_path <- "TestingSetToolOutputs/merged.nreuk.kaiju.names.out"
+kj_path <- "../ToolOutput/merged.nreuk.kaiju.names.out"
 # KB cutoff
@@
 
+this rule set is based on an original set of rules considered for analyzing the
+drinking water virome. upon analyzing the sequences, many false positive viruses
+were identified (particularly long sequences with no viral genes)
+note: checkV is used as a viral sorting tool
+note: VirFinder was used not DeepVirFinder
+```{r getting_viral_set_2}
+getting_viral_set_2 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_checkV=FALSE,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
+      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
+    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
+      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
+  }
+
+  if (include_deepvirfinder) {
+    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
+  }
+
+  if (include_checkV) {
+    keep_score[input_seqs$checkv_quality=="low"] <- keep_score[input_seqs$checkv_quality=="low"] + 0.3
+    keep_score[input_seqs$checkv_quality=="medium"] <- keep_score[input_seqs$checkv_quality=="medium"] + 0.5
+    keep_score[input_seqs$checkv_quality=="high"] <- keep_score[input_seqs$checkv_quality=="medium"] + 1
+
+    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>1] <- 0
+
+  }
+
+  return(keep_score)
+
+}
+
+```
+
+this is the set of rules used in the drinking water viromes paper
+note: VirFinder was used not DeepVirFinder
+```{r getting_viral_set_3}
+getting_viral_set_3 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_checkV=FALSE,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
+      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
+    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
+      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
+  }
+
+  if (include_deepvirfinder) {
+    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
+  }
+
+  if (include_checkV) {
+    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
+    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
+    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
+  }
+
+  return(keep_score)
+
+}
+
+```
+
 # Assessing performance against the "truth"
@@
 ```{r}
-build_score_list <- function(input_seqs, combos) {
+build_score_list <- function(input_seqs, combos, pipeline) {
   output <- data.frame(precision=rep(0, nrow(combos)),
@@
   for (i in 1:nrow(combos)) {
+    if (pipeline==1) {
       keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
@@
                                               include_deepvirfinder = combos$DVF[i])
+    }
+    if (pipeline==2) {
+      keep_score <- getting_viral_set_2(input_seqs, include_vibrant = combos$VIBRANT[i],
+                                              include_virsorter = combos$VS[i],
+                                              include_virsorter2 = combos$VS2[i],
+                                              include_checkV = combos$CheckV[i],
+                                              include_deepvirfinder = combos$DVF[i])
+    }
+
+    if (pipeline==3) {
+      keep_score <- getting_viral_set_3(input_seqs, include_vibrant = combos$VIBRANT[i],
+                                              include_virsorter = combos$VS[i],
+                                              include_virsorter2 = combos$VS2[i],
+                                              include_checkV = combos$CheckV[i],
+                                              include_deepvirfinder = combos$DVF[i])
+    }
 
@@
 ## Calculate the performance of each pipeline
+first rule set
 ```{r}
-accuracy_scores <- build_score_list(viruses[viruses$Index==1], combos_list)
-
 accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
@@
 accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==1], combos_list))
+                              build_score_list(viruses[viruses$Index==1], combos_list, pipeline=1))
 
@@
                            cbind(testing_set_index=rep(i, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==i], combos_list)))
+                              build_score_list(viruses[viruses$Index==i], combos_list, pipeline=1)))
 }
 
+```
+
+```{r}
+library("stringr")
+
+accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
+
+accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
+
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+
+accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
 ```
@@
 
-p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1)) +
+p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
+                                  color=numtools, fill=numtools)) +
   geom_point(alpha=0.5) +
@@
 ```
+## Calculate the performance of each pipeline
+original rule set from DW paper (didn't use)
+
+```{r}
+combos_list_3 <- read_csv("combinations_list_3.csv")
+```
+
+```{r}
+accuracy_scores_2 <- cbind(testing_set_index=rep(0, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==1], combos_list_3, pipeline=2))
+
+for (i in 2:10) {
+  accuracy_scores_2 <- rbind(accuracy_scores_2,
+                           cbind(testing_set_index=rep(i, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==i], combos_list_3, pipeline=2)))
+}
+
+```
+
+## Visualize how the precision, recall, and F1 scores change across pipelines.
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+ggplot(accuracy_scores_2, aes(x=toolcombo, y=F1)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("F1 Score")
+
+ggplot(accuracy_scores_2, aes(x=toolcombo, y=precision)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Precision")
+
+ggplot(accuracy_scores_2, aes(x=toolcombo, y=recall)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Recall")
+```
+
+## Calculate the performance of each pipeline
+rule set from DW paper
+
+```{r}
+combos_list_3 <- read_csv("combinations_list_3.csv")
+```
+
+```{r}
+accuracy_scores_3 <- data.frame(testing_set_index=rep(0, nrow(combos_list_3)*10),
+                      precision=rep(0, nrow(combos_list_3)*10),
+                       recall=rep(0, nrow(combos_list_3)*10),
+                       F1=rep(0, nrow(combos_list_3)*10))
+
+accuracy_scores_3 <- cbind(testing_set_index=rep(1, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==1], combos_list_3, pipeline=3))
+
+for (i in 2:10) {
+  accuracy_scores_3 <- rbind(accuracy_scores_3,
+                           cbind(testing_set_index=rep(i, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==i], combos_list_3, pipeline=3)))
+}
+
+```
+
+## Visualize how the precision, recall, and F1 scores change across pipelines.
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+p2 <- ggplot(accuracy_scores_3, aes(x=toolcombo, y=F1)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("F1 Score")
+
+p2
+
+ggplot(accuracy_scores_3, aes(x=toolcombo, y=precision)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Precision")
+
+ggplot(accuracy_scores_3, aes(x=toolcombo, y=recall)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Recall")
+```

### 2022-09-09 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updated VirSorter rule comment.
Playing around with other rule sets for DW data.

--- 2022-08-10 Data/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
@@
 Virsorter
-    category <= 5: +1
-    category == 6: +0.5
+    category ==  1,2,4,5: +1
+    category == 3,6: +0.5
 
@@
 
-## just vibrant
+## vibrant virsorter2 virsorter checkV
 ```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
+viruses$keep_score_vbvs2vscv <- getting_viral_set_1(viruses, include_vibrant = T,
+                                            include_virsorter2 = T,
+                                            include_checkV = T,
+                                            include_virsorter = T,
                                             include_kaiju = F,
@@
 
-## vibrant + checkV
 ```{r}
-viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
+table(viruses$keep_score_vbvs2vscv>=1)
+```
+
+```{r}
+getting_viral_set_3 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_checkV=FALSE,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
+      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
+    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
+      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
+  }
+
+  if (include_deepvirfinder) {
+    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
+  }
+
+  if (include_checkV) {
+    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
+    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
+    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
+  }
+
+  return(keep_score)
+
+}
+```
+
+## vibrant virsorter2 virsorter checkV
+```{r}
+viruses$keep_score_vbvs2vscv_3 <- getting_viral_set_3(viruses, include_vibrant = T,
+                                            include_virsorter2 = T,
                                             include_checkV = T,
-                                            include_kaiju = F,
+                                            include_virsorter = T,
                                             include_deepvirfinder = F)
@@
 
-## vibrant + checkV + kaiju
 ```{r}
-viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
+table(viruses$keep_score_vbvs2vscv_3>=1)
+```
+
+```{r}
+getting_viral_set_2 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_checkV=FALSE,
+                                include_virsorter=FALSE) {
+
+    keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
+      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
+    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
+      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
+  }
+
+  if (include_deepvirfinder) {
+    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
+  }
+
+  if (include_checkV) {
+    keep_score[input_seqs$checkv_quality=="low"] <- keep_score[input_seqs$checkv_quality=="low"] + 0.3
+    keep_score[input_seqs$checkv_quality=="medium"] <- keep_score[input_seqs$checkv_quality=="medium"] + 0.5
+    keep_score[input_seqs$checkv_quality=="high"] <- keep_score[input_seqs$checkv_quality=="medium"] + 1
+
+    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>1] <- 0
+  }
+
+  return(keep_score)
+
+}
+```
+
+## vibrant virsorter2 virsorter checkV
+```{r}
+viruses$keep_score_vbvs2vscv_2 <- getting_viral_set_2(viruses, include_vibrant = T,
+                                            include_virsorter2 = T,
                                             include_checkV = T,
-                                            include_kaiju = T,
+                                            include_virsorter = T,
                                             include_deepvirfinder = F)
@@
 
-## vibrant + checkV + kaiju + deepvirfinder
 ```{r}
-viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T)
-```
-
-## vibrant + virsorter2 + checkV + kaiju
-```{r}
-viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F)
-```
-
-#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
-```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-```
-
-## just virsorter2
-```{r}
-viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-```
-
-## virsorter2 + checkV
-```{r}
-viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-```
-
-## virsorter2 + checkV + kaiju
-```{r}
-viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F)
-```
-
-## virsorter2 + checkV + kaiju + deepvirfinder
-```{r}
-viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T)
-```
-
-## just checkV
-```{r}
-viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-```
-
-## checkV + kaiju
-```{r}
-viruses$keep_score_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F)
-```
-
-## just kaiju
-```{r}
-viruses$keep_score_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = F)
-```
-
-## just deepvirfinder
-```{r}
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = T)
-```
-
-## deepvirfinder + checkV
-```{r}
-viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = T)
-```
-
-## deepvirfinder + checkV + kaiju
-```{r}
-viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = F)
-```
-
-## virsorter2 + checkV + kaiju
-```{r}
-viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T)
-```
-
-## just virsorter
-```{r}
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-```
-
-## virsorter + checkV
-```{r}
-viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-```
-
-## virsorter + checkV + kaiju
-```{r}
-viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-```
-
-## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
-```{r}
-viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-```
-
-## virsorter + checkV + kaiju + vibrant
-```{r}
-viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-```
-
-## virsorter + checkV + kaiju + virsorter2 + vibrant
-```{r}
-viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-```
-
-## virsorter + checkV + kaiju + virsorter2
-```{r}
-viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
+table(viruses$keep_score_vbvs2vscv_2>=1)
 ```

### 2022-09-09 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Experimenting with new DVF rules accounting for length.

@@
   if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
-#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
+    keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<10000] <- keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<10000] + 0.5
+   keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<10000] <- keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<10000] + 1
   }
@@
 
+```{r}
+library("stringr")
+```
+
+```{r}
+accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
+accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
+```
+
 ## Visualize how the precision, recall, and F1 scores change across pipelines.
@@
 pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
-p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1)) +
+p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
+                                  color=numtools, fill=numtools)) +
   geom_point(alpha=0.5) +
@@
   ylab("F1 Score")
-
 p2
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=precision)) +
@@
   ylab("Precision")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=recall)) +

### 2022-09-13 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations of testing sets.

@@
     Viral >= 0.95: +1
-    Hallmark > 2: +1
 
@@
     Score >= 0.7: +0.5
+    Score >= 0.9: +0.5
 
@@
 CheckV
+    If %unknown >= 75: +0.5
+    Hallmark > 2: +1
     viral_genes == 0 and host_genes >= 1: keep_score = 0
@@
     If length > 50,000 and hallmark == 0: keep_score = 0
-    If %unknown >= 75: +0.5
 
@@
                                 include_deepvirfinder=FALSE,
-                                include_checkV=FALSE,
+                                include_tuning=FALSE,
                                 include_kaiju=FALSE,
@@
     keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+#    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
   }
@@
     keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
   }
@@
   if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<10000] <- keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<10000] + 0.5
-   keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<10000] <- keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<10000] + 1
+    keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<20000] + 0.5
+   keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<20000] + 0.5
   }
@@
 
-  if (include_checkV) {
+  if (include_tuning) {
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
     keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
+    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
     keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
@@
 
-  precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])
+  TP <- table(confusion_matrix)[4]
+  FP <- table(confusion_matrix)[2]
+  TN <- table(confusion_matrix)[3]
+  FN <- table(confusion_matrix)[1]
 
-  recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])
+  precision <- TP/(TP+FP)
+
+  recall <- TP/(TP+FN)
 
@@
 
-  performance <- c(precision, recall, F1)
-  names(performance) <- c("precision", "recall", "F1")
+  MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
+
+  auc <- round(auc(truepositive, keep_score),4)
+
+  #by type metrics
+  fungal_FP <- table(confusion_matrix[seqtype=="fungi"])[2]
+  bacterial_FP <- table(confusion_matrix[seqtype=="bacteria"])[2]
+  viral_FN <- table(confusion_matrix[seqtype=="virus"])[1]
+
+  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, bacterial_FP, viral_FN)
+  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
+                          "bacterial_FP", "viral_FN")
 
@@
                        recall=rep(0, nrow(combos)),
-                       F1=rep(0, nrow(combos)))
+                       F1=rep(0, nrow(combos)),
+                       MCC=rep(0, nrow(combos)),
+                       AUC=rep(0, nrow(combos)),
+                       fungal_FP=rep(0, nrow(combos)),
+                       bacterial_FP=rep(0, nrow(combos)),
+                       viral_FN=rep(0, nrow(combos)))
   for (i in 1:nrow(combos)) {
@@
                                             include_virsorter2 = combos$VS2[i],
-                                            include_checkV = combos$CheckV[i],
+                                            include_tuning = combos$CheckV[i],
                                             include_kaiju = combos$Kaiju[i],
@@
 
-    output[i,1:3] <- assess_performance(input_seqs$seqtype, keep_score)
+    output[i,1:8] <- assess_performance(input_seqs$seqtype, keep_score)
 
@@
                        recall=rep(0, nrow(combos_list)*10),
-                       F1=rep(0, nrow(combos_list)*10))
+                       F1=rep(0, nrow(combos_list)*10),
+                       MCC=rep(0, nrow(combos_list)*10), 
+                      AUC=rep(0, nrow(combos_list)*10),
+                      fungal_FP=rep(0, nrow(combos_list)*10),
+                      bacterial_FP=rep(0, nrow(combos_list)*10),
+                      viral_FN=rep(0, nrow(combos_list)*10))
 
@@
 accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
-accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
+#accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
+accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
 accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
@@
 p2
-ggplot(accuracy_scores, aes(x=toolcombo, y=precision)) +
+ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
+                                  color=numtools, fill=numtools)) +
   geom_point(alpha=0.5) +
@@
   ylab("Precision")
-ggplot(accuracy_scores, aes(x=toolcombo, y=recall)) +
+ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
+                                  color=numtools, fill=numtools)) +
   geom_point(alpha=0.5) +
@@
   ylab("Recall")
+
+ggplot(accuracy_scores, aes(x=precision, y=recall, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Precision") +
+  ylab("Recall")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Precision-Recall")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("MCC")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("AUC")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Fungal False Positives")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Bacterial False Positives")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Viral False Negatives")
+
 ```
+
+# Experimenting
+
+```{r}
+viruses$keep_score <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses$confusion_matrix <- "true negative"
+viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score<=1] <- "false negative"
+viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score>=1] <- "true positive"
+viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score>=1] <- "false positive"
+```
+
+visualizing confusion matrix by taxa
+```{r}
+confusion_by_taxa <- melt(table(viruses$confusion_matrix, viruses$seqtype, viruses$Index))
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+p2
+```
+note: are these fungal actually false negatives? At least one was classified as viral by Kaiju.
+
+# Visualizing confusion matrix by number of tools
+
+```{r}
+viruses$keep_score_visualize <- viruses$keep_score
+viruses$keep_score_visualize[viruses$keep_score>1] <- ">1"
+
+viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
+                                       levels=c("-0.5", "-1", "0", "0.5","1", ">1"))
+viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
+                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
+```
+
+```{r}
+levels(factor(viruses$keep_score_visualize))
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
+
+p1 <- ggplot(viruses, aes(x=as.factor(Index),
+                   fill=keep_score_visualize, color=keep_score_visualize)) +
+  geom_bar(stat="count", position="stack") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16)
+  ) +
+  scale_color_manual(name = 'Viral Score',
+                     values = alpha(c(pal(4)), 1)) +
+  scale_fill_manual(name = 'Viral Score',
+                     values = alpha(c(pal(4)), 0.5)) +
+  xlab("Index") +
+  ylab("Sequence Count") +
+  facet_wrap(~confusion_matrix, scales = "free")
+
+p1
+```
+
+# ROC 
+```{r}
+library(pROC)
+```
+
+```{r}
+viruses$truepositive <- rep(0, nrow(viruses))
+viruses$truepositive[viruses$seqtype=="virus"] <- 1
+```
+
+```{r}
+rocobj <- roc(viruses$truepositive, viruses$keep_score)
+rocobj_all <- roc(viruses$truepositive, viruses$keep_score_all)
+
+auc <- round(auc(viruses$truepositive, viruses$keep_score),4)
+auc_all <- round(auc(viruses$truepositive, viruses$keep_score_all),4)
+
+#create ROC plot
+ggroc(rocobj, colour = 'steelblue', size = 2) +
+  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')')) +
+  coord_equal()
+
+ggroc(rocobj_all, colour = 'green', size = 2) +
+  ggtitle(paste0('ROC Curve ', '(AUC = ', auc_all, ')'))
+```
+Sensitivity: The probability that the model predicts a positive outcome for an observation when indeed the outcome is positive.
+Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative.

### 2022-09-13 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updating rule tuning script visualizations.

@@
 KB_CUTOFF <- 3000
-
 ```
@@
 dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
-
 dvf_c <- separate(dvf_c, col = contig, into = c("seqtype", "contig"), sep="--")
@@
         )
-
 vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
@@
 vs_c$contig <- sub("\\.", "_", vs_c$contig)
-
 vs_c <- vs_c %>% drop_na(contig)
@@
 kj_c <- read_tsv(kj_path, col_names = T)
-
 kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
-
 kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
 kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
-
 kj_c$contig <- sub("\\.", "_", kj_c$contig)
@@
 viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
-
 # VIBRANT
 viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
-
 # DeepVirFinder
@@
 viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
-
 # VirSorter2
@@
 viruses$hallmark[is.na(viruses$hallmark)] <- 0
-
 #Virsorter
 viruses$category[is.na(viruses$category)] <- 0
-
 # Kaiju
@@
   }
-
   if (include_kaiju) {
@@
 }
-
 ```
@@
   precision <- TP/(TP+FP)
-
   recall <- TP/(TP+FN)
-
   F1 <- 2*precision*recall/(precision+recall)
@@
 accuracy_scores <- build_score_list(viruses[viruses$Index==1], combos_list)
-
 accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
@@
                       viral_FN=rep(0, nrow(combos_list)*10))
-
 accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                               build_score_list(viruses[viruses$Index==1], combos_list))
-
 for (i in 2:10) {
@@
   ylab("Recall")
-
 ggplot(accuracy_scores, aes(x=precision, y=recall, 
@@
   ylab("Recall")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
@@
   ylab("Precision-Recall")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
@@
   ylab("MCC")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
@@
   ylab("AUC")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
@@
   ylab("Fungal False Positives")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
@@
   ylab("Bacterial False Positives")
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
@@
   ylab("Viral False Negatives")
-
 ```
@@
   coord_flip()
-
 p2
@@
 viruses$keep_score_visualize[viruses$keep_score>1] <- ">1"
-
 viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
@@
 pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
-
 p1 <- ggplot(viruses, aes(x=as.factor(Index),
@@
   facet_wrap(~confusion_matrix, scales = "free")
-
 p1
@@
 rocobj_all <- roc(viruses$truepositive, viruses$keep_score_all)
-
 auc <- round(auc(viruses$truepositive, viruses$keep_score),4)
 auc_all <- round(auc(viruses$truepositive, viruses$keep_score_all),4)
-
 #create ROC plot
@@
   coord_equal()
-
 ggroc(rocobj_all, colour = 'green', size = 2) +

### 2022-09-15 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Comparing high precision vs high recall pipeline types.

@@
     Viral >= 50: +0.5
-    Viral >= 0.95: +1
+    Viral >= 0.95: +0.5
 
@@
 
+## high precision example
 ```{r}
-viruses$keep_score <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                               include_vibrant = T,
-                                              include_virsorter2 = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+viruses$confusion_matrix_high_precision <- "true negative"
+viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<=1] <- "false negative"
+viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision>=1] <- "true positive"
+viruses$confusion_matrix_high_precision[viruses$seqtype!="virus" & viruses$keep_score_high_precision>=1] <- "false positive"
+```
+
+visualizing confusion matrix by taxa
+```{r}
+confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_precision, viruses$seqtype, viruses$Index))
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+```
+
+```{r}
+ ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix,
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_boxplot(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Viral Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ ggplot(viruses, aes(x=percent_viral, y=confusion_matrix,
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_boxplot(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Percent Genes Viral") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ ggplot(viruses, aes(x=hallmark, y=confusion_matrix,
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_boxplot(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Hallmark Genes") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Number of Viral Genes") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+```
+
+```{r}
+table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)
+
+table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
+```
+
+## high recall example
+```{r}
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
                                               include_kaiju = T,
@@
 viruses$confusion_matrix <- "true negative"
-viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score<=1] <- "false negative"
-viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score>=1] <- "true positive"
-viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score>=1] <- "false positive"
+viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_high_recall<=1] <- "false negative"
+viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
+viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
 ```
@@
 ```
-note: are these fungal actually false negatives? At least one was classified as viral by Kaiju.
 

### 2022-09-21 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Plots to examine false positives and negatives.

@@
 checkV$contig <- sub("\\|", "_", checkV$contig)
+checkV <- checkV[!duplicated(paste(checkV$contig, checkV$Index)),]
 ```
@@
 vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
-vb_c <- vb_c[!duplicated(vb_c$contig),]
+vb_c <- vb_c[!duplicated(paste(vb_c$contig, vb_c$Index)),]
 ```
@@
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
+dvf_c <- dvf_c[!duplicated(paste(dvf_c$contig, dvf_c$Index)),]
 ```
@@
 vs_c <- vs_c %>% drop_na(contig)
-vs_c <- vs_c[!duplicated(vs_c$contig),]
+vs_c <- vs_c[!duplicated(paste(vs_c$contig, vs_c$Index)),]
 ```
@@
 vs2_c <- separate(vs2_c, col = contig, into = c("seqtype", "contig"), sep="--")
-vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
+vs2_c <- vs2_c[!duplicated(paste(vs2_c$contig, vs2_c$Index)),]
 ```
@@
 kj_c$contig <- sub("\\.", "_", kj_c$contig)
+kj_c <- kj_c[!duplicated(paste(kj_c$contig, kj_c$Index)),]
 ```
@@
   }
+
   if (include_kaiju) {
@@
     keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
+    #keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] <- keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] + 1 #add some ratio
     keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
+    keep_score[((input_seqs$viral_genes*3) <= input_seqs$host_genes) & input_seqs$checkv_provirus=="No"] <- 0 # consider accounting for provirus designation
+#    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0 # consider accounting for provirus designation
     keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
@@
 ```{r}
- ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix,
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
+ ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix_high_precision,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
   geom_boxplot(alpha=0.3) +
@@
 
- ggplot(viruses, aes(x=percent_viral, y=confusion_matrix,
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
+ ggplot(viruses, aes(x=percent_viral, y=confusion_matrix_high_precision,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
   geom_boxplot(alpha=0.3) +
@@
 
- ggplot(viruses, aes(x=hallmark, y=confusion_matrix,
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
+ ggplot(viruses, aes(x=hallmark, y=confusion_matrix_high_precision,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
   geom_boxplot(alpha=0.3) +
@@
 ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
   geom_point(alpha=0.3) +
@@
   facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+```
+
+```{r}
+viruses_false_positive <- viruses[viruses$confusion_matrix_high_precision=="false positive",]
+viruses_false_negative <- viruses[viruses$confusion_matrix_high_precision=="false negative",]
+```
+
+```{r}
+ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
+                   fill=checkv_length,
+                   color=checkv_length,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Number of Viral Genes") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive, aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive[viruses_false_positive$seqtype=="bacteria"], aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive[viruses_false_positive$seqtype=="fungi"], aes(x=hallmark, y=checkv_length,
+                   fill=keep_score_high_precision,
+                   color=keep_score_high_precision,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive[viruses_false_positive$seqtype=="protist"], aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
+                   fill=keep_score_high_precision,
+                   color=keep_score_high_precision,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
   coord_flip()

### 2022-09-23 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Compressed kaiju output file.

@@
 # DeepVirFinder
-dvf_path <- "DrinkingWaterToolOutput/merged_lake_michigan_dvf_assemblies.tsv"
+dvf_path <- "DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
 # Virsorter
@@
 # Kaiju
-kj_path <- "DrinkingWaterToolOutput/merged_lake_michigan_kaiju_assemblies.tsv"
+kj_path <- "DrinkingWaterToolOutput/kaju_outputs_merged.tsv"
 # KB cutoff
@@
 checkV$uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
+checkV <- checkV[!duplicated(checkV$uniq_contig),]
 ```
@@
 
-## DeepVirFinder - to run
+## DeepVirFinder
 ```{r}
@@
                select = c(
+                   'sample',
                    'name',
-                   'assembly',
                    'score',
@@
     rename(
-        contig = name
+        contig = name,
+        assembly = sample
         )
 dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
-
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
+dvf_c$uniq_contig <- paste(dvf_c$assembly, dvf_c$contig, sep = "--")
+dvf_c <- dvf_c[!duplicated(dvf_c$uniq_contig),]
 ```
@@
 
-## Kaiju - to run
-(parsing when reading the file in "fails" for sequences that weren't classified)
+## Kaiju
 ```{r}
-kj_c <- fread(kj_path)
-
-colnames(kj_c) <-c(
-    "Classified",
-    "Contig",
-    "NCBI_taxon",
-    "len",
-    "ID_best",
-    "IDs_all",
-    "Seq",
-    "Name"
-    )
-kj_c$Contig <- gsub('_', '', kj_c$Contig)
-kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
-
-kj_c <- setDT(kj_c)[, c("assembly", "contig") := tstrsplit(Contig, "--")]
-kj_c <- setDT(kj_c)[, c("Kaiju_Viral", "Kingdom") := tstrsplit(Name, ";", keep = c(1, 2))]
+kj_c <- read_tsv(kj_path, col_names = T)
+colnames(kj_c)[1] <- "assembly"
+colnames(kj_c)[3] <- "contig"
+kj_c$contig <- sub("\\.", "_", kj_c$contig)
+kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
+kj_c$uniq_contig <- paste(kj_c$assembly, kj_c$contig, sep="--")
+kj_c <- kj_c[!duplicated(kj_c$contig),]
 ```
@@
   if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
-#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
+    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
+   keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
   }
@@
 
-  if (include_checkV) {
+  if (include_tuning) {
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
+    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
+    #keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] <- keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] + 1 #add some ratio
     keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
+    keep_score[((input_seqs$viral_genes*3) <= input_seqs$host_genes) & input_seqs$checkv_provirus=="No"] <- 0 # consider accounting for provirus designation
+#    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0 # consider accounting for provirus designation
     keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
-    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
   }
@@
 ```{r}
-viruses$keep_score_vbvs2vscv <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_virsorter = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = F)
 ```
@@
 ```{r}
-table(viruses$keep_score_vbvs2vscv>=1)
+table(viruses$keep_score_high_precision>=1)
 ```
@@
 ```{r}
-getting_viral_set_3 <- function(input_seqs,
-                                include_vibrant=FALSE, 
-                                include_virsorter2=FALSE,
-                                include_deepvirfinder=FALSE,
-                                include_checkV=FALSE,
-                                include_virsorter=FALSE) {
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
 
-  keep_score <- rep(0, nrow(input_seqs))
+viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
 
-  if (include_vibrant) {
-    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
-      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
-    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
-  }
+viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
 
-  if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
-    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
-      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    }
+viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = T)
 
-  if (include_virsorter) {
-    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
-    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
-  }
+viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = T,
+                                              include_virsorter = F)
 
-  if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
-    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
-  }
-
-  if (include_checkV) {
-    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
-    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
-  }
-
-  return(keep_score)
-
-}
+viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = T,
+                                              include_tuning = F,
+                                              include_virsorter = F)
 ```
 
-## vibrant virsorter2 virsorter checkV
-```{r}
-viruses$keep_score_vbvs2vscv_3 <- getting_viral_set_3(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_virsorter = T,
-                                            include_deepvirfinder = F)
-```
-
-```{r}
-table(viruses$keep_score_vbvs2vscv_3>=1)
-```
-
-```{r}
-getting_viral_set_2 <- function(input_seqs,
-                                include_vibrant=FALSE, 
-                                include_virsorter2=FALSE,
-                                include_deepvirfinder=FALSE,
-                                include_checkV=FALSE,
-                                include_virsorter=FALSE) {
-
-    keep_score <- rep(0, nrow(input_seqs))
-
-  if (include_vibrant) {
-    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
-      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
-    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
-  }
-
-  if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
-    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
-      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    }
-
-  if (include_virsorter) {
-    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
-    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
-  }
-
-  if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
-    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
-  }
-
-  if (include_checkV) {
-    keep_score[input_seqs$checkv_quality=="low"] <- keep_score[input_seqs$checkv_quality=="low"] + 0.3
-    keep_score[input_seqs$checkv_quality=="medium"] <- keep_score[input_seqs$checkv_quality=="medium"] + 0.5
-    keep_score[input_seqs$checkv_quality=="high"] <- keep_score[input_seqs$checkv_quality=="medium"] + 1
-
-    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>1] <- 0
-  }
-
-  return(keep_score)
-
-}
-```
-
-## vibrant virsorter2 virsorter checkV
-```{r}
-viruses$keep_score_vbvs2vscv_2 <- getting_viral_set_2(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_virsorter = T,
-                                            include_deepvirfinder = F)
-```
-
-```{r}
-table(viruses$keep_score_vbvs2vscv_2>=1)
-```
-
- TODO WHAT DO I DO ABOUT THE ZEROS????
 # Plot of how many methods identify a contig
 ```{r}
-viruses_high <- viruses[viruses$keep_score_all>=1,]
-viruses_high_mod <- viruses_high %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_vs_cv, keep_score_cv_kj)
+#viruses_high <- viruses[viruses$keep_score_all>=1,]
+viruses_high_mod <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
 viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
@@
 
-sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb_cv>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf_cv>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_cv_kj>0,]))
+sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,]))
                           )
@@
     rbind(
-        virusHighCount(viruses_high_mod, "keep_score_vb_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_dvf_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_vs_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_cv_kj")
+        virusHighCount(viruses_high_mod, "keep_score_vb"),
+        virusHighCount(viruses_high_mod, "keep_score_dvf"),
+        virusHighCount(viruses_high_mod, "keep_score_vs2"),
+        virusHighCount(viruses_high_mod, "keep_score_vs"),
+        virusHighCount(viruses_high_mod, "keep_score_kj")
     )
@@
 ```
-
-# Visualizing which tools led to a contig being selected
-(alluvial plot)
-with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/
-
-```{r}
-alluvial_data <- viruses %>% select(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
-  apply(c(1,2), function(x) {if (x >= 1) {x <- 1} else if (x==0.5) {x <- 0.5} else {x <- 0}}) %>%
-  as_tibble() %>%
-  count(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
-```
-
-```{r}
-alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv, keep_score_vs2_cv, keep_score_cv_kj) %>% 
-  rowSums()
-```
-
-```{r}
-alluvial_data$keep_score <- "0-0.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=1] <- "1-1.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=2] <- "2-2.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=3] <- "3-3.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=4] <- "4-4.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=5] <- "5+"
-
-alluvial_data$keep_score <- factor(
-    alluvial_data$keep_score,
-    levels=c(
-        "0-0.5",
-        "1-1.5",
-        "2-2.5",
-        "3-3.5",
-        "4-4.5",
-        "5+"
-        )
-    )
-```
-
-```{r}
-alluvial_data <- alluvial_data %>% 
-  mutate(
-      across(
-          c(keep_score_vb_cv,
-            keep_score_vs_cv,
-            keep_score_dvf_cv,
-            keep_score_vs2_cv,
-            keep_score_cv_kj
-            ),
-          as.factor
-          )
-      )
-```
-
-```{r}
-head(alluvial_data)
-```
-
-```{r}
-library(ggalluvial)
-```
-
-```{r}
-ggplot(data = alluvial_data,
-       aes(axis1 = as.character(keep_score_vb_cv), 
-           axis2 = as.character(keep_score_vs_cv),
-           axis3 = as.character(keep_score_vs2_cv),
-           axis4 = as.character(keep_score_dvf_cv),
-           axis5 = as.character(keep_score_cv_kj),
-           y = n)) +
-  geom_alluvium(aes(fill = keep_score)) +
-  geom_stratum() +
-  geom_text(stat = "stratum",
-            aes(label = after_stat(stratum))) +
-  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "Virsorter", "VirSorter2", "DeepVirFinder", "Kaiju"),
-                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
-  ylab("Number of Sequences") +
-  xlab("method") +
-  theme_bw()
-```
-
-Get proportion of contigs from "viruses" classified as virus by combined tool:
-
-```{r}
-prop_viral <- paste(nrow(viruses_high) / nrow(viruses) * 100, "%", sep="")
-prop_viral
-```
-
-Distribution of keep_score across "viruses" dataframe.
-```{r}
-keep_scores <- reshape2::melt(table(viruses$keep_score_all))
-
-colnames(keep_scores) <- c('keep_score', 'count')
-keep_scores[,'keep_score'] <- as.factor(keep_scores[,'keep_score'])
-
-ggplot(data = keep_scores, aes(x = keep_score, y = count)) +
-    geom_bar(stat="identity", width = 0.3, fill = 'steelblue')
-```
-
-Determine number of viral sequences with kaiju taxonomic classification
-
-```{r}
-tax_counts <- viruses_high %>% count(Kaiju_Viral, name = "num")
-tax_counts$prop <- tax_counts$num / sum(tax_counts$num)
-```

### 2022-09-27 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
New visualizations of drinking water viruses.

@@
 # Kaiju
-kj_path <- "DrinkingWaterToolOutput/kaju_outputs_merged.tsv"
+kj_path <- "DrinkingWaterToolOutput/kaiju_outputs_merged.tsv"
 # KB cutoff
@@
 ## Merging
-to do: this step isn't working properly
 ```{r}
@@
 viruses_filename <- 'merged_viral_id_outputs.csv'
-write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
+#write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
 ```
@@
 ```
+
+# Comparing all of the tool methods
+
+```{r}
+combos_list <- data.frame(toolcombo=rep(0, 64),
+                          CheckV=rep(0, 64),
+                          DVF=rep(0, 64),
+                          Kaiju=rep(0, 64),
+                          VIBRANT=rep(0, 64),
+                          VS=rep(0, 64),
+                          VS2=rep(0, 64))
+p <- 1
+
+for (i in c(0,1)){
+  for (j in c(0,1)){
+    for (k in c(0,1)){
+      for (l in c(0,1)){
+        for (m in c(0,1)){
+          for (n in c(0,1)){
+            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
+            combos_list$CheckV[p] <- i
+            combos_list$DVF[p] <- j
+            combos_list$Kaiju[p] <- k
+            combos_list$VIBRANT[p] <- l
+            combos_list$VS[p] <- m
+            combos_list$VS2[p] <- n
+            p <- p+1
+          }
+        }
+      }
+    }
+  }
+}
+
+combos_list <- combos_list[-1,]
+```
+
+```{r}
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
+                          num_viruses=rep(0, nrow(combos_list)))
+
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning = combos_list$CheckV[i],
+                                            include_kaiju = combos_list$Kaiju[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+
+  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+}
+
+num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
+viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
+viral_scores_nozeros <- viral_scores_nozeros + 1
+viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
+
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo
+```
+
+```{r}
+library(phyloseq)
+```
+
+```{r}
+tooldata <- num_viruses
+rownames(tooldata) <- tooldata$toolcombo
+```
+
+```{r}
+physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
+                                     sample_data(tooldata))
+
+ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw() +
+  geom_label(label=tooldata$toolcombo)
+
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw()
+```
+to do: try coloring above based on the F1 scores of the testing set on each combination
+
+```{r}
+write_tsv(num_viruses, "dw_num_viruses_by_tool_combo.tsv")
+num_viruses <- read_tsv("dw_num_viruses_by_tool_combo.tsv")
+```
+
+```{r}
+accuracy_scores <- read_tsv("../../TestingPipeline/Scripts/20220927_accuracy_scores.tsv")
+```
+
+```{r}
+num_viruses_comp <- full_join(num_viruses, accuracy_scores[accuracy_scores$testing_set_index==1,], by=c("toolcombo", "numtools"))
+```
+
+```{r}
+num_viruses_comp$numtools <- as.factor(num_viruses_comp$numtools)
+```
+
+```{r}
+ggplot(num_viruses_comp, aes(x=num_viruses, y=F1, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("F1") +
+  xlab("Num Viruses Predicted")
+
+ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Precision") +
+  xlab("Num Viruses Predicted")
+
+ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Recall") +
+  xlab("Num Viruses Predicted")
+```

### 2022-09-27 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations of drinking water viruses.

@@
 ```{r}
-combos_list <- read_csv("combinations_list.csv", col_names = T)
+combos_list <- data.frame(toolcombo=rep(0, 64),
+                          CheckV=rep(0, 64),
+                          DVF=rep(0, 64),
+                          Kaiju=rep(0, 64),
+                          VIBRANT=rep(0, 64),
+                          VS=rep(0, 64),
+                          VS2=rep(0, 64))
+p <- 1
+
+for (i in c(0,1)){
+  for (j in c(0,1)){
+    for (k in c(0,1)){
+      for (l in c(0,1)){
+        for (m in c(0,1)){
+          for (n in c(0,1)){
+            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
+            combos_list$CheckV[p] <- i
+            combos_list$DVF[p] <- j
+            combos_list$Kaiju[p] <- k
+            combos_list$VIBRANT[p] <- l
+            combos_list$VS[p] <- m
+            combos_list$VS2[p] <- n
+            p <- p+1
+          }
+        }
+      }
+    }
+  }
+}
 ```
@@
 ```{r}
-accuracy_scores <- build_score_list(viruses[viruses$Index==1], combos_list)
 accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),

### 2022-10-04 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
More dw visualizations.

@@
             combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
+            combos_list$toolcombo2[p] <- paste(if(i){"C"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"K"}else{"0"},if(l){"VB"}else{"0"},
+                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
             combos_list$CheckV[p] <- i
@@
   num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+
+  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
 }
@@
 num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
 num_viruses$numtools <- as.factor(num_viruses$numtools)
@@
 ```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
 ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+
+ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
                                   color=numtools, fill=numtools)) +
@@
 
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
 ```
@@
 tooldata <- num_viruses
-rownames(tooldata) <- tooldata$toolcombo
+
+rownames(tooldata) <- tooldata$toolcombo2
 ```
@@
                                      sample_data(tooldata))
+```
 
+```{r}
 ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
@@
 to do: try coloring above based on the F1 scores of the testing set on each combination
+
+```{r}
+bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
+clusters <- hclust(dist(bray_dist))
+plot(clusters)
+
+myclusters <- cutree(clusters, h=0.8)
+```
+
+```{r}
+names(myclusters[myclusters==1])
+names(myclusters[myclusters==2])
+names(myclusters[myclusters==3])
+names(myclusters[myclusters==4])
+names(myclusters[myclusters==5])
+
+```
 

### 2022-10-05 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updated figures in the DW analysis R notebook.

@@
 # Kaiju
-kj_path <- "DrinkingWaterToolOutput/kaiju_outputs_merged.tsv"
+kj_path <- "/Users/hegartyb/Documents/BlueSkyPostdoc/ViralSortingToolsEvaluation/Temp/kaiju_outputs_merged.tsv"
 # KB cutoff
@@
 
+## 
+```{r}
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+table(viruses$keep_score_high_precision>=1)
+```
+
 ## vibrant virsorter2 virsorter checkV
@@
                                               include_virsorter = F)
-```
-
-```{r}
-table(viruses$keep_score_high_precision>=1)
 ```
@@
                                               include_virsorter = F)
+viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = T,
+                                              include_tuning = F,
+                                              include_virsorter = F)
 ```
@@
 ```{r}
-#viruses_high <- viruses[viruses$keep_score_all>=1,]
-viruses_high_mod <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
+viruses_high <- viruses[viruses$keep_score_high_precision>=1,] #uncomment this line if high precision example
+#viruses_high <- viruses[viruses$keep_score_high_recall>=1,] #uncomment this line if want high recall example 
+#viruses_high <- viruses[viruses$keep_score_all>=1,] #uncomment this line if want to use all 6 tools
+#viruses_high <- viruses[,] #comment this line if want only the "high precision contigs"
+viruses_high_mod <- viruses_high %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
 viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
@@
 ```{r}
+ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
+  geom_boxplot(aes(color=numtools)) +
+  geom_point(aes(color=numtools, fill=numtools)) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
 viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]

### 2022-10-05 EnvironmentalData/LakeMichiganViromes/identifying_viral_contigs_environmental_lm.Rmd
Added lake michigan and TARA identifying viral contigs directories

--- 2022-09-27 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
@@
 Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
-Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.
-
-TO DO:
-- add kaiju and DVF
+Bridget Hegarty (email) regarding any issues, or open an issue on github.
 
@@
 library(data.table)
+library(here)
+library(stringr)
 ```
@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "drinking_water"
+dataset_name <- "lake_michigan"
 # checkV
-checkV_path <- "DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
+checkV_path <- "merged_outputs/merged_checkv_assemblies.tsv"
 # VIBRANT
-vibrant_path <- "DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
+vibrant_path <- "merged_outputs/merged_vibrant_assemblies.tsv"
 # DeepVirFinder
-dvf_path <- "DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
+dvf_path <- "merged_outputs/merged_dvf_assemblies.tsv"
 # Virsorter
-vs_path <- "DrinkingWaterToolOutput/vs_merged_all.txt"
+vs_path <- "merged_outputs/merged_virsorter_assemblies.tsv"
 # Virsorter2
-vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
+vs2_path <- "merged_outputs/merged_vs2_assemblies.tsv"
 # Kaiju
-kj_path <- "DrinkingWaterToolOutput/kaiju_outputs_merged.tsv"
+kj_path <- "merged_outputs/merged_kaiju_assemblies.tsv"
 # KB cutoff
@@
                     'contig_id',
-                    'assembly',
                     'provirus',
@@
 checkV$method = 'checkv'
-checkV$contig <- sub("\\.", "_", checkV$contig)
-checkV$uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
-checkV <- checkV[!duplicated(checkV$uniq_contig),]
+checkV$contig <- gsub('_', '', checkV$contig)
 ```
@@
                   'scaffold',
-                  'assembly',
                   'type',
-                  'quality'
+                  'Quality'
                   )
@@
         contig = scaffold,
-        vibrant_quality = quality
+        vibrant_quality = Quality
         )
@@
 vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
-vb_c$contig <- sub("\\.", "_", vb_c$contig)
-vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
-vb_c <- vb_c[!duplicated(vb_c$uniq_contig),]
+vb_c$contig <- gsub('_', '', vb_c$contig)
 ```
@@
                select = c(
-                   'sample',
                    'name',
+                   'assembly',
                    'score',
@@
     rename(
-        contig = name,
-        assembly = sample
+        contig = name
         )
-dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
+dvf_c$contig <- gsub('_', '', dvf_c$contig)
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
-dvf_c$uniq_contig <- paste(dvf_c$assembly, dvf_c$contig, sep = "--")
-dvf_c <- dvf_c[!duplicated(dvf_c$uniq_contig),]
 ```
@@
               select = c(
-                  'assembly',
-                  'contig_id',
-                  'category'
-                  ),
-              ) %>% 
-  rename(contig = contig_id) %>%
-  drop_na(contig)
-vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
-vs_c$contig <- sub("-circula", "", vs_c$contig)
-vs_c$contig <- sub("uth", "South", vs_c$contig)
-vs_c$contig <- sub("pannenburg", "Spannenburg", vs_c$contig)
-vs_c$contig <- sub("\\.", "_", vs_c$contig)
-vs_c$uniq_contig <- paste(vs_c$assembly, vs_c$contig, sep = "--")
+                  '## Contig_id',
+                  'Nb genes contigs',
+                  'Nb genes',
+                  'Category',
+                  'Nb phage hallmark genes'
+                  )
+              ) %>% drop_na(`## Contig_id`)
+vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
+vs_c$`## Contig_id` <- sub("-circular", "", vs_c$`## Contig_id`)
+vs_c$`## Contig_id` <- gsub("_", "", vs_c$`## Contig_id`)
+vs_c <- separate(vs_c, col = `## Contig_id`, into = c("assembly", "contig"), sep="--")
 
-vs_c <- vs_c[!duplicated(vs_c$uniq_contig),]
-vs_c <- vs_c %>% drop_na(uniq_contig)
+# convert to numeric
+vs_c$category <- as.numeric(vs_c$Category)
+vs_c <- vs_c %>% drop_na(category)
 ```
@@
                         'seqname',
-                        'assembly',
                         'dsDNAphage',
+                        'RNA',
                         'ssDNA',
@@
                   )
-vs2_c$assembly <- sub("_3000", "", vs2_c$assembly)
-vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
-vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
-vs2_c <- vs2_c[!duplicated(vs2_c$uniq_contig),]
+vs2_c$contig <- gsub("_", "", vs2_c$contig)
 ```
@@
 ## Kaiju
+(parsing when reading the file in "fails" for sequences that weren't classified)
 ```{r}
-kj_c <- read_tsv(kj_path, col_names = T)
-colnames(kj_c)[1] <- "assembly"
-colnames(kj_c)[3] <- "contig"
-kj_c$contig <- sub("\\.", "_", kj_c$contig)
-kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
-kj_c$uniq_contig <- paste(kj_c$assembly, kj_c$contig, sep="--")
-kj_c <- kj_c[!duplicated(kj_c$contig),]
+kj_c <- fread(kj_path)
+
+colnames(kj_c) <-c(
+    "Classified",
+    "Contig",
+    "NCBI_taxon",
+    "len",
+    "ID_best",
+    "IDs_all",
+    "Seq",
+    "Name",
+    "assembly"
+    )
+kj_c$Contig <- gsub('_', '', kj_c$Contig)
+kj_c <- setDT(kj_c)[, c("assembly", "contig") := tstrsplit(Contig, "--")]
+kj_c <- setDT(kj_c)[, c("Kaiju_Viral", "Kingdom") := tstrsplit(Name, ";", keep = c(1, 2))]
 ```
@@
 ## Merging
+
 ```{r}
-viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=vb_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=kj_c, by = c("uniq_contig", "assembly"))
+viruses <- full_join(x=vs2_c, y=dvf_c, by = "contig")
+viruses <- full_join(x=viruses, y=checkV, by = "contig")
+viruses <- full_join(x=viruses, y=vb_c, by = "contig")
+viruses <- full_join(x=viruses, y=vs_c, by = "contig")
+viruses <- full_join(x=viruses, y=kj_c, by = "contig")
 ```
@@
 viruses_filename <- 'merged_viral_id_outputs.csv'
-#write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
+write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
 ```
@@
 Virsorter
-    category ==  1,2,4,5: +1
-    category == 3,6: +0.5
+    Category <= 5: +1
+    Category == 6: +0.5
 
@@
                                               include_virsorter = F)
-
 viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
@@
                                               include_virsorter = F)
-
 viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
@@
                                               include_virsorter = F)
-
 viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
@@
                                               include_virsorter = T)
-
 viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
@@
                                               include_virsorter = F)
-
 viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
@@
                                               include_virsorter = F)
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
 ```
@@
 ```{r}
-#viruses_high <- viruses[viruses$keep_score_all>=1,]
+viruses_high <- viruses[viruses$keep_score_all>=1,]
 viruses_high_mod <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
@@
 viruses_high_mod <- as_tibble(viruses_high_mod)
-
 virusHighCount <- function(df, tool_combo) {
@@
 }
-
 sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
@@
 )
-
 sm <- as.data.frame(
@@
 )
-
 sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "kaiju")
@@
 p <- 1
-
 for (i in c(0,1)){
@@
 }
-
 combos_list <- combos_list[-1,]
@@
                           num_viruses=rep(0, nrow(combos_list)))
-
 for (i in 1:nrow(combos_list)) {
@@
 }
-
 num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
@@
 viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo
+```
 
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo
+```{r}
+# histogram of contig lengths
+ggplot(viruses, aes(x=checkv_length)) + geom_histogram(binwidth=10000)
 ```
@@
                                      sample_data(tooldata))
-
 ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
@@
   geom_label(label=tooldata$toolcombo)
-
 phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
@@
 ```{r}
-write_tsv(num_viruses, "dw_num_viruses_by_tool_combo.tsv")
-num_viruses <- read_tsv("dw_num_viruses_by_tool_combo.tsv")
+write_tsv(num_viruses, "lake_michigan_num_viruses_by_tool_combo.tsv")
+num_viruses <- read_tsv("lake_michigan_num_viruses_by_tool_combo.tsv")
 ```
@@
   xlab("Num Viruses Predicted")
-
 ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
@@
   xlab("Num Viruses Predicted")
-
 ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 

### 2022-10-05 EnvironmentalData/TARAMetagenomes/identifying_viral_contigs_environmental_tara.Rmd
Added lake michigan and TARA identifying viral contigs directories

--- 2022-10-05 EnvironmentalData/LakeMichiganViromes/identifying_viral_contigs_environmental_lm.Rmd
@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "lake_michigan"
+dataset_name <- "tara"
 # checkV
@@
                     'contig_length',
-                    'checkv_quality'
+                    'checkv_quality',
+                    'assembly'
                     )
@@
 checkV$contig <- gsub('_', '', checkV$contig)
+checkV$uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
 ```
@@
                   'scaffold',
+                  'assembly',
                   'type',
@@
 vb_c$contig <- gsub('_', '', vb_c$contig)
+vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
 ```
@@
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
+dvf_c$uniq_contig <- paste(dvf_c$assembly, dvf_c$contig, sep = "--")
 ```
@@
 vs_c$`## Contig_id` <- gsub("_", "", vs_c$`## Contig_id`)
-vs_c <- separate(vs_c, col = `## Contig_id`, into = c("assembly", "contig"), sep="--")
-
+vs_c <- vs_c %>% rename(uniq_contig = `## Contig_id`)
 # convert to numeric
-vs_c$category <- as.numeric(vs_c$Category)
-vs_c <- vs_c %>% drop_na(category)
+vs_c$Category <- as.numeric(vs_c$Category)
+vs_c <- vs_c %>% drop_na(Category)
 ```
@@
                         'viral',
-                        'cellular'
+                        'cellular',
+                        'assembly'
                     )
@@
 vs2_c$contig <- gsub("_", "", vs2_c$contig)
+vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
 ```
@@
     "Classified",
-    "Contig",
+    "uniq_contig",
     "NCBI_taxon",
@@
 kj_c$Contig <- gsub('_', '', kj_c$Contig)
-kj_c <- setDT(kj_c)[, c("assembly", "contig") := tstrsplit(Contig, "--")]
 kj_c <- setDT(kj_c)[, c("Kaiju_Viral", "Kingdom") := tstrsplit(Name, ";", keep = c(1, 2))]
@@
 ```{r}
-viruses <- full_join(x=vs2_c, y=dvf_c, by = "contig")
-viruses <- full_join(x=viruses, y=checkV, by = "contig")
-viruses <- full_join(x=viruses, y=vb_c, by = "contig")
-viruses <- full_join(x=viruses, y=vs_c, by = "contig")
-viruses <- full_join(x=viruses, y=kj_c, by = "contig")
+viruses <- full_join(x=vs2_c, y=dvf_c, by = "uniq_contig")
+viruses <- full_join(x=viruses, y=checkV, by = "uniq_contig")
+viruses <- full_join(x=viruses, y=vb_c, by = "uniq_contig")
+viruses <- full_join(x=viruses, y=vs_c, by = "uniq_contig")
+viruses <- full_join(x=viruses, y=kj_c, by = "uniq_contig")
 ```
 
-Remove contigs not greater than the basepair length cutoff (3000)
+Remove contigs greater than the basepair length cutoff (3000)
 
@@
 #Virsorter
-viruses$category[is.na(viruses$category)] <- 0
+viruses$Category[is.na(viruses$Category)] <- 0
 
@@
   if (include_virsorter) {
-    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
-    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
+    keep_score[input_seqs$Category==1] <- keep_score[input_seqs$Category==1] + 1
+    keep_score[input_seqs$Category==2] <- keep_score[input_seqs$Category==2] + 1
+    keep_score[input_seqs$Category==3] <- keep_score[input_seqs$Category==3] + 0.5
+    keep_score[input_seqs$Category==4] <- keep_score[input_seqs$Category==4] + 1
+    keep_score[input_seqs$Category==5] <- keep_score[input_seqs$Category==5] + 1
+    keep_score[input_seqs$Category==6] <- keep_score[input_seqs$Category==6] + 0.5
   }
@@
 # histogram of contig lengths
-ggplot(viruses, aes(x=checkv_length)) + geom_histogram(binwidth=10000)
+ggplot(viruses, aes(x=checkv_length)) + 
+    geom_histogram(bins=1000) +
+    scale_y_log10()
+ggplot(checkV, aes(x=checkv_length)) + 
+    geom_histogram(bins=1000) +
+    scale_y_log10()
+
+```
+
+```{r}
+# histogram of contig lengths
+ggplot(checkV, aes(x=checkv_length)) + 
+    geom_histogram(bins=1000) +
+    scale_y_log10() +
 ```

### 2022-10-05 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
Added lake michigan and TARA identifying viral contigs directories

--- 2022-07-26 identifying_viral_contigs_environmental.Rmd
@@
   if (include_vibrant) {
-    keep_score[input_seqs$Quality=="high quality draft"] <- keep_score[input_seqs$Quality=="high quality draft"] + 1
-    keep_score[input_seqs$Quality=="medium quality draft"] <- keep_score[input_seqs$Quality=="medium quality draft"] + 1
-    keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$Quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+#    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
   }
@@
 }
-
 ```
 
-## just vibrant
+## vibrant virsorter2 virsorter checkV
 ```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_vibrant = T,
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
                                               include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb)
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = F)
 ```
 
-## vibrant + checkV
 ```{r}
-viruses$keep_score_vb_cv <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv)
+table(viruses$keep_score_high_precision>=1)
 ```
 
-## vibrant + checkV + kaiju
 ```{r}
-viruses$keep_score_vb_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
                                               include_virsorter2 = F,
-                                            include_checkV = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = T)
+viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = T,
+                                              include_virsorter = F)
+viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
                                               include_kaiju = T,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj)
+                                              include_tuning = F,
+                                              include_virsorter = F)
 ```
 
-## vibrant + checkV + kaiju + deepvirfinder
-```{r}
-viruses$keep_score_vb_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T)
-
-# # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_cv_kj_dvf)
-```
-
-## vibrant + virsorter2 + checkV + kaiju
-```{r}
-viruses$keep_score_vb_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F)
-
-# # # # assess_performance(viruses$seqtype, viruses$keep_score_vb_vs2_cv_kj)
-```
-
-#all: vibrant + virsorter2 + checkV + kaiju + deepvirfinder
-```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_all)
-```
-
-## just virsorter2
-```{r}
-viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2)
-```
-
-## virsorter2 + checkV
-```{r}
-viruses$keep_score_vs2_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv)
-```
-
-## virsorter2 + checkV + kaiju
-```{r}
-viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
-```
-
-## virsorter2 + checkV + kaiju + deepvirfinder
-```{r}
-viruses$keep_score_vs2_cv_kj_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj_dvf)
-```
-
-## just checkV
-```{r}
-viruses$keep_score_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_cv)
-```
-
-## checkV + kaiju
-```{r}
-viruses$keep_score_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_cv_kj)
-```
-
-## just kaiju
-```{r}
-viruses$keep_score_k <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_k)
-```
-
-## just deepvirfinder
-```{r}
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf)
-```
-
-## deepvirfinder + checkV
-```{r}
-viruses$keep_score_dvf_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv)
-```
-
-## deepvirfinder + checkV + kaiju
-```{r}
-viruses$keep_score_dvf_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = F)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_dvf_cv_kj)
-```
-
-## virsorter2 + checkV + kaiju
-```{r}
-viruses$keep_score_vs2_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs2_cv_kj)
-```
-
-## just virsorter
-```{r}
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = F,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs)
-```
-
-## virsorter + checkV
-```{r}
-viruses$keep_score_vs_cv <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = F,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv)
-```
-
-## virsorter + checkV + kaiju
-```{r}
-viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
-```
-
-## virsorter + checkV + kaiju + vibrant + deepvirfinder + virsorter2
-```{r}
-viruses$keep_score_vs_cv_kj <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj)
-```
-
-## virsorter + checkV + kaiju + vibrant
-```{r}
-viruses$keep_score_vs_cv_kj_vb <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = F,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb)
-```
-
-## virsorter + checkV + kaiju + virsorter2 + vibrant
-```{r}
-viruses$keep_score_vs_cv_kj_vb_vs2 <- getting_viral_set_1(viruses, include_vibrant = T,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = F,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vb_vs2)
-```
-
-## virsorter + checkV + kaiju + virsorter2
-```{r}
-viruses$keep_score_vs_cv_kj_vs2_dvf <- getting_viral_set_1(viruses, include_vibrant = F,
-                                            include_virsorter2 = T,
-                                            include_checkV = T,
-                                            include_kaiju = T,
-                                            include_deepvirfinder = T,
-                                            include_virsorter = T)
-
-# # # # # assess_performance(viruses$seqtype, viruses$keep_score_vs_cv_kj_vs2_dvf)
-```
-
- TODO WHAT DO I DO ABOUT THE ZEROS????
 # Plot of how many methods identify a contig
 ```{r}
-viruses_high <- viruses[viruses$keep_score_all>=1,]
-viruses_high_mod <- viruses_high %>% select(keep_score_vb_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_vs_cv, keep_score_cv_kj)
+#viruses_high <- viruses[viruses$keep_score_all>=1,]
+viruses_high_mod <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
 viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
 viruses_high_mod <- as_tibble(viruses_high_mod)
-
 virusHighCount <- function(df, tool_combo) {
@@
 }
-
-sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb_cv>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf_cv>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_cv_kj>0,]))
+sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,]))
                           )
 )
-
 sm <- as.data.frame(
     rbind(
-        virusHighCount(viruses_high_mod, "keep_score_vb_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_dvf_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_vs_cv"),
-        virusHighCount(viruses_high_mod, "keep_score_cv_kj")
+        virusHighCount(viruses_high_mod, "keep_score_vb"),
+        virusHighCount(viruses_high_mod, "keep_score_dvf"),
+        virusHighCount(viruses_high_mod, "keep_score_vs2"),
+        virusHighCount(viruses_high_mod, "keep_score_vs"),
+        virusHighCount(viruses_high_mod, "keep_score_kj")
     )
 )
-
 sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "kaiju")
@@
 
-# Visualizing which tools led to a contig being selected
-(alluvial plot)
-with help from: https://www.bioinfo-scrounger.com/archives/alluvial-and-Sankey/
+# Comparing all of the tool methods
 
 ```{r}
-alluvial_data <- viruses %>% select(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj) %>% 
-  apply(c(1,2), function(x) {if (x >= 1) {x <- 1} else if (x==0.5) {x <- 0.5} else {x <- 0}}) %>%
-  as_tibble() %>%
-  count(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv,keep_score_vs2_cv, keep_score_cv_kj)
+combos_list <- data.frame(toolcombo=rep(0, 64),
+                          CheckV=rep(0, 64),
+                          DVF=rep(0, 64),
+                          Kaiju=rep(0, 64),
+                          VIBRANT=rep(0, 64),
+                          VS=rep(0, 64),
+                          VS2=rep(0, 64))
+p <- 1
+for (i in c(0,1)){
+  for (j in c(0,1)){
+    for (k in c(0,1)){
+      for (l in c(0,1)){
+        for (m in c(0,1)){
+          for (n in c(0,1)){
+            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
+            combos_list$CheckV[p] <- i
+            combos_list$DVF[p] <- j
+            combos_list$Kaiju[p] <- k
+            combos_list$VIBRANT[p] <- l
+            combos_list$VS[p] <- m
+            combos_list$VS2[p] <- n
+            p <- p+1
+          }
+        }
+      }
+    }
+  }
+}
+combos_list <- combos_list[-1,]
 ```
@@
 ```{r}
-alluvial_data$cum_sum <- alluvial_data %>% select(keep_score_vb_cv, keep_score_vs_cv, keep_score_dvf_cv, keep_score_vs2_cv, keep_score_cv_kj) %>% 
-  rowSums()
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
+                          num_viruses=rep(0, nrow(combos_list)))
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning = combos_list$CheckV[i],
+                                            include_kaiju = combos_list$Kaiju[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+
+  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+}
+num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
 ```
@@
 ```{r}
-alluvial_data$keep_score <- "0-0.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=1] <- "1-1.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=2] <- "2-2.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=3] <- "3-3.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=4] <- "4-4.5"
-alluvial_data$keep_score[alluvial_data$cum_sum>=5] <- "5+"
-
-alluvial_data$keep_score <- factor(
-    alluvial_data$keep_score,
-    levels=c(
-        "0-0.5",
-        "1-1.5",
-        "2-2.5",
-        "3-3.5",
-        "4-4.5",
-        "5+"
-        )
-    )
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
 ```
@@
 ```{r}
-alluvial_data <- alluvial_data %>% 
-  mutate(
-      across(
-          c(keep_score_vb_cv,
-            keep_score_vs_cv,
-            keep_score_dvf_cv,
-            keep_score_vs2_cv,
-            keep_score_cv_kj
-            ),
-          as.factor
-          )
-      )
+viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
+viral_scores_nozeros <- viral_scores_nozeros + 1
+viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo
 ```
@@
 ```{r}
-head(alluvial_data)
+library(phyloseq)
 ```
@@
 ```{r}
-library(ggalluvial)
+tooldata <- num_viruses
+rownames(tooldata) <- tooldata$toolcombo
 ```
@@
 ```{r}
-ggplot(data = alluvial_data,
-       aes(axis1 = as.character(keep_score_vb_cv), 
-           axis2 = as.character(keep_score_vs_cv),
-           axis3 = as.character(keep_score_vs2_cv),
-           axis4 = as.character(keep_score_dvf_cv),
-           axis5 = as.character(keep_score_cv_kj),
-           y = n)) +
-  geom_alluvium(aes(fill = keep_score)) +
-  geom_stratum() +
-  geom_text(stat = "stratum",
-            aes(label = after_stat(stratum))) +
-  scale_fill_viridis_d() + scale_x_discrete(limits = c("VIBRANT", "Virsorter", "VirSorter2", "DeepVirFinder", "Kaiju"),
-                                  expand = c(0.05, 0.05, 0.05, 0.05)) +
-  ylab("Number of Sequences") +
-  xlab("method")
-  theme_classic()
+physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
+                                     sample_data(tooldata))
+ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw() +
+  geom_label(label=tooldata$toolcombo)
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw()
+```
+to do: try coloring above based on the F1 scores of the testing set on each combination
+
+```{r}
+write_tsv(num_viruses, "dw_num_viruses_by_tool_combo.tsv")
+num_viruses <- read_tsv("dw_num_viruses_by_tool_combo.tsv")
 ```
 
-Get proportion of contigs from "viruses" classified as virus by combined tool:
+```{r}
+accuracy_scores <- read_tsv("../../TestingPipeline/Scripts/20220927_accuracy_scores.tsv")
+```
 
 ```{r}
-prop_viral <- paste(nrow(viruses_high) / nrow(viruses) * 100, "%", sep="")
-prop_viral
+num_viruses_comp <- full_join(num_viruses, accuracy_scores[accuracy_scores$testing_set_index==1,], by=c("toolcombo", "numtools"))
 ```
 
-Distribution of keep_score across "viruses" dataframe.
 ```{r}
-keep_scores <- reshape2::melt(table(viruses$keep_score_all))
-
-colnames(keep_scores) <- c('keep_score', 'count')
-keep_scores[,'keep_score'] <- as.factor(keep_scores[,'keep_score'])
-
-ggplot(data = keep_scores, aes(x = keep_score, y = count)) +
-    geom_bar(stat="identity", width = 0.3, fill = 'steelblue')
+num_viruses_comp$numtools <- as.factor(num_viruses_comp$numtools)
 ```
 
-Determine number of viral sequences with kaiju taxonomic classification
-
 ```{r}
-tax_counts <- viruses_high %>% count(Kaiju_Viral, name = "num")
-tax_counts$prop <- tax_counts$num / sum(tax_counts$num)
+ggplot(num_viruses_comp, aes(x=num_viruses, y=F1, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("F1") +
+  xlab("Num Viruses Predicted")
+ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Precision") +
+  xlab("Num Viruses Predicted")
+ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Recall") +
+  xlab("Num Viruses Predicted")
 ```

### 2022-10-05 TestingPipeline/Scripts/comparing_rule_sets.Rmd
Added lake michigan and TARA identifying viral contigs directories

--- 2022-09-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 library(data.table)
+options(warn=-1)
 ```
@@
 # checkV
-checkV_path <- "TestingSetToolOutputs/testing_sets_checkv_output.tsv"
+checkV_path <- "../ToolOutput/testing_sets_checkv_output.tsv"
 # VIBRANT
-vibrant_path <- "TestingSetToolOutputs/testing_sets_vibrant_output.tsv"
+vibrant_path <- "../ToolOutput/testing_sets_vibrant_output.tsv"
 # DeepVirFinder
-dvf_path <- "TestingSetToolOutputs/testing_sets_deepvirfinder_output.tsv"
+dvf_path <- "../ToolOutput/testing_sets_deepvirfinder_output.tsv"
 # Virsorter
-vs_path <- "TestingSetToolOutputs/testing_sets_virsorter_output.tsv"
+vs_path <- "../ToolOutput/testing_sets_virsorter_output.tsv"
 # Virsorter2
-vs2_path <- "TestingSetToolOutputs/testing_sets_virsorter2_output.tsv"
+vs2_path <- "../ToolOutput/testing_sets_virsorter2_output.tsv"
 # Kaiju
-kj_path <- "TestingSetToolOutputs/merged.nreuk.kaiju.names.out"
+kj_path <- "../ToolOutput/merged.nreuk.kaiju.names.out"
 # KB cutoff
 KB_CUTOFF <- 3000
-
 ```
@@
 }
+```
 
+this rule set is based on an original set of rules considered for analyzing the
+drinking water virome. upon analyzing the sequences, many false positive viruses
+were identified (particularly long sequences with no viral genes)
+note: checkV is used as a viral sorting tool
+note: VirFinder was used not DeepVirFinder
+```{r getting_viral_set_2}
+getting_viral_set_2 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_checkV=FALSE,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
+      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
+    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
+      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
+  }
+
+  if (include_deepvirfinder) {
+    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
+  }
+
+  if (include_checkV) {
+    keep_score[input_seqs$checkv_quality=="low"] <- keep_score[input_seqs$checkv_quality=="low"] + 0.3
+    keep_score[input_seqs$checkv_quality=="medium"] <- keep_score[input_seqs$checkv_quality=="medium"] + 0.5
+    keep_score[input_seqs$checkv_quality=="high"] <- keep_score[input_seqs$checkv_quality=="medium"] + 1
+
+    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>1] <- 0
+  }
+
+  return(keep_score)
+
+}
+```
+
+this is the set of rules used in the drinking water viromes paper
+note: VirFinder was used not DeepVirFinder
+```{r getting_viral_set_3}
+getting_viral_set_3 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_checkV=FALSE,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- 
+      keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.3
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.3
+    keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] <- 
+      keep_score[(input_seqs$viral>=50 & input_seqs$host_genes==0) & input_seqs$viral_genes==0] + 0.5
+    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.3
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.3
+  }
+
+  if (include_deepvirfinder) {
+    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
+    keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 1
+  }
+
+  if (include_checkV) {
+    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
+    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
+    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
+  }
+
+  return(keep_score)
+
+}
 ```
@@
   precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])
-
   recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])
-
   F1 <- 2*precision*recall/(precision+recall)
@@
                                               include_deepvirfinder = combos$DVF[i])
+    }
+    if (pipeline==2) {
+      keep_score <- getting_viral_set_2(input_seqs, include_vibrant = combos$VIBRANT[i],
+                                              include_virsorter = combos$VS[i],
+                                              include_virsorter2 = combos$VS2[i],
+                                              include_checkV = combos$CheckV[i],
+                                              include_deepvirfinder = combos$DVF[i])
+    }
+    if (pipeline==3) {
+      keep_score <- getting_viral_set_3(input_seqs, include_vibrant = combos$VIBRANT[i],
+                                              include_virsorter = combos$VS[i],
+                                              include_virsorter2 = combos$VS2[i],
+                                              include_checkV = combos$CheckV[i],
+                                              include_deepvirfinder = combos$DVF[i])
+    }
 
@@
                        F1=rep(0, nrow(combos_list)*10))
-
 accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==1], combos_list))
-
+                              build_score_list(viruses[viruses$Index==1], combos_list, pipeline=1))
 for (i in 2:10) {
@@
                            cbind(testing_set_index=rep(i, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==i], combos_list)))
+                              build_score_list(viruses[viruses$Index==i], combos_list, pipeline=1)))
 }
+```
 
+```{r}
+library("stringr")
+accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
+accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
 ```
@@
 ```
+## Calculate the performance of each pipeline
+original rule set from DW paper (didn't use)
+
+```{r}
+combos_list_3 <- read_csv("combinations_list_3.csv")
+```
+
+```{r}
+options(warn=-1)
+accuracy_scores_2 <- cbind(testing_set_index=rep(0, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==1], combos_list_3, pipeline=2))
+for (i in 2:10) {
+  accuracy_scores_2 <- rbind(accuracy_scores_2,
+                           cbind(testing_set_index=rep(i, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==i], combos_list_3, pipeline=2)))
+}
+```
+
+## Visualize how the precision, recall, and F1 scores change across pipelines.
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+ggplot(accuracy_scores_2, aes(x=toolcombo, y=F1)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("F1 Score")
+ggplot(accuracy_scores_2, aes(x=toolcombo, y=precision)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Precision")
+ggplot(accuracy_scores_2, aes(x=toolcombo, y=recall)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Recall")
+```
+
+## Calculate the performance of each pipeline
+rule set from DW paper
+
+```{r}
+combos_list_3 <- read_csv("combinations_list_3.csv")
+```
+
+```{r}
+options(warn=-1)
+accuracy_scores_3 <- data.frame(testing_set_index=rep(0, nrow(combos_list_3)*10),
+                      precision=rep(0, nrow(combos_list_3)*10),
+                       recall=rep(0, nrow(combos_list_3)*10),
+                       F1=rep(0, nrow(combos_list_3)*10))
+accuracy_scores_3 <- cbind(testing_set_index=rep(1, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==1], combos_list_3, pipeline=3))
+for (i in 2:10) {
+  accuracy_scores_3 <- rbind(accuracy_scores_3,
+                           cbind(testing_set_index=rep(i, nrow(combos_list_3)),
+                              build_score_list(viruses[viruses$Index==i], combos_list_3, pipeline=3)))
+}
+```
+
+## Visualize how the precision, recall, and F1 scores change across pipelines.
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+p2 <- ggplot(accuracy_scores_3, aes(x=toolcombo, y=F1)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("F1 Score")
+p2
+ggplot(accuracy_scores_3, aes(x=toolcombo, y=precision)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Precision")
+ggplot(accuracy_scores_3, aes(x=toolcombo, y=recall)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, VB, VS, VS2)") +
+  ylab("Recall")
+```

### 2022-10-05 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updates to rule tuning script visualizations.
Added some to dos to the rule_tuning script.
Added lake michigan and TARA identifying viral contigs directories

--- 2022-09-21 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 library(data.table)
+library(pROC)
+library(here)
 ```
@@
 
-  auc <- round(auc(truepositive, keep_score),4)
+  AUC <- round(auc(truepositive, keep_score),4)
 
@@
 
-  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, bacterial_FP, viral_FN)
+  performance <- c(precision, recall, F1, MCC, AUC, fungal_FP, bacterial_FP, viral_FN)
   names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
@@
 rocobj_all <- roc(viruses$truepositive, viruses$keep_score_all)
-auc <- round(auc(viruses$truepositive, viruses$keep_score),4)
+AUC <- round(auc(viruses$truepositive, viruses$keep_score),4)
 auc_all <- round(auc(viruses$truepositive, viruses$keep_score_all),4)
@@
 ggroc(rocobj, colour = 'steelblue', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')')) +
+  ggtitle(paste0('ROC Curve ', '(AUC = ', AUC, ')')) +
   coord_equal()
@@
 Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative.
+
+## Pull out Fungi
+
+```{r}
+fungi_1 <- viruses[(viruses$seqtype == "fungi") & (viruses$Kaiju_Viral == "Viruses")]
+vb_1 <- read_tsv("/Users/riddellj/Documents/Research/VSTE/Troubleshoot_Fungi/VIBRANT_annotations_metagenomic_testing_set_1.tsv")
+vb_1 <- separate(vb_1, col = "scaffold", into = c("seqtype", "contig"), sep="--")
+vb_1$contig <- sub("\\.", "_", vb_1$contig)
+vb_1_vb_c <- inner_join(vb_1, vb_c, on='contig')
+vb_1_vb_c_fungi <- vb_1_vb_c %>% filter(seqtype == "fungi")
+
+fungi_1_vb_1_vb_c_fungi <- inner_join(fungi_1,vb_1_vb_c_fungi, on='contig')
+
+#Examples of "false positive" fungi
+# https://www.ncbi.nlm.nih.gov/nuccore/599124765
+
+```

### 2022-10-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Added pipeline clustering figures to tuning script.

--- 2022-09-27 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
   if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$contig_length<20000] + 0.5
-   keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$contig_length<20000] + 0.5
+    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
+   keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
   }
@@
 }
+
+combos_list <- combos_list[-1,]
 ```
@@
 
+```{r}
+write_tsv(accuracy_scores, "20220927_accuracy_scores.tsv")
+```
+
+to do: add in clustering and ordination like in the drinking water R notebook
+
 # Experimenting
@@
 Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative.
+
+# Comparing behavior of all testing sets combined (clustering analyses)
+
+```{r}
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
+                          num_viruses=rep(0, nrow(combos_list)))
+
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning = combos_list$CheckV[i],
+                                            include_kaiju = combos_list$Kaiju[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+
+  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+
+  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
+}
+
+num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
+```
+
+```{r}
+ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+
+ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
+ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
+  geom_boxplot(aes(color=numtools)) +
+  geom_point(aes(color=numtools, fill=numtools)) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
+viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
+viral_scores_nozeros <- viral_scores_nozeros + 1
+viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
+
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
+```
+
+```{r}
+library(phyloseq)
+```
+
+```{r}
+tooldata <- num_viruses
+
+rownames(tooldata) <- tooldata$toolcombo2
+```
+
+```{r}
+physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
+                                     sample_data(tooldata))
+```
+
+```{r}
+ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw() +
+  geom_label(label=tooldata$toolcombo)
+
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw()
+```
+to do: try coloring above based on the F1 scores of the testing set on each combination
+
+```{r}
+bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
+clusters <- hclust(dist(bray_dist))
+plot(clusters)
+
+myclusters <- cutree(clusters, h=1.1)
+```
+
+```{r}
+names(myclusters[myclusters==1])
+names(myclusters[myclusters==2])
+names(myclusters[myclusters==3])
+names(myclusters[myclusters==4])
+names(myclusters[myclusters==5])
+
+myclusters_df <- tibble(combo=names(myclusters),
+                            cluster_index=myclusters)
+
+myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
+                                                            "Kaiju", "VIBRANT",
+                                                            "VirSorter", "VirSorter2"),
+                          sep=" ", remove = F)
+
+tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
+                    )
+
+tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
+
+tool_count <- melt(tool_count)
+
+colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+ggplot(tool_count, aes(x=cluster_index, y=tool_count,
+                   fill=cluster_index,
+                   color=cluster_index)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  xlab("Cluster") +
+  ylab("Number of Times in Cluster") + 
+  facet_wrap(~tool, scales = "free")
+```

### 2022-10-07 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Added figure to compare clusters of tools to dw script.

@@
 
+myclusters_df <- tibble(combo=names(myclusters),
+                            cluster_index=myclusters)
+
+myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
+                                                            "Kaiju", "VIBRANT",
+                                                            "VirSorter", "VirSorter2"),
+                          sep=" ", remove = F)
+
+tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
+                    )
+
+tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
+
+tool_count <- melt(tool_count)
+
+colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+ggplot(tool_count, aes(x=cluster_index, y=tool_count,
+                   fill=cluster_index,
+                   color=cluster_index)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  xlab("Cluster") +
+  ylab("Number of Times in Cluster") + 
+  facet_wrap(~tool, scales = "free")
 ```

### 2022-10-16 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updated rules and add high MCC example.

@@
 library(data.table)
-```
-
-# Inputs
-1) For this script's inputs, the user must merge each viral identification 
-tool output file into one tab-separated file. 
-Replace the filenames with their paths if needed into the variables in the 
-'inputs' chunk.
-
-    General filenames for each tool:
-        Checkv: quality_summary.tsv
-        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
-        Virsorter: VIRSorter_global-phage-signal.csv
-        Virsorter2: final-viral-score.tsv
-        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
-        Kaiju: ${assembly}.kaiju.names.out
-
-2) Make sure all fasta headers (assembly, contig) are consistent within and
-across tools. Chunks for each tool do contain some lines for cleaning up these 
-features, but due to their variability it will be the user's responsibility to 
-make sure they match across tools.
-
-3) Check each chunk and ensure all columns are accounted for.
-
-4) This script is designed for contigs > 3000 basepairs. It can be modified
-to be higher or lower, but going lower will greatly increase the size of the
-dataframe and memory usage.
-
-```{r inputs}
-# dataset name for file organization and outputs for metrics file
-dataset_name <- "testing_set"
-# checkV
-checkV_path <- "../ToolOutput/testing_sets_checkv_output.tsv"
-# VIBRANT
-vibrant_path <- "../ToolOutput/testing_sets_vibrant_output.tsv"
-# DeepVirFinder
-dvf_path <- "../ToolOutput/testing_sets_deepvirfinder_output.tsv"
-# Virsorter
-vs_path <- "../ToolOutput/testing_sets_virsorter_output.tsv"
-# Virsorter2
-vs2_path <- "../ToolOutput/testing_sets_virsorter2_output.tsv"
-# Kaiju
-kj_path <- "../ToolOutput/merged.nreuk.kaiju.names.out"
-# KB cutoff
-KB_CUTOFF <- 3000
-```
-
-# All Viral Contigs
-
-## checkV
-```{r}
-checkV <- fread(checkV_path, 
-                sep="\t",
-                header = T, 
-                select = c(
-                    'Index',
-                    'contig_id',
-                    'provirus',
-                    'completeness',
-                    'contamination',
-                    'viral_genes',
-                    'host_genes',
-                    'gene_count',
-                    'contig_length',
-                    'checkv_quality'
-                    )
-                ) %>% 
-    rename(
-        contig = contig_id,
-        checkv_provirus = provirus,
-        checkv_completeness = completeness,
-        checkv_contamination = contamination,
-        checkv_viral_genes = viral_genes,
-        checkv_host_genes = host_genes,
-        checkv_total_genes = gene_count,
-        checkv_length = contig_length
-        )
-checkV$method = 'checkv'
-checkV <- separate(checkV, col = contig, into = c("seqtype", "contig"), sep="--")
-checkV$contig <- sub("\\.", "_", checkV$contig)
-checkV$contig <- sub("\\|", "_", checkV$contig)
-checkV <- checkV[!duplicated(paste(checkV$contig, checkV$Index)),]
-```
-
-## VIBRANT
-```{r}
-vb_c <- fread(vibrant_path,
-              header = T,
-              sep = "\t",
-              select = c(
-                  'Index',
-                  'scaffold',
-                  'type',
-                  'Quality'
-                  )
-              ) %>%
-    rename(
-        contig = scaffold,
-        vibrant_quality = Quality
-        )
-vb_c$method <- "vibrant"
-vb_c$vibrant_prophage <- "No"
-vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
-vb_c <- separate(vb_c, col = contig, into = c("seqtype", "contig"), sep="--")
-vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
-vb_c$contig <- sub("\\.", "_", vb_c$contig)
-vb_c <- separate(vb_c, col=contig, into="contig", remove=T, sep = " ")
-vb_c <- vb_c[!duplicated(paste(vb_c$contig, vb_c$Index)),]
-```
-
-## DeepVirFinder
-```{r}
-dvf_c <- fread(dvf_path,
-               header = T,
-               sep = "\t",
-               select = c(
-                   'Index',
-                   'name',
-                   'score',
-                   'pvalue'
-                   )
-               ) %>% 
-    rename(
-        contig = name
-        )
-dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
-dvf_c <- separate(dvf_c, col = contig, into = c("seqtype", "contig"), sep="--")
-dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
-dvf_c <- dvf_c[!duplicated(paste(dvf_c$contig, dvf_c$Index)),]
-```
-
-## VirSorter
-```{r}
-vs_c <- fread(vs_path,
-              select = c(
-                  'index',
-                  'contig',
-                  'category'
-                  )
-              ) %>% 
-    rename(
-        Index = index
-        )
-vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
-vs_c$contig <- sub("-circular", "", vs_c$contig)
-vs_c <- separate(vs_c, col = `contig`, into = c("seqtype", "contig"), sep="--")
-vs_c$contig <- sub("\\.", "_", vs_c$contig)
-vs_c <- vs_c %>% drop_na(contig)
-vs_c <- vs_c[!duplicated(paste(vs_c$contig, vs_c$Index)),]
-```
-
-## VirSorter2 
-```{r}
-vs2_c <- fread(vs2_path,
-                    header = T,
-                    sep = '\t',
-                    select = c(
-                        'Index',
-                        'seqname',
-                        'dsDNAphage',
-                        'ssDNA',
-                        'max_score',
-                        'max_score_group',
-                        'hallmark',
-                        'viral',
-                        'cellular'
-                    )
-               ) %>% 
-              separate(
-                  col = seqname,
-                  into = c("contig", "vs2type"), 
-                  sep = "\\|\\|",
-                  remove = T
-                  )
-vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
-vs2_c <- separate(vs2_c, col = contig, into = c("seqtype", "contig"), sep="--")
-vs2_c <- vs2_c[!duplicated(paste(vs2_c$contig, vs2_c$Index)),]
-```
-
-## Kaiju 
-```{r}
-kj_c <- read_tsv(kj_path, col_names = T)
-kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
-kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
-kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
-kj_c$contig <- sub("\\.", "_", kj_c$contig)
-kj_c <- kj_c[!duplicated(paste(kj_c$contig, kj_c$Index)),]
-```
-Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
-```{r}
-table(kj_c$seqtype, kj_c$Kaiju_Viral)
-```
-
-## Merging
-```{r}
-viruses <- full_join(x=checkV, y=kj_c, by = c("Index", "contig", "seqtype"))
-viruses <- full_join(x=viruses, y=dvf_c, by = c("Index", "contig", "seqtype"))
-viruses <- full_join(x=viruses, y=vb_c, by = c("Index", "contig", "seqtype"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("Index", "contig", "seqtype"))
-viruses <- full_join(x=viruses, y=vs2_c, by = c("Index", "contig", "seqtype"))
-```
-
-Remove contigs not greater than the basepair length cutoff (3000)
-```{r}
-viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
-```
-
-Use the next code chunk to check if any contigs are missing from checkv.
-If yes, then there is a mix up in contig names since checkv should contain
-all contig names even if they are identified as non-viral.
-```{r}
-v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
-```
-
-calculate percent viral/host/unknown
-```{r}
-viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
-viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
-viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
-```
-
-get rid of NAs for downstream processing
-```{r}
-# checkV
-viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
-viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
-# VIBRANT
-viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
-# DeepVirFinder
-viruses$score[is.na(viruses$score)] <- 0
-viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
-# VirSorter2
-viruses$viral[is.na(viruses$viral)] <- 0
-viruses$hallmark[is.na(viruses$hallmark)] <- 0
-#Virsorter
-viruses$category[is.na(viruses$category)] <- 0
-# Kaiju
-viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
-viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
+library(pROC)
+```
+
+Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
+```{r}
+viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
 ```
@@
     keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] + 0.5
 #    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
@@
     keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
+    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
   }
@@
     keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
+    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
+    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
     keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
     #keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] <- keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] + 1 #add some ratio
-    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[((input_seqs$viral_genes*3) <= input_seqs$host_genes) & input_seqs$checkv_provirus=="No"] <- 0 # consider accounting for provirus designation
-#    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0 # consider accounting for provirus designation
-    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
+    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for provirus designation
+#    keep_score[(input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes] <- 0 # consider accounting for provirus designation
+    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark==0] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark==0] - 1
   }
@@
   fungal_FP <- table(confusion_matrix[seqtype=="fungi"])[2]
+  protist_FP <- table(confusion_matrix[seqtype=="protist"])[2]
   bacterial_FP <- table(confusion_matrix[seqtype=="bacteria"])[2]
@@
 
-  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, bacterial_FP, viral_FN)
+  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, 
+                   protist_FP, bacterial_FP, viral_FN)
   names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
-                          "bacterial_FP", "viral_FN")
+                          "protist_FP", "bacterial_FP", "viral_FN")
 
@@
                        fungal_FP=rep(0, nrow(combos)),
+                       protist_FP=rep(0, nrow(combos)),
                        bacterial_FP=rep(0, nrow(combos)),
@@
 
-    output[i,1:8] <- assess_performance(input_seqs$seqtype, keep_score)
+    output[i,1:9] <- assess_performance(input_seqs$seqtype, keep_score)
 
@@
 
-  #return(keep_score)
   return (output)
@@
                       fungal_FP=rep(0, nrow(combos_list)*10),
+                      protist_FP=rep(0, nrow(combos_list)*10),
                       bacterial_FP=rep(0, nrow(combos_list)*10),
                       viral_FN=rep(0, nrow(combos_list)*10))
+
 accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==1], combos_list))
+                              build_score_list(viruses[viruses$Index==1,], combos_list))
 for (i in 2:10) {
@@
                            cbind(testing_set_index=rep(i, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==i], combos_list)))
+                              build_score_list(viruses[viruses$Index==i,], combos_list)))
 }
@@
   ylab("Fungal False Positives")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=protist_FP, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  ylab("Protist False Positives")
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
@@
 viruses$confusion_matrix_high_precision <- "true negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<=1] <- "false negative"
+viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<1] <- "false negative"
 viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision>=1] <- "true positive"
@@
 
-## high recall example
-```{r}
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+## high MCC example
+```{r}
+viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                               include_vibrant = T,
@@
 ```{r}
-viruses$confusion_matrix <- "true negative"
-viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_high_recall<=1] <- "false negative"
-viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
-viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
+viruses$confusion_matrix_high_MCC <- "true negative"
+viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
+viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
+viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
 ```
@@
 ```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix, viruses$seqtype, viruses$Index))
+confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_MCC, viruses$seqtype, viruses$Index))
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+```
+
+## high recall example
+```{r}
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses$confusion_matrix_high_recall <- "true negative"
+viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall<1] <- "false negative"
+viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
+viruses$confusion_matrix_high_recall[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
+```
+
+visualizing confusion matrix by taxa
+```{r}
+confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_recall, viruses$seqtype, viruses$Index))
 colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
@@
 p2
+```
+
+looking at false negatives
+```{r}
+viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<=1),]
 ```

### 2022-10-16 TestingPipeline/Scripts/combining_tool_output.Rmd
Pulled out lines combining tool output into own script.

--- 2022-09-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
-author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+title: "Combining Tool Output"
 output: html_notebook
 ---
-This Rmarkdown file assesses the output of CheckV, DeepVirFinder, Kaiju,
+
+This Rmarkdown file combines the output of CheckV, DeepVirFinder, Kaiju,
 VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
@@
 dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
-
 dvf_c <- separate(dvf_c, col = contig, into = c("seqtype", "contig"), sep="--")
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method="BH")
+dvf_c <- dvf_c[!duplicated(paste(dvf_c$contig, dvf_c$Index)),]
 ```
@@
 vs_c$contig <- sub("\\.", "_", vs_c$contig)
-
 vs_c <- vs_c %>% drop_na(contig)
-vs_c <- vs_c[!duplicated(vs_c$contig),]
+vs_c <- vs_c[!duplicated(paste(vs_c$contig, vs_c$Index)),]
 ```
@@
 kj_c <- read_tsv(kj_path, col_names = T)
-
 kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
-
 kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
 kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
-
 kj_c$contig <- sub("\\.", "_", kj_c$contig)
+kj_c <- kj_c[!duplicated(paste(kj_c$contig, kj_c$Index)),]
 ```
@@
 
-This section defines a viralness score "keep_score" based on the tool classifications. 
-A final keep_score above 1 indicates we will keep that sequence and call it viral.
-
-VIBRANT
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus == TRUE: +0.5
-
-Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +1
-    Hallmark > 2: +1
-
-Virsorter
-    category ==  1,2,4,5: +1
-    category == 3,6: +0.5
-
-DeepVirFinder:
-    Score >= 0.7: +0.5
-
-Kaiju:
-    Kaiju_viral = "cellular organisms": -1
-    Kaiju_viral = "Viruses": +1
-
-CheckV
-    viral_genes == 0 and host_genes >= 1: keep_score = 0
-    If 3*viral_genes <= host_genes: keep_score = 0
-    If length > 50,000 and hallmark == 0: keep_score = 0
-    If %unknown >= 75: +0.5
-
-This script produces visualizations of these combined viral scorings and
-includes ecological metrics like alpha diversity.
-
-You can decide which combination is appropriate for them and only need use the
-tools appropriate for your data.
-
-```{r getting_viral_set_1}
-getting_viral_set_1 <- function(input_seqs,
-                                include_vibrant=FALSE, 
-                                include_virsorter2=FALSE,
-                                include_deepvirfinder=FALSE,
-                                include_checkV=FALSE,
-                                include_kaiju=FALSE,
-                                include_virsorter=FALSE) {
-
-  keep_score <- rep(0, nrow(input_seqs))
-
-  if (include_vibrant) {
-    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
-  }
-
-  if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 1
-  }
-
-  if (include_virsorter) {
-    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
-    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
-  }
-
-  if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7] <- keep_score[input_seqs$score>=0.7] + 0.5
-#   keep_score[input_seqs$score>=0.9] <- keep_score[input_seqs$score>=0.9] + 0.5
-  }
-
-  if (include_kaiju) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
-  }
-
-  if (include_checkV) {
-    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
-    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0
-    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
-  }
-
-  return(keep_score)
-
-}
-
+```{r}
+write_tsv(viruses, "../IntermediaryFiles/viral_tools_combined.tsv")
 ```
-
-# Assessing performance against the "truth"
-note that this is only as accurate as the annotations of the input sequences
-
-this function calculates the precision, recall, and F1 score for each pipeline
-```{r}
-assess_performance <- function(seqtype, keep_score) {
-
-  truepositive <- rep("not viral", length(seqtype))
-  truepositive[seqtype=="virus"] <- "viral"
-
-  #make confusion matrix
-  confusion_matrix <- rep("true negative", length(keep_score))
-  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
-  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
-  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"
-
-  precision <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[2])
-
-  recall <- table(confusion_matrix)[4]/(table(confusion_matrix)[4]+table(confusion_matrix)[1])
-
-  F1 <- 2*precision*recall/(precision+recall)
-
-  performance <- c(precision, recall, F1)
-  names(performance) <- c("precision", "recall", "F1")
-
-  return(performance)
-}
-```
-
-combination of tools list
-```{r}
-combos_list <- read_csv("combinations_list.csv", col_names = T)
-```
-
-this function builds a list of all of the combinations that the user wants to 
-test. 
-In this case, we're comparing the performance of all unique combinations of the 
-six tools.
-```{r}
-build_score_list <- function(input_seqs, combos) {
-  output <- data.frame(precision=rep(0, nrow(combos)),
-                       recall=rep(0, nrow(combos)),
-                       F1=rep(0, nrow(combos)))
-  for (i in 1:nrow(combos)) {
-    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
-                                            include_virsorter = combos$VS[i],
-                                            include_virsorter2 = combos$VS2[i],
-                                            include_checkV = combos$CheckV[i],
-                                            include_kaiju = combos$Kaiju[i],
-                                            include_deepvirfinder = combos$DVF[i])
-
-    output[i,1:3] <- assess_performance(input_seqs$seqtype, keep_score)
-
-    output$toolcombo[i] <- paste(combos$CheckV[i],combos$DVF[i],
-                                 combos$Kaiju[i], combos$VIBRANT[i],
-                                 combos$VS[i], combos$VS2[i])
-  }
-
-  output[is.na(output)] <- 0
-
-  #return(keep_score)
-  return (output)
-}
-```
-
-## Calculate the performance of each pipeline
-```{r}
-accuracy_scores <- build_score_list(viruses[viruses$Index==1], combos_list)
-
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
-                      precision=rep(0, nrow(combos_list)*10),
-                       recall=rep(0, nrow(combos_list)*10),
-                       F1=rep(0, nrow(combos_list)*10))
-
-accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==1], combos_list))
-
-for (i in 2:10) {
-  accuracy_scores <- rbind(accuracy_scores,
-                           cbind(testing_set_index=rep(i, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==i], combos_list)))
-}
-
-```
-
-## Visualize how the precision, recall, and F1 scores change across pipelines.
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
-p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("F1 Score")
-
-p2
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=precision)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Precision")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=recall)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Recall")
-```

### 2022-10-16 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Visualizing sequence features based on sequence category.

---
title: "Testing Set Features Visualization"
output: html_notebook
---

This notebook examines the relationship between different features of the data for distinguishing viral from nonviral sequences.

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r}
colnames(viruses)
```

# Important features by sequence type

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Hallmark Genes") +
  ylab("Number of Viral Genes")
```

```{r}
ggplot(viruses, aes(x=checkv_host_genes, y=checkv_viral_genes)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Host Genes") +
  ylab("Number of Viral Genes")
```

```{r}
ggplot(viruses, aes(x=percent_unknown, y=percent_viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Unknown") +
  ylab("Percentage of Genes Viral")
```

```{r}
ggplot(viruses, aes(x=percent_unknown, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Unknown") +
  ylab("Length of Sequence")
```

```{r}
ggplot(viruses, aes(x=percent_viral, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Viral") +
  ylab("Length of Sequence") 
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Hallmark Genes") +
  ylab("Length of Sequence") 
```

```{r}
ggplot(viruses, aes(x=checkv_length, y=checkv_completeness)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Length") +
  ylab("Completeness") 
```

```{r}
table(viruses$seqtype[viruses$checkv_length>50000 & viruses$hallmark==0])/table(viruses$seqtype)
table(viruses$seqtype[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"])/table(viruses$seqtype)
table(viruses$seqtype[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1])/table(viruses$seqtype)

table(viruses$seqtype[viruses$percent_viral>=50])/table(viruses$seqtype)
table(viruses$seqtype[viruses$percent_unknown>=75])/table(viruses$seqtype)
table(viruses$seqtype[viruses$percent_unknown>=75 & viruses$checkv_length<50000])/table(viruses$seqtype)
table(viruses$seqtype[viruses$hallmark>2])/table(viruses$seqtype[viruses$seqtype %in% unique(viruses$seqtype[viruses$hallmark>2])])
```

```{r}
table(viruses$seqtype, viruses$Kaiju_Viral)
```

```{r}
seqdata <- data.frame(seqtype=viruses$seqtype[!duplicated(viruses$contig)])

rownames(seqdata) <- viruses$contig[!duplicated(viruses$contig)]
```

```{r}
library(phyloseq)
```

```{r}
features_table <- viruses[viruses$Index==1,]
features_table <- features_table[,colnames(features_table) %in% c( 
                                              "checkv_viral_genes",
                                              "checkv_host_genes",
                                              "checkv_unknown_genes",
                                              "checkv_length",
                                              "checkv_completeness",
                                              "checkv_total_genes",
                                              "percent_host",
                                              "percent_viral",
                                              "hallmark",
                                              "percent_unknown"
                                              )]

features_table[is.na(features_table)] <- 0
ft_colnames <- colnames(features_table)
features_table <- t(features_table)
rownames(features_table) <- ft_colnames
colnames(features_table) <- rownames(seqdata)

physeq_pooled <- phyloseq(otu_table(features_table, taxa_are_rows = T))
```

```{r}
ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw() +
  geom_label(label=seqdata$toolcombo)

phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw()
```

### 2022-10-16 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Contribution of individual tuning rules.

@@
 ```
+
+use of each tuning rule
+```{r}
+keep_score <- rep(0, nrow(viruses))
+
+keep_score[viruses$hallmark>2] <- keep_score[viruses$hallmark>2] + 1
+table(keep_score, viruses$seqtype)
+keep_score[viruses$checkv_host_genes>50 & viruses$checkv_provirus=="No"] <- keep_score[viruses$checkv_host_genes>50 & viruses$checkv_provirus=="No"] - 1
+table(keep_score, viruses$seqtype)
+keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
+table(keep_score, viruses$seqtype)
+keep_score[viruses$percent_viral>=50] <- keep_score[viruses$percent_viral>=50] + 0.5
+table(keep_score, viruses$seqtype)
+    #keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] <- keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] + 1 #add some ratio
+keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1
+table(keep_score, viruses$seqtype)
+keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"] <- keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"] - 1 # consider accounting for provirus designation
+table(keep_score, viruses$seqtype)
+#    keep_score[(viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes] <- 0 # consider accounting for provirus designation
+keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] <- keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] - 1
+table(keep_score, viruses$seqtype)
+```

### 2022-10-24 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Contribution of individual tuning rules.
Adjusting rules to help with false positive protists.

@@
 
+```{r}
+viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0
+```
+
 This section defines a viralness score "keep_score" based on the tool classifications. 
@@
 tools appropriate for your data.
+
+#seeing about a length cutoff
+```{r}
+viruses <- viruses[viruses$checkv_length>=5000,]
+```
 
@@
     keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+#    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
     keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
+#    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
   }
@@
 #    keep_score[(input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes] <- 0 # consider accounting for provirus designation
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark==0] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark==0] - 1
+    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
+    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 1
   }
@@
                                               include_vibrant = T,
-                                              include_virsorter2 = F,
+                                              include_virsorter2 = T,
                                               include_kaiju = T,
@@
   coord_flip()
+```
+
+```{r}
+ggplot(viruses, aes(x=checkv_length, y=keep_score_high_MCC,
+                   fill=confusion_matrix_high_MCC,
+                   color=confusion_matrix_high_MCC)) +
+  geom_point(stat="identity", shape=21) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Sequence Length (bp)") +
+  ylab("Pipeline Viral Score") + 
+  facet_wrap(~seqtype) + 
+  scale_x_log10()
 ```
@@
 
+accuracy:
+```{r}
+length(grep("true", viruses$confusion_matrix_high_recall))/nrow(viruses)
+```
+
 visualizing confusion matrix by taxa
@@
 
+```{r}
+ggplot(viruses, aes(x=checkv_completeness, y=hallmark,
+                   fill=confusion_matrix_high_recall,
+                   color=confusion_matrix_high_recall)) +
+  geom_point(stat="identity", shape=21) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("CheckV Completeness") +
+  ylab("Number of Hallmark Genes") + 
+  facet_wrap(~seqtype) + 
+  scale_x_log10()
+```
+
+```{r}
+ggplot(viruses, aes(x=checkv_completeness, y=keep_score_high_MCC,
+                   fill=confusion_matrix_high_recall,
+                   color=confusion_matrix_high_recall)) +
+  geom_point(stat="identity", shape=21) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("CheckV Completeness") +
+  ylab("Pipeline Viral Score") + 
+  facet_wrap(~seqtype) + 
+  scale_x_log10()
+```
+
+```{r}
+ggplot(viruses, aes(x=confusion_matrix_high_recall, y=checkv_length,
+                   fill=confusion_matrix_high_recall,
+                   color=confusion_matrix_high_recall)) +
+  geom_boxplot() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Sequence Length (bp)") +
+  ylab("Pipeline Viral Score") +
+  scale_y_log10()
+```
+
 looking at false negatives
 ```{r}
-viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<=1),]
+viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<1),]
+```
+
+looking at protists calling viral
+```{r}
+viruses_false_pos_protists <- viruses[(viruses$seqtype=="protist" & viruses$keep_score_high_recall>=1),]
 ```

### 2022-10-24 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Splitting refining rules based on whether they add or take away viruses.

@@
 
-#seeing about a length cutoff
-```{r}
-viruses <- viruses[viruses$checkv_length>=5000,]
-```
-
 ```{r getting_viral_set_1}
@@
                                 include_deepvirfinder=FALSE,
-                                include_tuning=FALSE,
-                                include_kaiju=FALSE,
+                                include_tuning_viral=FALSE,
+                                include_tuning_not_viral=FALSE,
                                 include_virsorter=FALSE) {
@@
 
-  if (include_kaiju) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
+  if (include_tuning_viral) {
     keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
-  }
-
-  if (include_tuning) {
     keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
     keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
     keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
-    #keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] <- keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] + 1 #add some ratio
+  }
+
+  if (include_tuning_not_viral) {
+    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
+    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
     keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for provirus designation
-#    keep_score[(input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes] <- 0 # consider accounting for provirus designation
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
     keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 1
+    keep_score[input_seqs$checkv_completeness>=75 & input_seqs$checkV_length<=5000] <- keep_score[input_seqs$checkv_completeness>=75 & input_seqs$checkV_length<=5000] - 1 # helped with protist contamination
   }
@@
 combos_list <- data.frame(toolcombo=rep(0, 64),
-                          CheckV=rep(0, 64),
+                          tune_not_viral=rep(0, 64),
                           DVF=rep(0, 64),
-                          Kaiju=rep(0, 64),
+                          tune_viral=rep(0, 64),
                           VIBRANT=rep(0, 64),
@@
             combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$toolcombo2[p] <- paste(if(i){"C"}else{"0"},if(j){"DVF"}else{"0"},
-                                               if(k){"K"}else{"0"},if(l){"VB"}else{"0"},
+            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
                                                if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
-            combos_list$CheckV[p] <- i
+            combos_list$tune_not_viral[p] <- i
             combos_list$DVF[p] <- j
-            combos_list$Kaiju[p] <- k
+            combos_list$tune_viral[p] <- k
             combos_list$VIBRANT[p] <- l
@@
                                             include_virsorter2 = combos$VS2[i],
-                                            include_tuning = combos$CheckV[i],
-                                            include_kaiju = combos$Kaiju[i],
+                                            include_tuning_viral = combos$tune_viral[i],
+                                            include_tuning_not_viral = combos$tune_not_viral[i],
                                             include_deepvirfinder = combos$DVF[i])
@@
 
-    output$toolcombo[i] <- paste(combos$CheckV[i],combos$DVF[i],
-                                 combos$Kaiju[i], combos$VIBRANT[i],
+    output$toolcombo[i] <- paste(combos$tune_viral[i],combos$DVF[i],
+                                 combos$tune_not_viral[i], combos$VIBRANT[i],
                                  combos$VS[i], combos$VS2[i])
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("F1 Score")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Precision")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Recall")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Precision-Recall")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("MCC")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("AUC")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Fungal False Positives")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Protist False Positives")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Bacterial False Positives")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Viral False Negatives")
@@
 ```{r}
-viruses$keep_score_visualize <- viruses$keep_score
-viruses$keep_score_visualize[viruses$keep_score>1] <- ">1"
+viruses$keep_score_visualize <- viruses$keep_score_high_MCC
+viruses$keep_score_visualize[viruses$keep_score_high_MCC>1] <- "> 1"
+viruses$keep_score_visualize[viruses$keep_score_high_MCC==1] <- "1"
+viruses$keep_score_visualize[viruses$keep_score_high_MCC==0.5] <- "0.5"
+viruses$keep_score_visualize[viruses$keep_score_high_MCC==0] <- "0"
+viruses$keep_score_visualize[viruses$keep_score_high_MCC==-0.5] <- "-0.5"
+viruses$keep_score_visualize[viruses$keep_score_high_MCC==-1] <- "-1"
+viruses$keep_score_visualize[viruses$keep_score_high_MCC<=-1] <- "< -1"
+
 viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-                                       levels=c("-0.5", "-1", "0", "0.5","1", ">1"))
-viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
+                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
+#viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
+#                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
 ```
@@
 pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
-p1 <- ggplot(viruses, aes(x=as.factor(Index),
+ggplot(viruses, aes(x=as.factor(Index),
                    fill=keep_score_visualize, color=keep_score_visualize)) +
@@
   scale_color_manual(name = 'Viral Score',
-                     values = alpha(c(pal(4)), 1)) +
+                     values = alpha(c(pal(6)), 1)) +
   scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(pal(4)), 0.5)) +
+                     values = alpha(c(pal(6)), 0.5)) +
   xlab("Index") +
   ylab("Sequence Count") +
-  facet_wrap(~confusion_matrix, scales = "free")
-p1
+  facet_wrap(~confusion_matrix_high_MCC, scales = "free")
+
+```
+
+# Plot of how many methods identify a contig
+
+```{r}
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_kaiju = T,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+```
+
+```{r}
+#viruses_high <- viruses[viruses$keep_score_high_precision>=1,] #uncomment this line if high precision example
+#viruses_high <- viruses[viruses$keep_score_high_recall>=1,] #uncomment this line if want high recall example 
+viruses_high <- viruses[viruses$keep_score_all>=1,] #uncomment this line if want to use all 6 tools
+#viruses_high <- viruses[,] #comment this line if want only the "high precision contigs"
+viruses_high_mod <- viruses_high %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_tuning)
+viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
+viruses_high_mod <- as_tibble(viruses_high_mod)
+
+virusHighCount <- function(df, tool_combo) {
+    tb <- rowSums(df[df[,tool_combo]>0,])
+    if (length(tb) == 0) {
+        return (0)
+    }
+    else {
+        return (table(tb))
+    }
+}
+
+sm <- as.data.frame(
+    rbind(
+        virusHighCount(viruses_high_mod, "keep_score_vb"),
+        virusHighCount(viruses_high_mod, "keep_score_dvf"),
+        virusHighCount(viruses_high_mod, "keep_score_vs2"),
+        virusHighCount(viruses_high_mod, "keep_score_vs"),
+        virusHighCount(viruses_high_mod, "keep_score_tuning")
+    )
+)
+
+sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "tuning")
+sm
+```
+
+```{r}
+sm_m <- reshape2::melt(sm)
+colnames(sm_m) <- c("primary_method", "number_methods", "number")
+```
+
+```{r}
+ggplot(sm_m, aes(x=primary_method, y=number,
+                   fill=number_methods)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name = 'Number of Methods',
+                     values = alpha(c(viridis(5)), 1)) +
+  xlab("Primary Method") +
+  ylab("Count of Viral Contigs")
+```
+
+# Considering how each method contributes to the final prediction
+```{r}
+#viruses_high <- viruses[viruses$keep_score_high_precision>=1,] #uncomment this line if high precision example
+#viruses_high <- viruses[viruses$keep_score_high_recall>=1,] #uncomment this line if want high recall example 
+#viruses_high <- viruses[viruses$keep_score_all>=1,] #uncomment this line if want to use all 6 tools
+viruses_high <- viruses[viruses$keep_score_high_MCC>=1,] #uncomment this line if want to use high MCC/AUC/F1 example
+#viruses_high <- viruses[,] #comment this line if want only the "high precision contigs"
+viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vs2, keep_score_vs, keep_score_kj, keep_score_tuning)
+viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 0.5) {x <- 1} else {x <- 0}})
+viruses_high_mod <- as_tibble(viruses_high_mod)
+
+virusHighCount <- function(df, tool_combo) {
+    tb <- rowSums(df[df[,tool_combo]>0,])
+    if (length(tb) == 0) {
+        return (0)
+    }
+    else {
+        return (table(tb))
+    }
+}
+
+sm <- as.data.frame(
+    rbind(
+        virusHighCount(viruses_high_mod, "keep_score_tuning"),
+        virusHighCount(viruses_high_mod, "keep_score_kj"),
+        virusHighCount(viruses_high_mod, "keep_score_vb"),
+        virusHighCount(viruses_high_mod, "keep_score_vs"),
+        virusHighCount(viruses_high_mod, "keep_score_vs2"),
+    )
+)
+
+sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,])),
+                          c(0,table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,]))),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
+                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_tuning>0,]))
+                          )
+)
+
+sm$method <- c("tuning", "deepvirfinder", "vibrant", "virsorter", "virsorter2")
+sm
+```
+
+```{r}
+sm_m <- reshape2::melt(sm)
+colnames(sm_m) <- c("primary_method", "number_methods", "number")
+```
+
+```{r}
+ggplot(sm_m, aes(x=primary_method, y=number,
+                   fill=number_methods)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name = 'Number of Methods',
+                     values = alpha(c(viridis(6)), 1)) +
+  xlab("Primary Method") +
+  ylab("Count of Viral Contigs")
+```
+
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tuning_kj <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+```
+
+# Considering how each method contributes to the final prediction
+```{r}
+viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tuning_kj>=1,] #uncomment this line if want to use all 6 tools
+viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
+                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
+                                            keep_score_vb_dvf_vs2_vs_tuning, keep_score_vb_dvf_vs2_vs_tuning_kj)
+#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
+viruses_high_mod <- as_tibble(viruses_high_mod)
+
+```
+
+```{r}
+sm_m <- reshape2::melt(viruses_high_mod)
+colnames(sm_m) <- c("method", "score")
+```
+
+```{r}
+ggplot(sm_m, aes(x=method, y=score,
+                   fill=as.factor(score))) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name = 'Number of Methods',
+                     values = alpha(c(viridis(14)), 1)) +
+  xlab("Primary Method") +
+  ylab("Count of Viral Contigs") +
+  coord_flip()
+```
+
+# Considering how each method contributes to the final prediction (high MCC)
+
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_vs2_vs_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = F,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_vs2_vs_tuning_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_kaiju = T,
+                                              include_tuning = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses_high <- viruses[viruses$keep_score_high_MCC>=1,] #uncomment this line if want to use all 6 tools
+viruses_high_mod <- viruses_high %>% select(keep_score_vb, 
+                                            keep_score_vb_vs2, keep_score_vb_vs2_vs, 
+                                            keep_score_vb_vs2_vs_tuning, keep_score_vb_vs2_vs_tuning_kj)
+#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
+viruses_high_mod <- as_tibble(viruses_high_mod)
+
+```
+
+```{r}
+sm_m <- reshape2::melt(viruses_high_mod)
+colnames(sm_m) <- c("method", "score")
+```
+
+```{r}
+ggplot(sm_m, aes(x=method, y=score,
+                   fill=as.factor(score))) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name = 'Number of Methods',
+                     values = alpha(c(viridis(14)), 1)) +
+  xlab("Primary Method") +
+  ylab("Count of Viral Contigs") +
+  coord_flip()
 ```

### 2022-10-25 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Minor figure adjustments.

@@
                                               include_virsorter2 = T,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = T)
+                                              include_virsorter = T,
+                                              include_tuning_viral= T,
+                                              include_tuning_not_viral = T)
 ```
@@
                                               include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = F)
@@
                                               include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = F)
@@
                                               include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = T)
@@
                                               include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = F)
@@
                                               include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = F)
@@
                                               include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = T)
 
-viruses$keep_score_vb_dvf_vs2_vs_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
                                               include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = T)
 
-viruses$keep_score_vb_dvf_vs2_vs_tuning_kj <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
                                               include_virsorter2 = T,
-                                              include_kaiju = T,
-                                              include_tuning = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = T)
@@
 ```{r}
-viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tuning_kj>=1,] #uncomment this line if want to use all 6 tools
+viruses_high <- viruses[viruses$keep_score_all>=1,] 
 viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
                                             keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
-                                            keep_score_vb_dvf_vs2_vs_tuning, keep_score_vb_dvf_vs2_vs_tuning_kj)
+                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
 #viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
@@
 sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "score")
+colnames(sm_m) <- c("method", "viral_score")
+```
+
+```{r}
+sm_m <- sm_m[sm_m$viral_score>0,]
+
+sm_m$score <- sm_m$viral_score
+
+sm_m$score[sm_m$viral_score==0.5] <- "0.5"
+sm_m$score[sm_m$viral_score>=1] <- "1"
+sm_m$score[sm_m$viral_score>=2] <- "2"
+sm_m$score[sm_m$viral_score>=3] <- "3"
+sm_m$score[sm_m$viral_score>=4] <- "4"
+sm_m$score[sm_m$viral_score>=5] <- "5"
+
+sm_m$score <- factor(sm_m$score, 
+                                       levels=c("0.5", "1", "2","3","4","5"))
 ```
@@
 ggplot(sm_m, aes(x=method, y=score,
-                   fill=as.factor(score))) +
+                   fill=score)) +
   geom_bar(stat="identity") +
@@
   scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(14)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs") +
+                     values = alpha(c(viridis(6)), 1)) +
+  xlab("") +
+  ylab("Viral Score") +
   coord_flip()
@@
                                               include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = F)
 
-viruses$keep_score_vb_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+viruses$keep_score_vb_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                               include_vibrant = T,
                                               include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = F)
-
-viruses$keep_score_vb_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_vs2_vs_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = T,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_vs2_vs_tuning_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses_high <- viruses[viruses$keep_score_high_MCC>=1,] #uncomment this line if want to use all 6 tools
+```
+
+```{r}
+viruses_high <- viruses[viruses$keep_score_vb_tv>=1,] #uncomment this line if want to use all 6 tools
 viruses_high_mod <- viruses_high %>% select(keep_score_vb, 
-                                            keep_score_vb_vs2, keep_score_vb_vs2_vs, 
-                                            keep_score_vb_vs2_vs_tuning, keep_score_vb_vs2_vs_tuning_kj)
+                                            keep_score_vb_tv)
 #viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})

### 2022-10-25 TestingPipeline/Scripts/Tools/vs2.sh
Including all virus types in vs2.

@@ -12,6 +12,6 @@ threads=$2
 path="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/VS2/VS2_${1}"
 threads=10
 
-virsorter run -w $path -i $samplename -j $threads all
+virsorter run -w $path -i $samplename --include-groups all -j $threads all
 
 echo done

### 2022-10-25 TestingPipeline/Scripts/Tools/vs2_slurm.sh

@@ -37,11 +37,11 @@ echo $SLURM_JOB_NODELIST
 echo start
 
 echo "activate conda environment"
-source activate  /home/hegartyb/miniconda3/envs/vs2_2
+source activate  /home/hegartyb/miniconda3/envs/vs2_20221025
 
 echo "run virsorter on all testing sets"
-mkdir ../VS2
-cd ../VS2
-bash ../Scripts/vs2.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
+mkdir /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/VS2
+cd /scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/VS2
+bash /scratch/duhaimem_root/duhaimem/shared_data/VSTE/Scripts/vs2.sh ${SLURM_ARRAY_TASK_ID} ${SLURM_CPUS_ON_NODE}
 
 echo done

### 2022-10-25 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Removed some old figures.

@@
   if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
     keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
@@
     keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] + 0.5
-#    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
   }
@@
     keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
-#    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
     keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
     keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
-#    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
   }
@@
   if (include_tuning_not_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
+    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
     keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for provirus designation
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for RNA viruses
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness>=75 & input_seqs$checkV_length<=5000] <- keep_score[input_seqs$checkv_completeness>=75 & input_seqs$checkV_length<=5000] - 1 # helped with protist contamination
+    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 0.5 
+    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
   }
@@
                                               include_virsorter2 = T,
-                                              include_kaiju = T,
-                                              include_tuning = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = T)
@@
 
-# Plot of how many methods identify a contig
-
-```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_virsorter = T,
-                                              include_tuning_viral= T,
-                                              include_tuning_not_viral = T)
-```
-
 ```{r}
@@
 
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-```
-
-```{r}
-#viruses_high <- viruses[viruses$keep_score_high_precision>=1,] #uncomment this line if high precision example
-#viruses_high <- viruses[viruses$keep_score_high_recall>=1,] #uncomment this line if want high recall example 
-viruses_high <- viruses[viruses$keep_score_all>=1,] #uncomment this line if want to use all 6 tools
-#viruses_high <- viruses[,] #comment this line if want only the "high precision contigs"
-viruses_high_mod <- viruses_high %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_tuning)
-viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-virusHighCount <- function(df, tool_combo) {
-    tb <- rowSums(df[df[,tool_combo]>0,])
-    if (length(tb) == 0) {
-        return (0)
-    }
-    else {
-        return (table(tb))
-    }
-}
-
-sm <- as.data.frame(
-    rbind(
-        virusHighCount(viruses_high_mod, "keep_score_vb"),
-        virusHighCount(viruses_high_mod, "keep_score_dvf"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2"),
-        virusHighCount(viruses_high_mod, "keep_score_vs"),
-        virusHighCount(viruses_high_mod, "keep_score_tuning")
-    )
-)
-
-sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "tuning")
-sm
-```
-
-```{r}
-sm_m <- reshape2::melt(sm)
-colnames(sm_m) <- c("primary_method", "number_methods", "number")
-```
-
-```{r}
-ggplot(sm_m, aes(x=primary_method, y=number,
-                   fill=number_methods)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(5)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs")
-```
-
-# Considering how each method contributes to the final prediction
-```{r}
-#viruses_high <- viruses[viruses$keep_score_high_precision>=1,] #uncomment this line if high precision example
-#viruses_high <- viruses[viruses$keep_score_high_recall>=1,] #uncomment this line if want high recall example 
-#viruses_high <- viruses[viruses$keep_score_all>=1,] #uncomment this line if want to use all 6 tools
-viruses_high <- viruses[viruses$keep_score_high_MCC>=1,] #uncomment this line if want to use high MCC/AUC/F1 example
-#viruses_high <- viruses[,] #comment this line if want only the "high precision contigs"
-viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vs2, keep_score_vs, keep_score_kj, keep_score_tuning)
-viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 0.5) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-virusHighCount <- function(df, tool_combo) {
-    tb <- rowSums(df[df[,tool_combo]>0,])
-    if (length(tb) == 0) {
-        return (0)
-    }
-    else {
-        return (table(tb))
-    }
-}
-
-sm <- as.data.frame(
-    rbind(
-        virusHighCount(viruses_high_mod, "keep_score_tuning"),
-        virusHighCount(viruses_high_mod, "keep_score_kj"),
-        virusHighCount(viruses_high_mod, "keep_score_vb"),
-        virusHighCount(viruses_high_mod, "keep_score_vs"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2"),
-    )
-)
-
-sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,])),
-                          c(0,table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,]))),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_tuning>0,]))
-                          )
-)
-
-sm$method <- c("tuning", "deepvirfinder", "vibrant", "virsorter", "virsorter2")
-sm
-```
-
-```{r}
-sm_m <- reshape2::melt(sm)
-colnames(sm_m) <- c("primary_method", "number_methods", "number")
-```
-
-```{r}
-ggplot(sm_m, aes(x=primary_method, y=number,
-                   fill=number_methods)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(6)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs")
-```
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
 viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
@@
 ```{r}
-viruses_high <- viruses[viruses$keep_score_all>=1,] 
+viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
 viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 

### 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set.Rmd
Splitting building viral set and identifying viral contigs scripts in the generic versions.

--- 2022-10-04 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "Building Environmental Data Viral Features Set"
 author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+date: 10-26-2022
 output: html_notebook
@@
 Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.
-
-TO DO:
-- add kaiju and DVF
 
@@
 viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
+viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0
 
@@
 viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
-viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
-
 ```
@@
 ```{r}
-viruses_filename <- 'merged_viral_id_outputs.csv'
-#write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
+viruses_filename <- 'merged_viral_feature_outputs.csv'
+write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
 ```
@@
 The construction of the final dataframe essentially doubles the memory being used, and even with garbage collector can make Rstudio unresponsive.
-
----
-title: "identifying_viral_metrics_environmental"
-author: Bridget Hegarty, James Riddell
-date: 7/23/2022
-output: html_document
----
-
-TODO: ADD ENVIRONMENTAL METRICS LIKE SHANNON, ALPHA DIVERSITY.
-
-This section defines a viralness score "keep_score" based on the tool classifications. 
-A final keep_score above 1 indicates we will keep that sequence and call it viral.
-
-VIBRANT
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus == TRUE: +0.5
-
-Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +1
-    Hallmark > 2: +1
-
-Virsorter
-    category ==  1,2,4,5: +1
-    category == 3,6: +0.5
-
-DeepVirFinder:
-    Score >= 0.7: +0.5
-
-Kaiju:
-    Kaiju_viral = "cellular organisms": -1
-    Kaiju_viral = "Viruses": +1
-
-CheckV
-    viral_genes == 0 and host_genes >= 1: keep_score = 0
-    If 3*viral_genes <= host_genes: keep_score = 0
-    If length > 50,000 and hallmark == 0: keep_score = 0
-    If %unknown >= 75: +0.5
-
-This script produces visualizations of these combined viral scorings and
-includes ecological metrics like alpha diversity.
-
-Users can decide which combination is appropriate for them and only need to run
-the chunks for the appropriate tools.
-
-```{r getting_viral_set_1}
-getting_viral_set_1 <- function(input_seqs,
-                                include_vibrant=FALSE, 
-                                include_virsorter2=FALSE,
-                                include_deepvirfinder=FALSE,
-                                include_tuning=FALSE,
-                                include_kaiju=FALSE,
-                                include_virsorter=FALSE) {
-
-  keep_score <- rep(0, nrow(input_seqs))
-
-  if (include_vibrant) {
-    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
-#    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
-  }
-
-  if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
-  }
-
-  if (include_virsorter) {
-    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
-    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
-  }
-
-  if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
-   keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
-  }
-
-  if (include_kaiju) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
-  }
-
-  if (include_tuning) {
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
-    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
-    #keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] <- keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] + 1 #add some ratio
-    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[((input_seqs$viral_genes*3) <= input_seqs$host_genes) & input_seqs$checkv_provirus=="No"] <- 0 # consider accounting for provirus designation
-#    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0 # consider accounting for provirus designation
-    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
-  }
-
-  return(keep_score)
-
-}
-
-```
-
-## vibrant virsorter2 virsorter checkV
-```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-table(viruses$keep_score_high_precision>=1)
-```
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-```
-
-# Plot of how many methods identify a contig
-```{r}
-#viruses_high <- viruses[viruses$keep_score_all>=1,]
-viruses_high_mod <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
-viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-virusHighCount <- function(df, tool_combo) {
-    tb <- rowSums(df[df[,tool_combo]>0,])
-    if (length(tb) == 0) {
-        return (0)
-    }
-    else {
-        return (table(tb))
-    }
-}
-
-sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,]))
-                          )
-)
-
-sm <- as.data.frame(
-    rbind(
-        virusHighCount(viruses_high_mod, "keep_score_vb"),
-        virusHighCount(viruses_high_mod, "keep_score_dvf"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2"),
-        virusHighCount(viruses_high_mod, "keep_score_vs"),
-        virusHighCount(viruses_high_mod, "keep_score_kj")
-    )
-)
-
-sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "kaiju")
-sm
-```
-
-Create list of contigs considered viral
-```{r}
-c_list <- data.frame(viruses_high$checkv_uniq_contig)
-write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
-```
-
-Export keep_score_high and list of contigs considered viral by the combined tool.
-
-```{r}
-write_tsv(viruses_high, paste(dataset_name, "viruses_high.tsv", sep='_'))
-```
-
-```{r}
-sm_m <- reshape2::melt(sm)
-colnames(sm_m) <- c("primary_method", "number_methods", "number")
-```
-
-```{r}
-ggplot(sm_m, aes(x=primary_method, y=number,
-                   fill=number_methods)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(5)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs")
-```
-
-# Comparing all of the tool methods
-
-```{r}
-combos_list <- data.frame(toolcombo=rep(0, 64),
-                          CheckV=rep(0, 64),
-                          DVF=rep(0, 64),
-                          Kaiju=rep(0, 64),
-                          VIBRANT=rep(0, 64),
-                          VS=rep(0, 64),
-                          VS2=rep(0, 64))
-p <- 1
-
-for (i in c(0,1)){
-  for (j in c(0,1)){
-    for (k in c(0,1)){
-      for (l in c(0,1)){
-        for (m in c(0,1)){
-          for (n in c(0,1)){
-            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$toolcombo2[p] <- paste(if(i){"C"}else{"0"},if(j){"DVF"}else{"0"},
-                                               if(k){"K"}else{"0"},if(l){"VB"}else{"0"},
-                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
-            combos_list$CheckV[p] <- i
-            combos_list$DVF[p] <- j
-            combos_list$Kaiju[p] <- k
-            combos_list$VIBRANT[p] <- l
-            combos_list$VS[p] <- m
-            combos_list$VS2[p] <- n
-            p <- p+1
-          }
-        }
-      }
-    }
-  }
-}
-
-combos_list <- combos_list[-1,]
-```
-
-```{r}
-viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
-num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
-                          num_viruses=rep(0, nrow(combos_list)))
-
-for (i in 1:nrow(combos_list)) {
-  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
-                                            include_virsorter = combos_list$VS[i],
-                                            include_virsorter2 = combos_list$VS2[i],
-                                            include_tuning = combos_list$CheckV[i],
-                                            include_kaiju = combos_list$Kaiju[i],
-                                            include_deepvirfinder = combos_list$DVF[i])
-
-  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
-
-  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
-
-  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
-}
-
-num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
-num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
-num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
-num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
-num_viruses$numtools <- as.factor(num_viruses$numtools)
-```
-
-```{r}
-ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-
-ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
-viral_scores_nozeros <- viral_scores_nozeros + 1
-viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
-
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
-```
-
-```{r}
-library(phyloseq)
-```
-
-```{r}
-tooldata <- num_viruses
-
-rownames(tooldata) <- tooldata$toolcombo2
-```
-
-```{r}
-physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
-                                     sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  geom_label(label=tooldata$toolcombo)
-
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw()
-```
-to do: try coloring above based on the F1 scores of the testing set on each combination
-
-```{r}
-bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
-clusters <- hclust(dist(bray_dist))
-plot(clusters)
-
-myclusters <- cutree(clusters, h=0.8)
-```
-
-```{r}
-names(myclusters[myclusters==1])
-names(myclusters[myclusters==2])
-names(myclusters[myclusters==3])
-names(myclusters[myclusters==4])
-names(myclusters[myclusters==5])
-
-```
-
-```{r}
-write_tsv(num_viruses, "dw_num_viruses_by_tool_combo.tsv")
-num_viruses <- read_tsv("dw_num_viruses_by_tool_combo.tsv")
-```
-
-```{r}
-accuracy_scores <- read_tsv("../../TestingPipeline/Scripts/20220927_accuracy_scores.tsv")
-```
-
-```{r}
-num_viruses_comp <- full_join(num_viruses, accuracy_scores[accuracy_scores$testing_set_index==1,], by=c("toolcombo", "numtools"))
-```
-
-```{r}
-num_viruses_comp$numtools <- as.factor(num_viruses_comp$numtools)
-```
-
-```{r}
-ggplot(num_viruses_comp, aes(x=num_viruses, y=F1, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("F1") +
-  xlab("Num Viruses Predicted")
-
-ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("Precision") +
-  xlab("Num Viruses Predicted")
-
-ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("Recall") +
-  xlab("Num Viruses Predicted")
-```

### 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
Splitting building viral set and identifying viral contigs scripts in the generic versions.

--- 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set.Rmd
@@
 ---
-title: "Building Environmental Data Viral Features Set"
+title: "Building Environmental Data Viral Features Set Drinking Water"
 author: Bridget Hegarty, James Riddell

### 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Splitting building viral set and identifying viral contigs scripts in the generic versions.

@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "Building Drinking Water Viral Set and Visualization"
 author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+date: 10-26-2022
 output: html_notebook
@@
 
-# Inputs
-1) For this script's inputs, the user must merge each viral identification 
-tool output with the following specified filenames into one tab-separated file. 
-Replace the filenames with their paths if needed into the variables in the 
-'inputs' chunk.
-
-    General filenames for each tool:
-        Checkv: quality_summary.tsv
-        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
-        Virsorter: VIRSorter_global-phage-signal.csv
-        Virsorter2: final-viral-score.tsv
-        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
-        Kaiju: ${assembly}.kaiju.names.out
-
-2) Make sure all fasta headers (assembly, contig) are consistent within and
-across tools. Chunks for each tool do contain some lines for cleaning up these features, but due to their variability it will be the user's responsibility to 
-make sure they match across tools.
-
-3) Check each chunk and ensure all columns are accounted for. For example,
-if you don't have "assembly" and called it something else, please either
-remove it from this script or modify your input files.
-
-4) This script is designed for contigs > 3000 basepairs. It can be modified
-to be higher or lower, but going lower will greatly increase the size of the
-dataframe and memory usage.
-
-```{r inputs}
+```{r}
 # dataset name for file organization and outputs for metrics file
 dataset_name <- "drinking_water"
-# checkV
-checkV_path <- "DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
-# VIBRANT
-vibrant_path <- "DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
-# DeepVirFinder
-dvf_path <- "DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
-# Virsorter
-vs_path <- "DrinkingWaterToolOutput/vs_merged_all.txt"
-# Virsorter2
-vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
-# Kaiju
-kj_path <- "/Users/hegartyb/Documents/BlueSkyPostdoc/ViralSortingToolsEvaluation/Temp/kaiju_outputs_merged.tsv"
-# KB cutoff
-KB_CUTOFF <- 3000
+#name from building_env_data_viral_features_set.Rmd
+viruses_filename <- 'merged_viral_feature_outputs.csv'
 
+viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
 ```
-
-# All Viral Contigs
-
-## checkV
-```{r}
-checkV <- fread(checkV_path, 
-                sep="\t",
-                header = T, 
-                select = c(
-                    'contig_id',
-                    'assembly',
-                    'provirus',
-                    'completeness',
-                    'contamination',
-                    'viral_genes',
-                    'host_genes',
-                    'gene_count',
-                    'contig_length',
-                    'checkv_quality'
-                    )
-                ) %>% 
-    rename(
-        contig = contig_id,
-        checkv_provirus = provirus,
-        checkv_completeness = completeness,
-        checkv_contamination = contamination,
-        checkv_viral_genes = viral_genes,
-        checkv_host_genes = host_genes,
-        checkv_total_genes = gene_count,
-        checkv_length = contig_length
-        )
-checkV$method = 'checkv'
-checkV$contig <- sub("\\.", "_", checkV$contig)
-checkV$uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
-checkV <- checkV[!duplicated(checkV$uniq_contig),]
-```
-
-## VIBRANT
-```{r}
-vb_c <- fread(vibrant_path,
-              header = T,
-              sep = "\t",
-              select = c(
-                  'scaffold',
-                  'assembly',
-                  'type',
-                  'quality'
-                  )
-              ) %>%
-    rename(
-        contig = scaffold,
-        vibrant_quality = quality
-        )
-vb_c$method <- "vibrant"
-vb_c$vibrant_prophage <- "No"
-vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
-vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
-vb_c$contig <- sub("\\.", "_", vb_c$contig)
-vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
-vb_c <- vb_c[!duplicated(vb_c$uniq_contig),]
-```
-
-## DeepVirFinder
-```{r}
-dvf_c <- fread(dvf_path,
-               header = T,
-               sep = "\t",
-               select = c(
-                   'sample',
-                   'name',
-                   'score',
-                   'pvalue'
-                   )
-               ) %>% 
-    rename(
-        contig = name,
-        assembly = sample
-        )
-dvf_c$contig <- sub("\\.", "_", dvf_c$contig)
-dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
-dvf_c$uniq_contig <- paste(dvf_c$assembly, dvf_c$contig, sep = "--")
-dvf_c <- dvf_c[!duplicated(dvf_c$uniq_contig),]
-```
-
-## VirSorter
-```{r}
-vs_c <- fread(vs_path,
-              select = c(
-                  'assembly',
-                  'contig_id',
-                  'category'
-                  ),
-              ) %>% 
-  rename(contig = contig_id) %>%
-  drop_na(contig)
-vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
-vs_c$contig <- sub("-circula", "", vs_c$contig)
-vs_c$contig <- sub("uth", "South", vs_c$contig)
-vs_c$contig <- sub("pannenburg", "Spannenburg", vs_c$contig)
-vs_c$contig <- sub("\\.", "_", vs_c$contig)
-vs_c$uniq_contig <- paste(vs_c$assembly, vs_c$contig, sep = "--")
-
-vs_c <- vs_c[!duplicated(vs_c$uniq_contig),]
-vs_c <- vs_c %>% drop_na(uniq_contig)
-```
-
-## VirSorter2 
-```{r}
-vs2_c <- fread(vs2_path,
-                    header = T,
-                    sep = '\t',
-                    select = c(
-                        'seqname',
-                        'assembly',
-                        'dsDNAphage',
-                        'ssDNA',
-                        'max_score',
-                        'max_score_group',
-                        'hallmark',
-                        'viral',
-                        'cellular'
-                    )
-               ) %>% 
-              separate(
-                  col = seqname,
-                  into = c("contig", "type"), 
-                  sep = "\\|\\|",
-                  remove = T
-                  )
-vs2_c$assembly <- sub("_3000", "", vs2_c$assembly)
-vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
-vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
-vs2_c <- vs2_c[!duplicated(vs2_c$uniq_contig),]
-```
-
-## Kaiju
-```{r}
-kj_c <- read_tsv(kj_path, col_names = T)
-colnames(kj_c)[1] <- "assembly"
-colnames(kj_c)[3] <- "contig"
-kj_c$contig <- sub("\\.", "_", kj_c$contig)
-kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
-kj_c$uniq_contig <- paste(kj_c$assembly, kj_c$contig, sep="--")
-kj_c <- kj_c[!duplicated(kj_c$contig),]
-```
-
-```{r}
-# Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
-table(kj_c$assembly, kj_c$Kaiju_Viral)
-```
-
-## Merging
-```{r}
-viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=vb_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=kj_c, by = c("uniq_contig", "assembly"))
-```
-
-Remove contigs not greater than the basepair length cutoff (3000)
-
-```{r}
-viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
-```
-
-Use the next code chunk to check if any contigs are missing from checkv.
-If yes, then there is a mix up in contig names since checkv should contain
-all contig names even if they are identified as non-viral.
-```{r}
-v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
-```
-
-calculate percent viral/host/unknown
-```{r}
-viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
-viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
-viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
-```
-
-get rid of NAs for downstream processing
-```{r}
-
-# checkV
-viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
-viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
-
-# VIBRANT
-viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
-
-# DeepVirFinder
-viruses$score[is.na(viruses$score)] <- 0
-viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
-
-# VirSorter2
-viruses$viral[is.na(viruses$viral)] <- 0
-viruses$hallmark[is.na(viruses$hallmark)] <- 0
-
-#Virsorter
-viruses$category[is.na(viruses$category)] <- 0
-
-# Kaiju
-viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
-viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
-
-```
-
-```{r}
-viruses_filename <- 'merged_viral_id_outputs.csv'
-#write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
-```
-
-Data cleaning is done!
-Export viruses to tsv and load into identifying_viral_metrics.Rmd for metrics on tool outputs.
-The code for metrics is also pasted below, but it is not recommended running it in the same R session due to memory issues.
-The construction of the final dataframe essentially doubles the memory being used, and even with garbage collector can make Rstudio unresponsive.
-
----
-title: "identifying_viral_metrics_environmental"
-author: Bridget Hegarty, James Riddell
-date: 7/23/2022
-output: html_document
----
 
@@
   if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
     keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
     keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus=="Yes"] + 0.5
-#    keep_score[input_seqs$vibrant_quality=="low quality draft"] <- keep_score[input_seqs$vibrant_quality=="low quality draft"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] + 0.5
   }
@@
     keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 1
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
     keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 1
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
   }
@@
 
-  if (include_kaiju) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
+  if (include_tuning_viral) {
+    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
+    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
   }
 
-  if (include_tuning) {
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$percent_unknown>=75] <- keep_score[input_seqs$percent_unknown>=75] + 0.5
-    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
-    #keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] <- keep_score[input_seqs$hallmark>=(input_seqs$checkv_viral_genes/5)] + 1 #add some ratio
-    keep_score[input_seqs$viral_genes==0 & input_seqs$host_genes>=1] <- 0
-    keep_score[((input_seqs$viral_genes*3) <= input_seqs$host_genes) & input_seqs$checkv_provirus=="No"] <- 0 # consider accounting for provirus designation
-#    keep_score[(input_seqs$viral_genes*3) <= input_seqs$host_genes] <- 0 # consider accounting for provirus designation
-    keep_score[input_seqs$contig_length>50000 & input_seqs$hallmark==0] <- 0
+  if (include_tuning_not_viral) {
+    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
+    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
+    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for RNA viruses
+    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
+    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 0.5 
+    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
   }
@@
 }
-
-```
-
-## 
-```{r}
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = T)
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = T)
-```
-
-```{r}
-table(viruses$keep_score_high_precision>=1)
-```
-
-## vibrant virsorter2 virsorter checkV
-```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-```
-
-# Plot of how many methods identify a contig
-```{r}
-viruses_high <- viruses[viruses$keep_score_high_precision>=1,] #uncomment this line if high precision example
-#viruses_high <- viruses[viruses$keep_score_high_recall>=1,] #uncomment this line if want high recall example 
-#viruses_high <- viruses[viruses$keep_score_all>=1,] #uncomment this line if want to use all 6 tools
-#viruses_high <- viruses[,] #comment this line if want only the "high precision contigs"
-viruses_high_mod <- viruses_high %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
-viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-virusHighCount <- function(df, tool_combo) {
-    tb <- rowSums(df[df[,tool_combo]>0,])
-    if (length(tb) == 0) {
-        return (0)
-    }
-    else {
-        return (table(tb))
-    }
-}
-
-sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,]))
-                          )
-)
-
-sm <- as.data.frame(
-    rbind(
-        virusHighCount(viruses_high_mod, "keep_score_vb"),
-        virusHighCount(viruses_high_mod, "keep_score_dvf"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2"),
-        virusHighCount(viruses_high_mod, "keep_score_vs"),
-        virusHighCount(viruses_high_mod, "keep_score_kj")
-    )
-)
-
-sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "kaiju")
-sm
 ```
@@
 ```{r}
-c_list <- data.frame(viruses_high$checkv_uniq_contig)
-write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
+#c_list <- data.frame(viruses_high$checkv_uniq_contig)
+#write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
 ```
@@
 ```{r}
-write_tsv(viruses_high, paste(dataset_name, "viruses_high.tsv", sep='_'))
-```
-
-```{r}
-sm_m <- reshape2::melt(sm)
-colnames(sm_m) <- c("primary_method", "number_methods", "number")
-```
-
-```{r}
-ggplot(sm_m, aes(x=primary_method, y=number,
-                   fill=number_methods)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(5)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs")
+#write_tsv(viruses_high, paste(dataset_name, "viruses_high.tsv", sep='_'))
 ```
@@
             combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$toolcombo2[p] <- paste(if(i){"C"}else{"0"},if(j){"DVF"}else{"0"},
-                                               if(k){"K"}else{"0"},if(l){"VB"}else{"0"},
+            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
                                                if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
-            combos_list$CheckV[p] <- i
+            combos_list$tune_not_viral[p] <- i
             combos_list$DVF[p] <- j
-            combos_list$Kaiju[p] <- k
+            combos_list$tune_viral[p] <- k
             combos_list$VIBRANT[p] <- l
@@
                                             include_virsorter2 = combos_list$VS2[i],
-                                            include_tuning = combos_list$CheckV[i],
-                                            include_kaiju = combos_list$Kaiju[i],
+                                            include_tuning_viral = combos_list$tune_viral[i],
+                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                             include_deepvirfinder = combos_list$DVF[i])
 
+  if (max(viral_scores[,i])<=0) {
+    num_viruses$num_viruses[i] <- 0
+  }
+  else {
     num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
-
+  }
   num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
   ylab("Num Viruses Predicted")
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
   ylab("Num Viruses Predicted")
@@
 ```
+
+# Visualizing how tools build on each other
+
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+```
+
+# Considering how each method contributes to the final prediction
+```{r}
+viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
+viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
+                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
+                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
+#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
+viruses_high_mod <- as_tibble(viruses_high_mod)
+
+```
+
+```{r}
+sm_m <- reshape2::melt(viruses_high_mod)
+colnames(sm_m) <- c("method", "viral_score")
+```
+
+```{r}
+sm_m <- sm_m[sm_m$viral_score>0,]
+
+sm_m$score <- sm_m$viral_score
+
+sm_m$score[sm_m$viral_score==0.5] <- "0.5"
+sm_m$score[sm_m$viral_score>=1] <- "1"
+sm_m$score[sm_m$viral_score>=2] <- "2"
+sm_m$score[sm_m$viral_score>=3] <- "3"
+sm_m$score[sm_m$viral_score>=4] <- "4"
+sm_m$score[sm_m$viral_score>=5] <- "5"
+
+sm_m$score <- factor(sm_m$score, 
+                                       levels=c("0.5", "1", "2","3","4","5"))
+```
+
+```{r}
+ggplot(sm_m, aes(x=method, y=score,
+                   fill=score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name = 'Number of Methods',
+                     values = alpha(c(viridis(6)), 1)) +
+  xlab("") +
+  ylab("Viral Score") +
+  coord_flip()
+```

### 2022-10-26 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
Splitting building viral set and identifying viral contigs scripts in the generic versions.

@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "Building Environmental Viral Set and Visualization"
 author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+date: 10-26-2022
 output: html_notebook
 ---
-This Rmarkdown file takes outputs from viral identification tools and merges
-them into one file, "{dataset}_merged_viral_id_outputs.csv", which is the input
-for identifying_viral_metrics_environmental.Rmd.
+This Rmarkdown file takes the output from "building_env_data_viral_features_set.Rmd" and creates a set of viral contigs and visualizations.
 
 Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
-Bridget Hegarty (email) regarding any issues, or open an issue on github.
+Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.
+
+TO DO:
+- add kaiju and DVF
 
@@
 
-# Inputs
-1) For this script's inputs, the user must merge each viral identification 
-tool output with the following specified filenames into one tab-separated file. 
-Replace the filenames with their paths if needed into the variables in the 
-'inputs' chunk.
+```{r}
+# dataset name for file organization and outputs for metrics file
+dataset_name <- "drinking_water" #replace with the dataset_name from "building_env_data_viral_features_set.Rmd"
+#name from building_env_data_viral_features_set.Rmd
+viruses_filename <- 'merged_viral_feature_outputs.csv'
 
-    General filenames for each tool:
-        Checkv: quality_summary.tsv
-        VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
-        Virsorter: VIRSorter_global-phage-signal.csv
-        Virsorter2: final-viral-score.tsv
-        DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
-        Kaiju: ${assembly}.kaiju.names.out
-
-2) Make sure all fasta headers (assembly, contig) are consistent within and
-across tools. Chunks for each tool do contain some lines for cleaning up these features, but due to their variability it will be the user's responsibility to 
-make sure they match across tools.
-
-3) Check each chunk and ensure all columns are accounted for. For example,
-if you don't have "assembly" and called it something else, please either
-remove it from this script or modify your input files.
-
-4) This script is designed for contigs > 3000 basepairs. It can be modified
-to be higher or lower, but going lower will greatly increase the size of the
-dataframe and memory usage.
-
-```{r inputs}
-# dataset name for file organization and outputs for metrics file
-dataset_name <- "lake_michigan"
-# checkV
-checkV_path <- "merged_lake_michigan_checkv_assemblies.tsv"
-# VIBRANT
-vibrant_path <- "merged_lake_michigan_vibrant_assemblies.tsv"
-# DeepVirFinder
-dvf_path <- "merged_lake_michigan_dvf_assemblies.tsv"
-# Virsorter
-vs_path <- "merged_lake_michigan_virsorter_assemblies.tsv"
-# Virsorter2
-vs2_path <- "merged_lake_michigan_vs2_assemblies.tsv"
-# Kaiju
-kj_path <- "merged_lake_michigan_kaiju_assemblies.tsv"
-# KB cutoff
-KB_CUTOFF <- 3000
-
+viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
 ```
-
-# All Viral Contigs
-
-## checkV
-```{r}
-checkV <- fread(checkV_path, 
-                sep="\t",
-                header = T, 
-                select = c(
-                    'contig_id',
-                    'assembly',
-                    'provirus',
-                    'completeness',
-                    'contamination',
-                    'viral_genes',
-                    'host_genes',
-                    'gene_count',
-                    'contig_length',
-                    'checkv_quality'
-                    )
-                ) %>% 
-    rename(
-        contig = contig_id,
-        checkv_provirus = provirus,
-        checkv_completeness = completeness,
-        checkv_contamination = contamination,
-        checkv_viral_genes = viral_genes,
-        checkv_host_genes = host_genes,
-        checkv_total_genes = gene_count,
-        checkv_length = contig_length
-        )
-checkV$method = 'checkv'
-checkV$contig <- gsub('_', '', checkV$contig)
-checkV$checkv_uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
-```
-
-## VIBRANT
-```{r}
-vb_c <- fread(vibrant_path,
-              header = T,
-              sep = "\t",
-              select = c(
-                  'scaffold',
-                  'assembly',
-                  'type',
-                  'Quality'
-                  )
-              ) %>%
-    rename(
-        contig = scaffold,
-        vibrant_quality = Quality
-        )
-vb_c$method <- "vibrant"
-vb_c$vibrant_prophage <- "No"
-vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
-vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
-vb_c$contig <- gsub('_', '', vb_c$contig)
-vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
-vb_c <- vb_c[!duplicated(vb_c$uniq_contig),]
-```
-
-## DeepVirFinder
-```{r}
-dvf_c <- fread(dvf_path,
-               header = T,
-               sep = "\t",
-               select = c(
-                   'name',
-                   'assembly',
-                   'score',
-                   'pvalue'
-                   )
-               ) %>% 
-    rename(
-        contig = name
-        )
-dvf_c$contig <- gsub('_', '', dvf_c$contig)
-dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
-```
-
-## VirSorter
-```{r}
-vs_c <- fread(vs_path,
-              select = c(
-                  '## Contig_id',
-                  'Nb genes contigs',
-                  'Nb genes',
-                  'Category',
-                  'Nb phage hallmark genes'
-                  )
-              ) %>% drop_na(`## Contig_id`)
-vs_c$`## Contig_id` <- sub("VIRSorter_", "", vs_c$`## Contig_id`)
-vs_c$`## Contig_id` <- sub("-circular", "", vs_c$`## Contig_id`)
-vs_c$`## Contig_id` <- gsub("_", "", vs_c$`## Contig_id`)
-vs_c <- separate(vs_c, col = `## Contig_id`, into = c("assembly", "contig"), sep="--")
-
-vs_c <- vs_c %>% drop_na(contig)
-```
-
-## VirSorter2 
-```{r}
-vs2_c <- fread(vs2_path,
-                    header = T,
-                    sep = '\t',
-                    select = c(
-                        'seqname',
-                        'assembly',
-                        'dsDNAphage',
-                        'RNA',
-                        'ssDNA',
-                        'max_score',
-                        'max_score_group',
-                        'hallmark',
-                        'viral',
-                        'cellular'
-                    )
-               ) %>% 
-              separate(
-                  col = seqname,
-                  into = c("contig", "type"), 
-                  sep = "\\|\\|",
-                  remove = T
-                  )
-vs2_c$contig <- gsub("_", "", vs2_c$contig)
-vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
-vs2_c <- vs2_c[!duplicated(vs2_c$uniq_contig),]
-```
-
-## Kaiju
-(parsing when reading the file in "fails" for sequences that weren't classified)
-```{r}
-kj_c <- fread(kj_path)
-
-colnames(kj_c) <-c(
-    "Classified",
-    "Contig",
-    "NCBI_taxon",
-    "len",
-    "ID_best",
-    "IDs_all",
-    "Seq",
-    "Name"
-    )
-kj_c$Contig <- gsub('_', '', kj_c$Contig)
-kj_c <- setDT(kj_c)[, c("assembly", "contig") := tstrsplit(Contig, "--")]
-kj_c <- setDT(kj_c)[, c("Kaiju_Viral", "Kingdom") := tstrsplit(Name, ";", keep = c(1, 2))]
-```
-
-```{r}
-# Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
-table(kj_c$assembly, kj_c$Kaiju_Viral)
-```
-
-## Merging
-
-```{r}
-viruses <- full_join(x=vs2_c, y=dvf_c, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=checkV, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=vb_c, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("contig", "assembly"))
-viruses <- full_join(x=viruses, y=kj_c, by = c("contig", "assembly"))
-```
-
-Remove contigs not greater than the basepair length cutoff (3000)
-
-```{r}
-viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
-```
-
-Use the next code chunk to check if any contigs are missing from checkv.
-If yes, then there is a mix up in contig names since checkv should contain
-all contig names even if they are identified as non-viral.
-```{r}
-v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
-```
-
-calculate percent viral/host/unknown
-```{r}
-viruses$percent_host <- viruses$checkv_host_genes/viruses$checkv_total_genes*100
-viruses$percent_viral <- viruses$checkv_viral_genes/viruses$checkv_total_genes*100
-viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
-```
-
-get rid of NAs for downstream processing
-```{r}
-
-# checkV
-viruses$percent_viral[is.na(viruses$percent_viral)] <- 0
-viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
-
-# VIBRANT
-viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
-
-# DeepVirFinder
-viruses$score[is.na(viruses$score)] <- 0
-viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
-
-# VirSorter2
-viruses$viral[is.na(viruses$viral)] <- 0
-viruses$hallmark[is.na(viruses$hallmark)] <- 0
-
-#Virsorter
-viruses$Category[is.na(viruses$Category)] <- 0
-
-# Kaiju
-viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
-viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
-
-```
-
-```{r}
-viruses_filename <- 'merged_viral_id_outputs.csv'
-write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
-```
-
-Data cleaning is done!
-Export viruses to tsv and load into identifying_viral_metrics.Rmd for metrics on tool outputs.
-The code for metrics is also pasted below, but it is not recommended running it in the same R session due to memory issues.
-The construction of the final dataframe essentially doubles the memory being used, and even with garbage collector can make Rstudio unresponsive.
-
----
-title: "identifying_viral_metrics_environmental"
-author: Bridget Hegarty, James Riddell
-date: 7/23/2022
-output: html_document
----
 
@@
 A final keep_score above 1 indicates we will keep that sequence and call it viral.
+
+TO: these rules are not written up correctly
 
@@
 Virsorter
-    Category <= 5: +1
-    Category == 6: +0.5
+    category ==  1,2,4,5: +1
+    category == 3,6: +0.5
 
@@
 
-## vibrant virsorter2 virsorter checkV
-```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-table(viruses$keep_score_high_precision>=1)
-```
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-viruses$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = F,
-                                              include_virsorter = T)
-viruses$keep_score_tuning <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = F,
-                                              include_tuning = T,
-                                              include_virsorter = F)
-viruses$keep_score_kj <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = F,
-                                              include_virsorter = F)
-```
-
-# Plot of how many methods identify a contig
-```{r}
-#viruses_high <- viruses[viruses$keep_score_all>=1,]
-viruses_high_mod <- viruses %>% select(keep_score_vb, keep_score_dvf,keep_score_vs2, keep_score_vs, keep_score_kj)
-viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-virusHighCount <- function(df, tool_combo) {
-    tb <- rowSums(df[df[,tool_combo]>0,])
-    if (length(tb) == 0) {
-        return (0)
-    }
-    else {
-        return (table(tb))
-    }
-}
-sm <- as.data.frame(rbind(table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vb>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_dvf>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs2>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_vs>0,])),
-                          table(rowSums(viruses_high_mod[viruses_high_mod$keep_score_kj>0,]))
-                          )
-)
-sm <- as.data.frame(
-    rbind(
-        virusHighCount(viruses_high_mod, "keep_score_vb"),
-        virusHighCount(viruses_high_mod, "keep_score_dvf"),
-        virusHighCount(viruses_high_mod, "keep_score_vs2"),
-        virusHighCount(viruses_high_mod, "keep_score_vs"),
-        virusHighCount(viruses_high_mod, "keep_score_kj")
-    )
-)
-sm$method <- c("vibrant", "deepvirfinder", "virsorter2", "virsorter", "kaiju")
-sm
-```
-
 Create list of contigs considered viral
 ```{r}
-c_list <- data.frame(viruses_high$checkv_uniq_contig)
-write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
+#c_list <- data.frame(viruses_high$checkv_uniq_contig)
+#write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
 ```
@@
 p <- 1
+
 for (i in c(0,1)){
@@
             combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$CheckV[p] <- i
+            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
+                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
+            combos_list$tune_not_viral[p] <- i
             combos_list$DVF[p] <- j
-            combos_list$Kaiju[p] <- k
+            combos_list$tune_viral[p] <- k
             combos_list$VIBRANT[p] <- l
@@
 }
+
 combos_list <- combos_list[-1,]
@@
                           num_viruses=rep(0, nrow(combos_list)))
+
 for (i in 1:nrow(combos_list)) {
@@
                                             include_virsorter2 = combos_list$VS2[i],
-                                            include_tuning = combos_list$CheckV[i],
-                                            include_kaiju = combos_list$Kaiju[i],
+                                            include_tuning_viral = combos_list$tune_viral[i],
+                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                             include_deepvirfinder = combos_list$DVF[i])
 
+  if (max(viral_scores[,i])<=0) {
+    num_viruses$num_viruses[i] <- 0
+  }
+  else {
     num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+  }
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
 
-  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
 }
+
 num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
@@
 ```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
 ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
@@
   ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
+  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+
+ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
+ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
+  geom_boxplot(aes(color=numtools)) +
+  geom_point(aes(color=numtools, fill=numtools)) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
   ylab("Num Viruses Predicted")
@@
 viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo
+
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
 ```
@@
                                      sample_data(tooldata))
+```
+
+```{r}
 ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
@@
   geom_label(label=tooldata$toolcombo)
+
 phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
@@
 to do: try coloring above based on the F1 scores of the testing set on each combination
+
+```{r}
+bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
+clusters <- hclust(dist(bray_dist))
+plot(clusters)
+
+myclusters <- cutree(clusters, h=0.8)
+```
+
+```{r}
+names(myclusters[myclusters==1])
+names(myclusters[myclusters==2])
+names(myclusters[myclusters==3])
+names(myclusters[myclusters==4])
+names(myclusters[myclusters==5])
+
+myclusters_df <- tibble(combo=names(myclusters),
+                            cluster_index=myclusters)
+
+myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
+                                                            "Kaiju", "VIBRANT",
+                                                            "VirSorter", "VirSorter2"),
+                          sep=" ", remove = F)
+
+tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
+                    )
+
+tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
+
+tool_count <- melt(tool_count)
+
+colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+ggplot(tool_count, aes(x=cluster_index, y=tool_count,
+                   fill=cluster_index,
+                   color=cluster_index)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  xlab("Cluster") +
+  ylab("Number of Times in Cluster") + 
+  facet_wrap(~tool, scales = "free")
+```
 
@@
   xlab("Num Viruses Predicted")
+
 ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
@@
   xlab("Num Viruses Predicted")
+
 ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 

### 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Visualizing viral totals by assembly.

@@
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 0.5 
+    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] - 0.5 
     keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
@@
 viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
-num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
-                          num_viruses=rep(0, nrow(combos_list)))
 
+num_assemblies <- dim(table(viruses$assembly))
+assemblies <- names(table(viruses$assembly))
+
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)*num_assemblies),
+                          toolcombo2=rep(0, nrow(combos_list)*num_assemblies),
+                          assembly=rep(0,nrow(combos_list)*num_assemblies),
+                          num_viruses=rep(0, nrow(combos_list)*num_assemblies),
+                          prop_vir=rep(0, nrow(combos_list)*num_assemblies)
+                          )
+
+n <- 1
 for (i in 1:nrow(combos_list)) {
@@
                                             include_deepvirfinder = combos_list$DVF[i])
-
+  for (j in 1:num_assemblies) {
     if (max(viral_scores[,i])<=0) {
-    num_viruses$num_viruses[i] <- 0
+      num_viruses$num_viruses[n] <- 0
     }
     else {
-    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+      num_viruses$num_viruses[n] <- sum(viral_scores[viruses$assembly==assemblies[j],]>=1)
+      num_viruses$prop_vir[n] <- num_viruses$num_viruses[n]/table(viruses$assembly)[[j]]
     }
-  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
 
-  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
+    num_viruses$toolcombo[n] <- combos_list$toolcombo[i]
+    num_viruses$toolcombo2[n] <- combos_list$toolcombo2[i]
+    num_viruses$assembly[n] <- assemblies[j]
+
+    n <- n+1
+  }
 }
@@
 num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
-num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses <- num_viruses[order(num_viruses$prop_vir, decreasing=F),]
 num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
@@
 ```{r}
-ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
+ggplot(num_viruses, aes(x=toolcombo, y=prop_vir, 
+                                  shape=numtools, color=assembly, fill=assembly)) +
   geom_point() +
@@
   xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
+  ylab("Proportion of Contigs >3kb Viral")
 

### 2022-10-28 TestingPipeline/Scripts/combining_tool_output.Rmd
Updated with new vs2 file.

@@
                         'ssDNA',
+                        'NCLDV',
+                        'RNA',
+                        'lavidaviridae',
                         'max_score',
@@
 kj_c <- kj_c[!duplicated(paste(kj_c$contig, kj_c$Index)),]
-```
-Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
-```{r}
-table(kj_c$seqtype, kj_c$Kaiju_Viral)
 ```
@@
 viruses$percent_unknown[is.na(viruses$percent_unknown)] <- 0
+viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0
 # VIBRANT

### 2022-10-28 TestingPipeline/Scripts/combining_tool_output.Rmd
Updating rules with different viral types from VS2.

@@
 viruses$hallmark[is.na(viruses$hallmark)] <- 0
+viruses$RNA[is.na(viruses$RNA)] <- 0
+viruses$lavidaviridae[is.na(viruses$lavidaviridae)] <- 0
+viruses$NCLDV[is.na(viruses$NCLDV)] <- 0
+viruses$ssDNA[is.na(viruses$ssDNA)] <- 0
+viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
+
 #Virsorter

### 2022-10-28 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updating rules with different viral types from VS2.

@@
 
-```{r}
-viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0
-```
-
 This section defines a viralness score "keep_score" based on the tool classifications. 
@@
     keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
+    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 0.5
+    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 0.5
+    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 0.5
   }

### 2022-10-28 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Adjusting rules to help with false positive protists.
Updating rules with different viral types from VS2.

@@
 colnames(viruses)
+```
+
+```{r}
+plot(viruses$RNA, viruses$viral)
+plot(viruses$lavidaviridae, viruses$viral)
+plot(viruses$NCLDV, viruses$viral)
+plot(viruses$ssDNA, viruses$viral)
 ```
@@
   xlab("Length") +
+  ylab("Completeness") 
+```
+
+```{r}
+ggplot(viruses, aes(x=hallmark, y=checkv_completeness)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("Hallmark Genes") +
   ylab("Completeness") 

### 2022-10-28 TestingPipeline/Scripts/MergingScripts/merge_virsorter2.py

@@ -6,7 +6,7 @@ create a list of all the vs2 viral contigs
 with open("/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/testing_sets_virsorter2_output.tsv", 'wt') as outfile:
     j=0
     for index in range(1,11):
-        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/TestingSetToolOutputs/Testing_Set_%s/final-viral-score.tsv" %(index)
+        filename="/scratch/duhaimem_root/duhaimem/shared_data/VSTE/RunningToolsOnTestingSet/VS2/VS2_%s/final-viral-score.tsv" %(index)
         with open(filename, 'rt') as v:
             if (j==0):
                 i=0

### 2022-10-29 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Tuning output plots with new rules.

@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "drinking_water"
+#dataset_name <- "drinking_water"
 #name from building_env_data_viral_features_set.Rmd
-viruses_filename <- 'merged_viral_feature_outputs.csv'
+#viruses_filename <- 'merged_viral_feature_outputs.csv'
 
-viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+#viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
 ```
@@
     keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
+    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 1
+    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 1
+    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 1
   }
@@
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] - 0.5 
     keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
@@
 ```
+
+```{r}
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+sum(viruses$keep_score_all>=1)
+sum(viruses$keep_score_high_recall>=1)
+sum(viruses$keep_score_high_precision>=1)
+sum(viruses$keep_score_high_MCC>=1)
+
+sum(viruses$keep_score_all>=0.5)
+sum(viruses$keep_score_high_recall>=0.5)
+sum(viruses$keep_score_high_precision>=0.5)
+sum(viruses$keep_score_high_MCC>=0.5)
+```
+
+not having the rule, gives the same
 
@@
 viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
+                          num_viruses=rep(0, nrow(combos_list)))
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning_viral = combos_list$tune_viral[i],
+                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+
+  if (max(viral_scores[,i])<=0) {
+    num_viruses$num_viruses[i] <- 0
+  }
+  else {
+    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+  }
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+
+  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
+}
+num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
+```
+
+```{r}
+ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
+                                  color=numtools, fill=numtools)) +
+  geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
+ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
+  geom_boxplot(aes()) +
+  geom_point(aes(color=assembly, fill=assembly)) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Num Viruses Predicted")
+```
+
+```{r}
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
 
@@
     else {
-      num_viruses$num_viruses[n] <- sum(viral_scores[viruses$assembly==assemblies[j],]>=1)
+      num_viruses$num_viruses[n] <- sum(viral_scores[viruses$assembly==assemblies[j],i]>=1)
       num_viruses$prop_vir[n] <- num_viruses$num_viruses[n]/table(viruses$assembly)[[j]]
@@
   ylab("Proportion of Contigs >3kb Viral")
+```
 
-ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
+by only a subset of the assemblies (UK/NL)
+
+```{r}
+num_viruses_uknl <- num_viruses[num_viruses$assembly %in% c("Aberdeen_Man",
+                                                            "Amsterdam_Leid",
+                                                            "Amsterdam_Wee",
+                                                            "Birmingham_Maria",
+                                                            "Cardiff_Maria",
+                                                            "Dundee_Clat",
+                                                            "Dundee_Maria",
+                                                            "Edinburgh_Glen",
+                                                            "Exeter_Maria",
+                                                            "Glasgow_Bal",
+                                                            "Glasgow_Mil",
+                                                            "London_Maria",
+                                                            "Nieuwkoop_DHB",
+                                                            "South_Holland_Dla",
+                                                            "Spannenburg"),]
+```
+
+```{r}
+ggplot(num_viruses_uknl, aes(x=toolcombo, y=prop_vir, 
+                                  shape=numtools, color=assembly, fill=assembly)) +
   geom_point() +
@@
   ) +
-  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
+  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral")
+```
+
+ordination: sample by tool combo
+
+```{r}
+library(phyloseq)
+```
+
+make the sample by tool combo matrix
+```{r}
+num_viruses_uknl_wide <- num_viruses_uknl %>% 
+  select(toolcombo, assembly, num_viruses) %>%
+  spread(assembly, num_viruses)
+rownames(num_viruses_uknl_wide) <- num_viruses_uknl_wide$toolcombo
+num_viruses_uknl_wide <- num_viruses_uknl_wide[,-1]
+
+num_viruses_uknl_wide <- num_viruses_uknl_wide[rowSums(num_viruses_uknl_wide)>0,]
 ```
@@
 ```{r}
-ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
-  geom_boxplot(aes(color=numtools)) +
-  geom_point(aes(color=numtools, fill=numtools)) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Tools") +
-  ylab("Num Viruses Predicted")
+tooldata <- combos_list
+rownames(tooldata) <- tooldata$toolcombo
+
+tooldata <- tooldata[c(rownames(num_viruses_uknl_wide)),]
+
+tooldata$numtools <- str_count(tooldata$toolcombo, "1")
+
+tooldata$numtools <- as.factor(tooldata$numtools)
+
+tooldata$total_viruses <- rowSums(num_viruses_uknl_wide)
 ```
+
+to reduce arch effect:
+```{r}
+num_viruses_uknl_wide <- num_viruses_uknl_wide/rowSums(num_viruses_uknl_wide)
+```
+
+```{r}
+physeq_pooled <- phyloseq(otu_table(num_viruses_uknl_wide, taxa_are_rows = F),
+                                     sample_data(tooldata))
+```
+
+```{r}
+ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
+```
+
+```{r}
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", 
+                          color="total_viruses") + 
+  geom_point(size = 3) +
+  theme_bw() +
+  scale_shape_manual(name="Num Tools",
+                     values = c(21,22,23,24,4,8)) +
+  scale_color_viridis_c()
+```
+each point is tool combination
 

### 2022-10-29 TestingPipeline/Scripts/combining_tool_output.Rmd
Tuning output plots with new rules.

--- 2022-10-28 TestingPipeline/Scripts/combining_tool_output.Rmd
@@
 viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0
+
 # VIBRANT
 viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
+
 # DeepVirFinder
@@
 viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
+
 # VirSorter2
@@
 viruses$hallmark[is.na(viruses$hallmark)] <- 0
+viruses$RNA[is.na(viruses$RNA)] <- 0
+viruses$lavidaviridae[is.na(viruses$lavidaviridae)] <- 0
+viruses$NCLDV[is.na(viruses$NCLDV)] <- 0
+viruses$ssDNA[is.na(viruses$ssDNA)] <- 0
+viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
+
 #Virsorter
 viruses$category[is.na(viruses$category)] <- 0
+
 # Kaiju

### 2022-10-29 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Tuning output plots with new rules.

@@
     keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
-    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 0.5
-    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 0.5
-    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 0.5
+    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 1
+    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 1
+    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 1
   }
@@
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 0.5 
     keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
@@
 ```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
                                               include_virsorter2 = F,
-                                              include_kaiju = T,
-                                              include_tuning = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = F)
@@
 ```{r}
+length(grep("true", viruses$confusion_matrix_high_precision))/nrow(viruses)
+```
+
+```{r}
 pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
@@
 ```
-
-```{r}
- ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Viral Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
- ggplot(viruses, aes(x=percent_viral, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Percent Genes Viral") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
- ggplot(viruses, aes(x=hallmark, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-viruses_false_positive <- viruses[viruses$confusion_matrix_high_precision=="false positive",]
-viruses_false_negative <- viruses[viruses$confusion_matrix_high_precision=="false negative",]
-```
-
-```{r}
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=checkv_length,
-                   color=checkv_length,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="bacteria"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="fungi"], aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="protist"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)
-
-table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
-```
+this rule set had the highest precision, but as you can see, this comes with a big sacrifice in recall
 
@@
 
+accuracy:
+```{r}
+length(grep("true", viruses$confusion_matrix_high_MCC))/nrow(viruses)
+```
+
+recall
+```{r}
+length(grep("true positive", viruses$confusion_matrix_high_MCC))/length(grep("virus", viruses$seqtype))
+```
+
+```{r}
+viruses$size_class <- "3-10kb"
+viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
+```
+
 visualizing confusion matrix by taxa
 ```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_MCC, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
+confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, Index)
+
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","index", "count")
 ```
@@
 ```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+ggplot(confusion_by_taxa, aes(x=count, y=as.factor(index),
                    fill=confusion_matrix,
@@
 ```
-
-```{r}
-ggplot(viruses, aes(x=checkv_length, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_MCC,
-                   color=confusion_matrix_high_MCC)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
+percent viral
+
+visualizing confusion matrix by taxa
+```{r}
+confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, size_class, Index)
+
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "index", "count")
+```
+
+```{r}
+confusion_vir_called <- confusion_by_taxa %>% filter(confusion_matrix=="true positive" | confusion_matrix=="false positive")
+
+type_count <- viruses %>% count(seqtype, size_class, Index)
+
+confusion_vir_called$per_viral <- 0
+
+for (i in c(1:nrow(confusion_vir_called))) {
+  confusion_vir_called$per_viral[i] <- confusion_vir_called$count[i]/type_count$n[type_count$seqtype==confusion_vir_called$seqtype[i] & 
+                                                                                    type_count$Index==confusion_vir_called$index[i] &
+                                                                                    type_count$size_class==confusion_vir_called$size[i]]*100
+}
+
+confusion_vir_called <- confusion_vir_called %>% group_by(seqtype, size) %>%
+  summarise(mean=mean(per_viral), 
+            sd=sd(per_viral))
+```
+
+```{r}
+ggplot(confusion_vir_called, aes(y=mean, x=size,
+                   fill=seqtype,
+                   color=seqtype)) +
+  geom_bar(stat="identity", position=position_dodge()) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
+                 position=position_dodge(.9)) +
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
+                     values = alpha(rev(pal(6)), 0.5)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-## high recall example
-```{r}
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                     values = alpha(rev(pal(6)), 1)) +
+  xlab("Length (bp)") +
+  ylab("Sequences Called Viral (%)") 
+```
+
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
 
-```{r}
-viruses$confusion_matrix_high_recall <- "true negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall<1] <- "false negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
-viruses$confusion_matrix_high_recall[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$confusion_matrix_high_recall))/nrow(viruses)
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_recall, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
+# Considering how each method contributes to the final prediction
+```{r}
+viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
+viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
+                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
+                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
+#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
+viruses_high_mod <- as_tibble(viruses_high_mod)
+
+```
+
+```{r}
+sm_m <- reshape2::melt(viruses_high_mod)
+colnames(sm_m) <- c("method", "viral_score")
+```
+
+```{r}
+sm_m <- sm_m[sm_m$viral_score>0,]
+
+sm_m$score <- sm_m$viral_score
+
+sm_m$score[sm_m$viral_score==0.5] <- "0.5"
+sm_m$score[sm_m$viral_score>=1] <- "1"
+sm_m$score[sm_m$viral_score>=2] <- "2"
+sm_m$score[sm_m$viral_score>=3] <- "3"
+sm_m$score[sm_m$viral_score>=4] <- "4"
+sm_m$score[sm_m$viral_score>=5] <- "5"
+
+sm_m$score <- factor(sm_m$score, 
+                                       levels=c("0.5", "1", "2","3","4","5"))
+```
+
+```{r}
+ggplot(sm_m, aes(x=method, y=score,
+                   fill=score)) +
   geom_bar(stat="identity") +
   theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name = 'Viral Score',
+                     values = alpha(c(viridis(6)), 1)) +
+  xlab("") +
+  ylab("Number of Sequences") +
   coord_flip()
-p2
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=hallmark,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Number of Hallmark Genes") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=confusion_matrix_high_recall, y=checkv_length,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") +
-  scale_y_log10()
-```
-
-looking at false negatives
-```{r}
-viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<1),]
-```
-
-looking at protists calling viral
-```{r}
-viruses_false_pos_protists <- viruses[(viruses$seqtype=="protist" & viruses$keep_score_high_recall>=1),]
-```
-
-# Visualizing confusion matrix by number of tools
+```
+## Visualizing confusion matrix by number of tools
 
@@
 ```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
 ggplot(viruses, aes(x=as.factor(Index),
@@
   scale_color_manual(name = 'Viral Score',
-                     values = alpha(c(pal(6)), 1)) +
+                     values = alpha(c(viridis(6)), 1)) +
   scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(pal(6)), 0.5)) +
+                     values = alpha(c(viridis(6)), 0.5)) +
   xlab("Index") +
@@
 
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+## all 6 tools example
+```{r}
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
                                               include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+```
+
+```{r}
+viruses$confusion_matrix_all <- "true negative"
+viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
+viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
+viruses$confusion_matrix_all[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
+```
+
+visualizing confusion matrix by taxa
+```{r}
+confusion_by_taxa <- melt(table(viruses$confusion_matrix_all, viruses$seqtype, viruses$Index))
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
+```
+
+```{r}
+table(viruses$confusion_matrix_all)
+
+length(grep("true", viruses$confusion_matrix_all))/nrow(viruses)
+```
+
+```{r}
+length(grep("true positive", viruses$confusion_matrix_all))/length(grep("virus", viruses$seqtype))
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+```
+
+## high recall example
+```{r}
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
                                               include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-```
-
-# Considering how each method contributes to the final prediction
-```{r}
-viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
-viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
-                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
-                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "viral_score")
-```
-
-```{r}
-sm_m <- sm_m[sm_m$viral_score>0,]
-
-sm_m$score <- sm_m$viral_score
-
-sm_m$score[sm_m$viral_score==0.5] <- "0.5"
-sm_m$score[sm_m$viral_score>=1] <- "1"
-sm_m$score[sm_m$viral_score>=2] <- "2"
-sm_m$score[sm_m$viral_score>=3] <- "3"
-sm_m$score[sm_m$viral_score>=4] <- "4"
-sm_m$score[sm_m$viral_score>=5] <- "5"
-
-sm_m$score <- factor(sm_m$score, 
-                                       levels=c("0.5", "1", "2","3","4","5"))
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=score)) +
+```
+
+```{r}
+viruses$confusion_matrix_high_recall <- "true negative"
+viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall<1] <- "false negative"
+viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
+viruses$confusion_matrix_high_recall[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
+```
+
+visualizing confusion matrix by taxa
+```{r}
+confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_recall, viruses$seqtype, viruses$Index))
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+                   fill=confusion_matrix,
+                   color=confusion_matrix)) +
   geom_bar(stat="identity") +
   theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(6)), 1)) +
-  xlab("") +
-  ylab("Viral Score") +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
   coord_flip()
+p2
+```
+
+accuracy:
+```{r}
+length(grep("true", viruses$confusion_matrix_high_recall))/nrow(viruses)
+```
+
+recall
+```{r}
+length(grep("true positive", viruses$confusion_matrix_high_recall))/length(grep("virus", viruses$seqtype))
+```
+recover almost all of the viruses this way, but more protist contamination
+
+######################################################################
+Extra Stuff
+#####################################################################
+
+```{r}
+ggplot(viruses, aes(x=checkv_length, y=keep_score_high_MCC,
+                   fill=confusion_matrix_high_MCC,
+                   color=confusion_matrix_high_MCC)) +
+  geom_point(stat="identity", shape=21) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Sequence Length (bp)") +
+  ylab("Pipeline Viral Score") + 
+  facet_wrap(~seqtype) + 
+  scale_x_log10()
+```
+
+```{r}
+ggplot(viruses, aes(x=checkv_completeness, y=hallmark,
+                   fill=confusion_matrix_high_recall,
+                   color=confusion_matrix_high_recall)) +
+  geom_point(stat="identity", shape=21) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("CheckV Completeness") +
+  ylab("Number of Hallmark Genes") + 
+  facet_wrap(~seqtype) + 
+  scale_x_log10()
+```
+
+```{r}
+ggplot(viruses, aes(x=checkv_completeness, y=keep_score_high_MCC,
+                   fill=confusion_matrix_high_recall,
+                   color=confusion_matrix_high_recall)) +
+  geom_point(stat="identity", shape=21) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("CheckV Completeness") +
+  ylab("Pipeline Viral Score") + 
+  facet_wrap(~seqtype) + 
+  scale_x_log10()
+```
+
+```{r}
+ggplot(viruses, aes(x=confusion_matrix_high_recall, y=checkv_length,
+                   fill=confusion_matrix_high_recall,
+                   color=confusion_matrix_high_recall)) +
+  geom_boxplot() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Sequence Length (bp)") +
+  ylab("Pipeline Viral Score") +
+  scale_y_log10()
+```
+
+looking at false negatives
+```{r}
+viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<1),]
+```
+
+looking at protists calling viral
+```{r}
+viruses_false_pos_protists <- viruses[(viruses$seqtype=="protist" & viruses$keep_score_high_recall>=1),]
 ```
@@
 ```
+
+```{r}
+ ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix_high_precision,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
+  geom_boxplot(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Viral Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ ggplot(viruses, aes(x=percent_viral, y=confusion_matrix_high_precision,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
+  geom_boxplot(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Percent Genes Viral") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ ggplot(viruses, aes(x=hallmark, y=confusion_matrix_high_precision,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
+  geom_boxplot(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Hallmark Genes") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
+                   fill=confusion_matrix_high_precision,
+                   color=confusion_matrix_high_precision)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Number of Viral Genes") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+```
+
+```{r}
+viruses_false_positive <- viruses[viruses$confusion_matrix_high_precision=="false positive",]
+viruses_false_negative <- viruses[viruses$confusion_matrix_high_precision=="false negative",]
+```
+
+```{r}
+ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
+                   fill=checkv_length,
+                   color=checkv_length,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Number of Viral Genes") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive, aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~seqtype, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive[viruses_false_positive$seqtype=="bacteria"], aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive[viruses_false_positive$seqtype=="fungi"], aes(x=hallmark, y=checkv_length,
+                   fill=keep_score_high_precision,
+                   color=keep_score_high_precision,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_positive[viruses_false_positive$seqtype=="protist"], aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
+                   fill=checkv_viral_genes,
+                   color=checkv_viral_genes,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+
+ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
+                   fill=keep_score_high_precision,
+                   color=keep_score_high_precision,
+                   shape=checkv_provirus)) +
+  geom_point(alpha=0.3) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Hallmark Genes") +
+  ylab("Contig Length") + 
+  facet_wrap(~Kaiju_Viral, scales = "free") +
+  coord_flip()
+```
+
+```{r}
+table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)
+
+table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
+```

### 2022-10-29 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Tuning output plots with new rules.

@@
 plot(viruses$ssDNA, viruses$viral)
+plot(viruses$checkv_completeness, viruses$hallmark)
 ```

### 2022-10-29 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updating provirus rules to pull from all tools.

@@
 VIBRANT
+    Quality == "Complete Circular": +1
     Quality == "High Quality Draft": +1
@@
     Viral >= 50: +0.5
-    Viral >= 0.95: +1
-    Hallmark > 2: +1
+    Viral >= 0.95: +0.5
+    RNA >= 0.9: +1
+    lavidaviridae >= 0.9: +1
+    NCLDV >= 0.9: +1
 
 Virsorter
-    category ==  1,2,4,5: +1
-    category == 3,6: +0.5
+    category ==  1,4: +1
+    category == 2,5: +0.5
 
@@
 
-Kaiju:
-    Kaiju_viral = "cellular organisms": -1
+Tuning - No Viral Signature:
+    Kaiju_viral = "cellular organisms": -0.5
+    If host_genes >50 and NOT checkv_provirus: -1 
+    If viral_genes == 0 and host_genes >= 1: -1
+    If 3*viral_genes <= host_genes and NOT checkv_provirus: -1
+    If length > 50,000 and hallmark <=1: -1
+    If length < 5000 and checkv completeness <= 75: -0.5
+
+Tuning - Viral Signature:
     Kaiju_viral = "Viruses": +1
-
-CheckV
-    viral_genes == 0 and host_genes >= 1: keep_score = 0
-    If 3*viral_genes <= host_genes: keep_score = 0
-    If length > 50,000 and hallmark == 0: keep_score = 0
-    If %unknown >= 75: +0.5
+    If %unknown >= 75 and length < 50000: +0.5
+    If %viral >= 50: +0.5
+    Hallmark > 2: +1
 
@@
   xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
-  geom_boxplot(aes()) +
-  geom_point(aes(color=assembly, fill=assembly)) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Tools") +
   ylab("Num Viruses Predicted")

### 2022-10-29 TestingPipeline/Scripts/combining_tool_output.Rmd
Updating provirus rules to pull from all tools.

@@
 viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
+
+#adding a combined provirus category
+viruses$provirus <- FALSE
+viruses$provirus[viruses$checkv_provirus=="Yes" | viruses$vibrant_prophage=="Yes" | viruses$category==4] <- TRUE
 ```

### 2022-10-29 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updating provirus rules to pull from all tools.

@@
 VIBRANT
+    Quality == "Complete Circular": +1
     Quality == "High Quality Draft": +1
     Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus == TRUE: +0.5
+    Quality == "Low Quality Draft" & provirus: +0.5
 
@@
     Viral >= 0.95: +0.5
+    RNA >= 0.9: +1
+    lavidaviridae >= 0.9: +1
+    NCLDV >= 0.9: +1
 
 Virsorter
-    category ==  1,2,4,5: +1
-    category == 3,6: +0.5
+    category ==  1,4: +1
+    category == 2,5: +0.5
 
@@
     Score >= 0.7: +0.5
-    Score >= 0.9: +0.5
-
-Kaiju:
-    Kaiju_viral = "cellular organisms": -1
+
+Tuning - No Viral Signature:
+    Kaiju_viral = "cellular organisms": -0.5
+    If host_genes >50 and NOT provirus: -1 
+    If viral_genes == 0 and host_genes >= 1: -1
+    If 3*viral_genes <= host_genes and NOT provirus: -1
+    If length > 50,000 and hallmark <=1: -1
+    If length < 5000 and checkv completeness <= 75: -0.5
+
+Tuning - Viral Signature:
     Kaiju_viral = "Viruses": +1
-
-CheckV
-    If %unknown >= 75: +0.5
+    If %unknown >= 75 and length < 50000: +0.5
+    If %viral >= 50: +0.5
     Hallmark > 2: +1
-    viral_genes == 0 and host_genes >= 1: keep_score = 0
-    If 3*viral_genes <= host_genes: keep_score = 0
-    If length > 50,000 and hallmark == 0: keep_score = 0
 
@@
     keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
   }
@@
     keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
+    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for RNA viruses
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 # consider accounting for RNA viruses
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
@@
 ```{r}
-write_tsv(accuracy_scores, "20220927_accuracy_scores.tsv")
+write_tsv(accuracy_scores, "20221029_accuracy_scores.tsv")
 ```
@@
 ```
+0.887
 
@@
 recover almost all of the viruses this way, but more protist contamination
+
+0.960
 

### 2022-10-30 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
New testing set visualizations.

@@
   ) +
-  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
   ylab("Num Viruses Predicted")
@@
   ) +
-  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
   ylab("Proportion of Contigs >3kb Viral")
+```
+
+only plotting the average
+
+```{r}
+num_viruses$numtools <- as.numeric(num_viruses$numtools)
+
+median_prop_vir_by_tool <- num_viruses %>% 
+  select(toolcombo, prop_vir, numtools) %>%
+  group_by(toolcombo) %>%
+  summarize(median=median(prop_vir),
+         #sd=sd(prop_vir),
+         numtools=mean(numtools))
+
+median_prop_vir_by_tool$toolcombo <- factor(median_prop_vir_by_tool$toolcombo, 
+                                          levels = median_prop_vir_by_tool$toolcombo[order(median_prop_vir_by_tool$median, decreasing = F)])
+
+median_prop_vir_by_tool$numtools <- as.factor(median_prop_vir_by_tool$numtools)
+```
+
+```{r}
+ggplot(median_prop_vir_by_tool, aes(y=median, x=toolcombo,
+                                  color=numtools, fill=numtools, shape=numtools))  +
+  geom_hline(yintercept=0, color="grey") +
+  geom_point() +
+#  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
+#                 position=position_dodge(.9)) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral")
+```
+```{r}
+ggplot(median_prop_vir_by_tool, aes(x=numtools, y=median, 
+                                  color=numtools, fill=numtools)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Proportion Viruses (%)") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
 ```
@@
                                                             "Spannenburg"),]
+
+num_viruses_uknl$residual <- "residual"
+num_viruses_uknl$residual[num_viruses_uknl$assembly=="Amsterdam_Leid" |
+                          num_viruses_uknl$assembly=="Amsterdam_Wee" |
+                          num_viruses_uknl$assembly=="Nieuwkoop_DHB" |
+                          num_viruses_uknl$assembly=="South_Holland_Dla" |
+                          num_viruses_uknl$assembly=="Spannenburg"
+                          ] <- "no residual"
 ```
@@
 ```{r}
-ggplot(num_viruses_uknl, aes(x=toolcombo, y=prop_vir, 
-                                  shape=numtools, color=assembly, fill=assembly)) +
-  geom_point() +
+num_viruses_uknl$numtools <- as.numeric(num_viruses_uknl$numtools)
+
+mean_prop_vir_by_tool <- num_viruses_uknl %>% 
+  select(toolcombo, prop_vir, numtools, residual) %>%
+  group_by(toolcombo, residual) %>%
+  summarize(prop_vir=mean(prop_vir),
+         #sd_per_vir=sd(prop_vir),
+         numtools=mean(numtools))
+
+mean_uknl <- mean_prop_vir_by_tool %>% 
+  group_by(toolcombo) %>%
+  summarize(mean=mean(prop_vir))
+
+mean_prop_vir_by_tool$toolcombo <- factor(mean_prop_vir_by_tool$toolcombo, 
+                                          levels = mean_uknl$toolcombo[order(mean_uknl$mean, decreasing = F)])
+
+mean_prop_vir_by_tool$numtools <- as.factor(mean_prop_vir_by_tool$numtools)
+```
+
+```{r}
+num_viruses_uknl$toolcombo <- factor(num_viruses_uknl$toolcombo, 
+                                          levels = levels(mean_prop_vir_by_tool$toolcombo))
+num_viruses_uknl$numtools <- as.factor(num_viruses_uknl$numtools)
+
+```
+
+```{r}
+ggplot(num_viruses_uknl) +
+  geom_hline(yintercept=0, color="grey") +
+  geom_point(aes(color=numtools, fill=numtools, y=prop_vir, x=toolcombo, shape=residual), alpha=0.5) +
+  geom_point(data=mean_prop_vir_by_tool, aes(x=toolcombo, 
+                                               y=prop_vir,
+                                             shape=residual),
+             color="black") +
   theme_light() +
@@
   ) +
-  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
-  ylab("Proportion of Contigs >3kb Viral")
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
+```
+```{r}
+t.test(num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="residual"],
+       num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="no residual"]
+       )
 ```

### 2022-10-30 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New testing set visualizations.

@@
 ```{r}
+accuracy_scores_melt <- accuracy_scores %>% 
+  select(testing_set_index, precision, recall, MCC, numtools, toolcombo) %>%
+  pivot_longer(cols=c(precision, recall, MCC), 
+               names_to="performance_metric",
+               values_to="performance_metric_score")
+```
+
+```{r}
+ggplot(accuracy_scores_melt, aes(x=numtools, y=performance_metric_score, 
+                                  color=numtools, fill=numtools)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+```
+comparing metric with and without tuning rules
+```{r}
+accuracy_scores_melt$tuning_inc <- "no"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1] <- "tv"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "tnv"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1 &
+                                  substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "both"
+```
+
+```{r}
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+  geom_boxplot() +
+  geom_point(aes(color=numtools, fill=numtools), alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score,
+                                 color=numtools, fill=numtools)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+  geom_boxplot() +
+  geom_boxplot(aes(color=numtools, fill=numtools)) +
+ # geom_point(aes(color=numtools, fill=numtools), alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+```
+
+```{r}
 write_tsv(accuracy_scores, "20221029_accuracy_scores.tsv")
@@
 
-```{r}
-viruses$size_class <- "3-10kb"
-viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
-```
-
 visualizing confusion matrix by taxa
@@
 ```
-percent viral
-
-visualizing confusion matrix by taxa
+
+differences based on genome size
+```{r}
+viruses$size_class <- "3-5kb"
+viruses$size_class[viruses$checkv_length>5000] <- "5-10kb"
+viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
+```
+
 ```{r}
@@
             sd=sd(per_viral))
+
+confusion_vir_called$size <- factor(confusion_vir_called$size,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
 ```
@@
                      values = alpha(rev(pal(6)), 1)) +
-  xlab("Length (bp)") +
+  xlab("Length") +
   ylab("Sequences Called Viral (%)") 
@@
 
+```
+
+## clustering
+```{r}
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
+                          num_viruses=rep(0, nrow(combos_list)))
+
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning_viral = combos_list$tune_viral[i],
+                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+
+  if (max(viral_scores[,i])<=0) {
+    num_viruses$num_viruses[i] <- 0
+  }
+  else {
+    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+  }
+
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+
+  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
+}
+
+num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
+```
+
+```{r}
+viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
+viral_scores_nozeros <- viral_scores_nozeros + 1
+viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
+
+colnames(viral_scores_nozeros) <- num_viruses$toolcombo
+```
+
+```{r}
+library(phyloseq)
+```
+
+```{r}
+tooldata <- num_viruses
+
+rownames(tooldata) <- tooldata$toolcombo
+```
+
+```{r}
+physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
+                                     sample_data(tooldata))
+```
+
+```{r}
+ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw() +
+  geom_label(label=tooldata$toolcombo)
+
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", color="num_viruses") + 
+  geom_point(size = 3) +
+  theme_bw()
+```
+
+```{r}
+bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
+clusters <- hclust(dist(bray_dist))
+plot(clusters)
+
+myclusters <- cutree(clusters, h=0.8)
+```
+
+```{r}
+#names(myclusters[myclusters==1])
+#names(myclusters[myclusters==2])
+#names(myclusters[myclusters==3])
+#names(myclusters[myclusters==4])
+#names(myclusters[myclusters==5])
+
+myclusters_df <- tibble(combo=names(myclusters),
+                            cluster_index=myclusters)
+
+myclusters_df <- separate(myclusters_df, col=combo, into=c("tnv", "DVF",
+                                                            "tv", "VB",
+                                                            "VS", "VS2"),
+                          sep=" ", remove = F)
+
+tool_count <- as.data.frame(rbind(table(myclusters_df$tnv, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$tv, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VB, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VS, myclusters_df$cluster_index)[2,],
+                          table(myclusters_df$VS2, myclusters_df$cluster_index)[2,])
+                    )
+
+tool_count <- data.frame(t(apply(tool_count, c(1), function(x) {x <- x/table(myclusters_df$cluster_index)})))
+
+tool_count$method <- c("tnv", "DVF", "tv", "VB", "VS", "VS2")
+
+tool_count <- melt(tool_count)
+
+colnames(tool_count) <- c("tool", "cluster_index", "tool_count_norm")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+ggplot(tool_count, aes(x=tool, y=tool_count_norm,
+                   fill=tool,
+                   color=tool)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "none",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    #legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  xlab("Tool") +
+  ylab("Proportion of Times in Cluster") + 
+  facet_wrap(~cluster_index, nrow=1)
+```
+
+```{r}
+accuracy_scores_melt <- accuracy_scores %>% 
+  select(precision, recall, MCC, numtools, toolcombo) %>%
+  group_by(numtools, toolcombo) %>%
+  summarise(precision=mean(precision),
+            recall=mean(recall),
+            MCC=mean(MCC)) %>%
+  pivot_longer(cols=c(precision, recall, MCC), 
+               names_to="performance_metric",
+               values_to="performance_metric_score")
+```
+
+```{r}
+myclusters_df <- inner_join(accuracy_scores_melt, myclusters_df, 
+                            by=c("toolcombo"="combo"))
+
+myclusters_df$cluster_index <- as.factor(myclusters_df$cluster_index)
+```
+
+```{r}
+ggplot(myclusters_df, aes(x=cluster_index, y=performance_metric_score, 
+                                  color=cluster_index, fill=cluster_index)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Cluster") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(9)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(9)), 1)) +
+  facet_wrap(~performance_metric)
 ```

### 2022-11-01 EnvironmentalData/DrinkingWaterMetagenomes/Diversity_Metrics/Snakefile-overarching-slurm.sh
Scripts for read mapping and diversity comparisons.

@@ -1,0 +1,55 @@
+#!/bin/bash
+
+###############################
+#                             #
+#  1) Job Submission Options  #
+#                             #
+###############################
+
+# Name
+#SBATCH --job-name=overarching
+
+# Resources
+# For MPI, increase ntasks-per-node
+# For multithreading, increase cpus-per-task
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --cpus-per-task=1
+#SBATCH --mem-per-cpu=300mb
+#SBATCH --time=04-00:00:00
+
+# Account
+#SBATCH --account=kwigg1
+#SBATCH --partition=standard
+
+# Logs
+#SBATCH --mail-user=hegartyb@umich.edu
+#SBATCH --mail-type=BEGIN,END,FAIL
+#SBATCH --output=Logs/%x-%j.out
+
+# Environment
+##SBATCH --export=ALL
+
+source /etc/profile.d/http_proxy.sh
+
+# List compute nodes allocated to the job
+if [[ $SLURM_JOB_NODELIST ]] ; then
+    echo "Running on"
+    scontrol show hostnames $SLURM_JOB_NODELIST
+    echo -e "\n"
+fi
+
+
+
+#####################
+#                   #
+#  2) Job Commands  #
+#                   #
+#####################
+
+# Initiating snakemake and running workflow in cluster mode
+snakemake --profile /scratch/kwigg_root/kwigg/hegartyb/SnakemakeAssemblies3000/CompetitiveMapping/Config --latency-wait 60 --use-conda --conda-prefix /home/hegartyb/miniconda3/envs/ --snakefile Snakefile-overarching --keep-going --rerun-incomplete
+
+
+
+

### 2022-11-01 EnvironmentalData/DrinkingWaterMetagenomes/Diversity_Metrics/get_trimmed_viral_contigs.py

@@ -1,0 +1,70 @@
+import Bio
+from Bio import SeqIO
+
+with open(snakemake.output["fasta"], 'wt') as outfile:
+    with open(snakemake.input["clist"], 'rt') as viralcontigslist:
+        contigs = []
+        i = 0
+        j = 0
+        for line in viralcontigslist:
+            if (i==0):
+                i += 1
+            else:
+                i+=1
+                contig = line.split('\t')[int(snakemake.params["contig_column"])]
+                contigs.append(contig)
+                j+=1
+                #print("here")
+        print(i)
+        print("contigs count: %s" % j)
+
+        i = 0
+
+        #virsorter2
+        filename = snakemake.input["virsorter2_fasta"]
+        for seqrecord in SeqIO.parse(filename,"fasta"):
+            comp = seqrecord.id
+            comp2 = comp.replace(".", "_")
+            comp = comp2.split("|")[0]
+            if (comp in contigs):
+                i+=1
+                contigs.remove(comp)
+                outfile.write(">A2--%s\n%s\n" % (comp2, seqrecord.seq))
+        print("matches after virsorter2count: %s" % i)
+
+        #checkV
+        filename = snakemake.input["checkv_fasta"]
+        for seqrecord in SeqIO.parse(filename,"fasta"):
+            comp = seqrecord.id
+            comp2 = comp.replace(".", "_")
+            comp = comp2.split(" ")[0].rsplit("_",1)[0]
+            if (comp in contigs):
+                i+=1
+                contigs.remove(comp)
+                outfile.write(">A2--%s\n%s\n" % (comp2, seqrecord.seq))
+        print("matches from checkv count: %s" % i)
+
+        #vibrant
+        filename = snakemake.input["vibrant_fasta"]
+        for seqrecord in SeqIO.parse(filename,"fasta"):
+            comp = seqrecord.id
+            comp = comp.replace(".", "_")
+            if (comp in contigs):
+                i+=1
+                contigs.remove(comp)
+                outfile.write(">A2--%s\n%s\n" % (comp, seqrecord.seq))
+        print("matches after vibrant count: %s" % i)
+
+        #free viruses
+        filename = snakemake.input["free_fasta"]
+        for seqrecord in SeqIO.parse(filename,"fasta"):
+            comp = seqrecord.id
+            comp = comp.replace(".", "_")
+            if (comp in contigs):
+                i+=1
+                contigs.remove(comp)
[…]

### 2022-11-01 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
New visualizations of testing data.

@@
   select(toolcombo, assembly, num_viruses) %>%
-  spread(assembly, num_viruses)
+  pivot_wider(assembly, num_viruses)
 rownames(num_viruses_uknl_wide) <- num_viruses_uknl_wide$toolcombo
@@
 ```
+
+Comparing three recommendations
+
+```{r}
+viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+table(viruses$keep_score_all>=1)/nrow(viruses)
+table(viruses$keep_score_high_MCC>=1)/nrow(viruses)
+table(viruses$keep_score_few_tools>=1)/nrow(viruses)
+
+```

### 2022-11-01 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
New visualizations of testing data.

--- 2022-10-29 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "VSTE - Comparing Methods Visualization"
 author: Bridget Hegarty, James Riddell
@@
 library(data.table)
-library(pROC)
+library("stringr")
 ```
@@
 
-# Assessing performance against the "truth"
-note that this is only as accurate as the annotations of the input sequences
+# Comparing all of the tool methods
 
-this function calculates the precision, recall, and F1 score for each pipeline
-```{r}
-assess_performance <- function(seqtype, keep_score) {
-
-  truepositive <- rep("not viral", length(seqtype))
-  truepositive[seqtype=="virus"] <- "viral"
-
-  #make confusion matrix
-  confusion_matrix <- rep("true negative", length(keep_score))
-  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
-  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
-  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"
-
-  TP <- table(confusion_matrix)[4]
-  FP <- table(confusion_matrix)[2]
-  TN <- table(confusion_matrix)[3]
-  FN <- table(confusion_matrix)[1]
-
-  precision <- TP/(TP+FP)
-  recall <- TP/(TP+FN)
-  F1 <- 2*precision*recall/(precision+recall)
-
-  MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
-
-  auc <- round(auc(truepositive, keep_score),4)
-
-  #by type metrics
-  fungal_FP <- table(confusion_matrix[seqtype=="fungi"])[2]
-  protist_FP <- table(confusion_matrix[seqtype=="protist"])[2]
-  bacterial_FP <- table(confusion_matrix[seqtype=="bacteria"])[2]
-  viral_FN <- table(confusion_matrix[seqtype=="virus"])[1]
-
-  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, 
-                   protist_FP, bacterial_FP, viral_FN)
-  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
-                          "protist_FP", "bacterial_FP", "viral_FN")
-
-  return(performance)
-}
-```
-
-combination of tools list
 ```{r}
@@
             combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
-                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
+            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
                                                if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
@@
 
-this function builds a list of all of the combinations that the user wants to 
-test. 
-In this case, we're comparing the performance of all unique combinations of the 
-six tools.
-```{r}
-build_score_list <- function(input_seqs, combos) {
-  output <- data.frame(precision=rep(0, nrow(combos)),
-                       recall=rep(0, nrow(combos)),
-                       F1=rep(0, nrow(combos)),
-                       MCC=rep(0, nrow(combos)),
-                       AUC=rep(0, nrow(combos)),
-                       fungal_FP=rep(0, nrow(combos)),
-                       protist_FP=rep(0, nrow(combos)),
-                       bacterial_FP=rep(0, nrow(combos)),
-                       viral_FN=rep(0, nrow(combos)))
-  for (i in 1:nrow(combos)) {
-    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
-                                            include_virsorter = combos$VS[i],
-                                            include_virsorter2 = combos$VS2[i],
-                                            include_tuning_viral = combos$tune_viral[i],
-                                            include_tuning_not_viral = combos$tune_not_viral[i],
-                                            include_deepvirfinder = combos$DVF[i])
-
-    output[i,1:9] <- assess_performance(input_seqs$seqtype, keep_score)
-
-    output$toolcombo[i] <- paste(combos$tune_viral[i],combos$DVF[i],
-                                 combos$tune_not_viral[i], combos$VIBRANT[i],
-                                 combos$VS[i], combos$VS2[i])
-  }
-
-  output[is.na(output)] <- 0
-
-  return (output)
-}
-```
-
-## Calculate the performance of each pipeline
-```{r}
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
-                      precision=rep(0, nrow(combos_list)*10),
-                       recall=rep(0, nrow(combos_list)*10),
-                       F1=rep(0, nrow(combos_list)*10),
-                       MCC=rep(0, nrow(combos_list)*10), 
-                      AUC=rep(0, nrow(combos_list)*10),
-                      fungal_FP=rep(0, nrow(combos_list)*10),
-                      protist_FP=rep(0, nrow(combos_list)*10),
-                      bacterial_FP=rep(0, nrow(combos_list)*10),
-                      viral_FN=rep(0, nrow(combos_list)*10))
-
-accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==1,], combos_list))
-for (i in 2:10) {
-  accuracy_scores <- rbind(accuracy_scores,
-                           cbind(testing_set_index=rep(i, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==i,], combos_list)))
-}
-```
-
-```{r}
-library("stringr")
-```
-
-```{r}
-accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
-#accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
-accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
-accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
-accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
-```
-
-## Visualize how the precision, recall, and F1 scores change across pipelines.
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("F1 Score")
-p2
-ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Precision")
-ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Recall")
-ggplot(accuracy_scores, aes(x=precision, y=recall, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Precision") +
-  ylab("Recall")
-ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Precision-Recall")
-ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("MCC")
-ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("AUC")
-ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Fungal False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=protist_FP, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Protist False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Bacterial False Positives")
-ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Viral False Negatives")
-```
-
-```{r}
-write_tsv(accuracy_scores, "20221029_accuracy_scores.tsv")
-```
-
-to do: add in clustering and ordination like in the drinking water R notebook
-
-# Experimenting
-
-## high precision example
-```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses$confusion_matrix_high_precision <- "true negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<1] <- "false negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision>=1] <- "true positive"
-viruses$confusion_matrix_high_precision[viruses$seqtype!="virus" & viruses$keep_score_high_precision>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_precision, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-length(grep("true", viruses$confusion_matrix_high_precision))/nrow(viruses)
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-this rule set had the highest precision, but as you can see, this comes with a big sacrifice in recall
-
-## high MCC example
-```{r}
-viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_high_MCC <- "true negative"
-viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
-viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
-viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$confusion_matrix_high_MCC))/nrow(viruses)
-```
-
-recall
-```{r}
-length(grep("true positive", viruses$confusion_matrix_high_MCC))/length(grep("virus", viruses$seqtype))
-```
-
-```{r}
-viruses$size_class <- "3-10kb"
-viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, Index)
-
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-percent viral
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, size_class, Index)
-
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "index", "count")
-```
-
-```{r}
-confusion_vir_called <- confusion_by_taxa %>% filter(confusion_matrix=="true positive" | confusion_matrix=="false positive")
-
-type_count <- viruses %>% count(seqtype, size_class, Index)
-
-confusion_vir_called$per_viral <- 0
-
-for (i in c(1:nrow(confusion_vir_called))) {
-  confusion_vir_called$per_viral[i] <- confusion_vir_called$count[i]/type_count$n[type_count$seqtype==confusion_vir_called$seqtype[i] & 
-                                                                                    type_count$Index==confusion_vir_called$index[i] &
-                                                                                    type_count$size_class==confusion_vir_called$size[i]]*100
-}
-
-confusion_vir_called <- confusion_vir_called %>% group_by(seqtype, size) %>%
-  summarise(mean=mean(per_viral), 
-            sd=sd(per_viral))
-```
-
-```{r}
-ggplot(confusion_vir_called, aes(y=mean, x=size,
-                   fill=seqtype,
-                   color=seqtype)) +
-  geom_bar(stat="identity", position=position_dodge()) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
-                 position=position_dodge(.9)) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Length (bp)") +
-  ylab("Sequences Called Viral (%)") 
-```
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-```
-
-# Considering how each method contributes to the final prediction
-```{r}
-viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
-viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
-                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
-                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "viral_score")
-```
-
-```{r}
-sm_m <- sm_m[sm_m$viral_score>0,]
-
-sm_m$score <- sm_m$viral_score
-
-sm_m$score[sm_m$viral_score==0.5] <- "0.5"
-sm_m$score[sm_m$viral_score>=1] <- "1"
-sm_m$score[sm_m$viral_score>=2] <- "2"
-sm_m$score[sm_m$viral_score>=3] <- "3"
-sm_m$score[sm_m$viral_score>=4] <- "4"
-sm_m$score[sm_m$viral_score>=5] <- "5"
-
-sm_m$score <- factor(sm_m$score, 
-                                       levels=c("0.5", "1", "2","3","4","5"))
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  xlab("") +
-  ylab("Number of Sequences") +
-  coord_flip()
-```
-## Visualizing confusion matrix by number of tools
-
-```{r}
-viruses$keep_score_visualize <- viruses$keep_score_high_MCC
-viruses$keep_score_visualize[viruses$keep_score_high_MCC>1] <- "> 1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==1] <- "1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==0.5] <- "0.5"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==0] <- "0"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==-0.5] <- "-0.5"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==-1] <- "-1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC<=-1] <- "< -1"
-
-viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
-#viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-#                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
-```
-
-```{r}
-levels(factor(viruses$keep_score_visualize))
-```
-
-```{r}
-ggplot(viruses, aes(x=as.factor(Index),
-                   fill=keep_score_visualize, color=keep_score_visualize)) +
-  geom_bar(stat="count", position="stack") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16)
-  ) +
-  scale_color_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 0.5)) +
-  xlab("Index") +
-  ylab("Sequence Count") +
-  facet_wrap(~confusion_matrix_high_MCC, scales = "free")
-
-```
-
-## all 6 tools example
-```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_all <- "true negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
-viruses$confusion_matrix_all[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_all, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-table(viruses$confusion_matrix_all)
-
-length(grep("true", viruses$confusion_matrix_all))/nrow(viruses)
-```
-
-```{r}
-length(grep("true positive", viruses$confusion_matrix_all))/length(grep("virus", viruses$seqtype))
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-## high recall example
-```{r}
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_high_recall <- "true negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall<1] <- "false negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
-viruses$confusion_matrix_high_recall[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_recall, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-p2
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$confusion_matrix_high_recall))/nrow(viruses)
-```
-0.887
-
-recall
-```{r}
-length(grep("true positive", viruses$confusion_matrix_high_recall))/length(grep("virus", viruses$seqtype))
-```
-recover almost all of the viruses this way, but more protist contamination
-
-0.960
-
-######################################################################
-Extra Stuff
-#####################################################################
-
-```{r}
-ggplot(viruses, aes(x=checkv_length, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_MCC,
-                   color=confusion_matrix_high_MCC)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=hallmark,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Number of Hallmark Genes") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=confusion_matrix_high_recall, y=checkv_length,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") +
-  scale_y_log10()
-```
-
-looking at false negatives
-```{r}
-viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<1),]
-```
-
-looking at protists calling viral
-```{r}
-viruses_false_pos_protists <- viruses[(viruses$seqtype=="protist" & viruses$keep_score_high_recall>=1),]
-```
-
-# Considering how each method contributes to the final prediction (high MCC)
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses_high <- viruses[viruses$keep_score_vb_tv>=1,] #uncomment this line if want to use all 6 tools
-viruses_high_mod <- viruses_high %>% select(keep_score_vb, 
-                                            keep_score_vb_tv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "score")
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=as.factor(score))) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(14)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs") +
-  coord_flip()
-```
-
-# ROC 
-```{r}
-library(pROC)
-```
-
-```{r}
-viruses$truepositive <- rep(0, nrow(viruses))
-viruses$truepositive[viruses$seqtype=="virus"] <- 1
-```
-
-```{r}
-rocobj <- roc(viruses$truepositive, viruses$keep_score)
-rocobj_all <- roc(viruses$truepositive, viruses$keep_score_all)
-auc <- round(auc(viruses$truepositive, viruses$keep_score),4)
-auc_all <- round(auc(viruses$truepositive, viruses$keep_score_all),4)
-#create ROC plot
-ggroc(rocobj, colour = 'steelblue', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')')) +
-  coord_equal()
-ggroc(rocobj_all, colour = 'green', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc_all, ')'))
-```
-Sensitivity: The probability that the model predicts a positive outcome for an observation when indeed the outcome is positive.
-Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative.
-
-# Comparing behavior of all testing sets combined (clustering analyses)
-
 ```{r}
@@
 ```{r}
-ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-
-ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
+dim(viral_scores)
 ```
 
+##matrix of number in agreement between tools
+### number predicted together
 ```{r}
-ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
-  geom_boxplot(aes(color=numtools)) +
-  geom_point(aes(color=numtools, fill=numtools)) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Tools") +
-  ylab("Num Viruses Predicted")
-```
+num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
 
-```{r}
-viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
-viral_scores_nozeros <- viral_scores_nozeros + 1
-viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
+for (i in 1:63) {
+  for (j in 1:63) {
+    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+  }
+}
 
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
-```
+rownames(num_high_viruses) <- combos_list$toolcombo
+colnames(num_high_viruses) <- combos_list$toolcombo
 
-```{r}
-library(phyloseq)
-```
+prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
 
-```{r}
-tooldata <- num_viruses
+for (i in 1:63) {
+  for (j in 1:63) {
+    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
+  }
+}
 
-rownames(tooldata) <- tooldata$toolcombo2
-```
-
-```{r}
-physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
-                                     sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  geom_label(label=tooldata$toolcombo)
-
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw()
-```
-to do: try coloring above based on the F1 scores of the testing set on each combination
-
-```{r}
-bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
-clusters <- hclust(dist(bray_dist))
-plot(clusters)
-
-myclusters <- cutree(clusters, h=1.1)
-```
-
-```{r}
-names(myclusters[myclusters==1])
-names(myclusters[myclusters==2])
-names(myclusters[myclusters==3])
-names(myclusters[myclusters==4])
-names(myclusters[myclusters==5])
-
-myclusters_df <- tibble(combo=names(myclusters),
-                            cluster_index=myclusters)
-
-myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
-                                                            "Kaiju", "VIBRANT",
-                                                            "VirSorter", "VirSorter2"),
-                          sep=" ", remove = F)
-
-tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
-                    )
-
-tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
-
-tool_count <- melt(tool_count)
-
-colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
+rownames(prop_high_viruses) <- combos_list$toolcombo
+colnames(prop_high_viruses) <- combos_list$toolcombo
 ```
@@
 
-ggplot(tool_count, aes(x=cluster_index, y=tool_count,
-                   fill=cluster_index,
-                   color=cluster_index)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Cluster") +
-  ylab("Number of Times in Cluster") + 
-  facet_wrap(~tool, scales = "free")
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2,
+    num_vir = colSums(viral_scores>=1)+1
+)
+
+ann_colors = list(
+  num_vir=magma(100),
+  tnv = c("black", pal(6)[1]),
+  tv = c("black", pal(6)[3]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
+
+rownames(ann_col) <- combos_list$toolcombo
+
+pheatmap::pheatmap(num_high_viruses,
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = "black",
+                  color = magma(100),
+                  cutree_rows = 4,
+                  cutree_cols = 4,
+                  show_rownames = F,
+                  filename = "../IntermediaryFiles/pheatmap_num_viruses_in_common.png",
+                  annotation_legend = F)
 ```
@@
 ```{r}
- ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Viral Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
 
- ggplot(viruses, aes(x=percent_viral, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Percent Genes Viral") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2,
+    num_vir = colSums(viral_scores>=1)+1
+)
 
- ggplot(viruses, aes(x=hallmark, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
+ann_colors = list(
+  num_vir=magma(100),
+  tnv = c("black", pal(6)[1]),
+  tv = c("black", pal(6)[3]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
 
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
+rownames(ann_col) <- combos_list$toolcombo
+
+pheatmap::pheatmap(prop_high_viruses,
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = "black",
+                  color = viridis(100),
+                  cutree_rows = 4,
+                  cutree_cols = 4,
+                  show_rownames = F,
+                  filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common.png",
+                  annotation_legend = F)
 ```
@@
 ```{r}
-viruses_false_positive <- viruses[viruses$confusion_matrix_high_precision=="false positive",]
-viruses_false_negative <- viruses[viruses$confusion_matrix_high_precision=="false negative",]
+prop_high_viruses_long <- as.data.frame(prop_high_viruses)
+prop_high_viruses_long$tool <- combos_list$toolcombo
+
+prop_high_viruses_long <- as.data.frame(prop_high_viruses_long) %>% 
+  pivot_longer(!tool,
+               values_to="proportion",
+               names_to="tool2")
+
+prop_high_viruses_long$binned_prop <- "0-10%"
+prop_high_viruses_long$binned_prop[prop_high_viruses_long$proportion>=0.10] <- "10-50%"
+prop_high_viruses_long$binned_prop[prop_high_viruses_long$proportion>=0.50] <- "50-90%"
+prop_high_viruses_long$binned_prop[prop_high_viruses_long$proportion>=0.90] <- "90-100%"
+
 ```
@@
 ```{r}
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=checkv_length,
-                   color=checkv_length,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="bacteria"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="fungi"], aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="protist"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
+table(prop_high_viruses_long$binned_prop)
 ```
@@
 ```{r}
-table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)
+prop_high_viruses_long$toolcombo <- paste(prop_high_viruses_long$tool,
+                                          prop_high_viruses_long$tool2,
+                                          sep="-")
+prop_high_viruses_long$toolcombo <- gsub(" ", "", prop_high_viruses_long$toolcombo)
+prop_high_viruses_long$toolcombo <- unique(prop_high_viruses_long$toolcombo)
+```
 
-table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
+```{r}
+head(prop_high_viruses_long)
 ```
+
+```{r}
+hist(prop_high_viruses_long$proportion,
+     )
+```
+
+# numbers in common between two tool combos
+```{r}
+combos_list_two <- combos_list
+combos_list_two$numtools <- str_count(combos_list_two$toolcombo, "1")
+combos_list_two <- combos_list_two[combos_list_two$numtools==1,]
+
+two_tool_pairs <- num_high_viruses[rownames(num_high_viruses) %in% combos_list_two$toolcombo,
+                                   colnames(num_high_viruses) %in% combos_list_two$toolcombo]
+
+colnames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+rownames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+```
+
+```{r}
+viral_scores_two <- viral_scores[,colnames(num_high_viruses) %in% combos_list_two$toolcombo]
+
+n <- 6
+viral_scores_two_comb <- matrix(data=NA, nrow=n, ncol=n)
+
+for (i in 1:n) {
+  for (j in 1:n) {
+    if (i==j) {
+      viral_scores_two_comb[i,j] <- sum(viral_scores_two[,i]>=1)
+    }
+    else if (i<j) {
+      viral_scores_two_comb[i,j] <- sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)
+    }
+    else if (i>j) {
+      viral_scores_two_comb[i,j] <- paste(round(sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)/sum(viral_scores_two[,i]>=1 | viral_scores_two[,j]>=1)*100, digits=0), "%", sep="")
+    }
+  }
+}
+
+colnames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+rownames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+```

### 2022-11-01 TestingPipeline/Scripts/examining_false_positives.Rmd
New visualizations of testing data.

--- 2022-10-29 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "Examining False Negatives"
 author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+date: 10-30-2022
 output: html_notebook
 ---
-This Rmarkdown file assesses the output of CheckV, DeepVirFinder, Kaiju,
+This Rmarkdown file examines the false positives from the output of CheckV, DeepVirFinder, Kaiju,
 VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
@@
 
-# Assessing performance against the "truth"
-note that this is only as accurate as the annotations of the input sequences
-
-this function calculates the precision, recall, and F1 score for each pipeline
-```{r}
-assess_performance <- function(seqtype, keep_score) {
-
-  truepositive <- rep("not viral", length(seqtype))
-  truepositive[seqtype=="virus"] <- "viral"
-
-  #make confusion matrix
-  confusion_matrix <- rep("true negative", length(keep_score))
-  confusion_matrix[truepositive=="viral" & keep_score<=1] <- "false negative"
-  confusion_matrix[truepositive=="viral" & keep_score>=1] <- "true positive"
-  confusion_matrix[truepositive=="not viral" & keep_score>=1] <- "false positive"
-
-  TP <- table(confusion_matrix)[4]
-  FP <- table(confusion_matrix)[2]
-  TN <- table(confusion_matrix)[3]
-  FN <- table(confusion_matrix)[1]
-
-  precision <- TP/(TP+FP)
-  recall <- TP/(TP+FN)
-  F1 <- 2*precision*recall/(precision+recall)
-
-  MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
-
-  auc <- round(auc(truepositive, keep_score),4)
-
-  #by type metrics
-  fungal_FP <- table(confusion_matrix[seqtype=="fungi"])[2]
-  protist_FP <- table(confusion_matrix[seqtype=="protist"])[2]
-  bacterial_FP <- table(confusion_matrix[seqtype=="bacteria"])[2]
-  viral_FN <- table(confusion_matrix[seqtype=="virus"])[1]
-
-  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, 
-                   protist_FP, bacterial_FP, viral_FN)
-  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
-                          "protist_FP", "bacterial_FP", "viral_FN")
-
-  return(performance)
-}
-```
-
-combination of tools list
-```{r}
-combos_list <- data.frame(toolcombo=rep(0, 64),
-                          tune_not_viral=rep(0, 64),
-                          DVF=rep(0, 64),
-                          tune_viral=rep(0, 64),
-                          VIBRANT=rep(0, 64),
-                          VS=rep(0, 64),
-                          VS2=rep(0, 64))
-p <- 1
-
-for (i in c(0,1)){
-  for (j in c(0,1)){
-    for (k in c(0,1)){
-      for (l in c(0,1)){
-        for (m in c(0,1)){
-          for (n in c(0,1)){
-            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
-                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
-                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
-            combos_list$tune_not_viral[p] <- i
-            combos_list$DVF[p] <- j
-            combos_list$tune_viral[p] <- k
-            combos_list$VIBRANT[p] <- l
-            combos_list$VS[p] <- m
-            combos_list$VS2[p] <- n
-            p <- p+1
-          }
-        }
-      }
-    }
-  }
-}
-
-combos_list <- combos_list[-1,]
-```
-
-this function builds a list of all of the combinations that the user wants to 
-test. 
-In this case, we're comparing the performance of all unique combinations of the 
-six tools.
-```{r}
-build_score_list <- function(input_seqs, combos) {
-  output <- data.frame(precision=rep(0, nrow(combos)),
-                       recall=rep(0, nrow(combos)),
-                       F1=rep(0, nrow(combos)),
-                       MCC=rep(0, nrow(combos)),
-                       AUC=rep(0, nrow(combos)),
-                       fungal_FP=rep(0, nrow(combos)),
-                       protist_FP=rep(0, nrow(combos)),
-                       bacterial_FP=rep(0, nrow(combos)),
-                       viral_FN=rep(0, nrow(combos)))
-  for (i in 1:nrow(combos)) {
-    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
-                                            include_virsorter = combos$VS[i],
-                                            include_virsorter2 = combos$VS2[i],
-                                            include_tuning_viral = combos$tune_viral[i],
-                                            include_tuning_not_viral = combos$tune_not_viral[i],
-                                            include_deepvirfinder = combos$DVF[i])
-
-    output[i,1:9] <- assess_performance(input_seqs$seqtype, keep_score)
-
-    output$toolcombo[i] <- paste(combos$tune_viral[i],combos$DVF[i],
-                                 combos$tune_not_viral[i], combos$VIBRANT[i],
-                                 combos$VS[i], combos$VS2[i])
-  }
-
-  output[is.na(output)] <- 0
-
-  return (output)
-}
-```
-
-## Calculate the performance of each pipeline
-```{r}
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
-                      precision=rep(0, nrow(combos_list)*10),
-                       recall=rep(0, nrow(combos_list)*10),
-                       F1=rep(0, nrow(combos_list)*10),
-                       MCC=rep(0, nrow(combos_list)*10), 
-                      AUC=rep(0, nrow(combos_list)*10),
-                      fungal_FP=rep(0, nrow(combos_list)*10),
-                      protist_FP=rep(0, nrow(combos_list)*10),
-                      bacterial_FP=rep(0, nrow(combos_list)*10),
-                      viral_FN=rep(0, nrow(combos_list)*10))
-
-accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==1,], combos_list))
-for (i in 2:10) {
-  accuracy_scores <- rbind(accuracy_scores,
-                           cbind(testing_set_index=rep(i, nrow(combos_list)),
-                              build_score_list(viruses[viruses$Index==i,], combos_list)))
-}
-```
-
-```{r}
-library("stringr")
-```
-
-```{r}
-accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
-#accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
-accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
-accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
-accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
-```
-
-## Visualize how the precision, recall, and F1 scores change across pipelines.
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("F1 Score")
-p2
-ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Precision")
-ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Recall")
-ggplot(accuracy_scores, aes(x=precision, y=recall, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Precision") +
-  ylab("Recall")
-ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Precision-Recall")
-ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("MCC")
-ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("AUC")
-ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Fungal False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=protist_FP, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Protist False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Bacterial False Positives")
-ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Viral False Negatives")
-```
-
-```{r}
-write_tsv(accuracy_scores, "20221029_accuracy_scores.tsv")
-```
-
-to do: add in clustering and ordination like in the drinking water R notebook
-
-# Experimenting
-
-## high precision example
-```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses$confusion_matrix_high_precision <- "true negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<1] <- "false negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision>=1] <- "true positive"
-viruses$confusion_matrix_high_precision[viruses$seqtype!="virus" & viruses$keep_score_high_precision>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_precision, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-length(grep("true", viruses$confusion_matrix_high_precision))/nrow(viruses)
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-this rule set had the highest precision, but as you can see, this comes with a big sacrifice in recall
-
-## high MCC example
 ```{r}
@@
 
-accuracy:
+getting a list of false positive bacteria and plasmids from testing set 1
 ```{r}
-length(grep("true", viruses$confusion_matrix_high_MCC))/nrow(viruses)
+fp_bacteria_list <- viruses$contig[viruses$Index==1 & 
+                                     viruses$seqtype=="bacteria" &
+                                     viruses$confusion_matrix_high_MCC=="false positive"]
+
+fp_plasmid_list <- viruses$contig[viruses$Index==1 & 
+                                     viruses$seqtype=="plasmid" &
+                                     viruses$confusion_matrix_high_MCC=="false positive"]
+
+tp_virus_list <- viruses$contig[viruses$Index==1 & 
+                                     viruses$seqtype=="virus" &
+                                     viruses$confusion_matrix_high_MCC=="true positive"]
 ```
 
-recall
+vibrant annotations
 ```{r}
-length(grep("true positive", viruses$confusion_matrix_high_MCC))/length(grep("virus", viruses$seqtype))
+annotations_vb <- read_tsv("../ToolOutput/VIBRANT_annotations_metagenomic_testing_set_1.tsv")
+```
+```{r}
+annotations_vb <- separate(annotations_vb, col="scaffold", into=c("seqtype", "contig"),
+                           sep="--")
+annotations_vb$contig <- sub("\\.", "_", annotations_vb$contig)
 ```
@@
 ```{r}
-viruses$size_class <- "3-10kb"
-viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
+head(annotations_vb)
+```
+```{r}
+table(annotations_vb$contig %in% fp_bacteria_list)
+table(annotations_vb$contig %in% fp_plasmid_list)
 ```
 
-visualizing confusion matrix by taxa
 ```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, Index)
+annotations_vb_fp <- rbind(annotations_vb[annotations_vb$contig %in% fp_bacteria_list,],
+                           annotations_vb[annotations_vb$contig %in% fp_plasmid_list,],
+                           annotations_vb[annotations_vb$contig %in% tp_virus_list,])
+```
 
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","index", "count")
+```{r}
+annotations_vb_fp_vog_count <- annotations_vb_fp %>%
+  select(contig, VOG, seqtype) %>%
+  group_by(contig, seqtype) %>%
+  summarize(num_vog=table(is.na(VOG))[[1]],
+            prop_vog=table(is.na(VOG))[[1]]/length(VOG),
+            num_orfs=length(VOG))
+```
+
+```{r}
+annotations_vb_fp_vog_count$confusion_matrix <- "false positive"
+annotations_vb_fp_vog_count$confusion_matrix[annotations_vb_fp_vog_count$seqtype=="virus"] <- "true positive"
 ```
@@
 ```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
+ggplot(annotations_vb_fp_vog_count, 
+       aes(x=seqtype, y=prop_vog, 
+           color=seqtype, fill=seqtype)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
   theme_light() +
@@
   ) +
+  ylab("Proportion Genes VOGs") +
+  xlab("Sequence Type") +
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
+                     values = alpha(rev(pal(3)), 0.5)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
+                     values = alpha(rev(pal(3)), 1)) +
+  facet_wrap(~confusion_matrix, scales="free_x")
 ```
-percent viral
-
-visualizing confusion matrix by taxa
 ```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, size_class, Index)
-
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "index", "count")
+t.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="bacteria"],
+       annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
+```
+```{r}
+t.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="plasmid"],
+       annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
 ```
@@
 ```{r}
-confusion_vir_called <- confusion_by_taxa %>% filter(confusion_matrix=="true positive" | confusion_matrix=="false positive")
-
-type_count <- viruses %>% count(seqtype, size_class, Index)
-
-confusion_vir_called$per_viral <- 0
-
-for (i in c(1:nrow(confusion_vir_called))) {
-  confusion_vir_called$per_viral[i] <- confusion_vir_called$count[i]/type_count$n[type_count$seqtype==confusion_vir_called$seqtype[i] & 
-                                                                                    type_count$Index==confusion_vir_called$index[i] &
-                                                                                    type_count$size_class==confusion_vir_called$size[i]]*100
-}
-
-confusion_vir_called <- confusion_vir_called %>% group_by(seqtype, size) %>%
-  summarise(mean=mean(per_viral), 
-            sd=sd(per_viral))
-```
-
-```{r}
-ggplot(confusion_vir_called, aes(y=mean, x=size,
-                   fill=seqtype,
-                   color=seqtype)) +
-  geom_bar(stat="identity", position=position_dodge()) +
+ggplot(annotations_vb_fp_vog_count, 
+       aes(x=num_vog, y=prop_vog, 
+           color=seqtype, fill=seqtype)) +
+  geom_point(alpha=0.5) +
   theme_light() +
@@
   ) +
-  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
-                 position=position_dodge(.9)) +
+  ylab("Proportion Genes VOGs") +
+  xlab("Number of VOGs") +
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(pal(3)), 0.5)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Length (bp)") +
-  ylab("Sequences Called Viral (%)") 
+                     values = alpha(rev(pal(3)), 1))
+```
+
+# Thinking about short protists
+```{r}
+viruses[viruses$confusion_matrix_high_MCC=="false positive" & viruses$seqtype=="protist",]
+
+table(viruses$confusion_matrix_high_MCC[viruses$seqtype=="protist"], 
+      viruses$max_score_group[viruses$seqtype=="protist"])
+
+table(viruses$Kingdom[viruses$seqtype=="protist" &  viruses$confusion_matrix_high_MCC=="false positive"], 
+      viruses$max_score_group[viruses$seqtype=="protist" & viruses$confusion_matrix_high_MCC=="false positive"])
+
+table(viruses$percent_viral[viruses$seqtype=="protist" &  viruses$confusion_matrix_high_MCC=="false positive"]>20, 
+      viruses$max_score_group[viruses$seqtype=="protist" & viruses$confusion_matrix_high_MCC=="false positive"])
 ```
@@
 ```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
+ggplot(viruses, aes(y=percent_viral, x=confusion_matrix_high_MCC)) +
+  geom_boxplot() +
+  geom_point() +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") 
 ```
-
-# Considering how each method contributes to the final prediction
 ```{r}
-viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
-viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
-                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
-                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
+table(viruses$confusion_matrix_high_MCC[viruses$checkv_length>=5000])
 ```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "viral_score")
-```
-
-```{r}
-sm_m <- sm_m[sm_m$viral_score>0,]
-
-sm_m$score <- sm_m$viral_score
-
-sm_m$score[sm_m$viral_score==0.5] <- "0.5"
-sm_m$score[sm_m$viral_score>=1] <- "1"
-sm_m$score[sm_m$viral_score>=2] <- "2"
-sm_m$score[sm_m$viral_score>=3] <- "3"
-sm_m$score[sm_m$viral_score>=4] <- "4"
-sm_m$score[sm_m$viral_score>=5] <- "5"
-
-sm_m$score <- factor(sm_m$score, 
-                                       levels=c("0.5", "1", "2","3","4","5"))
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  xlab("") +
-  ylab("Number of Sequences") +
-  coord_flip()
-```
-## Visualizing confusion matrix by number of tools
-
-```{r}
-viruses$keep_score_visualize <- viruses$keep_score_high_MCC
-viruses$keep_score_visualize[viruses$keep_score_high_MCC>1] <- "> 1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==1] <- "1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==0.5] <- "0.5"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==0] <- "0"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==-0.5] <- "-0.5"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==-1] <- "-1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC<=-1] <- "< -1"
-
-viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
-#viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-#                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
-```
-
-```{r}
-levels(factor(viruses$keep_score_visualize))
-```
-
-```{r}
-ggplot(viruses, aes(x=as.factor(Index),
-                   fill=keep_score_visualize, color=keep_score_visualize)) +
-  geom_bar(stat="count", position="stack") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16)
-  ) +
-  scale_color_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 0.5)) +
-  xlab("Index") +
-  ylab("Sequence Count") +
-  facet_wrap(~confusion_matrix_high_MCC, scales = "free")
-
-```
-
-## all 6 tools example
-```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_all <- "true negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
-viruses$confusion_matrix_all[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_all, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-table(viruses$confusion_matrix_all)
-
-length(grep("true", viruses$confusion_matrix_all))/nrow(viruses)
-```
-
-```{r}
-length(grep("true positive", viruses$confusion_matrix_all))/length(grep("virus", viruses$seqtype))
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-## high recall example
-```{r}
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_high_recall <- "true negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall<1] <- "false negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
-viruses$confusion_matrix_high_recall[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_recall, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-p2
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$confusion_matrix_high_recall))/nrow(viruses)
-```
-0.887
-
-recall
-```{r}
-length(grep("true positive", viruses$confusion_matrix_high_recall))/length(grep("virus", viruses$seqtype))
-```
-recover almost all of the viruses this way, but more protist contamination
-
-0.960
-
-######################################################################
-Extra Stuff
-#####################################################################
-
-```{r}
-ggplot(viruses, aes(x=checkv_length, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_MCC,
-                   color=confusion_matrix_high_MCC)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=hallmark,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Number of Hallmark Genes") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=confusion_matrix_high_recall, y=checkv_length,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") +
-  scale_y_log10()
-```
-
-looking at false negatives
-```{r}
-viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<1),]
-```
-
-looking at protists calling viral
-```{r}
-viruses_false_pos_protists <- viruses[(viruses$seqtype=="protist" & viruses$keep_score_high_recall>=1),]
-```
-
-# Considering how each method contributes to the final prediction (high MCC)
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses_high <- viruses[viruses$keep_score_vb_tv>=1,] #uncomment this line if want to use all 6 tools
-viruses_high_mod <- viruses_high %>% select(keep_score_vb, 
-                                            keep_score_vb_tv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "score")
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=as.factor(score))) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(14)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs") +
-  coord_flip()
-```
-
-# ROC 
-```{r}
-library(pROC)
-```
-
-```{r}
-viruses$truepositive <- rep(0, nrow(viruses))
-viruses$truepositive[viruses$seqtype=="virus"] <- 1
-```
-
-```{r}
-rocobj <- roc(viruses$truepositive, viruses$keep_score)
-rocobj_all <- roc(viruses$truepositive, viruses$keep_score_all)
-auc <- round(auc(viruses$truepositive, viruses$keep_score),4)
-auc_all <- round(auc(viruses$truepositive, viruses$keep_score_all),4)
-#create ROC plot
-ggroc(rocobj, colour = 'steelblue', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')')) +
-  coord_equal()
-ggroc(rocobj_all, colour = 'green', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc_all, ')'))
-```
-Sensitivity: The probability that the model predicts a positive outcome for an observation when indeed the outcome is positive.
-Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative.
-
-# Comparing behavior of all testing sets combined (clustering analyses)
-
-```{r}
-viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
-num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
-                          num_viruses=rep(0, nrow(combos_list)))
-
-for (i in 1:nrow(combos_list)) {
-  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
-                                            include_virsorter = combos_list$VS[i],
-                                            include_virsorter2 = combos_list$VS2[i],
-                                            include_tuning = combos_list$CheckV[i],
-                                            include_kaiju = combos_list$Kaiju[i],
-                                            include_deepvirfinder = combos_list$DVF[i])
-
-  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
-
-  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
-
-  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
-}
-
-num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
-num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
-num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
-num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
-num_viruses$numtools <- as.factor(num_viruses$numtools)
-```
-
-```{r}
-ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-
-ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
-  geom_boxplot(aes(color=numtools)) +
-  geom_point(aes(color=numtools, fill=numtools)) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Tools") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
-viral_scores_nozeros <- viral_scores_nozeros + 1
-viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
-
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
-```
-
-```{r}
-library(phyloseq)
-```
-
-```{r}
-tooldata <- num_viruses
-
-rownames(tooldata) <- tooldata$toolcombo2
-```
-
-```{r}
-physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
-                                     sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  geom_label(label=tooldata$toolcombo)
-
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw()
-```
-to do: try coloring above based on the F1 scores of the testing set on each combination
-
-```{r}
-bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
-clusters <- hclust(dist(bray_dist))
-plot(clusters)
-
-myclusters <- cutree(clusters, h=1.1)
-```
-
-```{r}
-names(myclusters[myclusters==1])
-names(myclusters[myclusters==2])
-names(myclusters[myclusters==3])
-names(myclusters[myclusters==4])
-names(myclusters[myclusters==5])
-
-myclusters_df <- tibble(combo=names(myclusters),
-                            cluster_index=myclusters)
-
-myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
-                                                            "Kaiju", "VIBRANT",
-                                                            "VirSorter", "VirSorter2"),
-                          sep=" ", remove = F)
-
-tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
-                    )
-
-tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
-
-tool_count <- melt(tool_count)
-
-colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
-ggplot(tool_count, aes(x=cluster_index, y=tool_count,
-                   fill=cluster_index,
-                   color=cluster_index)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Cluster") +
-  ylab("Number of Times in Cluster") + 
-  facet_wrap(~tool, scales = "free")
-```
-
-```{r}
- ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Viral Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
- ggplot(viruses, aes(x=percent_viral, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Percent Genes Viral") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
- ggplot(viruses, aes(x=hallmark, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-viruses_false_positive <- viruses[viruses$confusion_matrix_high_precision=="false positive",]
-viruses_false_negative <- viruses[viruses$confusion_matrix_high_precision=="false negative",]
-```
-
-```{r}
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=checkv_length,
-                   color=checkv_length,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="bacteria"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="fungi"], aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="protist"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)
-
-table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
-```

### 2022-11-01 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations of testing data.

@@
 library(pROC)
+library("stringr")
 ```
@@
 ```{r}
-library("stringr")
-```
-
-```{r}
 accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
@@
   ylab("Bacterial False Positives")
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
@@
   ylab("Viral False Negatives")
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Viral False Negatives")
+
+ggplot(accuracy_scores, aes(x=protist_FP, y=fungal_FP, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Protist FP") +
+  ylab("Fungal FP")
+
+ggplot(accuracy_scores, aes(x=recall, y=fungal_FP, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Recall") +
+  ylab("Fungal FP")
+
+ggplot(accuracy_scores, aes(x=recall, y=protist_FP, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Recall") +
+  ylab("Protist FP")
 ```
@@
 ```
+
+```{r}
+TP <- table(viruses$confusion_matrix_high_MCC)[4]
+FP <- table(viruses$confusion_matrix_high_MCC)[2]
+TN <- table(viruses$confusion_matrix_high_MCC)[3]
+FN <- table(viruses$confusion_matrix_high_MCC)[1]
+
+precision <- TP/(TP+FP)
+precision
+recall <- TP/(TP+FN)
+recall
+F1 <- 2*precision*recall/(precision+recall)
+F1
+
+MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
+MCC
+```
+
+precision=66%, recall=91%, MCC=74%
+
+precision adjusting size to be equal viral/not viral
+```{r}
+TP <- table(viruses$confusion_matrix_high_MCC)[4]
+FP <- table(viruses$confusion_matrix_high_MCC)[2]*.11
+TN <- table(viruses$confusion_matrix_high_MCC)[3]*.11
+FN <- table(viruses$confusion_matrix_high_MCC)[1]
+
+precision <- TP/(TP+FP)
+precision
+recall <- TP/(TP+FN)
+recall
+F1 <- 2*precision*recall/(precision+recall)
+F1
+
+MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
+MCC
+```
+
+precision=0.94, recall=0.90, F1=0.92, MCC=0.85
 
@@
 
+```{r}
+TP <- table(viruses$confusion_matrix_all)[4]
+FP <- table(viruses$confusion_matrix_all)[2]
+TN <- table(viruses$confusion_matrix_all)[3]
+FN <- table(viruses$confusion_matrix_all)[1]
+
+precision <- TP/(TP+FP)
+precision
+recall <- TP/(TP+FN)
+recall
+F1 <- 2*precision*recall/(precision+recall)
+F1
+
+MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
+MCC
+```
+precision=62%, recall=92%, MCC=73%
+
 visualizing confusion matrix by taxa
@@
 0.960
+
+## few tools, high MCC example
+```{r}
+viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+viruses$confusion_matrix_few_tools <- "true negative"
+viruses$confusion_matrix_few_tools[viruses$seqtype=="virus" & viruses$keep_score_few_tools<1] <- "false negative"
+viruses$confusion_matrix_few_tools[viruses$seqtype=="virus" & viruses$keep_score_few_tools>=1] <- "true positive"
+viruses$confusion_matrix_few_tools[viruses$seqtype!="virus" & viruses$keep_score_few_tools>=1] <- "false positive"
+```
+
+```{r}
+TP <- table(viruses$confusion_matrix_few_tools)[4]
+FP <- table(viruses$confusion_matrix_few_tools)[2]
+TN <- table(viruses$confusion_matrix_few_tools)[3]
+FN <- table(viruses$confusion_matrix_few_tools)[1]
+
+precision <- TP/(TP+FP)
+precision
+recall <- TP/(TP+FN)
+recall
+F1 <- 2*precision*recall/(precision+recall)
+F1
+
+MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
+MCC
+```
+precision=77%, recall=76%, MCC=74%
 

### 2022-11-01 TestingPipeline/Scripts/combining_tool_output.Rmd
Stats for RefSeq vs NonRefSeq viral recovery.

@@
 viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
+viruses$viral[is.na(viruses$max_score)] <- 0
 

### 2022-11-01 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Stats for RefSeq vs NonRefSeq viral recovery.

@@
 ```
+
+# Comparing RefSeq vs VS2 database
+VS2 sequences used: (/scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/NCLDV/env/NCLDV_env.fna /scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/lavidaviridae/env/HQ_virophages.fna /scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/dsDNAphage/env/* /scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/dsDNAphage/provirus/VirSorterCuratedDataset_HQ_contigs.fna)
+
+```{r}
+refseq_vir <- read_tsv("../IntermediaryFiles/refseq_viral_genome_ids.txt",col_names = F)
+refseq_vir$X1 <- sub(">", "", refseq_vir$X1)
+refseq_vir <- separate(refseq_vir, col=X1, sep=" ", remove=T, into="id")
+refseq_vir$id <- sub("\\.", "_", refseq_vir$id)
+```
+
+```{r}
+table(refseq_vir$id %in% viruses$contig)
+```
+
+```{r}
+viruses_refseq_vir <- viruses[viruses$contig %in% refseq_vir$id,]
+viruses_not_refseq_vir <- viruses[!(viruses$contig %in% refseq_vir$id) & viruses$seqtype=="virus",]
+```
+
+```{r}
+table(viruses_refseq_vir$max_score_group)
+table(viruses_not_refseq_vir$max_score_group)
+```
+recall
+```{r}
+length(grep("true positive", viruses_refseq_vir$confusion_matrix_high_MCC))/nrow(viruses_refseq_vir)
+```
+
+```{r}
+length(grep("true positive", viruses_not_refseq_vir$confusion_matrix_high_MCC))/nrow(viruses_not_refseq_vir)
+```

### 2022-11-01 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Stats for RefSeq vs NonRefSeq viral recovery.

@@
   if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
+    keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] <- keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] + 0.5
     keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
-    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 1
-    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 1
-    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 1
+    keep_score[input_seqs$RNA>=0.95] <- keep_score[input_seqs$RNA>=0.95] + 1
   }
@@
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 # consider accounting for RNA viruses
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
+    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 
   }
@@
   ylab("Recall")
+
+ggplot(accuracy_scores[accuracy_scores$testing_set_index==1,], aes(x=precision, y=recall)) +
+  geom_label(alpha=0.7, label=accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1]) +
+  geom_point(alpha=0.5, aes(color=numtools, fill=numtools)) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Precision") +
+  ylab("Recall")
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
@@
 
-precision <- TP/(TP+FP)
+precision <- as.numeric(TP/(TP+FP))
 precision
-recall <- TP/(TP+FN)
+recall <- as.numeric(TP/(TP+FN))
 recall
-F1 <- 2*precision*recall/(precision+recall)
+F1 <- as.numeric(2*precision*recall/(precision+recall))
 F1
 
-MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
+MCC <- as.numeric((TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN)))
 MCC
@@
 
-precision=66%, recall=91%, MCC=74%
+precision=69%, recall=87%, MCC=77%
 
@@
 
-precision=0.94, recall=0.90, F1=0.92, MCC=0.85
+precision=0.95, recall=0.87, F1=0.91, MCC=0.82
+
+stats for different types of sequences
+```{r}
+TP <- table(viruses$confusion_matrix_high_MCC)[4]
+FN <- table(viruses$confusion_matrix_high_MCC)[1]
+
+recall <- as.numeric(TP/(TP+FN))
+recall
+```
+
+precision=0.95, recall=0.87, F1=0.91, MCC=0.82
+
+```{r}
+head(viruses$contig[viruses$seqtype=="virus"])
+tail(viruses$contig[viruses$seqtype=="virus"])
+```
 

### 2022-11-01 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Stats for RefSeq vs NonRefSeq viral recovery.

--- 2022-10-28 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
@@
 ```{r}
-plot(viruses$RNA, viruses$viral)
-plot(viruses$lavidaviridae, viruses$viral)
-plot(viruses$NCLDV, viruses$viral)
-plot(viruses$ssDNA, viruses$viral)
+ggplot(viruses, aes(x=RNA, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("RNA VS2 Score") +
+  ylab("VS2 Viral Score")
+
+ggplot(viruses, aes(x=RNA, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~confusion_matrix_high_MCC, scales = "free") +
+  xlab("RNA VS2 Score") +
+  ylab("VS2 Viral Score")
+
+ggplot(viruses, aes(x=NCLDV, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("NCLDV VS2 Score") +
+  ylab("VS2 Viral Score")
+
+ggplot(viruses, aes(x=NCLDV, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~confusion_matrix_high_MCC, scales = "free") +
+  xlab("NCLDV VS2 Score") +
+  ylab("VS2 Viral Score")
+
+ggplot(viruses, aes(x=lavidaviridae, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("Lavidaviridae VS2 Score") +
+  ylab("VS2 Viral Score")
+
+ggplot(viruses, aes(x=lavidaviridae, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~confusion_matrix_high_MCC, scales = "free") +
+  xlab("Lavidaviridae VS2 Score") +
+  ylab("VS2 Viral Score")
 ```
@@
   ylab("Length of Sequence") 
+```
+
+```{r}
+ggplot(viruses, aes(x=checkv_host_genes, y=checkv_length)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("Length of Sequence") +
+  ylab("Number of Host Genes")
+```
+```{r}
+table(viruses$checkv_host_genes>=50, viruses$seqtype)
 ```
@@
 table(keep_score, viruses$seqtype)
-keep_score[viruses$checkv_host_genes>50 & viruses$checkv_provirus=="No"] <- keep_score[viruses$checkv_host_genes>50 & viruses$checkv_provirus=="No"] - 1
+keep_score[viruses$checkv_host_genes>50 & viruses$provirus=="No"] <- keep_score[viruses$checkv_host_genes>50 & viruses$provirus=="No"] - 1
 table(keep_score, viruses$seqtype)
@@
 table(keep_score, viruses$seqtype)
-keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"] <- keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"] - 1 # consider accounting for provirus designation
+keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus=="No"] <- keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus=="No"] - 1 # consider accounting for provirus designation
 table(keep_score, viruses$seqtype)

### 2022-11-01 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
Visualizing viral totals by assembly.
Drinking water rule stats.

@@
 ```{r}
-
 # checkV
@@
 viruses$checkv_completeness[is.na(viruses$checkv_completeness)] <- 0
-
 # VIBRANT
 viruses$vibrant_quality[is.na(viruses$vibrant_quality)] <- 0
-
 # DeepVirFinder
@@
 viruses$hallmark[is.na(viruses$hallmark)] <- 0
-
+viruses$RNA <- 0
+viruses$RNA[is.na(viruses$RNA)] <- 0
+viruses$lavidaviridae <- 0
+viruses$lavidaviridae[is.na(viruses$lavidaviridae)] <- 0
+viruses$NCLDV <- 0
+viruses$NCLDV[is.na(viruses$NCLDV)] <- 0
+viruses$ssDNA[is.na(viruses$ssDNA)] <- 0
+viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
+viruses$viral[is.na(viruses$max_score)] <- 0
 #Virsorter
 viruses$category[is.na(viruses$category)] <- 0
-
 # Kaiju
 viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
+viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
+#adding a combined provirus category
+viruses$provirus <- FALSE
+viruses$provirus[viruses$checkv_provirus=="Yes" | viruses$vibrant_prophage=="Yes" | viruses$category==4] <- TRUE
 ```
@@
 Data cleaning is done!
-Export viruses to tsv and load into identifying_viral_metrics.Rmd for metrics on tool outputs.
+Export viruses to tsv and load into identifying_viral_contigs_environmental.Rmd for metrics on tool outputs.
 The code for metrics is also pasted below, but it is not recommended running it in the same R session due to memory issues.

### 2022-11-01 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Drinking water rule stats.

@@
     keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 0.5
     keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
@@
     keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
+    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for RNA viruses
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
+    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 
   }
@@
 ```
+
+# numbers in common between two tool combos
+```{r}
+num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+
+for (i in 1:63) {
+  for (j in 1:63) {
+    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+  }
+}
+
+rownames(num_high_viruses) <- combos_list$toolcombo
+colnames(num_high_viruses) <- combos_list$toolcombo
+```
+
+```{r}
+combos_list_two <- combos_list
+combos_list_two$numtools <- str_count(combos_list_two$toolcombo, "1")
+combos_list_two <- combos_list_two[combos_list_two$numtools==1,]
+
+two_tool_pairs <- num_high_viruses[rownames(num_high_viruses) %in% combos_list_two$toolcombo,
+                                   colnames(num_high_viruses) %in% combos_list_two$toolcombo]
+
+colnames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+rownames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+```
+
+```{r}
+viral_scores_two <- viral_scores[,colnames(num_high_viruses) %in% combos_list_two$toolcombo]
+
+n <- 6
+viral_scores_two_comb <- matrix(data=NA, nrow=n, ncol=n)
+
+for (i in 1:n) {
+  for (j in 1:n) {
+    if (i==j) {
+      viral_scores_two_comb[i,j] <- sum(viral_scores_two[,i]>=1)
+    }
+    else if (i<j) {
+      viral_scores_two_comb[i,j] <- sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)
+    }
+    else if (i>j) {
+      viral_scores_two_comb[i,j] <- paste(round(sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)/sum(viral_scores_two[,i]>=1 | viral_scores_two[,j]>=1)*100, digits=0), "%", sep="")
+    }
+  }
+}
+
+colnames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+rownames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+```
+
+```{r}
+viral_scores_two_comb
+```

### 2022-11-01 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Drinking water rule stats.

@@
 Tuning - Viral Signature:
-    Kaiju_viral = "Viruses": +1
+    Kaiju_viral = "Viruses": +0.5
     If %unknown >= 75 and length < 50000: +0.5
     If %viral >= 50: +0.5
-    Hallmark > 2: +1
+    Hallmark > 2: +0.5
 
@@
 
-stats for different types of sequences
-```{r}
-TP <- table(viruses$confusion_matrix_high_MCC)[4]
-FN <- table(viruses$confusion_matrix_high_MCC)[1]
-
-recall <- as.numeric(TP/(TP+FN))
-recall
-```
-
-precision=0.95, recall=0.87, F1=0.91, MCC=0.82
-
-```{r}
-head(viruses$contig[viruses$seqtype=="virus"])
-tail(viruses$contig[viruses$seqtype=="virus"])
-```
-
 visualizing confusion matrix by taxa

### 2022-11-02 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updates to figure labeling.

@@
 ```{r}
-accuracy_scores$numtools <- str_count(accuracy_scores$toolcombo, "1")
-#accuracy_scores <- accuracy_scores[order(accuracy_scores$numtools, decreasing=F),]
+accuracy_scores$numrules <- str_count(accuracy_scores$toolcombo, "1")
+#accuracy_scores <- accuracy_scores[order(accuracy_scores$numrules, decreasing=F),]
 accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
 accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
-accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
+accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
 ```
@@
 p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=precision, y=recall, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
   geom_label(alpha=0.7, label=accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1]) +
-  geom_point(alpha=0.5, aes(color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5, aes(color=numrules, fill=numrules)) +
   theme_light() +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=protist_FP, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=protist_FP, y=fungal_FP, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=recall, y=fungal_FP, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 ggplot(accuracy_scores, aes(x=recall, y=protist_FP, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point(alpha=0.5) +
@@
 accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numtools, toolcombo) %>%
+  select(testing_set_index, precision, recall, MCC, numrules, toolcombo) %>%
   pivot_longer(cols=c(precision, recall, MCC), 
@@
 ```{r}
-ggplot(accuracy_scores_melt, aes(x=numtools, y=performance_metric_score, 
-                                  color=numtools, fill=numtools)) +
+ggplot(accuracy_scores_melt, aes(x=numrules, y=performance_metric_score, 
+                                  color=numrules, fill=numrules)) +
   geom_boxplot() +
   geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Rule Sets") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+```
+comparing metric with and without tuning rules
+```{r}
+accuracy_scores_melt$tuning_inc <- "no"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1] <- "tv"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "tnv"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1 &
+                                  substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "both"
+```
+
+```{r}
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+  geom_boxplot() +
+  geom_point(aes(color=numrules, fill=numrules), alpha=0.5) +
   theme_light() +
@@
 
-```
-comparing metric with and without tuning rules
-```{r}
-accuracy_scores_melt$tuning_inc <- "no"
-accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1] <- "tv"
-accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "tnv"
-accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1 &
-                                  substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "both"
-```
-
-```{r}
-ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score,
+                                 color=numrules, fill=numrules)) +
   geom_boxplot() +
-  geom_point(aes(color=numtools, fill=numtools), alpha=0.5) +
+  geom_point(alpha=0.5) +
   theme_light() +
@@
 
-ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score,
-                                 color=numtools, fill=numtools)) +
-  geom_boxplot() +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("MCC") +
-  xlab("Number of Tools") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  facet_wrap(~performance_metric)
-
 ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
   geom_boxplot() +
-  geom_boxplot(aes(color=numtools, fill=numtools)) +
- # geom_point(aes(color=numtools, fill=numtools), alpha=0.5) +
+  geom_boxplot(aes(color=numrules, fill=numrules)) +
+ # geom_point(aes(color=numrules, fill=numrules), alpha=0.5) +
   theme_light() +
@@
   ylab("MCC") +
-  xlab("Number of Tools") +
+  xlab("") +
   scale_fill_manual(name="",
@@
 
-num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses$numrules <- str_count(num_viruses$toolcombo, "1")
 num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
@@
 num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
-num_viruses$numtools <- as.factor(num_viruses$numtools)
+num_viruses$numrules <- as.factor(num_viruses$numrules)
 ```
@@
 phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
+                          shape="numrules", color="num_viruses") + 
   geom_point(size = 3) +
@@
 accuracy_scores_melt <- accuracy_scores %>% 
-  select(precision, recall, MCC, numtools, toolcombo) %>%
-  group_by(numtools, toolcombo) %>%
+  select(precision, recall, MCC, numrules, toolcombo) %>%
+  group_by(numrules, toolcombo) %>%
   summarise(precision=mean(precision),
@@
 ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point() +
@@
 ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
+                                  color=numrules, fill=numrules)) +
   geom_point() +
@@
 ```{r}
-ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
-  geom_boxplot(aes(color=numtools)) +
-  geom_point(aes(color=numtools, fill=numtools)) +
+ggplot(num_viruses, aes(x=numrules, y=num_viruses)) +
+  geom_boxplot(aes(color=numrules)) +
+  geom_point(aes(color=numrules, fill=numrules)) +
   theme_light() +

### 2022-11-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
New visualizations for SI.

@@
 ```
+## Another way of visualizing between rule sets
+
+```{r}
+viruses_mcc_alluvial <- data.frame(seqtype=viruses$seqtype,
+                                   keep_score_high_MCC=viruses$keep_score_high_MCC,
+                                   confusion_matrix_high_MCC=viruses$confusion_matrix_high_MCC)
+
+viruses_mcc_alluvial$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses_mcc_alluvial$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses_mcc_alluvial$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses_mcc_alluvial$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses_mcc_alluvial$keep_score_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses_mcc_alluvial$keep_score_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+viruses_mcc_alluvial %>%
+  count(seqtype, keep_score_high_MCC) %>% 
+  spread(key = keep_score_high_MCC, value=n)
+```
+
+```{r}
+viruses_mcc_alluvial <- viruses_mcc_alluvial %>%
+  count(seqtype, keep_score_dvf, keep_score_vb, keep_score_vs,
+        keep_score_vs2, keep_score_tv, keep_score_tnv, keep_score_high_MCC) %>%
+  mutate(high_mcc_viral_score=factor(round(keep_score_high_MCC)))
+```
+
+```{r}
+ggplot(viruses_mcc_alluvial,
+       aes(axis1 = keep_score_dvf, axis2 = keep_score_vb, 
+           axis3 = keep_score_vs, axis4 = keep_score_vs2, 
+           axis5 = keep_score_tv, axis6 = keep_score_tnv, 
+           y=n)) +
+  geom_alluvium(aes(fill=high_mcc_viral_score),
+                width = 0, knot.pos = 0, reverse = FALSE) +
+  geom_stratum(width = 1/5) +
+  theme_bw() +
+  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
+            reverse = FALSE) +
+  theme(
+        axis.text.x=element_text(size=14, angle = 90)
+        ) +
+  scale_x_continuous(breaks=c(1,2,3,4,5,6),
+    labels=c("dvf", "kj", "vs", "vs2",
+             "tv", "tnv")) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  facet_wrap(~seqtype, scales="free_y") 
+```
+
 ## Visualizing confusion matrix by number of tools
@@
 0.960
+
+```{r}
+confusion_by_taxa <- viruses %>% count(confusion_matrix_high_recall, seqtype, size_class)
+
+colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "count")
+```
 

### 2022-11-07 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
New visualizations for SI.

@@
 
-use of each tuning rule
+# Viral Addition Tuning Rules
+
 ```{r}
-keep_score <- rep(0, nrow(viruses))
+viruses_sankey_tv <- data.frame(seqtype=viruses$seqtype,
+                             kj_cel=rep(0,nrow(viruses)),
+                             hall=rep(0,nrow(viruses)),
+                             pv=rep(0,nrow(viruses)),
+                             cvl_pu=rep(0, nrow(viruses)))
+```
 
-keep_score[viruses$hallmark>2] <- keep_score[viruses$hallmark>2] + 1
-table(keep_score, viruses$seqtype)
-keep_score[viruses$checkv_host_genes>50 & viruses$provirus=="No"] <- keep_score[viruses$checkv_host_genes>50 & viruses$provirus=="No"] - 1
-table(keep_score, viruses$seqtype)
-keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- keep_score[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
-table(keep_score, viruses$seqtype)
-keep_score[viruses$percent_viral>=50] <- keep_score[viruses$percent_viral>=50] + 0.5
-table(keep_score, viruses$seqtype)
-    #keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] <- keep_score[viruses$hallmark>=(viruses$checkv_viral_genes/5)] + 1 #add some ratio
-keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- keep_score[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1
-table(keep_score, viruses$seqtype)
-keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus=="No"] <- keep_score[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus=="No"] - 1 # consider accounting for provirus designation
-table(keep_score, viruses$seqtype)
-#    keep_score[(viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes] <- 0 # consider accounting for provirus designation
-keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] <- keep_score[viruses$checkv_length>500000 & viruses$hallmark==0] - 1
-table(keep_score, viruses$seqtype)
+```{r}
+viruses_sankey_tv$kj_cel[viruses$Kaiju_Viral=="Viruses"] <- 0.5
+viruses_sankey_tv$hall[viruses$hallmark>2] <- 0.5
+viruses_sankey_tv$pv[viruses$percent_viral>=50] <- 0.5
+viruses_sankey_tv$cvl_pu[viruses$checkv_length>50000 & viruses$percent_unknown<=75] <- 0.5
+
+viruses_sankey_tv$all <- rowSums(viruses_sankey_tv[,2:5])
 ```
+
+```{r}
+viruses_sankey_tv %>%
+  count(seqtype, all) %>% spread(key = all, value=n)
+```
+
+```{r}
+viruses_sankey_tv <- viruses_sankey_tv %>%
+  count(seqtype, kj_cel, hall, pv, cvl_pu, all) %>%
+  mutate(viral_score=factor(all))
+```
+
+```{r}
+ggplot(viruses_sankey_tv,
+       aes(axis1 = kj_cel, axis2 = hall, axis3 = pv, axis4 = cvl_pu, y=n)) +
+  geom_alluvium(aes(fill=viral_score),
+                width = 0, knot.pos = 0, reverse = FALSE) +
+  geom_stratum(width = 1/5) +
+  theme_bw() +
+  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
+            reverse = FALSE) +
+  theme(
+        axis.text.x=element_text(size=14, angle = 90)
+        ) +
+  scale_x_continuous(breaks=c(1,2,3,4),
+    labels=c("kaiju", "hallmark", "% viral", "% unknown")) +
+  facet_wrap(~seqtype, scales="free_y") 
+```
+
+# Viral Removal Tuning Rules
+
+```{r}
+viruses_sankey_tnv <- data.frame(seqtype=viruses$seqtype,
+                             kj_cel=rep(0,nrow(viruses)),
+                             hg_pro=rep(0,nrow(viruses)),
+                             vg_hg=rep(0,nrow(viruses)),
+                             vg_hg_pro=rep(0, nrow(viruses)),
+                             cvl_hm=rep(0, nrow(viruses)),
+                             cvl_cp=rep(0,nrow(viruses)))
+```
+
+```{r}
+viruses_sankey_tnv$kj_cel[viruses$Kaiju_Viral=="cellular organisms"] <- -0.5
+viruses_sankey_tnv$hg_pro[viruses$checkv_host_genes>50 & viruses$provirus==FALSE] <- -1
+viruses_sankey_tnv$vg_hg[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- -1
+viruses_sankey_tnv$vg_hg_pro[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus==FALSE] <- -1
+viruses_sankey_tnv$cvl_hm[viruses$checkv_length>500000 & viruses$hallmark<=1] <- -1
+viruses_sankey_tnv$cvl_cp[viruses$checkv_length>5000 & viruses$checkv_completeness<=75] <- -0.5             
+viruses_sankey_tnv$all <- rowSums(viruses_sankey_tnv[,2:7])
+```
+
+```{r}
+viruses_sankey_tnv %>%
+  count(seqtype, all) %>% spread(key = all, value=n)
+```
+
+```{r}
+viruses_sankey_tnv <- viruses_sankey_tnv %>%
+  count(seqtype, kj_cel, hg_pro, vg_hg, vg_hg_pro, cvl_hm, cvl_cp, all) %>%
+  mutate(viral_score=factor(all))
+```
+
+```{r}
+ggplot(viruses_sankey_tnv,
+       aes(axis1 = kj_cel, axis2 = hg_pro, axis3 = vg_hg, axis4 = cvl_hm, axis5=cvl_hm, axis6=cvl_cp, 
+           y=n)) +
+  geom_alluvium(aes(fill=viral_score),
+                width = 0, knot.pos = 0, reverse = FALSE) +
+  geom_stratum(width = 1/5) +
+  theme_bw() +
+  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
+            reverse = FALSE) +
+  theme(
+        axis.text.x=element_text(size=14, angle = 90)
+        ) +
+  scale_x_continuous(breaks=c(1,2,3,4,5,6),
+    labels=c("kaiju", "host genes", "viral and host", "provirus",
+             "hallmark", "complete")) +
+  facet_wrap(~seqtype, scales="free_y") 
+```

### 2022-11-07 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updating drinking water scripts.

@@
 ```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+# add this
 ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
@@
 ```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
 ggplot(tool_count, aes(x=cluster_index, y=tool_count,

### 2022-11-17 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Refining figures.

@@
 ```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
+ggplot(confusion_by_taxa, aes(y=count, x=confusion_matrix,
                    fill=confusion_matrix,
                    color=confusion_matrix)) +
-  geom_bar(stat="identity") +
+  geom_boxplot() +
   theme_light() +
@@
   ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
+  facet_wrap(~seqtype, scales = "free")
 ```
@@
 
+## Visualizing the different sets
+
+```{r}
+confusion_by_taxa_method <- viruses %>% 
+  select(contains("confusion_matrix"), seqtype, Index) %>%
+  pivot_longer(cols=contains("confusion_matrix"), 
+               names_to="confusion_matrix_type",
+               values_to="confusion_matrix_value") %>%
+  count(seqtype, Index, confusion_matrix_type, confusion_matrix_value)
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+ggplot(confusion_by_taxa_method, aes(y=n, x=confusion_matrix_type,
+                   fill=confusion_matrix_value,
+                   color=confusion_matrix_value)) +
+  geom_boxplot() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4)), 0.5),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4)), 1),
+                    labels=c("false negative", "false positive", 
+                             "true negative", "true positive")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free")
+```
+
+```{r}
+confusion_by_taxa_method <- viruses %>% 
+  select(contains("confusion_matrix"), seqtype, Index) %>%
+  pivot_longer(cols=contains("confusion_matrix"), 
+               names_to="confusion_matrix_type",
+               values_to="confusion_matrix_value") %>%
+  count(seqtype, Index, confusion_matrix_type, confusion_matrix_value) %>%
+  filter(grepl("true", confusion_matrix_value)) %>%
+  mutate(confusion_matrix_type=sub("confusion_matrix_", "", confusion_matrix_type))
+```
+
+```{r}
+ggplot(confusion_by_taxa_method, aes(y=n, x=confusion_matrix_type,
+                   fill=confusion_matrix_value,
+                   color=confusion_matrix_value)) +
+  geom_boxplot() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=10, angle=90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(4))[3:4], 0.5),
+                    labels=c( 
+                             "true negative", "true positive")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(4))[3:4], 1),
+                    labels=c(
+                             "true negative", "true positive")) +
+  xlab("Tool Set") +
+  ylab("Contig Count") + 
+  facet_wrap(~seqtype, scales = "free")
+```
+
 ######################################################################
@@
 ```
+
+## Visualizing confusion matrix by number of tools
+
+```{r}
+confusion_by_keep_score <- viruses %>% 
+  select(contains("keep_score"), seqtype, Index) %>%
+  pivot_longer(cols=contains("keep_score"), 
+               names_to="pipeline_type",
+               values_to="pipeline_value") %>%
+  mutate(pipeline_type=sub("keep_score_", "", pipeline_type)) %>%
+  count(seqtype, Index, pipeline_type, pipeline_value)
+```
+
+```{r}
+confusion_by_keep_score$confusion_matrix <- "true negative"
+confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype=="virus" & confusion_by_keep_score$pipeline_value<1] <- "false negative"
+confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype=="virus" & confusion_by_keep_score$pipeline_value>=1] <- "true positive"
+confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype!="virus" & confusion_by_keep_score$pipeline_value>=1] <- "false positive"
+```
+
+```{r}
+confusion_by_keep_score$keep_score_visualize <- confusion_by_keep_score$pipeline_value
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value>1] <- "> 1"
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==1] <- "1"
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==0.5] <- "0.5"
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==0] <- "0"
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==-0.5] <- "-0.5"
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==-1] <- "-1"
+confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value<=-1] <- "< -1"
+
+confusion_by_keep_score$keep_score_visualize <- factor(confusion_by_keep_score$keep_score_visualize, 
+                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
+```
+
+```{r}
+ggplot(confusion_by_keep_score, aes(x=confusion_matrix, y=n,
+                   fill=keep_score_visualize, color=keep_score_visualize)) +
+  geom_boxplot() +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16)
+  ) +
+  scale_color_manual(name = 'Viral Score',
+                     values = alpha(c(viridis(6)), 1)) +
+  scale_fill_manual(name = 'Viral Score',
+                     values = alpha(c(viridis(6)), 0.5)) +
+  xlab("Index") +
+  ylab("Sequence Count") +
+  facet_wrap(~pipeline_type, scales = "free")
+
+```

### 2022-11-20 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
Updated dw vs2 output with all seq types.

--- 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "drinking_water"
+dataset_name <- "drinking_water_uknl"
 # checkV
@@
 # Virsorter2
-vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
+vs2_path <- "DrinkingWaterToolOutput/merged_vs2_all_types_uknl.tsv"
 # Kaiju
@@
 checkV$contig <- sub("\\.", "_", checkV$contig)
-checkV$uniq_contig <- paste(checkV$assembly, checkV$contig, sep = "--")
-checkV <- checkV[!duplicated(checkV$uniq_contig),]
+checkV <- checkV[!duplicated(checkV$contig),]
 ```
@@
                   'scaffold',
-                  'assembly',
                   'type',
@@
 vb_c$contig <- sub("\\.", "_", vb_c$contig)
-vb_c$uniq_contig <- paste(vb_c$assembly, vb_c$contig, sep = "--")
-vb_c <- vb_c[!duplicated(vb_c$uniq_contig),]
+vb_c <- vb_c[!duplicated(vb_c$contig),]
 ```
@@
 dvf_c$bh_pvalue <- p.adjust(dvf_c$pvalue, method = "BH")
-dvf_c$uniq_contig <- paste(dvf_c$assembly, dvf_c$contig, sep = "--")
-dvf_c <- dvf_c[!duplicated(dvf_c$uniq_contig),]
+dvf_c <- dvf_c[!duplicated(dvf_c$contig),]
 ```
@@
               select = c(
-                  'assembly',
                   'contig_id',
@@
 vs_c$contig <- sub("\\.", "_", vs_c$contig)
-vs_c$uniq_contig <- paste(vs_c$assembly, vs_c$contig, sep = "--")
+vs_c$contig <- paste(vs_c$assembly, vs_c$contig, sep = "--")
 
-vs_c <- vs_c[!duplicated(vs_c$uniq_contig),]
-vs_c <- vs_c %>% drop_na(uniq_contig)
+vs_c <- vs_c[!duplicated(vs_c$contig),]
+vs_c <- vs_c %>% drop_na(contig)
 ```
@@
                         'seqname',
-                        'assembly',
                         'dsDNAphage',
                         'ssDNA',
+                        'NCLDV',
+                        'RNA',
+                        'lavidaviridae',
                         'max_score',
@@
                   col = seqname,
-                  into = c("contig", "type"), 
+                  into = c("contig", "vs2type"), 
                   sep = "\\|\\|",
@@
                   )
-vs2_c$assembly <- sub("_3000", "", vs2_c$assembly)
 vs2_c$contig <- sub("\\.", "_", vs2_c$contig)
-vs2_c$uniq_contig <- paste(vs2_c$assembly, vs2_c$contig, sep="--")
-vs2_c <- vs2_c[!duplicated(vs2_c$uniq_contig),]
+vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
 ```
@@
 kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
-kj_c$uniq_contig <- paste(kj_c$assembly, kj_c$contig, sep="--")
 kj_c <- kj_c[!duplicated(kj_c$contig),]
-```
-
-```{r}
-# Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
-table(kj_c$assembly, kj_c$Kaiju_Viral)
 ```
@@
 ```{r}
-viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=vb_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=vs_c, by = c("uniq_contig", "assembly"))
-viruses <- full_join(x=viruses, y=kj_c, by = c("uniq_contig", "assembly"))
+viruses <- left_join(x=vs2_c, y=checkV, by = c("contig"))
+viruses <- left_join(x=viruses, y=dvf_c, by = c("contig", "assembly"))
+viruses <- left_join(x=viruses, y=vb_c, by = c("contig"))
+viruses <- left_join(x=viruses, y=vs_c, by = c("contig"))
+viruses <- left_join(x=viruses, y=kj_c, by = c("contig", "assembly"))
 ```
@@
 viruses$hallmark[is.na(viruses$hallmark)] <- 0
+viruses$RNA[is.na(viruses$RNA)] <- 0
+viruses$lavidaviridae[is.na(viruses$lavidaviridae)] <- 0
+viruses$NCLDV[is.na(viruses$NCLDV)] <- 0
+viruses$ssDNA[is.na(viruses$ssDNA)] <- 0
+viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
+viruses$viral[is.na(viruses$max_score)] <- 0
 
@@
 viruses$Kaiju_Viral[is.na(viruses$Kaiju_Viral)] <- "unknown"
+viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
+
+#adding a combined provirus category
+viruses$provirus <- FALSE
+viruses$provirus[viruses$checkv_provirus=="Yes" | viruses$vibrant_prophage=="Yes" | viruses$category==4] <- TRUE
+
+#remove contigs flagged as "less than 2 genes by vs2
+viruses <- viruses[!viruses$vs2type=="lt2gene",]
 ```
@@
 ```{r}
-viruses_filename <- 'merged_viral_feature_outputs.csv'
-write.csv(viruses, paste(dataset_name, viruses_filename, sep='_'))
+viruses_filename <- 'merged_viral_feature_outputs.tsv'
+write_tsv(viruses, paste(dataset_name, viruses_filename, sep='_'))
 ```

### 2022-11-20 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updated dw vs2 output with all seq types.

@@
 # dataset name for file organization and outputs for metrics file
-#dataset_name <- "drinking_water"
+dataset_name <- "drinking_water_uknl"
 #name from building_env_data_viral_features_set.Rmd
-#viruses_filename <- 'merged_viral_feature_outputs.csv'
-
-#viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+viruses_filename <- 'merged_viral_feature_outputs.tsv'
+viruses2 <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
 ```
-
-TODO: ADD ENVIRONMENTAL METRICS LIKE SHANNON, ALPHA DIVERSITY.
 
@@
   select(toolcombo, assembly, num_viruses) %>%
-  pivot_wider(assembly, num_viruses)
+  pivot_wider(names_from=assembly, values_from=num_viruses)
+num_viruses_uknl_wide <- as.data.frame(num_viruses_uknl_wide)
 rownames(num_viruses_uknl_wide) <- num_viruses_uknl_wide$toolcombo

### 2022-11-21 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
Updating figures summarizing testing data.

@@
 #name from building_env_data_viral_features_set.Rmd
-viruses_filename <- 'merged_viral_feature_outputs.csv'
+viruses_filename <- 'merged_viral_feature_outputs_nluk.csv'
 
-viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+viruses <- read_tsv(viruses_filename)
 ```

### 2022-11-21 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Updating figures summarizing testing data.

@@
 
+```{r}
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vb"), size_class, true_virus) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, true_virus) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~true_virus, scales = "free")
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(size_class~true_virus, scales = "free")
+```
+
 # Considering how each method contributes to the final prediction
@@
 
+## another way of visualizing the different tool sets scores
+```{r}
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_high"), contains("keep_score_all"), size_class, true_virus) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, true_virus) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~true_virus, scales = "free")
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(size_class~true_virus, scales = "free")
+```
+
 ######################################################################

### 2022-12-03 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
Some new visualizations and modifying VS2 rule.

---
title: "Comparing rule sets across environments"
output: html_notebook
---

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

```{r}
viruses_dw <- read_csv("../DrinkingWaterMetagenomes/drinking_water_num_viruses.csv")
```
```{r}
viruses_rest <- read_csv("/Users/hegartyb/Downloads/num_viruses_env.csv")
```

only keeping chlorinated drinking water samples
```{r}
viruses_dw <- viruses_dw[viruses_dw$assembly %in% c("Birmingham_Maria",
                                                       "Cardiff_Maria", "Dundee_Clat",
                                                       "Dundee_Maria",
                                                       "Exeter_Maria", "Glasgow_Bal",
                                                       "Glasgow_Mil", "London_Maria"),
                           ]
```

```{r}
viruses_dw$environment <- "drinking_water"

viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake_Erie"
viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake_Michigan"
viruses_rest$environment[viruses_rest$environment=="tara"] <- "global_oceans"

viruses_all <- rbind(viruses_dw, viruses_rest)
```

recoding AJ's samples to be able to separate virome from metagenome (and remove others)
```{r}
viruses_all$environment2 <- viruses_all$environment
viruses_all$environment2[viruses_all$assembly %in% c("49639", "49640", "49641", "49642")] <- "Lake_Erie_virome"
viruses_all$environment2[viruses_all$assembly %in% c("49613", "49614", "49618", "49621",
                                                     "49624", "49628", "49632", "49638")] <- "Lake_Erie_metagenome"
viruses_all <- viruses_all[viruses_all$environment2!="Lake_Erie",]
```

```{r}
viruses_all$environment2 <- factor(viruses_all$environment2,
                                    levels = c("drinking_water", "Lake_Erie_metagenome", 
                                               "global_oceans",
                                               "wastewater", "Lake_Erie_virome", 
                                               "Lake_Michigan"))
```

```{r}
viruses_all$best_coloring <- "others"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 1 1 1 1"] <- "all"
viruses_all$best_coloring[viruses_all$toolcombo=="0 1 1 1 1 1"] <- "high recall"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 1 1 1"] <- "high MCC"
viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 0 1"] <- "VS2 combo"
viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 0 0"] <- "high precision"
```

```{r}
ggplot(viruses_all, aes(x=toolcombo, y=prop_vir, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment2)
```

```{r}
viruses_all$type <- "metagenome"
viruses_all$type[viruses_all$environment2 %in% c("wastewater", "Lake_Erie_virome",
                                                 "Lake_Michigan")] <- "virome"
```

```{r}
ggplot(viruses_all[viruses_all$best_coloring %in% c("high MCC", "high recall"),], aes(x=environment2, y=prop_vir, 
                                  color=type, fill=type)) +
  geom_boxplot() +
  #geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Environment") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(magma(4)[2:3], 0.3)) +
  scale_color_manual(name="",
                     values = alpha(magma(4)[2:3], 0.7)) +
  facet_wrap(~best_coloring)
```

```{r}
imp_toolcombs <- viruses_all %>% group_by(best_coloring, environment) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            )
```

```{r}
ggplot(imp_toolcombs, aes(x=best_coloring, y=mean, 
                                  color=best_coloring, fill=best_coloring)) +
  geom_point(size=3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(viridis(6)), 0.3)) +
  scale_color_manual(name="",
                     values = alpha(rev(viridis(6)), 0.7)) +
  facet_wrap(~environment)
```

```{r}
imp_toolcombs_wide <- viruses_all %>% group_by(best_coloring, environment) %>%
  summarize(median=median(prop_vir),
            mean=mean(prop_vir),
            sd=sd(prop_vir)
            ) %>%
  select(mean, environment, best_coloring) %>%
  pivot_wider(names_from=environment, values_from=mean)
```

### 2022-12-03 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
Some new visualizations and modifying VS2 rule.

@@
 vs2_c <- vs2_c[!duplicated(vs2_c$contig),]
+vs2_c <- vs2_c[!grep("seqname", vs2_c$contig),]
 ```
@@
 viruses$RNA[is.na(viruses$RNA)] <- 0
+viruses$RNA[grep("nan", viruses$RNA)] <- 0
 viruses$lavidaviridae[is.na(viruses$lavidaviridae)] <- 0
+viruses$lavidaviridae[grep("nan", viruses$lavidaviridae)] <- 0
 viruses$NCLDV[is.na(viruses$NCLDV)] <- 0
+viruses$NCLDV[grep("nan", viruses$NCLDV)] <- 0
 viruses$ssDNA[is.na(viruses$ssDNA)] <- 0
+viruses$ssDNA[grep("nan", viruses$ssDNA)] <- 0
 viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
+viruses$dsDNAphage[grep("nan", viruses$dsDNAphage)] <- 0
 viruses$viral[is.na(viruses$max_score)] <- 0

### 2022-12-03 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Some new visualizations and modifying VS2 rule.

@@
 viruses_filename <- 'merged_viral_feature_outputs.tsv'
-viruses2 <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
 ```
@@
 
+```{r}
+write_csv(viruses[,c(1,43:46)], "drinking_water_top_viral_scores.csv")
+```
+
 not having the rule, gives the same
@@
 }
-num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
 num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
@@
   xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
+  ylab("Num Viruses Predicted") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
 ```
@@
 
-num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
 num_viruses <- num_viruses[order(num_viruses$prop_vir, decreasing=F),]
@@
 num_viruses$numtools <- as.factor(num_viruses$numtools)
+```
+
+```{r}
+write_csv(num_viruses, "drinking_water_num_viruses.csv")
 ```

### 2022-12-03 TestingPipeline/Scripts/combining_tool_output.Rmd
Added comments about why removed duplicates.
Some new visualizations and modifying VS2 rule.

--- 2022-10-29 TestingPipeline/Scripts/combining_tool_output.Rmd
@@
 dataframe and memory usage.
+
+5) This script only keeps each fragment once. This has two purposes: (1) it 
+deals with the duplicate entries for fragments labeled as "complete" (the second
+entry gives the quality); (2) it keeps only one instance of sequences that have
+multiple viral fragments on them.
 
@@
 # VirSorter2
+viruses$max_score[is.na(viruses$max_score)] <- 0
 viruses$viral[is.na(viruses$viral)] <- 0

### 2022-12-03 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Some new visualizations and modifying VS2 rule.

@@
 dim(viral_scores)
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+viruses_1 <- viruses[viruses$Index=="1",]
+
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2
+)
+
+ann_row = data.frame(
+    seqtype = viruses_1$seqtype
+)
+
+ann_colors = list(
+  tnv = c("black", pal(6)[1]),
+  tv = c("black", pal(6)[3]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6]),
+  seqtype = c("archaea"=pal(6)[1], "bacteria"=pal(6)[2], "fungi"=pal(6)[3],
+              "plasmid"=pal(6)[4], "protist"=pal(6)[5], "virus"=pal(6)[6])
+)
+
+rownames(ann_col) <- combos_list$toolcombo
+rownames(ann_row) <- viruses_1$contig
+
+viral_scores_1 <- viral_scores[viruses$Index=="1",]
+
+colnames(viral_scores_1) <- combos_list$toolcombo
+rownames(viral_scores_1) <- viruses_1$contig
+```
+
+```{r}
+pheatmap::pheatmap(t(viral_scores_1[(viruses_1$seqtype=="virus"),]),
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  #border_color = "black",
+                  color = c(rev(RColorBrewer::brewer.pal(3,"Oranges")),
+                            RColorBrewer::brewer.pal(5,"Purples") 
+                            ),
+                  show_rownames = F,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows=F#,
+                 # filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
+                  )
+```
+
+```{r}
+pheatmap::pheatmap(t(viral_scores_1[(viruses_1$seqtype=="bacteria"),]),
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  #border_color = "black",
+                  color = c(rev(RColorBrewer::brewer.pal(5,"Oranges")),
+                            RColorBrewer::brewer.pal(4,"Purples") 
+                            ),
+                  show_rownames = F,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows=F#,
+                 # filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
+                  )
+```
+
+```{r}
+pheatmap::pheatmap(t(viral_scores_1),
+                  annotation_col = ann_row,
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  #border_color = "black",
+                  color = c(rev(RColorBrewer::brewer.pal(5,"Oranges")),
+                            RColorBrewer::brewer.pal(5,"Purples") 
+                            ),
+                  show_rownames = F,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows = F,
+                  cluster_cols = F#,
+                 # filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
+                  )
 ```

### 2022-12-03 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Some new visualizations and modifying VS2 rule.

---
title: "Individual Rule Visualizations"
output: html_notebook
---

This is a notebook to look more at each rule.

# Looking at each rule

## VIBRANT
```{r}
vscore <- tibble(seqtype=viruses$seqtype,
                 vscore=rep(0, nrow(viruses)))

vscore$vscore[viruses$vibrant_quality=="high quality draft"] <- vscore$vscore[viruses$vibrant_quality=="high quality draft"] + 1
vscore$vscore[viruses$vibrant_quality=="medium quality draft"] <- vscore$vscore[viruses$vibrant_quality=="medium quality draft"] + 1
vscore$vscore[viruses$vibrant_quality=="low quality draft" & viruses$provirus] <- vscore$vscore[viruses$vibrant_quality=="low quality draft" & viruses$provirus] + 0.5

vscore$confusion_matrix <- "correct"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"

table(vscore$seqtype, vscore$confusion_matrix)

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10)
```

## VirSorter2
```{r}
vscore <- tibble(seqtype=viruses$seqtype,
                 vscore=rep(0, nrow(viruses)))

vscore$vscore[viruses$max_score>=0.5] <- vscore$vscore[viruses$max_score>=0.5] + 0.5
vscore$vscore[viruses$max_score>=0.95] <- vscore$vscore[viruses$max_score>=0.95] + 0.5

vscore$confusion_matrix <- "correct"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"

table(vscore$seqtype, vscore$confusion_matrix)

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10)
```

## VirSorter
```{r}
vscore <- tibble(seqtype=viruses$seqtype,
                 vscore=rep(0, nrow(viruses)))

vscore$vscore[viruses$category==1] <- vscore$vscore[viruses$category==1] + 1
vscore$vscore[viruses$category==2] <- vscore$vscore[viruses$category==2] + 0.5
vscore$vscore[viruses$category==4] <- vscore$vscore[viruses$category==4] + 1
vscore$vscore[viruses$category==5] <- vscore$vscore[viruses$category==5] + 0.5

vscore$confusion_matrix <- "correct"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"

table(vscore$seqtype, vscore$confusion_matrix)

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10)
```

## DeepVirFinder
```{r}
vscore <- tibble(seqtype=viruses$seqtype,
                 vscore=rep(0, nrow(viruses)))

vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] + 0.5
vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] + 0.5

vscore$confusion_matrix <- "correct"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"

table(vscore$seqtype, vscore$confusion_matrix)

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10)
```

## Tuning Addition
```{r}
vscore <- tibble(seqtype=viruses$seqtype,
                 vscore=rep(0, nrow(viruses)))

vscore$vscore[viruses$Kaiju_Viral=="Viruses"] <- vscore$vscore[viruses$Kaiju_Viral=="Viruses"] + 0.5
vscore$vscore[viruses$hallmark>2] <- vscore$vscore[viruses$hallmark>2] + 0.5
vscore$vscore[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- vscore$vscore[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
vscore$vscore[viruses$viral>=50] <- vscore$vscore[viruses$viral>=50] + 0.5

vscore$confusion_matrix <- "correct"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"

table(vscore$seqtype, vscore$confusion_matrix)

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10,
                   number_format = "%.3f")
```

## Tuning Removal
```{r}
vscore <- tibble(seqtype=viruses$seqtype,
                 vscore=rep(1, nrow(viruses)))

#vscore$vscore[viruses$Kaiju_Viral=="cellular organisms"] <- vscore$vscore[viruses$Kaiju_Viral=="cellular organisms"] - 0.5
vscore$vscore[viruses$checkv_host_genes>50 & !viruses$provirus] <- vscore$vscore[viruses$checkv_host_genes>50 & !viruses$provirus] - 1
vscore$vscore[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- vscore$vscore[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1
vscore$vscore[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & !viruses$provirus] <- vscore$vscore[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & !viruses$provirus] - 1 
vscore$vscore[viruses$checkv_length>500000 & viruses$hallmark<=1] <- vscore$vscore[viruses$checkv_length>500000 & viruses$hallmark<=1] - 1
vscore$vscore[(viruses$checkv_completeness<=75 | viruses$vibrant_quality=="complete circular")& viruses$checkv_length<=5000] <- vscore$vscore[viruses$checkv_completeness<=75 & viruses$checkv_length<=5000] - 0.5

vscore$confusion_matrix <- "correct"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"

table(vscore$seqtype, vscore$confusion_matrix)

pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
                   color=c("black", viridis(1000)),
                   cluster_rows = F,
                   cluster_cols = F,
                   display_numbers = T,
                   number_color = "white", 
                   fontsize_number = 10)
```

### 2022-12-03 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Some new visualizations and modifying VS2 rule.

@@
 VIBRANT
-    Quality == "Complete Circular": +1
     Quality == "High Quality Draft": +1
@@
   if (include_vibrant) {
-    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
     keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
@@
   if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] <- keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
-    keep_score[input_seqs$RNA>=0.95] <- keep_score[input_seqs$RNA>=0.95] + 1
+    #keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] <- keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] + 0.5
+    keep_score[input_seqs$max_score>=50] <- keep_score[input_seqs$max_score>=50] + 0.5
+    keep_score[input_seqs$max_score>=95] <- keep_score[input_seqs$max_score>=95] + 0.5
+    #keep_score[input_seqs$RNA>=0.95] <- keep_score[input_seqs$RNA>=0.95] + 1
   }
@@
     keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
+    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
   }
@@
     keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 
+    keep_score[(input_seqs$checkv_completeness<=75 | input_seqs$vibrant_quality=="complete circular")& input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5
   }
@@
 
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
 ## Visualize how the precision, recall, and F1 scores change across pipelines.
 ```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
 p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
@@
 p2
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
@@
   ylab("Recall")
+
 ggplot(accuracy_scores, aes(x=precision, y=recall, 
@@
     legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
+    axis.title=element_text(size=20),
   ) +
   xlab("Precision") +
-  ylab("Recall")
+  ylab("Recall") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
 
@@
     legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
+    axis.title=element_text(size=20),
   ) +
   xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("MCC")
+  ylab("MCC") +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(pal(6)), 1))
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
@@
 ```{r}
-ggplot(accuracy_scores_melt, aes(x=numrules, y=performance_metric_score, 
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="MCC",], aes(x=numrules, y=performance_metric_score, 
                                   color=numrules, fill=numrules)) +
@@
   ylab("MCC") +
+  xlab("Number of Rule Sets") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
+
+```
+
+```{r}
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="recall",], aes(x=numrules, y=performance_metric_score, 
+                                  color=numrules, fill=numrules)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("recall") +
+  xlab("Number of Rule Sets") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
+
+```
+
+```{r}
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
+                                  color=numrules, fill=numrules)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Score") +
   xlab("Number of Rule Sets") +
@@
 
-ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=tuning_inc, y=performance_metric_score)) +
   geom_boxplot() +
@@
   ) +
-  ylab("MCC") +
+  ylab("Score") +
   xlab("") +
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(pal(6)), 0.3)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
+                     values = alpha(rev(pal(6)), 0.7)) +
   facet_wrap(~performance_metric)
@@
 ```{r}
+viruses_long_scores_addition <- viruses_long_scores[(viruses_long_scores$true_virus=="virus" &  (viruses_long_scores$rule_combination!="keep_score_vb_dvf_vs2_vs_tv_tnv") & viruses_long_scores$viral_score!="0"),]
+ggplot(viruses_long_scores_addition, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "Purples") +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
+                            "VB+DVF+VS2+VS+addition")) +
+  facet_grid(~true_virus, scales = "free")
+```
+
+```{r}
 ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
@@
   ylab("Number of Sequences") + 
+  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
+                            "VB+DVF+VS2+VS+addition", "VB+DVF+VS2+VS+addition-removal")) +
   facet_grid(~true_virus, scales = "free")

### 2022-12-03 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Some new visualizations and modifying VS2 rule.

@@
 ```
+
+# Looking at each rule
+
+  if (include_virsorter) {
+    vscore$vscore[viruses$category==1] <- vscore$vscore[viruses$category==1] + 1
+    vscore$vscore[viruses$category==2] <- vscore$vscore[viruses$category==2] + 0.5
+    vscore$vscore[viruses$category==4] <- vscore$vscore[viruses$category==4] + 1
+    vscore$vscore[viruses$category==5] <- vscore$vscore[viruses$category==5] + 0.5
+  }
+
+  if (include_deepvirfinder) {
+    vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] + 0.5
+    vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] + 0.5
+  }
+
+  if (include_tuning_viral) {
+    vscore$vscore[viruses$Kaiju_Viral=="Viruses"] <- vscore$vscore[viruses$Kaiju_Viral=="Viruses"] + 0.5
+    vscore$vscore[viruses$hallmark>2] <- vscore$vscore[viruses$hallmark>2] + 0.5
+    vscore$vscore[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- vscore$vscore[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
+    vscore$vscore[viruses$viral>=50] <- vscore$vscore[viruses$viral>=50] + 0.5
+  }
+
+  if (include_tuning_not_viral) {
+    vscore$vscore[viruses$Kaiju_Viral=="cellular organisms"] <- vscore$vscore[viruses$Kaiju_Viral=="cellular organisms"] - 0.5
+    vscore$vscore[viruses$checkv_host_genes>50 & !viruses$provirus] <- vscore$vscore[viruses$checkv_host_genes>50 & !viruses$provirus] - 1
+    vscore$vscore[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- vscore$vscore[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1
+    vscore$vscore[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & !viruses$provirus] <- vscore$vscore[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & !viruses$provirus] - 1
+    vscore$vscore[viruses$checkv_length>500000 & viruses$hallmark<=1] <- vscore$vscore[viruses$checkv_length>500000 & viruses$hallmark<=1] - 1
+    vscore$vscore[(viruses$checkv_completeness<=75 | viruses$vibrant_quality=="complete circular")& viruses$checkv_length<=5000] <- vscore$vscore[viruses$checkv_completeness<=75 & viruses$checkv_length<=5000] - 0.5
+  }
+
+  return(keep_score)
+
+}
+
+# VIBRANT
+```{r}
+vscore <- tibble(seqtype=viruses$seqtype,
+                 vscore=rep(0, nrow(viruses)))
+
+vscore$vscore[viruses$vibrant_quality=="high quality draft"] <- vscore$vscore[viruses$vibrant_quality=="high quality draft"] + 1
+vscore$vscore[viruses$vibrant_quality=="medium quality draft"] <- vscore$vscore[viruses$vibrant_quality=="medium quality draft"] + 1
+vscore$vscore[viruses$vibrant_quality=="low quality draft" & viruses$provirus] <- vscore$vscore[viruses$vibrant_quality=="low quality draft" & viruses$provirus] + 0.5
+
+vscore$confusion_matrix <- "true negative"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false negative"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "true positive"
+vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false positive"
+
+table(vscore$seqtype, vscore$confusion_matrix)
+
+pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
+                   color=c("black", viridis(100)))
+```
+
+# VIBRANT
+```{r}
+vscore <- tibble(seqtype=viruses$seqtype,
+                 vscore=rep(0, nrow(viruses)))
+
+vscore$vscore[viruses$max_score>=0.5] <- vscore$vscore[viruses$max_score>=0.5] + 0.5
+vscore$vscore[viruses$max_score>=0.95] <- vscore$vscore[viruses$max_score>=0.95] + 0.5
+
+vscore$confusion_matrix <- "true negative"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false negative"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "true positive"
+vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false positive"
+
+table(vscore$seqtype, vscore$confusion_matrix)
+
+pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
+                   color=c("black", viridis(100)))
+```
+
+# Looking at Features
 
@@
   ylab("VS2 Viral Score")
+
+ggplot(viruses, aes(x=viral, y=percent_viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("VS2 Percent Viral") +
+  ylab("CheckV Percent Viral")
 ```

### 2022-12-04 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Adding DVF as a tuning removal piece.

@@
 
-This is a notebook to look more at each rule.
+This is a notebook to look more at each rule set.
+
+```{r setup-library}
+library(ggplot2)
+library(plyr)
+library(reshape2)
+library(viridis)
+library(tidyr)
+library(dplyr)
+library(readr)
+library(data.table)
+library(pROC)
+library("stringr")
+```
+
+Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
+```{r}
+viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
+```
 
 # Looking at each rule
+
+```{r}
+#viruses2 <- viruses
+#viruses <- viruses[viruses$checkv_length<=10000,]
+```
+
+```{r}
+#viruses <- viruses2
+```
 
@@
 
-vscore$vscore[viruses$vibrant_quality=="high quality draft"] <- vscore$vscore[viruses$vibrant_quality=="high quality draft"] + 1
-vscore$vscore[viruses$vibrant_quality=="medium quality draft"] <- vscore$vscore[viruses$vibrant_quality=="medium quality draft"] + 1
+#vscore$vscore[viruses$vibrant_quality=="high quality draft"] <- vscore$vscore[viruses$vibrant_quality=="high quality draft"] + 1
+#vscore$vscore[viruses$vibrant_quality=="medium quality draft"] <- vscore$vscore[viruses$vibrant_quality=="medium quality draft"] + 1
 vscore$vscore[viruses$vibrant_quality=="low quality draft" & viruses$provirus] <- vscore$vscore[viruses$vibrant_quality=="low quality draft" & viruses$provirus] + 0.5
@@
                    number_color = "white", 
-                   fontsize_number = 10)
+                   fontsize_number = 10,
+                   number_format = "%.3f")
 ```
@@
 
-vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] + 0.5
-vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] + 0.5
+#vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.7 & viruses$checkv_length<20000] + 0.5
+#vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] <- vscore$vscore[viruses$score>=0.9 & viruses$checkv_length<20000] + 0.5
+vscore$vscore[viruses$score>=0.9] <- vscore$vscore[viruses$score>=0.9] + 0.5
+vscore$vscore[viruses$score>=0.7] <- vscore$vscore[viruses$score>=0.7] + 0.5
 
@@
                    number_color = "white", 
-                   fontsize_number = 10)
+                   fontsize_number = 10,
+                   number_format = "%.3f")
+```
+## DVF for removal
+```{r}
+vscore <- tibble(seqtype=viruses$seqtype,
+                 vscore=rep(1, nrow(viruses)))
+
+vscore$vscore[viruses$score<=0.9] <- vscore$vscore[viruses$score<=0.9] - 0.5
+vscore$vscore[viruses$score<=0.7] <- vscore$vscore[viruses$score<=0.7] - 0.5
+
+vscore$confusion_matrix <- "correct"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
+vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"
+
+table(vscore$seqtype, vscore$confusion_matrix)
+
+pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
+                   color=c("black", viridis(1000)),
+                   cluster_rows = F,
+                   cluster_cols = F,
+                   display_numbers = T,
+                   number_color = "white", 
+                   fontsize_number = 10,
+                   number_format = "%.3f")
 ```
@@
 
+## Tuning Addition
+by sub rule
+```{r}
+vscore <- tibble(seqtype=viruses$seqtype,
+                 vscore=rep(0, nrow(viruses)))
+
+vscore$vscore[viruses$Kaiju_Viral=="Viruses"] <- vscore$vscore[viruses$Kaiju_Viral=="Viruses"] + 0.25
+vscore$vscore[viruses$hallmark>2] <- vscore$vscore[viruses$hallmark>2] + 0.5
+vscore$vscore[viruses$percent_unknown>=75 & viruses$checkv_length<50000] <- vscore$vscore[viruses$percent_unknown>=75 & viruses$checkv_length<50000] + 0.5
+vscore$vscore[viruses$viral>=50] <- vscore$vscore[viruses$viral>=50] + 0.25
+
+vscore$confusion_matrix <- "correct"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
+vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"
+
+table(vscore$seqtype, vscore$confusion_matrix)
+
+pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
+                   color=c("black", viridis(1000)),
+                   cluster_rows = F,
+                   cluster_cols = F,
+                   display_numbers = T,
+                   number_color = "white", 
+                   fontsize_number = 10,
+                   number_format = "%.3f")
+```
+
+notes: hallmark+vs2 % viral: no archael, fungal, protists wrong
+- hallmark+length based percent unknown: no archaea, bacteria, fungi, protists wrong
+- vs2 % viral+length based percent unknown: no fungi wrong
+
 ## Tuning Removal
@@
                    number_color = "white", 
-                   fontsize_number = 10)
+                   fontsize_number = 10,
+                   number_format = "%.3f")
 ```
+
+## Tuning Removal by rule
+```{r}
+vscore <- tibble(seqtype=viruses$seqtype,
+                 vscore=rep(1, nrow(viruses)))
+
+#vscore$vscore[viruses$Kaiju_Viral=="cellular organisms"] <- vscore$vscore[viruses$Kaiju_Viral=="cellular organisms"] - 0.5
+vscore$vscore[viruses$checkv_host_genes>50 & !viruses$provirus] <- 0
+vscore$vscore[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- vscore$vscore[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] - 1 # maybe make 0
+vscore$vscore[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & !viruses$provirus] <- vscore$vscore[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & !viruses$provirus] - 1 # maybe make 0
+vscore$vscore[viruses$checkv_length>500000 & viruses$hallmark<=1] <- 0
+vscore$vscore[(viruses$checkv_completeness<=75 | viruses$vibrant_quality=="complete circular")& viruses$checkv_length<=5000] <- vscore$vscore[viruses$checkv_completeness<=75 & viruses$checkv_length<=5000] - 0.5
+
+vscore$confusion_matrix <- "correct"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore<1] <- "false"
+vscore$confusion_matrix[vscore$seqtype=="virus" & vscore$vscore>=1] <- "correct"
+vscore$confusion_matrix[vscore$seqtype!="virus" & vscore$vscore>=1] <- "false"
+
+table(vscore$seqtype, vscore$confusion_matrix)
+
+pheatmap::pheatmap(table(vscore$seqtype, vscore$confusion_matrix)/c(rowSums(table(vscore$seqtype, vscore$confusion_matrix))),
+                   color=c("black", viridis(1000)),
+                   cluster_rows = F,
+                   cluster_cols = F,
+                   display_numbers = T,
+                   number_color = "white", 
+                   fontsize_number = 10,
+                   number_format = "%.4f")
+```

### 2022-12-04 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Adding DVF as a tuning removal piece.

@@
 ```
+(this is still the old rules)
 
@@
 
+# Trying new rule set
+this set is based on optimizing for the 
 ```{r getting_viral_set_1}
@@
   if (include_virsorter2) {
-    #keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] <- keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] + 0.5
     keep_score[input_seqs$max_score>=50] <- keep_score[input_seqs$max_score>=50] + 0.5
     keep_score[input_seqs$max_score>=95] <- keep_score[input_seqs$max_score>=95] + 0.5  
-    #keep_score[input_seqs$RNA>=0.95] <- keep_score[input_seqs$RNA>=0.95] + 1
   }
@@
     keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
     keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
     keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
   }
@@
   if (include_deepvirfinder) {
+    # remove if DVF calls it not viral
+    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
+    # add if DVF calls viral
+    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
     keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
-    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
   }
@@
   if (include_tuning_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 0.5
+  # tuning addition
+    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
     keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
+    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
+    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5
   }
@@
   if (include_tuning_not_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
+    # tuning removal 
+    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
     keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[(input_seqs$checkv_completeness<=75 | input_seqs$vibrant_quality=="complete circular")& input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5
+    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
+    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
   }
@@
 }
+```
+
+## Trying with all of the rules
+```{r}
+viruses$keep_score_all_2 <- getting_viral_set_1(viruses,
+                                                include_vibrant=TRUE, 
+                                                include_virsorter2=TRUE,
+                                                include_deepvirfinder=TRUE,
+                                                include_tuning_viral=TRUE,
+                                                include_tuning_not_viral=TRUE,
+                                                include_virsorter=TRUE)
+
+viruses$confusion_matrix_all_2 <- "true negative"
+viruses$confusion_matrix_all_2[viruses$seqtype=="virus" & viruses$keep_score_all_2<1] <- "false negative"
+viruses$confusion_matrix_all_2[viruses$seqtype=="virus" & viruses$keep_score_all_2>=1] <- "true positive"
+viruses$confusion_matrix_all_2[viruses$seqtype!="virus" & viruses$keep_score_all_2>=1] <- "false positive"
+
+viruses$accurate_all_2 <- "false"
+viruses$accurate_all_2[grep("true", viruses$confusion_matrix_all_2)] <- "true"
+```
+
+accuracy:
+```{r}
+length(grep("true", viruses$accurate_all_2))/nrow(viruses)
+```
+
+precision
+```{r}
+length(grep("true positive", viruses$confusion_matrix_all_2))/length(grep("positive", viruses$confusion_matrix_all_2))
+```
+
+recall
+```{r}
+length(grep("true positive", viruses$confusion_matrix_all_2))/(length(grep("true positive", viruses$confusion_matrix_all_2))+length(grep("false negative", viruses$confusion_matrix_all_2)))
+```
+
+visualizing accuracy by taxa
+```{r}
+accurate_by_taxa <- melt(table(viruses$accurate_all_2, viruses$seqtype, viruses$Index))
+colnames(accurate_by_taxa) <- c("accurate", "seqtype","Index", "count")
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
+ggplot(accurate_by_taxa, aes(y=count, x=accurate,
+                   fill=accurate,
+                   color=accurate)) +
+  geom_boxplot() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(2)), 0.5),
+                    labels=c("false", "true")) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(2)), 1),
+                    labels=c("false", "true")) +
+  xlab("Number of Sequences") +
+  ylab("") + 
+  facet_wrap(~seqtype, scales = "free")
 ```
@@
 
-  #by type metrics
-  fungal_FP <- table(confusion_matrix[seqtype=="fungi"])[2]
-  protist_FP <- table(confusion_matrix[seqtype=="protist"])[2]
-  bacterial_FP <- table(confusion_matrix[seqtype=="bacteria"])[2]
-  viral_FN <- table(confusion_matrix[seqtype=="virus"])[1]
-
-  performance <- c(precision, recall, F1, MCC, auc, fungal_FP, 
-                   protist_FP, bacterial_FP, viral_FN)
-  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC", "fungal_FP",
-                          "protist_FP", "bacterial_FP", "viral_FN")
+  performance <- c(precision, recall, F1, MCC, auc)
+  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC")
 
@@
                        MCC=rep(0, nrow(combos)),
-                       AUC=rep(0, nrow(combos)),
-                       fungal_FP=rep(0, nrow(combos)),
-                       protist_FP=rep(0, nrow(combos)),
-                       bacterial_FP=rep(0, nrow(combos)),
-                       viral_FN=rep(0, nrow(combos)))
+                       AUC=rep(0, nrow(combos)))
   for (i in 1:nrow(combos)) {
@@
 
-    output[i,1:9] <- assess_performance(input_seqs$seqtype, keep_score)
+    output[i,1:5] <- assess_performance(input_seqs$seqtype, keep_score)
 
@@
                        MCC=rep(0, nrow(combos_list)*10), 
-                      AUC=rep(0, nrow(combos_list)*10),
-                      fungal_FP=rep(0, nrow(combos_list)*10),
-                      protist_FP=rep(0, nrow(combos_list)*10),
-                      bacterial_FP=rep(0, nrow(combos_list)*10),
-                      viral_FN=rep(0, nrow(combos_list)*10))
+                      AUC=rep(0, nrow(combos_list)*10))
 
@@
 accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+```
+
+accounting for rules that have multiple tools in them
+```{r}
+accuracy_scores$numtools <- accuracy_scores$numrules
+
+for (i in 1:nrow(accuracy_scores)) {
+  toolcombo <- as.character(accuracy_scores$toolcombo[i])
+  if (substr(toolcombo, 1, 1)=="1") {
+    accuracy_scores$numtools[i] <- accuracy_scores$numtools[i] + 3
+  }
+
+  if (substr(toolcombo, 5, 5)=="1") {
+    accuracy_scores$numtools[i] <- accuracy_scores$numtools[i] + 4
+  }
+}
+
+accuracy_scores$numtools[accuracy_scores$numtools>6] <- 6
+```
+
+```{r}
 accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
+accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
 ```
@@
   xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("F1 Score")
+  ylab("F1 Score") +
+  scale_fill_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 1))
 p2
@@
   xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Precision")
+  ylab("Precision") +
+  scale_fill_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 1))
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
@@
   xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Recall")
+  ylab("Recall") +
+  scale_fill_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 1))
 
@@
   ylab("Recall") +
-  scale_fill_manual(name="",
+  scale_fill_manual(name="Number of Rules",
                      values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
+  scale_color_manual(name="Number of Rules",
                      values = alpha(rev(pal(6)), 1))
 
-ggplot(accuracy_scores[accuracy_scores$testing_set_index==1,], aes(x=precision, y=recall)) +
-  geom_label(alpha=0.7, label=accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1]) +
-  geom_point(alpha=0.5, aes(color=numrules, fill=numrules)) +
+ggplot(accuracy_scores, aes(x=precision, y=recall, 
+                                  color=numtools, fill=numtools)) +
+  geom_point(alpha=0.5) +
   theme_light() +
@@
     legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
+    axis.title=element_text(size=20),
   ) +
   xlab("Precision") +
-  ylab("Recall")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=abs(precision-recall), 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Precision-Recall")
+  ylab("Recall") +
+  scale_fill_manual(name="Number of Tools",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Tools",
+                     values = alpha(rev(pal(6)), 1))
+
 ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
@@
 
-ggplot(accuracy_scores, aes(x=toolcombo, y=AUC, 
-                                  color=numrules, fill=numrules)) +
+ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
+                                  color=numtools, fill=numtools)) +
   geom_point(alpha=0.5) +
@@
     legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
+    axis.title=element_text(size=20),
   ) +
   xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("AUC")
-ggplot(accuracy_scores, aes(x=toolcombo, y=fungal_FP, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Fungal False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=protist_FP, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Protist False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=bacterial_FP, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Bacterial False Positives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Viral False Negatives")
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=viral_FN, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Viral False Negatives")
-
-ggplot(accuracy_scores, aes(x=protist_FP, y=fungal_FP, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Protist FP") +
-  ylab("Fungal FP")
-
-ggplot(accuracy_scores, aes(x=recall, y=fungal_FP, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Recall") +
-  ylab("Fungal FP")
-
-ggplot(accuracy_scores, aes(x=recall, y=protist_FP, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Recall") +
-  ylab("Protist FP")
+  ylab("MCC") +
+  scale_fill_manual(name="Number of Tools",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Tools",
+                     values = alpha(rev(pal(6)), 1))
 ```
@@
 accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numrules, toolcombo) %>%
+  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
   pivot_longer(cols=c(precision, recall, MCC), 
@@
 
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="MCC",], aes(x=numrules, y=performance_metric_score, 
+                                  color=numtools, fill=numtools)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Rule Sets") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) 
 ```

### 2022-12-04 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Adding DVF as a tuning removal piece.

--- 2022-11-07 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
@@
 ```
+
+# Looking at Features
 
@@
        aes(axis1 = kj_cel, axis2 = hall, axis3 = pv, axis4 = cvl_pu, y=n)) +
-  geom_alluvium(aes(fill=viral_score),
+  ggalluvial::geom_alluvium(aes(fill=viral_score),
                 width = 0, knot.pos = 0, reverse = FALSE) +
-  geom_stratum(width = 1/5) +
+  ggalluvial::geom_stratum(width = 1/5) +
   theme_bw() +

### 2022-12-04 TestingPipeline/Scripts/comparing_to_VS2_SOP.Rmd
Testing rule set based on VS2 SOP.

---
title: "Comparing to VS2 SOP"
output: html_notebook
---

# VS2 SOP testing

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
library("stringr")
```

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r virsorter2_SOP}
virsorter2_SOP <- function(input_seqs) {
  keep_score <- rep(0, nrow(input_seqs))
  # Keep1: viral_gene >0
  keep_score[input_seqs$checkv_viral_genes > 0] <- 1

  # Keep2: viral_gene =0 AND (host_gene =0 OR score >=0.95 OR hallmark >2)
  keep_score[(input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))] <- 1

  # Manual check: (NOT in Keep1 OR Keep2) AND viral_gene =0 AND host_gene =1 AND length >=10kb
  keep_score[!(input_seqs$checkv_viral_genes > 0) & !((input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))) & (input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length > 10000)] <- 0.5

  # Discard: the rest
  # implied by starting with zero matrix, unchanged will stay zero

  return(keep_score)
}

```

```{r}
viruses$keep_score_virsorter2_SOP <- virsorter2_SOP(viruses)
```

```{r}
viruses$confusion_matrix_vs2_sop <- "true negative"
viruses$confusion_matrix_vs2_sop[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP<1] <- "false negative"
viruses$confusion_matrix_vs2_sop[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP>=1] <- "true positive"
viruses$confusion_matrix_vs2_sop[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP>=1] <- "false positive"

viruses$accurate_vs2_SOP <- "false"
viruses$accurate_vs2_SOP[grep("true", viruses$confusion_matrix_vs2_sop)] <- "true"
```

accuracy:
```{r}
length(grep("true", viruses$accurate_vs2_SOP))/nrow(viruses)
```

precision
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop))/length(grep("positive", viruses$confusion_matrix_vs2_sop))
```

recall
```{r}
length(grep("true positive", viruses$confusion_matrix_vs2_sop))/(length(grep("true positive", viruses$confusion_matrix_vs2_prsop))+length(grep("false negative", viruses$confusion_matrix_vs2_sop)))
```

### 2022-12-04 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Testing rule set based on VS2 SOP.

@@
   facet_wrap(~seqtype, scales = "free")
+```
+
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vb"), true_virus) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score,true_virus) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
+                            "VB+DVF+VS2+VS+addition", "VB+DVF+VS2+VS+addition-removal")) +
+  facet_grid(~true_virus, scales = "free")
 ```

### 2022-12-07 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
Rmd for figures in manuscript.

@@
 viruses$dsDNAphage[grep("nan", viruses$dsDNAphage)] <- 0
-viruses$viral[is.na(viruses$max_score)] <- 0
+viruses$max_score[is.na(viruses$max_score)] <- 0
 

### 2022-12-07 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Rmd for figures in manuscript.

@@
   xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Proportion of Contigs >3kb Viral")
+  ylab("Proportion of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
 ```
@@
 ```{r}
-physeq_pooled <- phyloseq(otu_table(num_viruses_uknl_wide, taxa_are_rows = F),
-                                     sample_data(tooldata))
+physeq_pooled <- phyloseq::phyloseq(phyloseq::otu_table(num_viruses_uknl_wide, taxa_are_rows = F),
+                                     phyloseq::sample_data(tooldata))
 ```

### 2022-12-07 TestingPipeline/Scripts/combining_tool_output.Rmd
Rmd for figures in manuscript.

@@
 viruses$dsDNAphage[is.na(viruses$dsDNAphage)] <- 0
-viruses$viral[is.na(viruses$max_score)] <- 0
+viruses$max_score[is.na(viruses$max_score)] <- 0
 

### 2022-12-07 TestingPipeline/Scripts/paper_figure_generation.Rmd
Rmd for figures in manuscript.

--- 2022-12-04 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
-author: Bridget Hegarty, James Riddell
-date: 07-22-2022
+title: "Pipeline Performance Figure Generation"
 output: html_notebook
 ---
+
+This R script creates all of the figures used to compare the performance across pipelines.
+
 This Rmarkdown file assesses the output of CheckV, DeepVirFinder, Kaiju,
@@
 ```
-(this is still the old rules)
-
-This section defines a viralness score "keep_score" based on the tool classifications. 
-A final keep_score above 1 indicates we will keep that sequence and call it viral.
-
-VIBRANT
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus: +0.5
-
-Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +0.5
-    RNA >= 0.9: +1
-    lavidaviridae >= 0.9: +1
-    NCLDV >= 0.9: +1
-
-Virsorter
-    category ==  1,4: +1
-    category == 2,5: +0.5
-
-DeepVirFinder:
-    Score >= 0.7: +0.5
-
-Tuning - No Viral Signature:
-    Kaiju_viral = "cellular organisms": -0.5
-    If host_genes >50 and NOT provirus: -1 
-    If viral_genes == 0 and host_genes >= 1: -1
-    If 3*viral_genes <= host_genes and NOT provirus: -1
-    If length > 50,000 and hallmark <=1: -1
-    If length < 5000 and checkv completeness <= 75: -0.5
-
-Tuning - Viral Signature:
-    Kaiju_viral = "Viruses": +0.5
-    If %unknown >= 75 and length < 50000: +0.5
-    If %viral >= 50: +0.5
-    Hallmark > 2: +0.5
-
-This script produces visualizations of these combined viral scorings and
-includes ecological metrics like alpha diversity.
-
-You can decide which combination is appropriate for them and only need use the
-tools appropriate for your data.
-
-# Trying new rule set
-this set is based on optimizing for the 
+
+Function that allows for the comparing pieces of the pipeline
 ```{r getting_viral_set_1}
@@
 }
-```
-
-## Trying with all of the rules
-```{r}
-viruses$keep_score_all_2 <- getting_viral_set_1(viruses,
-                                                include_vibrant=TRUE, 
-                                                include_virsorter2=TRUE,
-                                                include_deepvirfinder=TRUE,
-                                                include_tuning_viral=TRUE,
-                                                include_tuning_not_viral=TRUE,
-                                                include_virsorter=TRUE)
-
-viruses$confusion_matrix_all_2 <- "true negative"
-viruses$confusion_matrix_all_2[viruses$seqtype=="virus" & viruses$keep_score_all_2<1] <- "false negative"
-viruses$confusion_matrix_all_2[viruses$seqtype=="virus" & viruses$keep_score_all_2>=1] <- "true positive"
-viruses$confusion_matrix_all_2[viruses$seqtype!="virus" & viruses$keep_score_all_2>=1] <- "false positive"
-
-viruses$accurate_all_2 <- "false"
-viruses$accurate_all_2[grep("true", viruses$confusion_matrix_all_2)] <- "true"
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$accurate_all_2))/nrow(viruses)
-```
-
-precision
-```{r}
-length(grep("true positive", viruses$confusion_matrix_all_2))/length(grep("positive", viruses$confusion_matrix_all_2))
-```
-
-recall
-```{r}
-length(grep("true positive", viruses$confusion_matrix_all_2))/(length(grep("true positive", viruses$confusion_matrix_all_2))+length(grep("false negative", viruses$confusion_matrix_all_2)))
-```
-
-visualizing accuracy by taxa
-```{r}
-accurate_by_taxa <- melt(table(viruses$accurate_all_2, viruses$seqtype, viruses$Index))
-colnames(accurate_by_taxa) <- c("accurate", "seqtype","Index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(accurate_by_taxa, aes(y=count, x=accurate,
-                   fill=accurate,
-                   color=accurate)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(2)), 0.5),
-                    labels=c("false", "true")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(2)), 1),
-                    labels=c("false", "true")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free")
 ```
@@
 
-## Visualize how the precision, recall, and F1 scores change across pipelines.
-```{r}
-p2 <- ggplot(accuracy_scores, aes(x=toolcombo, y=F1, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("F1 Score") +
-  scale_fill_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 1))
-p2
+## Visualize how the precision, recall, and MCC changes across pipelines.
+precision
+```{r}
+accuracy_scores <- accuracy_scores[order(accuracy_scores$precision, decreasing=F),]
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
 
@@
                      values = alpha(rev(pal(6)), 1))
+```
+
+recall
+```{r}
+accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
 
@@
                      values = alpha(rev(pal(6)), 1))
-
+```
+
+precision vs recall
+```{r}
 ggplot(accuracy_scores, aes(x=precision, y=recall, 
@@
                      values = alpha(rev(pal(6)), 1))
+
+```
+MCC
+```{r}
+accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
 
@@
 ```{r}
-ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="MCC",], aes(x=numrules, y=performance_metric_score, 
-                                  color=numrules, fill=numrules)) +
-  geom_boxplot() +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("MCC") +
-  xlab("Number of Rule Sets") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1))
-
-ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="MCC",], aes(x=numrules, y=performance_metric_score, 
-                                  color=numtools, fill=numtools)) +
-  geom_boxplot() +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("MCC") +
-  xlab("Number of Rule Sets") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) 
-```
-
-```{r}
-ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="recall",], aes(x=numrules, y=performance_metric_score, 
-                                  color=numrules, fill=numrules)) +
-  geom_boxplot() +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("recall") +
-  xlab("Number of Rule Sets") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1))
-
-```
-
-```{r}
 ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
@@
 
-```
-comparing metric with and without tuning rules
-```{r}
-accuracy_scores_melt$tuning_inc <- "no"
-accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1] <- "tv"
-accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "tnv"
-accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1 &
-                                  substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "both"
-```
-
-```{r}
-ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
+                                  color=numtools, fill=numtools)) +
   geom_boxplot() +
-  geom_point(aes(color=numrules, fill=numrules), alpha=0.5) +
+  geom_point(alpha=0.5) +
   theme_light() +
@@
   ) +
-  ylab("MCC") +
-  xlab("Number of Tools") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
+  ylab("Score") +
+  xlab("Number of Rules") +
+  scale_fill_manual(name="Number of Tools",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Tools",
                      values = alpha(rev(pal(6)), 1)) +
@@
 
-ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score,
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numtools, y=performance_metric_score, 
                                   color=numrules, fill=numrules)) +
@@
   ) +
-  ylab("MCC") +
+  ylab("Score") +
   xlab("Number of Tools") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
+  scale_fill_manual(name="Number of Rules",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="Number of Rules",
                      values = alpha(rev(pal(6)), 1)) +
@@
 
-ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=tuning_inc, y=performance_metric_score)) +
-  geom_boxplot() +
-  geom_boxplot(aes(color=numrules, fill=numrules)) +
- # geom_point(aes(color=numrules, fill=numrules), alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("Score") +
-  xlab("") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.3)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 0.7)) +
-  facet_wrap(~performance_metric)
-
-```
-
-```{r}
-write_tsv(accuracy_scores, "20221029_accuracy_scores.tsv")
-```
-
-to do: add in clustering and ordination like in the drinking water R notebook
-
-# Experimenting
-
-## high precision example
-```{r}
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses$confusion_matrix_high_precision <- "true negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<1] <- "false negative"
-viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision>=1] <- "true positive"
-viruses$confusion_matrix_high_precision[viruses$seqtype!="virus" & viruses$keep_score_high_precision>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_precision, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-length(grep("true", viruses$confusion_matrix_high_precision))/nrow(viruses)
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(y=count, x=confusion_matrix,
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free")
-```
-this rule set had the highest precision, but as you can see, this comes with a big sacrifice in recall
-
-## high MCC example
-```{r}
-viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+```
+
+## differences based on genome size
+```{r}
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
                                               include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_high_MCC <- "true negative"
-viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
-viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
-viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$confusion_matrix_high_MCC))/nrow(viruses)
-```
-
-recall
-```{r}
-length(grep("true positive", viruses$confusion_matrix_high_MCC))/length(grep("virus", viruses$seqtype))
-```
-
-```{r}
-TP <- table(viruses$confusion_matrix_high_MCC)[4]
-FP <- table(viruses$confusion_matrix_high_MCC)[2]
-TN <- table(viruses$confusion_matrix_high_MCC)[3]
-FN <- table(viruses$confusion_matrix_high_MCC)[1]
-
-precision <- as.numeric(TP/(TP+FP))
-precision
-recall <- as.numeric(TP/(TP+FN))
-recall
-F1 <- as.numeric(2*precision*recall/(precision+recall))
-F1
-
-MCC <- as.numeric((TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN)))
-MCC
-```
-
-precision=69%, recall=87%, MCC=77%
-
-precision adjusting size to be equal viral/not viral
-```{r}
-TP <- table(viruses$confusion_matrix_high_MCC)[4]
-FP <- table(viruses$confusion_matrix_high_MCC)[2]*.11
-TN <- table(viruses$confusion_matrix_high_MCC)[3]*.11
-FN <- table(viruses$confusion_matrix_high_MCC)[1]
-
-precision <- as.numeric(TP/(TP+FP))
-precision
-recall <- as.numeric(TP/(TP+FN))
-recall
-F1 <- as.numeric(2*precision*recall/(precision+recall))
-F1
-
-MCC <- as.numeric((TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN)))
-MCC
-```
-
-precision=0.95, recall=0.87, F1=0.91, MCC=0.82
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, Index)
-
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-differences based on genome size
-```{r}
+
+viruses$confusion_matrix_all <- "true negative"
+viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
+viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
+viruses$confusion_matrix_all[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
+
 viruses$size_class <- "3-5kb"
@@
 viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
-```
-
-```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, size_class, Index)
+
+confusion_by_taxa <- viruses %>% count(confusion_matrix_all, seqtype, size_class, Index)
 
 colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "index", "count")
-```
-
-```{r}
+
 confusion_vir_called <- confusion_by_taxa %>% filter(confusion_matrix=="true positive" | confusion_matrix=="false positive")
@@
 
+## visualizing a select set of tools
+```{r}
+viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+```
+
+```{r}
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_"), size_class, true_virus) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, true_virus) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~true_virus, scales = "free")
+```
+### Effect of combining tools
 ```{r}
@@
                                               include_virsorter = T)
-```
-
-```{r}
+
 viruses$true_virus <- "not"
@@
 ```{r}
-viruses_long_scores_addition <- viruses_long_scores[(viruses_long_scores$true_virus=="virus" &  (viruses_long_scores$rule_combination!="keep_score_vb_dvf_vs2_vs_tv_tnv") & viruses_long_scores$viral_score!="0"),]
-ggplot(viruses_long_scores_addition, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_brewer(palette = "Purples") +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
-                            "VB+DVF+VS2+VS+addition")) +
-  facet_grid(~true_virus, scales = "free")
-```
-
-```{r}
 ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
@@
 ```
-
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_brewer(palette = "PuOr", ) +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  facet_grid(size_class~true_virus, scales = "free")
-```
-
-# Considering how each method contributes to the final prediction
-```{r}
-viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
-viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
-                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
-                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "viral_score")
-```
-
-```{r}
-sm_m <- sm_m[sm_m$viral_score>0,]
-
-sm_m$score <- sm_m$viral_score
-
-sm_m$score[sm_m$viral_score==0.5] <- "0.5"
-sm_m$score[sm_m$viral_score>=1] <- "1"
-sm_m$score[sm_m$viral_score>=2] <- "2"
-sm_m$score[sm_m$viral_score>=3] <- "3"
-sm_m$score[sm_m$viral_score>=4] <- "4"
-sm_m$score[sm_m$viral_score>=5] <- "5"
-
-sm_m$score <- factor(sm_m$score, 
-                                       levels=c("0.5", "1", "2","3","4","5"))
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  xlab("") +
-  ylab("Number of Sequences") +
-  coord_flip()
-```
-## Another way of visualizing between rule sets
-
-```{r}
-viruses_mcc_alluvial <- data.frame(seqtype=viruses$seqtype,
-                                   keep_score_high_MCC=viruses$keep_score_high_MCC,
-                                   confusion_matrix_high_MCC=viruses$confusion_matrix_high_MCC)
-
-viruses_mcc_alluvial$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses_mcc_alluvial$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses_mcc_alluvial$keep_score_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses_mcc_alluvial$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses_mcc_alluvial$keep_score_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses_mcc_alluvial$keep_score_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses_mcc_alluvial %>%
-  count(seqtype, keep_score_high_MCC) %>% 
-  spread(key = keep_score_high_MCC, value=n)
-```
-
-```{r}
-viruses_mcc_alluvial <- viruses_mcc_alluvial %>%
-  count(seqtype, keep_score_dvf, keep_score_vb, keep_score_vs,
-        keep_score_vs2, keep_score_tv, keep_score_tnv, keep_score_high_MCC) %>%
-  mutate(high_mcc_viral_score=factor(round(keep_score_high_MCC)))
-```
-
-```{r}
-ggplot(viruses_mcc_alluvial,
-       aes(axis1 = keep_score_dvf, axis2 = keep_score_vb, 
-           axis3 = keep_score_vs, axis4 = keep_score_vs2, 
-           axis5 = keep_score_tv, axis6 = keep_score_tnv, 
-           y=n)) +
-  geom_alluvium(aes(fill=high_mcc_viral_score),
-                width = 0, knot.pos = 0, reverse = FALSE) +
-  geom_stratum(width = 1/5) +
-  theme_bw() +
-  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
-            reverse = FALSE) +
-  theme(
-        axis.text.x=element_text(size=14, angle = 90)
-        ) +
-  scale_x_continuous(breaks=c(1,2,3,4,5,6),
-    labels=c("dvf", "kj", "vs", "vs2",
-             "tv", "tnv")) +
-  scale_fill_brewer(palette = "PuOr", ) +
-  facet_wrap(~seqtype, scales="free_y") 
-```
-
-## Visualizing confusion matrix by number of tools
-
-```{r}
-viruses$keep_score_visualize <- viruses$keep_score_high_MCC
-viruses$keep_score_visualize[viruses$keep_score_high_MCC>1] <- "> 1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==1] <- "1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==0.5] <- "0.5"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==0] <- "0"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==-0.5] <- "-0.5"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC==-1] <- "-1"
-viruses$keep_score_visualize[viruses$keep_score_high_MCC<=-1] <- "< -1"
-
-viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
-#viruses$keep_score_visualize <- factor(viruses$keep_score_visualize, 
-#                                       labels=c("≤ 0", "≤ 0", "≤ 0", "0.5","1", "> 1"))
-```
-
-```{r}
-levels(factor(viruses$keep_score_visualize))
-```
-
-```{r}
-ggplot(viruses, aes(x=as.factor(Index),
-                   fill=keep_score_visualize, color=keep_score_visualize)) +
-  geom_bar(stat="count", position="stack") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16)
-  ) +
-  scale_color_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 0.5)) +
-  xlab("Index") +
-  ylab("Sequence Count") +
-  facet_wrap(~confusion_matrix_high_MCC, scales = "free")
-
-```
-
-## clustering
-```{r}
-viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
-num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
-                          num_viruses=rep(0, nrow(combos_list)))
-
-for (i in 1:nrow(combos_list)) {
-  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
-                                            include_virsorter = combos_list$VS[i],
-                                            include_virsorter2 = combos_list$VS2[i],
-                                            include_tuning_viral = combos_list$tune_viral[i],
-                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
-                                            include_deepvirfinder = combos_list$DVF[i])
-
-  if (max(viral_scores[,i])<=0) {
-    num_viruses$num_viruses[i] <- 0
-  }
-  else {
-    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
-  }
-
-  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
-
-  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
-}
-
-num_viruses$numrules <- str_count(num_viruses$toolcombo, "1")
-num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
-num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
-num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
-num_viruses$numrules <- as.factor(num_viruses$numrules)
-```
-
-```{r}
-viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
-viral_scores_nozeros <- viral_scores_nozeros + 1
-viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
-
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo
-```
-
-```{r}
-library(phyloseq)
-```
-
-```{r}
-tooldata <- num_viruses
-
-rownames(tooldata) <- tooldata$toolcombo
-```
-
-```{r}
-physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
-                                     sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numrules", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  geom_label(label=tooldata$toolcombo)
-
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numrules", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw()
-```
-
-```{r}
-bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
-clusters <- hclust(dist(bray_dist))
-plot(clusters)
-
-myclusters <- cutree(clusters, h=0.8)
-```
-
-```{r}
-#names(myclusters[myclusters==1])
-#names(myclusters[myclusters==2])
-#names(myclusters[myclusters==3])
-#names(myclusters[myclusters==4])
-#names(myclusters[myclusters==5])
-
-myclusters_df <- tibble(combo=names(myclusters),
-                            cluster_index=myclusters)
-
-myclusters_df <- separate(myclusters_df, col=combo, into=c("tnv", "DVF",
-                                                            "tv", "VB",
-                                                            "VS", "VS2"),
-                          sep=" ", remove = F)
-
-tool_count <- as.data.frame(rbind(table(myclusters_df$tnv, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$tv, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VB, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VS, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VS2, myclusters_df$cluster_index)[2,])
-                    )
-
-tool_count <- data.frame(t(apply(tool_count, c(1), function(x) {x <- x/table(myclusters_df$cluster_index)})))
-
-tool_count$method <- c("tnv", "DVF", "tv", "VB", "VS", "VS2")
-
-tool_count <- melt(tool_count)
-
-colnames(tool_count) <- c("tool", "cluster_index", "tool_count_norm")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
-ggplot(tool_count, aes(x=tool, y=tool_count_norm,
-                   fill=tool,
-                   color=tool)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "none",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    #legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Tool") +
-  ylab("Proportion of Times in Cluster") + 
-  facet_wrap(~cluster_index, nrow=1)
-```
-
-```{r}
-accuracy_scores_melt <- accuracy_scores %>% 
-  select(precision, recall, MCC, numrules, toolcombo) %>%
-  group_by(numrules, toolcombo) %>%
-  summarise(precision=mean(precision),
-            recall=mean(recall),
-            MCC=mean(MCC)) %>%
-  pivot_longer(cols=c(precision, recall, MCC), 
-               names_to="performance_metric",
-               values_to="performance_metric_score")
-```
-
-```{r}
-myclusters_df <- inner_join(accuracy_scores_melt, myclusters_df, 
-                            by=c("toolcombo"="combo"))
-
-myclusters_df$cluster_index <- as.factor(myclusters_df$cluster_index)
-```
-
-```{r}
-ggplot(myclusters_df, aes(x=cluster_index, y=performance_metric_score, 
-                                  color=cluster_index, fill=cluster_index)) +
-  geom_boxplot() +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("MCC") +
-  xlab("Cluster") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(9)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(9)), 1)) +
-  facet_wrap(~performance_metric)
-```
-
-## all 6 tools example
-```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_all <- "true negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
-viruses$confusion_matrix_all[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
-```
-
-```{r}
-TP <- table(viruses$confusion_matrix_all)[4]
-FP <- table(viruses$confusion_matrix_all)[2]
-TN <- table(viruses$confusion_matrix_all)[3]
-FN <- table(viruses$confusion_matrix_all)[1]
-
-precision <- TP/(TP+FP)
-precision
-recall <- TP/(TP+FN)
-recall
-F1 <- 2*precision*recall/(precision+recall)
-F1
-
-MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
-MCC
-```
-precision=62%, recall=92%, MCC=73%
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_all, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-table(viruses$confusion_matrix_all)
-
-length(grep("true", viruses$confusion_matrix_all))/nrow(viruses)
-```
-
-```{r}
-length(grep("true positive", viruses$confusion_matrix_all))/length(grep("virus", viruses$seqtype))
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-## high recall example
-```{r}
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-```
-
-```{r}
-viruses$confusion_matrix_high_recall <- "true negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall<1] <- "false negative"
-viruses$confusion_matrix_high_recall[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
-viruses$confusion_matrix_high_recall[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
-```
-
-visualizing confusion matrix by taxa
-```{r}
-confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_recall, viruses$seqtype, viruses$Index))
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
-                   fill=confusion_matrix,
-                   color=confusion_matrix)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-p2
-```
-
-accuracy:
-```{r}
-length(grep("true", viruses$confusion_matrix_high_recall))/nrow(viruses)
-```
-0.887
-
-recall
-```{r}
-length(grep("true positive", viruses$confusion_matrix_high_recall))/length(grep("virus", viruses$seqtype))
-```
-recover almost all of the viruses this way, but more protist contamination
-
-0.960
-
-```{r}
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_recall, seqtype, size_class)
-
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "count")
-```
-
-## few tools, high MCC example
-```{r}
-viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses$confusion_matrix_few_tools <- "true negative"
-viruses$confusion_matrix_few_tools[viruses$seqtype=="virus" & viruses$keep_score_few_tools<1] <- "false negative"
-viruses$confusion_matrix_few_tools[viruses$seqtype=="virus" & viruses$keep_score_few_tools>=1] <- "true positive"
-viruses$confusion_matrix_few_tools[viruses$seqtype!="virus" & viruses$keep_score_few_tools>=1] <- "false positive"
-```
-
-```{r}
-TP <- table(viruses$confusion_matrix_few_tools)[4]
-FP <- table(viruses$confusion_matrix_few_tools)[2]
-TN <- table(viruses$confusion_matrix_few_tools)[3]
-FN <- table(viruses$confusion_matrix_few_tools)[1]
-
-precision <- TP/(TP+FP)
-precision
-recall <- TP/(TP+FN)
-recall
-F1 <- 2*precision*recall/(precision+recall)
-F1
-
-MCC <- (TP*TN-FP*FN)/sqrt(as.numeric(TP+FP)*as.numeric(TP+FN)*as.numeric(TN+FP)*as.numeric(TN+FN))
-MCC
-```
-precision=77%, recall=76%, MCC=74%
-
-## Visualizing the different sets
-
-```{r}
-confusion_by_taxa_method <- viruses %>% 
-  select(contains("confusion_matrix"), seqtype, Index) %>%
-  pivot_longer(cols=contains("confusion_matrix"), 
-               names_to="confusion_matrix_type",
-               values_to="confusion_matrix_value") %>%
-  count(seqtype, Index, confusion_matrix_type, confusion_matrix_value)
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-ggplot(confusion_by_taxa_method, aes(y=n, x=confusion_matrix_type,
-                   fill=confusion_matrix_value,
-                   color=confusion_matrix_value)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free")
-```
-
-```{r}
-confusion_by_taxa_method <- viruses %>% 
-  select(contains("confusion_matrix"), seqtype, Index) %>%
-  pivot_longer(cols=contains("confusion_matrix"), 
-               names_to="confusion_matrix_type",
-               values_to="confusion_matrix_value") %>%
-  count(seqtype, Index, confusion_matrix_type, confusion_matrix_value) %>%
-  filter(grepl("true", confusion_matrix_value)) %>%
-  mutate(confusion_matrix_type=sub("confusion_matrix_", "", confusion_matrix_type))
-```
-
-```{r}
-ggplot(confusion_by_taxa_method, aes(y=n, x=confusion_matrix_type,
-                   fill=confusion_matrix_value,
-                   color=confusion_matrix_value)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=10, angle=90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4))[3:4], 0.5),
-                    labels=c( 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4))[3:4], 1),
-                    labels=c(
-                             "true negative", "true positive")) +
-  xlab("Tool Set") +
-  ylab("Contig Count") + 
-  facet_wrap(~seqtype, scales = "free")
-```
-
-## another way of visualizing the different tool sets scores
-```{r}
-viruses$true_virus <- "not"
-viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
-
-viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_high"), contains("keep_score_all"), size_class, true_virus) %>%
-  pivot_longer(cols=contains("keep_score_"), 
-               names_to="rule_combination",
-               values_to="viral_score") %>% 
-  mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, true_virus) %>%
-  summarise(n = n())
-
-viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
-```
-
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_brewer(palette = "PuOr", ) +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  facet_grid(~true_virus, scales = "free")
-```
-
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_brewer(palette = "PuOr", ) +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  facet_grid(size_class~true_virus, scales = "free")
-```
-
-######################################################################
-Extra Stuff
-#####################################################################
-
-```{r}
-ggplot(viruses, aes(x=checkv_length, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_MCC,
-                   color=confusion_matrix_high_MCC)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=hallmark,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Number of Hallmark Genes") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=checkv_completeness, y=keep_score_high_MCC,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_point(stat="identity", shape=21) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("CheckV Completeness") +
-  ylab("Pipeline Viral Score") + 
-  facet_wrap(~seqtype) + 
-  scale_x_log10()
-```
-
-```{r}
-ggplot(viruses, aes(x=confusion_matrix_high_recall, y=checkv_length,
-                   fill=confusion_matrix_high_recall,
-                   color=confusion_matrix_high_recall)) +
-  geom_boxplot() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Sequence Length (bp)") +
-  ylab("Pipeline Viral Score") +
-  scale_y_log10()
-```
-
-looking at false negatives
-```{r}
-viruses_false_negs <- viruses[(viruses$seqtype=="virus" & viruses$keep_score_high_recall<1),]
-```
-
-looking at protists calling viral
-```{r}
-viruses_false_pos_protists <- viruses[(viruses$seqtype=="protist" & viruses$keep_score_high_recall>=1),]
-```
-
-# Considering how each method contributes to the final prediction (high MCC)
-
-```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-```
-
-```{r}
-viruses_high <- viruses[viruses$keep_score_vb_tv>=1,] #uncomment this line if want to use all 6 tools
-viruses_high_mod <- viruses_high %>% select(keep_score_vb, 
-                                            keep_score_vb_tv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
-```
-
-```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "score")
-```
-
-```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=as.factor(score))) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(14)), 1)) +
-  xlab("Primary Method") +
-  ylab("Count of Viral Contigs") +
-  coord_flip()
-```
-
-# ROC 
-```{r}
-library(pROC)
-```
-
-```{r}
-viruses$truepositive <- rep(0, nrow(viruses))
-viruses$truepositive[viruses$seqtype=="virus"] <- 1
-```
-
-```{r}
-rocobj <- roc(viruses$truepositive, viruses$keep_score)
-rocobj_all <- roc(viruses$truepositive, viruses$keep_score_all)
-auc <- round(auc(viruses$truepositive, viruses$keep_score),4)
-auc_all <- round(auc(viruses$truepositive, viruses$keep_score_all),4)
-#create ROC plot
-ggroc(rocobj, colour = 'steelblue', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')')) +
-  coord_equal()
-ggroc(rocobj_all, colour = 'green', size = 2) +
-  ggtitle(paste0('ROC Curve ', '(AUC = ', auc_all, ')'))
-```
-Sensitivity: The probability that the model predicts a positive outcome for an observation when indeed the outcome is positive.
-Specificity: The probability that the model predicts a negative outcome for an observation when indeed the outcome is negative.
-
-# Comparing behavior of all testing sets combined (clustering analyses)
-
-```{r}
-viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
-num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
-                          num_viruses=rep(0, nrow(combos_list)))
-
-for (i in 1:nrow(combos_list)) {
-  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
-                                            include_virsorter = combos_list$VS[i],
-                                            include_virsorter2 = combos_list$VS2[i],
-                                            include_tuning = combos_list$CheckV[i],
-                                            include_kaiju = combos_list$Kaiju[i],
-                                            include_deepvirfinder = combos_list$DVF[i])
-
-  num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
-
-  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
-
-  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
-}
-
-num_viruses$numrules <- str_count(num_viruses$toolcombo, "1")
-num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
-num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
-num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
-num_viruses$numrules <- as.factor(num_viruses$numrules)
-```
-
-```{r}
-ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
-                                  color=numrules, fill=numrules)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-
-ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numrules, fill=numrules)) +
-  geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (CV, DVF, KJ, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-ggplot(num_viruses, aes(x=numrules, y=num_viruses)) +
-  geom_boxplot(aes(color=numrules)) +
-  geom_point(aes(color=numrules, fill=numrules)) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Tools") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
-viral_scores_nozeros <- viral_scores_nozeros + 1
-viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
-
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
-```
-
-```{r}
-library(phyloseq)
-```
-
-```{r}
-tooldata <- num_viruses
-
-rownames(tooldata) <- tooldata$toolcombo2
-```
-
-```{r}
-physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
-                                     sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numrules", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  geom_label(label=tooldata$toolcombo)
-
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numrules", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw()
-```
-to do: try coloring above based on the F1 scores of the testing set on each combination
-
-```{r}
-bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
-clusters <- hclust(dist(bray_dist))
-plot(clusters)
-
-myclusters <- cutree(clusters, h=1.1)
-```
-
-```{r}
-names(myclusters[myclusters==1])
-names(myclusters[myclusters==2])
-names(myclusters[myclusters==3])
-names(myclusters[myclusters==4])
-names(myclusters[myclusters==5])
-
-myclusters_df <- tibble(combo=names(myclusters),
-                            cluster_index=myclusters)
-
-myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
-                                                            "Kaiju", "VIBRANT",
-                                                            "VirSorter", "VirSorter2"),
-                          sep=" ", remove = F)
-
-tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
-                    )
-
-tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
-
-tool_count <- melt(tool_count)
-
-colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
-ggplot(tool_count, aes(x=cluster_index, y=tool_count,
-                   fill=cluster_index,
-                   color=cluster_index)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Cluster") +
-  ylab("Number of Times in Cluster") + 
-  facet_wrap(~tool, scales = "free")
-```
-
-```{r}
- ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Viral Sequences") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
- ggplot(viruses, aes(x=percent_viral, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Percent Genes Viral") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
- ggplot(viruses, aes(x=hallmark, y=confusion_matrix_high_precision,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_boxplot(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=confusion_matrix_high_precision,
-                   color=confusion_matrix_high_precision)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(4)), 0.5),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(4)), 1),
-                    labels=c("false negative", "false positive", 
-                             "true negative", "true positive")) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-viruses_false_positive <- viruses[viruses$confusion_matrix_high_precision=="false positive",]
-viruses_false_negative <- viruses[viruses$confusion_matrix_high_precision=="false negative",]
-```
-
-```{r}
-ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
-                   fill=checkv_length,
-                   color=checkv_length,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Number of Viral Genes") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~seqtype, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="bacteria"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="fungi"], aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_positive[viruses_false_positive$seqtype=="protist"], aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=checkv_viral_genes,
-                   color=checkv_viral_genes,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-
-ggplot(viruses_false_negative, aes(x=hallmark, y=checkv_length,
-                   fill=keep_score_high_precision,
-                   color=keep_score_high_precision,
-                   shape=checkv_provirus)) +
-  geom_point(alpha=0.3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Number of Hallmark Genes") +
-  ylab("Contig Length") + 
-  facet_wrap(~Kaiju_Viral, scales = "free") +
-  coord_flip()
-```
-
-```{r}
-table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)
-
-table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
-```
-
-## Visualizing confusion matrix by number of tools
-
-```{r}
-confusion_by_keep_score <- viruses %>% 
-  select(contains("keep_score"), seqtype, Index) %>%
-  pivot_longer(cols=contains("keep_score"), 
-               names_to="pipeline_type",
-               values_to="pipeline_value") %>%
-  mutate(pipeline_type=sub("keep_score_", "", pipeline_type)) %>%
-  count(seqtype, Index, pipeline_type, pipeline_value)
-```
-
-```{r}
-confusion_by_keep_score$confusion_matrix <- "true negative"
-confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype=="virus" & confusion_by_keep_score$pipeline_value<1] <- "false negative"
-confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype=="virus" & confusion_by_keep_score$pipeline_value>=1] <- "true positive"
-confusion_by_keep_score$confusion_matrix[confusion_by_keep_score$seqtype!="virus" & confusion_by_keep_score$pipeline_value>=1] <- "false positive"
-```
-
-```{r}
-confusion_by_keep_score$keep_score_visualize <- confusion_by_keep_score$pipeline_value
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value>1] <- "> 1"
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==1] <- "1"
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==0.5] <- "0.5"
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==0] <- "0"
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==-0.5] <- "-0.5"
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value==-1] <- "-1"
-confusion_by_keep_score$keep_score_visualize[confusion_by_keep_score$pipeline_value<=-1] <- "< -1"
-
-confusion_by_keep_score$keep_score_visualize <- factor(confusion_by_keep_score$keep_score_visualize, 
-                                       levels=c("< -1", "-1", "-0.5", "0", "0.5","1", "> 1"))
-```
-
-```{r}
-ggplot(confusion_by_keep_score, aes(x=confusion_matrix, y=n,
-                   fill=keep_score_visualize, color=keep_score_visualize)) +
-  geom_boxplot() +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16)
-  ) +
-  scale_color_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 1)) +
-  scale_fill_manual(name = 'Viral Score',
-                     values = alpha(c(viridis(6)), 0.5)) +
-  xlab("Index") +
-  ylab("Sequence Count") +
-  facet_wrap(~pipeline_type, scales = "free")
-
-```

### 2022-12-07 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Rmd for figures in manuscript.

@@
 ---
-title: "Viral Sequence Sorting Tools Evaluation"
+title: "Viral Sequence Sorting Tools Rule Tuning"
 author: Bridget Hegarty, James Riddell

### 2022-12-08 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Most updated rules.

@@
   if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
-    keep_score[input_seqs$RNA>=0.9] <- keep_score[input_seqs$RNA>=0.9] + 1
-    keep_score[input_seqs$lavidaviridae>=0.9] <- keep_score[input_seqs$lavidaviridae>=0.9] + 1
-    keep_score[input_seqs$NCLDV>=0.9] <- keep_score[input_seqs$NCLDV>=0.9] + 1
+    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
   }
@@
   if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
-    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
+     # add if DVF calls viral
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
   }
@@
   if (include_tuning_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
+  # tuning addition
     keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
+    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
     keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
+    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
+    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5
   }
@@
   if (include_tuning_not_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
+    # tuning removal 
+    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
     keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 # consider accounting for RNA viruses
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
+    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
+    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
+    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
+    # remove if DVF calls it not viral
+    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
+
   }
@@
 
+# overlap in proportion of contigs for tool sets
+
+get number of tools and number of rules for each combination
+```{r}
+ann_col$numrules <- rowSums(ann_col[,1:6])
+ann_col$numtools <- ann_col$numrules
+
+for (i in 1:nrow(ann_col)) {
+  if (ann_col$tv[i]==1) {
+    ann_col$numtools[i] <- ann_col$numtools[i] + 3
+  }
+
+  if (ann_col$tnv[i]==1) {
+    ann_col$numtools[i] <- ann_col$numtools[i] + 5
+  }
+}
+
+ann_col$numtools[ann_col$numtools>6] <- 6
+
+ann_col$tool <- rownames(ann_col)
+```
+
+```{r}
+prop_high_viruses_long <- as.data.frame(prop_high_viruses)
+prop_high_viruses_long$tool <- combos_list$toolcombo
+
+prop_high_viruses_long <- as.data.frame(prop_high_viruses_long) %>% 
+  pivot_longer(!tool,
+               values_to="proportion",
+               names_to="tool2") %>%
+  inner_join(ann_col, by=c("tool"))
+
+```
+
+```{r}
+head(prop_high_viruses_long)
+```
+
+```{r}
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 0.7))
+```
+
+```{r}
+table(prop_high_viruses_long$numrules[prop_high_viruses_long$dvf==1])
+
+```
+
+```{r}
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numtools), y=proportion, 
+                                  color=as.factor(numtools), fill=as.factor(numtools))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numtools), y=proportion, 
+                                  color=as.factor(numtools), fill=as.factor(numtools))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numtools), y=proportion, 
+                                  color=as.factor(numtools), fill=as.factor(numtools))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numtools), y=proportion, 
+                                  color=as.factor(numtools), fill=as.factor(numtools))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+```
+rather than tool against all others, ones against ones, twos against twos...
+
+```{r}
+prop_high_viruses_long$numrules2 <- str_count(prop_high_viruses_long$tool, "1")
+
+# accounting for rules that have multiple tools in them
+prop_high_viruses_long$numtools2 <- prop_high_viruses_long$numrules2
+
+prop_high_viruses_long$numtoolscommon <- 0
+
+for (i in 1:nrow(prop_high_viruses_long)) {
+  toolcombo <- as.character(prop_high_viruses_long$tool[i])
+  if (substr(toolcombo, 1, 1)=="1") {
+    prop_high_viruses_long$numtools2[i] <- prop_high_viruses_long$numtools2[i] + 3
+  }
+
+  if (substr(toolcombo, 5, 5)=="1") {
+    prop_high_viruses_long$numtools2[i] <- prop_high_viruses_long$numtools2[i] + 4
+  }
+
+  if (prop_high_viruses_long$numtools2[i]==prop_high_viruses_long$numrules2[i]) {
+    prop_high_viruses_long$numtoolscommon[i] <- prop_high_viruses_long$numtools2[i] 
+  }
+
+}
+
+# can't actually have more than 6 tools
+prop_high_viruses_long$numtools2[prop_high_viruses_long$numtools2>6] <- 6
+
+prop_high_viruses_long$numrules2 <- as.factor(prop_high_viruses_long$numrules2)
+prop_high_viruses_long$numtools2 <- as.factor(prop_high_viruses_long$numtools2)
+prop_high_viruses_long$numtoolscommon <- as.factor(prop_high_viruses_long$numtoolscommon)
+```
+
+```{r}
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
+                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
+                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
+                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numtoolscommon), y=proportion, 
+                                  color=as.factor(numtoolscommon), fill=as.factor(numtoolscommon))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Tools") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(pal(6), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(pal(6), 0.7))
+```
+
+# distribution of proportion in common for each pair
+
 ```{r}
@@
 ```{r}
-hist(prop_high_viruses_long$proportion,
-     )
+hist(prop_high_viruses_long$proportion)
 ```

### 2022-12-08 TestingPipeline/Scripts/paper_figure_generation.Rmd
Most updated rules.

@@
   if (include_virsorter2) {
-    keep_score[input_seqs$max_score>=50] <- keep_score[input_seqs$max_score>=50] + 0.5
-    keep_score[input_seqs$max_score>=95] <- keep_score[input_seqs$max_score>=95] + 0.5  
+    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
   }
@@
   if (include_deepvirfinder) {
-    # remove if DVF calls it not viral
-    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
      # add if DVF calls viral
-    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
-    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
   }
@@
     keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
+    # remove if DVF calls it not viral
+    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
+
   }
@@
 ```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
                                               include_virsorter2 = T,
@@
                                               include_tuning_not_viral = T,
-                                              include_virsorter = T)
+                                              include_virsorter = F)
 
@@
 ```{r}
-viruses$keep_score_few_tools <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
                                               include_virsorter = T)
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
                                               include_virsorter2 = T,
@@
 viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
+                                              include_vibrant = F,
                                               include_virsorter2 = T,
@@
                                               include_tuning_not_viral = T,
-                                              include_virsorter = T)
+                                              include_virsorter = F)
 ```
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_"), size_class, true_virus) %>%
+  select(contains("keep_score_high"), size_class, true_virus) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
 ```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
+viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
                                               include_virsorter2 = F,
@@
 
-viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_dvf_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_dvf_vs_vb <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
                                               include_tuning_not_viral = F,
-                                              include_virsorter = F)
+                                              include_virsorter = T)
 
-viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_dvf_vs_vb_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
 
-viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_dvf_vs_vb_vs2_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
 
-viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_dvf_vs_vb_vs2_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_vb"), size_class, true_virus) %>%
+  select(contains("keep_score_dvf"), size_class, true_virus) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
   ylab("Number of Sequences") + 
-  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
-                            "VB+DVF+VS2+VS+addition", "VB+DVF+VS2+VS+addition-removal")) +
+  scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
+                            "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) +
   facet_grid(~true_virus, scales = "free")

### 2022-12-12 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Improved visualizations.

@@
 num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
 
@@
     num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+
+    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
   }
@@
 colnames(num_high_viruses) <- combos_list$toolcombo
-
-prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
-
-for (i in 1:63) {
-  for (j in 1:63) {
-    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
-    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
-  }
-}
 
@@
 
+```{r}
+viral_scores_two_comb
+```
+
+comparing VS2 and VIBRANT viruses
+```{r}
+vs2_vb <- data.frame(vs2=viral_scores_two[,1],
+                     vb=viral_scores_two[,3],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,3],
+                     comb="neither",
+                     seqtype=viruses$seqtype)
+
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb>=1] <- "either"
+vs2_vb$comb[vs2_vb$vs2==0.5 & vs2_vb$vb==0.5] <- "together"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0] <- "vs2_only"
+vs2_vb$comb[vs2_vb$vs2==0 & vs2_vb$vb>=1] <- "vb_only"
+
+#vs2_vb$true_virus <- "not"
+#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"
+
+vs2_vb_seqtype <- vs2_vb %>% 
+  group_by(comb, comb_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+table(vs2_vb$comb)
+```
+
+```{r}
+ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
+comparing rule set individually
+```{r}
+ind <- data.frame(vs2_scores=viral_scores_two[,1],
+                  vs_scores=viral_scores_two[,2],
+                  vb_scores=viral_scores_two[,3],
+                  tv_scores=viral_scores_two[,4],
+                  dvf_scores=viral_scores_two[,5],
+                  tnv_scores=viral_scores_two[,6],
+                  seqtype=viruses$seqtype
+                     )
+
+ind_scores <- ind %>% 
+  select(contains("scores"), seqtype) %>%
+  pivot_longer(cols=contains("_scores"), 
+               names_to="rule_set",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_set, viral_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(ind_scores, aes(y=n, x=rule_set,
+                   fill=as.factor(viral_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
 # Comparing RefSeq vs VS2 database

### 2022-12-12 TestingPipeline/Scripts/paper_figure_generation.Rmd
Improved visualizations.

@@
                                 include_tuning_viral=FALSE,
+                                tv_1=T, tv_2=T, tv_3=T, tv_4=T, tv_5=T,
                                 include_tuning_not_viral=FALSE,
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T, ntv_6=T,
+
                                 include_virsorter=FALSE) {
@@
   if (include_tuning_viral) {
-  # tuning addition
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
-    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
-    #note: having kaiju removal rule didn't help
-    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
-    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4,
+                         include_tv_5=tv_5)
   }
@@
     # tuning removal 
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
-    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
-    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
-    # remove if DVF calls it not viral
-    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5,
+                         include_tv_6=ntv_6)
 
@@
 
+}
+
+tuning_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=T) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+  }
+  if (include_tv_2) {
+    current_score[ins$Kaiju_Viral=="Viruses"] <- current_score[ins$Kaiju_Viral=="Viruses"] + 1
+  }
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
+  if (include_tv_3) {
+    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
+  }
+  if (include_tv_4) {
+    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
+  }
+  if (include_tv_5) {
+    current_score[(ins$checkv_completeness>=75 | ins$vibrant_quality=="complete circular") & (ins$viral>=50 | ins$percent_viral>=50)] <- current_score[(ins$checkv_completeness>=75 | ins$vibrant_quality=="complete circular") & (ins$viral>=50 | ins$percent_viral>=50)] + 0.5
+  }
+
+  return(current_score)
+}
+
+tuning_not_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=T,
+                         include_tv_6=T) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[ins$checkv_host_genes>50 & !ins$provirus] <- 0
+  }
+  if (include_tv_2) {
+    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
+  }
+  if (include_tv_3) {
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+  }
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
+  }
+  if (include_tv_5) {
+    current_score[(ins$checkv_completeness<=75) & ins$checkv_length<=5000] <- current_score[(ins$checkv_completeness<=75) & ins$checkv_length<=5000] - 0.5 
+  }
+
+  if (include_tv_6) {
+    # remove if DVF calls it not viral
+    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  return(current_score)
 }
@@
 
+```{r}
+accuracy_scores_melt <- accuracy_scores %>% 
+  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
+  pivot_longer(cols=c(precision, recall, MCC), 
+               names_to="performance_metric",
+               values_to="performance_metric_score")
+```
+
 ## Visualize how the precision, recall, and MCC changes across pipelines.
@@
 ```
-
-```{r}
-accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
-  pivot_longer(cols=c(precision, recall, MCC), 
-               names_to="performance_metric",
-               values_to="performance_metric_score")
-```
-
 ```{r}
@@
 
+```
+```{r}
+one.way <- aov(precision ~ numrules, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
+one.way <- aov(recall ~ numrules, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
+one.way <- aov(MCC ~ numrules, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
+one.way <- aov(MCC ~ numtools, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
 ```
@@
 ```{r}
-viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                               include_vibrant = T,
@@
                                               include_virsorter = F)
-```
 
-```{r}
 viruses$true_virus <- "not"
@@
 ```
+
 ### Effect of combining tools
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_dvf"), size_class, true_virus) %>%
+  select(contains("keep_score_dvf"), size_class, seqtype) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
   mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, true_virus) %>%
+  group_by(rule_combination, viral_score, size_class, seqtype) %>%
   summarise(n = n())
@@
                             "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) +
-  facet_grid(~true_virus, scales = "free")
+  facet_grid(~seqtype, scales = "free")
 ```
+
+### Effect of tuning addition rules
+```{r}
+viruses$keep_score_all_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F, tv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F, tv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F, tv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T, tv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=F, tv_5=T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=F, tv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_tv"), size_class, seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+ # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
+ #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
+  facet_grid(size_class~seqtype, scales = "free")
+```
+
+### Effect of tuning removal rules
+```{r}
+viruses$keep_score_all_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F, ntv_6=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F, ntv_6=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F, ntv_6=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F, ntv_6=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T, ntv_6=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses$keep_score_all_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T, ntv_6=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_6 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T, ntv_6=T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F, ntv_6=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_ntv"), size_class, seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+ # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
+ #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
+  facet_grid(size_class~seqtype, scales = "free")
+```

### 2023-02-14 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Visualizing fragments analyses.

@@
 
-VIBRANT
-    Quality == "Complete Circular": +1
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus == TRUE: +0.5
-
-Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +0.5
-    RNA >= 0.9: +1
-    lavidaviridae >= 0.9: +1
-    NCLDV >= 0.9: +1
-
-Virsorter
-    category ==  1,4: +1
-    category == 2,5: +0.5
-
-DeepVirFinder:
-    Score >= 0.7: +0.5
-
-Tuning - No Viral Signature:
-    Kaiju_viral = "cellular organisms": -0.5
-    If host_genes >50 and NOT checkv_provirus: -1 
-    If viral_genes == 0 and host_genes >= 1: -1
-    If 3*viral_genes <= host_genes and NOT checkv_provirus: -1
-    If length > 50,000 and hallmark <=1: -1
-    If length < 5000 and checkv completeness <= 75: -0.5
-
-Tuning - Viral Signature:
-    Kaiju_viral = "Viruses": +1
-    If %unknown >= 75 and length < 50000: +0.5
-    If %viral >= 50: +0.5
-    Hallmark > 2: +1
-
 This script produces visualizations of these combined viral scorings and
@@
 
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
+Function that allows for the comparing pieces of the pipeline
 ```{r getting_viral_set_1}
@@
                                 include_tuning_viral=FALSE,
+                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                 include_tuning_not_viral=FALSE,
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                 include_virsorter=FALSE) {
@@
   if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] <- keep_score[input_seqs$viral>=50 | input_seqs$lavidaviridae>=0.95 | input_seqs$NCLDV>=0.95] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
-    keep_score[input_seqs$RNA>=0.95] <- keep_score[input_seqs$RNA>=0.95] + 1
+    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
   }
@@
   if (include_deepvirfinder) {
-    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
-    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
+     # add if DVF calls viral
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
   }
@@
   if (include_tuning_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 0.5
-    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4)
   }
@@
   if (include_tuning_not_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 
+    # tuning removal 
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5)
+
   }
@@
 
+}
+
+tuning_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+  }
+  if (include_tv_2) {
+    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
+  }
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
+  if (include_tv_3) {
+    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
+  }
+  if (include_tv_4) {
+    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
+  }
+  return(current_score)
+}
+
+tuning_not_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=F) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- 0
+  }
+  if (include_tv_2) {
+    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
+  }
+  if (include_tv_3) {
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+  }
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
+  }
+
+  if (include_tv_5) {
+    # remove if DVF calls it not viral
+    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  return(current_score)
 }
@@
 ```
+
+comparing VS2 and VIBRANT viruses
+```{r}
+vs2_vb <- data.frame(vs2=viral_scores_two[,1],
+                     vb=viral_scores_two[,3],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,3],
+                     #seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_vb$comb[vs2_vb$comb_score<1] <- "neither"
+vs2_vb$comb[vs2_vb$vs2==0.5 & vs2_vb$vb==0.5] <- "both_low"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0.5] <- "vs2_high_vb_low"
+vs2_vb$comb[vs2_vb$vs2>=0.5 & vs2_vb$vb==1] <- "vs2_low_vb_high"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0] <- "vs2_only"
+vs2_vb$comb[vs2_vb$vs2==0 & vs2_vb$vb>=1] <- "vb_only"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb>=1] <- "both_high"
+
+#vs2_vb$true_virus <- "not"
+#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"
+
+vs2_vb_seqtype <- vs2_vb %>% 
+  group_by(comb, comb_score, max_score_group) %>%
+  summarise(n = n())
+```
+
+```{r}
+table(vs2_vb$comb)
+```
+
+```{r}
+ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~max_score_group, scales = "free")
+```
+
+comparing VS2 and DVF viruses
+```{r}
+vs2_dvf <- data.frame(vs2=viral_scores_two[,1],
+                     dvf=viral_scores_two[,5],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,5],
+                     max_score_group=viruses$max_score_group)
+
+vs2_dvf$comb[vs2_dvf$comb_score<1] <- "neither"
+vs2_dvf$comb[vs2_dvf$vs2==0.5 & vs2_dvf$dvf==0.5] <- "both_low"
+vs2_dvf$comb[vs2_dvf$vs2>=1 & vs2_dvf$dvf==0.5] <- "vs2_high_dvf_low"
+vs2_dvf$comb[vs2_dvf$vs2>=0.5 & vs2_dvf$dvf==1] <- "vs2_low_dvf_high"
+vs2_dvf$comb[vs2_dvf$vs2>=1 & vs2_dvf$dvf==0] <- "vs2_only"
+vs2_dvf$comb[vs2_dvf$vs2==0 & vs2_dvf$dvf>=1] <- "dvf_only"
+vs2_dvf$comb[vs2_dvf$vs2>=1 & vs2_dvf$dvf>=1] <- "both_high"
+
+#vs2_vb$true_virus <- "not"
+#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"
+
+vs2_dvf_seqtype <- vs2_dvf %>% 
+  group_by(comb, comb_score, max_score_group) %>%
+  summarise(n = n())
+```
+
+```{r}
+table(vs2_dvf$max_score_group, vs2_dvf$comb)
+```
+
+looking at the max_score_group
+```{r}
+ggplot(vs2_dvf_seqtype, aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~max_score_group, scales = "free")
+```
+
+```{r}
+ggplot(viruses, aes(y=max_score, x=log10(checkv_length),
+                   fill=as.factor(max_score_group))) +
+  geom_point(alpha=0.1) +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  xlab("Contig Length") +
+  ylab("VS2 Score") + 
+  facet_grid(~max_score_group, scales = "free")
+```
+
+# Effect of combining tools
+```{r}
+viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_dvf_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_dvf_vs_vb <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_dvf_vs_vb_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_dvf_vs_vb_vs2_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_dvf_vs_vb_vs2_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_dvf")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.9)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
+                            "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal"))
+```
+
+### Effect of tuning addition rules
+```{r}
+viruses$keep_score_all_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_tv")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences")
+```
+
+### Effect of tuning removal rules
+```{r}
+viruses$keep_score_all_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_ntv")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") 
+```
+
+### Effect of tuning removal rules - VS2 base
+```{r}
+viruses$keep_score_vs2_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vs2_ntv")) %>%
+  pivot_longer(cols=contains("keep_score_vs2"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") 
+```

### 2023-02-14 TestingPipeline/Scripts/combining_tool_output-fragments.Rmd
Visualizing fragments analyses.

--- 2022-12-07 TestingPipeline/Scripts/combining_tool_output.Rmd
@@
 ---
-title: "Combining Tool Output"
+title: "Combining Tool Output - Fragments"
 output: html_notebook
@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "testing_set"
+dataset_name <- "testing_set-fragments"
 # checkV
-checkV_path <- "../ToolOutput/testing_sets_checkv_output.tsv"
+checkV_path <- "../ToolOutput/Fragments_Results/merged_checkv_assemblies.tsv"
 # VIBRANT
-vibrant_path <- "../ToolOutput/testing_sets_vibrant_output.tsv"
+vibrant_path <- "../ToolOutput/Fragments_Results/merged_vibrant_assemblies.tsv"
 # DeepVirFinder
-dvf_path <- "../ToolOutput/testing_sets_deepvirfinder_output.tsv"
+dvf_path <- "../ToolOutput/Fragments_Results/merged_dvf_assemblies.tsv"
 # Virsorter
-vs_path <- "../ToolOutput/testing_sets_virsorter_output.tsv"
+vs_path <- "../ToolOutput/Fragments_Results/merged_virsorter_assemblies.tsv"
 # Virsorter2
-vs2_path <- "../ToolOutput/testing_sets_virsorter2_output.tsv"
+vs2_path <- "../ToolOutput/Fragments_Results/merged_vs2_assemblies.tsv"
 # Kaiju
-kj_path <- "../ToolOutput/merged.nreuk.kaiju.names.out"
+kj_path <- "../ToolOutput/Fragments_Results/merged_kaiju_assemblies.tsv"
 # KB cutoff
@@
                 select = c(
-                    'Index',
+                    'index',
                     'contig_id',
@@
     rename(
+        Index = index,
         contig = contig_id,
@@
               select = c(
-                  'Index',
+                  'index',
                   'scaffold',
@@
     rename(
+        Index = index,
         contig = scaffold,
@@
                select = c(
-                   'Index',
+                   'index',
                    'name',
@@
     rename(
+        Index = index,
         contig = name
@@
               select = c(
-                  'index',
-                  'contig',
-                  'category'
+                  '## Contig_id',
+                  'Category'
                   )
@@
     rename(
-        Index = index
+        contig = '## Contig_id',
+        category = Category,
         )
-vs_c$contig <- sub("VIRSorter_", "", vs_c$contig)
-vs_c$contig <- sub("-circular", "", vs_c$contig)
-vs_c <- separate(vs_c, col = `contig`, into = c("seqtype", "contig"), sep="--")
+vs_c <- vs_c[!grep("#", vs_c$contig),]
+vs_c <- separate(vs_c, col = `contig`, into = c("Index", "type", "contig"), sep="--")
+vs_c$seqtype <- sub("VIRSorter_", "", vs_c$type)
 vs_c$contig <- sub("\\.", "_", vs_c$contig)
@@
 vs_c <- vs_c[!duplicated(paste(vs_c$contig, vs_c$Index)),]
+vs_c$Index <- as.numeric(vs_c$Index)
+vs_c$category <- as.numeric(vs_c$category)
 ```
@@
                     select = c(
-                        'Index',
+                        'index',
                         'seqname',
@@
                     )
+               ) %>% 
+            rename(
+                Index = index
             ) %>%
@@
 kj_c <- read_tsv(kj_path, col_names = T)
+colnames(kj_c)[2] <- "Contig"
+colnames(kj_c)[ncol(kj_c)] <- "Index"
 kj_c$Contig <- sub("\\.", "_", kj_c$Contig)
 kj_c <- separate(kj_c, col = Contig, into = c("seqtype", "contig"), sep="--")
-kj_c <- separate(kj_c, col = Name, into = c("Kaiju_Viral","Kingdom"), sep=";")
+kj_c <- separate(kj_c, col = taxonomy, into = c("Kaiju_Viral","Kingdom"), sep=";")
 kj_c$contig <- sub("\\.", "_", kj_c$contig)
@@
 viruses$bh_pvalue[is.na(viruses$bh_pvalue)] <- 0
+viruses$pvalue[is.na(viruses$pvalue)] <- 0
 
@@
 viruses$Kingdom[is.na(viruses$Kingdom)] <- "unknown"
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
 
@@
 ```{r}
-write_tsv(viruses, "../IntermediaryFiles/viral_tools_combined.tsv")
+write_tsv(viruses, "../IntermediaryFiles/viral_tools_combined-fragments.tsv")
 ```

### 2023-02-14 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Visualizing fragments analyses.

@@
 
-VIBRANT
-    Quality == "Complete Circular": +1
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus: +0.5
-
-Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +0.5
-    RNA >= 0.9: +1
-    lavidaviridae >= 0.9: +1
-    NCLDV >= 0.9: +1
-
-Virsorter
-    category ==  1,4: +1
-    category == 2,5: +0.5
-
-DeepVirFinder:
-    Score >= 0.7: +0.5
-
-Tuning - No Viral Signature:
-    Kaiju_viral = "cellular organisms": -0.5
-    If host_genes >50 and NOT provirus: -1 
-    If viral_genes == 0 and host_genes >= 1: -1
-    If 3*viral_genes <= host_genes and NOT provirus: -1
-    If length > 50,000 and hallmark <=1: -1
-    If length < 5000 and checkv completeness <= 75: -0.5
-
-Tuning - Viral Signature:
-    Kaiju_viral = "Viruses": +1
-    If %unknown >= 75 and length < 50000: +0.5
-    If %viral >= 50: +0.5
-    Hallmark > 2: +1
+<add rules back in maybe>
 
@@
 tools appropriate for your data.
+
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
 
@@
                                 include_tuning_viral=FALSE,
+                                tv_1=T, tv_2=T, tv_3=T, tv_4=T, tv_5=T,
                                 include_tuning_not_viral=FALSE,
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F, ntv_6=T,
                                 include_virsorter=FALSE) {
@@
      # add if DVF calls viral
-    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
-    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
   }
@@
   if (include_tuning_viral) {
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4,
+                         include_tv_5=tv_5)
+  }
+
+  if (include_tuning_not_viral) {
+    # tuning removal 
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5,
+                         include_tv_6=ntv_6)
+
+  }
+
+  return(keep_score)
+
+}
+
+tuning_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=T) {
     # tuning addition
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
+  if (include_tv_1) {
+    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+  }
+  if (include_tv_2) {
+    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
+  }
     #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
     #note: having kaiju removal rule didn't help
-    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
-    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5
-  }
-
-  if (include_tuning_not_viral) {
-    # tuning removal 
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
-    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
-    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
+  if (include_tv_3) {
+    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
+  }
+  if (include_tv_4) {
+    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
+  }
+  if (include_tv_5) {
+    current_score[(ins$checkv_completeness>=75 | ins$vibrant_quality=="complete circular") & (ins$viral>=50 | ins$percent_viral>=50)] <- current_score[(ins$checkv_completeness>=75 | ins$vibrant_quality=="complete circular") & (ins$viral>=50 | ins$percent_viral>=50)] + 0.5
+  }
+
+  return(current_score)
+}
+
+tuning_not_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=F,
+                         include_tv_6=F) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- 0
+  }
+  if (include_tv_2) {
+    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
+  }
+  if (include_tv_3) {
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+  }
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
+  }
+  #if (include_tv_5) {
+  #  current_score[(ins$checkv_completeness<=75) & ins$checkv_length<=5000] <- current_score[(ins$checkv_completeness<=75) & ins$checkv_length<=5000] - 0.5 
+  #}
+
+  if (include_tv_6) {
     # remove if DVF calls it not viral
-    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
-
-  }
-
-  return(keep_score)
-
+    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  return(current_score)
 }
@@
                      comb_score=viral_scores_two[,1]+viral_scores_two[,3],
-                     comb="neither",
-                     seqtype=viruses$seqtype)
-
-vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb>=1] <- "either"
-vs2_vb$comb[vs2_vb$vs2==0.5 & vs2_vb$vb==0.5] <- "together"
+                     seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_vb$comb[vs2_vb$comb_score<1] <- "neither"
+vs2_vb$comb[vs2_vb$vs2==0.5 & vs2_vb$vb==0.5] <- "both_low"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0.5] <- "vs2_high_vb_low"
+vs2_vb$comb[vs2_vb$vs2>=0.5 & vs2_vb$vb==1] <- "vs2_low_vb_high"
 vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0] <- "vs2_only"
 vs2_vb$comb[vs2_vb$vs2==0 & vs2_vb$vb>=1] <- "vb_only"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb>=1] <- "both_high"
 
@@
 vs2_vb_seqtype <- vs2_vb %>% 
-  group_by(comb, comb_score, seqtype) %>%
+  group_by(comb, comb_score, seqtype, max_score_group) %>%
   summarise(n = n())
@@
 ```{r}
-ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
+ggplot(vs2_vb_seqtype, aes(y=n, x=seqtype,
                    fill=as.factor(comb_score))) +
@@
   ylab("Number of Sequences") + 
-  facet_grid(~seqtype, scales = "free")
+  facet_grid(comb~max_score_group, scales = "free")
+```
+
+looking at the max_score_group
+```{r}
+ggplot(vs2_vb_seqtype[vs2_vb_seqtype$seqtype=="virus",], aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~max_score_group, scales = "free")
+```
+
+comparing VS2 and DVF viruses
+```{r}
+vs2_dvf <- data.frame(vs2=viral_scores_two[,1],
+                     dvf=viral_scores_two[,5],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,5],
+                     seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_dvf$comb[vs2_dvf$comb_score<1] <- "neither"
+vs2_dvf$comb[vs2_dvf$vs2==0.5 & vs2_dvf$dvf==0.5] <- "both_low"
+vs2_dvf$comb[vs2_dvf$vs2>=1 & vs2_dvf$dvf==0.5] <- "vs2_high_dvf_low"
+vs2_dvf$comb[vs2_dvf$vs2>=0.5 & vs2_dvf$dvf==1] <- "vs2_low_dvf_high"
+vs2_dvf$comb[vs2_dvf$vs2>=1 & vs2_dvf$dvf==0] <- "vs2_only"
+vs2_dvf$comb[vs2_dvf$vs2==0 & vs2_dvf$dvf>=1] <- "dvf_only"
+vs2_dvf$comb[vs2_dvf$vs2>=1 & vs2_dvf$dvf>=1] <- "both_high"
+
+#vs2_vb$true_virus <- "not"
+#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"
+
+vs2_dvf_seqtype <- vs2_dvf %>% 
+  group_by(comb, comb_score, seqtype, max_score_group) %>%
+  summarise(n = n())
+```
+
+looking at the max_score_group
+```{r}
+ggplot(vs2_dvf_seqtype, aes(y=n, x=seqtype,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(comb~max_score_group, scales = "free")
+```
+
+comparing VS2 and tv viruses
+```{r}
+vs2_tv <- data.frame(vs2=viral_scores_two[,1],
+                     tv=viral_scores_two[,4],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,4],
+                     seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_tv$comb[vs2_tv$comb_score<1] <- "neither"
+vs2_tv$comb[vs2_tv$vs2==0.5 & vs2_tv$tv==0.5] <- "both_low"
+vs2_tv$comb[vs2_tv$vs2>=1 & vs2_tv$tv==0.5] <- "vs2_high_tv_low"
+vs2_tv$comb[vs2_tv$vs2>=0.5 & vs2_tv$tv>=1] <- "vs2_low_tv_high"
+vs2_tv$comb[vs2_tv$vs2>=1 & vs2_tv$tv==0] <- "vs2_only"
+vs2_tv$comb[vs2_tv$vs2==0 & vs2_tv$tv>=1] <- "tv_only"
+vs2_tv$comb[vs2_tv$vs2>=1 & vs2_tv$tv>=1] <- "both_high"
+
+vs2_tv_seqtype <- vs2_tv %>% 
+  group_by(comb, comb_score, seqtype, max_score_group) %>%
+  summarise(n = n())
+```
+
+looking at the max_score_group
+```{r}
+ggplot(vs2_tv_seqtype, aes(y=n, x=seqtype,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr") +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(comb~max_score_group, scales = "free")
+```
+
+comparing VS2 and tnv viruses
+```{r}
+vs2_tnv <- data.frame(vs2=viral_scores_two[,1],
+                     tnv=viral_scores_two[,6],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,6],
+                     seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_tnv$comb[vs2_tnv$comb_score<1] <- "both_not"
+vs2_tnv$comb[vs2_tnv$vs2>=1 & vs2_tnv$tnv<=0] <- "vs2_only"
+
+vs2_tnv_seqtype <- vs2_tnv %>% 
+  group_by(comb, comb_score, seqtype, max_score_group) %>%
+  summarise(n = n())
+```
+
+looking at the max_score_group
+```{r}
+ggplot(vs2_tnv_seqtype, aes(y=n, x=seqtype,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr") +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(comb~max_score_group, scales = "free")
 ```

### 2023-02-14 TestingPipeline/Scripts/examining_false_positives-short_fragments.Rmd
Visualizing fragments analyses.

---
title: "Examining False Negatives - Fragments"
author: Bridget Hegarty, James Riddell
date: 10-30-2022
output: html_notebook
---
This Rmarkdown file examines the false positives from the output of CheckV, DeepVirFinder, Kaiju,
VIBRANT, VirSorter, and VirSorter2 on multiple training sets of microbial DNA, 
primarily from NCBI. Created from fungal, viral, bacterial, archeael, protist,
and plasmid DNA sequences

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
library(pROC)
```

Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined-fragments.tsv")
```

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }

  if (include_deepvirfinder) {
     # add if DVF calls viral
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score <- tuning_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=tv_1,
                         include_tv_2=tv_2,
                         include_tv_3=tv_3,
                         include_tv_4=tv_4)
  }

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score <- tuning_not_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=ntv_1,
                         include_tv_2=ntv_2,
                         include_tv_3=ntv_3,
                         include_tv_4=ntv_4,
                         include_tv_5=ntv_5)

  }

  return(keep_score)

}

tuning_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T) {
    # tuning addition
  if (include_tv_1) {
    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
  }
  if (include_tv_2) {
    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
  }
    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
    #note: having kaiju removal rule didn't help
  if (include_tv_3) {
    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
  }
  if (include_tv_4) {
    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
  }
  return(current_score)
}

tuning_not_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T,
                         include_tv_5=F) {
    # tuning addition
  if (include_tv_1) {
    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- 0
  }
  if (include_tv_2) {
    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
  }
  if (include_tv_3) {
    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
  }
  if (include_tv_4) {
    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
  }

  if (include_tv_5) {
    # remove if DVF calls it not viral
    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
  }

  return(current_score)
}
```

```{r}
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
```

```{r}
viruses$confusion_matrix_high_MCC <- "true negative"
viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
```

getting a list of false positive bacteria and plasmids from testing set 1
```{r}
fp_bacteria_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="bacteria" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

fp_plasmid_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="plasmid" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

fp_protist_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="protist" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

fp_fungi_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="fungi" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

fp_archaea_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="archaea" &
                                     viruses$confusion_matrix_high_MCC=="false positive"]

tp_virus_list <- viruses$contig[viruses$Index==1 & 
                                     viruses$seqtype=="virus" &
                                     viruses$confusion_matrix_high_MCC=="true positive"]
```

vibrant annotations
```{r}
annotations_vb <- read_tsv("../ToolOutput/Fragments_Results/VIBRANT_annotations_fractionated_test_set_1.tsv")
```
```{r}
annotations_vb <- separate(annotations_vb, col="scaffold", into=c("seqtype", "contig"),
                           sep="--")
annotations_vb$contig <- sub("\\.", "_", annotations_vb$contig)
```

```{r}
head(annotations_vb)
```
```{r}
table(annotations_vb$contig %in% fp_bacteria_list)
table(annotations_vb$contig %in% fp_plasmid_list)
```

```{r}
annotations_vb_fp <- rbind(annotations_vb[annotations_vb$contig %in% fp_bacteria_list,],
                           annotations_vb[annotations_vb$contig %in% fp_plasmid_list,],
                           annotations_vb[annotations_vb$contig %in% tp_virus_list,],
                           annotations_vb[annotations_vb$contig %in% fp_archaea_list,],
                           )
```

```{r}
annotations_vb_fp
```

```{r}
annotations_vb_fp_vog_count <- annotations_vb_fp %>%
  select(contig, VOG, seqtype) %>%
  group_by(contig, seqtype) %>%
  summarize(num_vog=table(is.na(VOG))[[1]],
            prop_vog=table(is.na(VOG))[[1]]/length(VOG),
            num_orfs=length(VOG))
```

```{r}
annotations_vb_fp_vog_count$confusion_matrix <- "false positive"
annotations_vb_fp_vog_count$confusion_matrix[annotations_vb_fp_vog_count$seqtype=="virus"] <- "true positive"
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
```

```{r}
ggplot(annotations_vb_fp_vog_count, 
       aes(x=seqtype, y=prop_vog, 
           color=seqtype, fill=seqtype)) +
  geom_jitter(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Proportion Genes VOGs") +
  xlab("Sequence Type") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1)) +
  facet_wrap(~confusion_matrix, scales="free_x")
```
```{r}
t.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="bacteria"],
       annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
```
```{r}
t.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="plasmid"],
       annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
```

```{r}
ggplot(annotations_vb_fp_vog_count, 
       aes(x=num_vog, y=prop_vog, 
           color=seqtype, fill=seqtype)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Proportion Genes VOGs") +
  xlab("Number of VOGs") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1))
```

# Thinking about short protists
```{r}
viruses[viruses$confusion_matrix_high_MCC=="false positive" & viruses$seqtype=="protist",]

table(viruses$confusion_matrix_high_MCC[viruses$seqtype=="protist"], 
      viruses$max_score_group[viruses$seqtype=="protist"])

table(viruses$Kingdom[viruses$seqtype=="protist" &  viruses$confusion_matrix_high_MCC=="false positive"], 
      viruses$max_score_group[viruses$seqtype=="protist" & viruses$confusion_matrix_high_MCC=="false positive"])

table(viruses$percent_viral[viruses$seqtype=="protist" &  viruses$confusion_matrix_high_MCC=="false positive"]>20, 
      viruses$max_score_group[viruses$seqtype=="protist" & viruses$confusion_matrix_high_MCC=="false positive"])
```

```{r}
ggplot(viruses, aes(y=percent_viral, x=confusion_matrix_high_MCC)) +
  geom_violin() +
  geom_point(alpha=0.5) +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") 
```
```{r}
table(viruses$confusion_matrix_high_MCC[viruses$checkv_length>=5000])
```

```{r}
ggplot(viruses, aes(y=viral, x=confusion_matrix_high_MCC)) +
  geom_violin() +
  geom_point(alpha=0.2) +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") 
```

```{r}
table(viruses$seqtype)
```

### 2023-02-14 TestingPipeline/Scripts/paper_figure_generation.Rmd
Visualizing fragments analyses.

--- 2022-12-08 TestingPipeline/Scripts/paper_figure_generation.Rmd
@@
 
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
 Function that allows for the comparing pieces of the pipeline
@@
   if (include_tuning_viral) {
-  # tuning addition
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
-    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
-    #note: having kaiju removal rule didn't help
-    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] <- keep_score[input_seqs$viral>=50 | input_seqs$percent_viral>=50] + 0.5 
-    keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] <- keep_score[(input_seqs$checkv_completeness>=75 | input_seqs$vibrant_quality=="complete circular") & (input_seqs$viral>=50 | input_seqs$percent_viral>=50)] + 0.5
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4)
   }
@@
     # tuning removal 
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- 0
-    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- 0
-    keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] <- keep_score[(input_seqs$checkv_completeness<=75) & input_seqs$checkv_length<=5000] - 0.5 
-    # remove if DVF calls it not viral
-    keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] <- keep_score[input_seqs$score<=0.7 & input_seqs$pvalue<=0.05] - 1
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5)
 
@@
 
+```{r}
+mean_precision <- accuracy_scores %>% 
+  select(precision, toolcombo) %>%
+  group_by(toolcombo) %>%
+  summarise(mean = mean(precision))
+```
+
 recall
@@
                      values = alpha(rev(pal(6)), 1))
+```
+
+```{r}
+mean_recall <- accuracy_scores %>% 
+  select(recall, toolcombo) %>%
+  group_by(toolcombo) %>%
+  summarise(mean = mean(recall))
 ```
@@
 ```
-
 ```{r}
-accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
-  pivot_longer(cols=c(precision, recall, MCC), 
-               names_to="performance_metric",
-               values_to="performance_metric_score")
+mean_MCC <- accuracy_scores %>% 
+  select(MCC, toolcombo) %>%
+  group_by(toolcombo) %>%
+  summarise(mean = mean(MCC))
 ```
@@
 
+```{r}
+one.way <- aov(precision ~ numrules, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
+one.way <- aov(recall ~ numrules, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
+one.way <- aov(MCC ~ numrules, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
+one.way <- aov(MCC ~ numtools, data = accuracy_scores)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
 ## differences based on genome size
 ```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                               include_vibrant = F,
@@
 
-viruses$confusion_matrix_all <- "true negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all<1] <- "false negative"
-viruses$confusion_matrix_all[viruses$seqtype=="virus" & viruses$keep_score_all>=1] <- "true positive"
-viruses$confusion_matrix_all[viruses$seqtype!="virus" & viruses$keep_score_all>=1] <- "false positive"
+viruses$confusion_matrix_high_MCC <- "true negative"
+viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
+viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
+viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
 
@@
 
-confusion_by_taxa <- viruses %>% count(confusion_matrix_all, seqtype, size_class, Index)
+confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, size_class, Index)
 
@@
 }
+```
 
+```{r}
+confusion_vir_called_subset <- confusion_vir_called[confusion_vir_called$seqtype=="virus",]
+
+one.way <- aov(per_viral ~ size, data = confusion_vir_called_subset)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
+
+```{r}
 confusion_vir_called <- confusion_vir_called %>% group_by(seqtype, size) %>%
@@
 ```
+```{r}
+one.way <- aov(mean ~ seqtype, data = confusion_vir_called)
+
+summary(one.way)
+
+TukeyHSD(one.way)
+```
 
@@
                             "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) +
-  facet_grid(~true_virus, scales = "free")
+  facet_grid(~seqtype, scales = "free")
 ```
+
+### Effect of tuning addition rules
+```{r}
+viruses$keep_score_all_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_tv"), size_class, seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.5)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+ # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
+ #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
+  facet_grid(size_class~seqtype, scales = "free")
+```
+
+### Effect of tuning removal rules
+```{r}
+viruses$keep_score_all_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_ntv"), size_class, seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  summarise(n = n())
+
+viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
+                                    levels = c("3-5kb", "5-10kb", ">10kb"))
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 1)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+ # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
+ #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
+  facet_grid(size_class~seqtype, scales = "free")
+```
+
+same idea but with a VS2 base
+```{r}
+viruses$keep_score_vs2_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vs2_ntv"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_vs2"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```

### 2023-02-14 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Visualizing fragments analyses.

@@
 
+# 2023.01.15 Testing
+
 Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
@@
 ```
+
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
+```{r getting_viral_set_1}
+getting_viral_set_1 <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_tuning_viral=FALSE,
+                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
+                                include_tuning_not_viral=FALSE,
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F, ntv_6=T,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
+  }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
+  }
+
+  if (include_deepvirfinder) {
+     # add if DVF calls viral
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+  }
+
+  if (include_tuning_viral) {
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4,
+                         include_tv_5=tv_5)
+  }
+
+  if (include_tuning_not_viral) {
+    # tuning removal 
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5,
+                         include_tv_6=ntv_6)
+
+  }
+
+  return(keep_score)
+
+}
+
+tuning_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+  }
+  if (include_tv_2) {
+    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
+  }
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
+  if (include_tv_3) {
+    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
+  }
+  if (include_tv_4) {
+    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
+  }
+  return(current_score)
+}
+
+tuning_not_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=F,
+                         include_tv_6=F) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- 0
+  }
+  if (include_tv_2) {
+    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
+  }
+  if (include_tv_3) {
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+  }
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
+  }
+  #if (include_tv_5) {
+  #  current_score[(ins$checkv_completeness<=75) & ins$checkv_length<=5000] <- current_score[(ins$checkv_completeness<=75) & ins$checkv_length<=5000] - 0.5 
+  #}
+
+  if (include_tv_6) {
+    # remove if DVF calls it not viral
+    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  return(current_score)
+}
+```
+
+getting a 3kb set, equal virus and not virus
+```{r}
+set.seed(32)
+
+viruses_3to5kb <- viruses[viruses$checkv_length <= 5000,]
+
+viruses_3to5kb <- rbind(viruses_3to5kb[sample(grep("archaea", viruses_3to5kb$seqtype), size=100),],
+                        viruses_3to5kb[sample(grep("bacteria", viruses_3to5kb$seqtype), size=100),],
+                        viruses_3to5kb[sample(grep("fungi", viruses_3to5kb$seqtype), size=100),],
+                        viruses_3to5kb[sample(grep("plasmid", viruses_3to5kb$seqtype), size=100),],
+                        viruses_3to5kb[sample(grep("protist", viruses_3to5kb$seqtype), size=100),],
+                        viruses_3to5kb[sample(grep("virus", viruses_3to5kb$seqtype), size=500),])
+```
+
+```{r}
+table(viruses_3to5kb$seqtype)
+```
+
+## visualizing a select set of tools
+```{r}
+viruses_3to5kb$keep_score_high_all <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses_3to5kb$keep_score_high_recall <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+viruses_3to5kb$keep_score_high_precision <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses_3to5kb$keep_score_high_MCC <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses_3to5kb$keep_score_high_vs2 <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses_3to5kb$true_virus <- "not"
+viruses_3to5kb$true_virus[viruses_3to5kb$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses_3to5kb %>% 
+  select(contains("keep_score_high"), true_virus) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, true_virus) %>%
+  summarise(n = n())
+```
+
+```{r}
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~true_virus, scales = "free")
+```
+```{r}
+table(viruses_3to5kb$seqtype, viruses_3to5kb$keep_score_high_vs2)
+table(viruses_3to5kb$seqtype, viruses_3to5kb$keep_score_high_recall)
+table(viruses_3to5kb$seqtype, viruses_3to5kb$keep_score_high_MCC)
+```
+
+```{r}
+viruses_3to5kb$keep_score_tnv <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+table(viruses_3to5kb$seqtype, viruses_3to5kb$keep_score_tnv)
+```
+
+note: after these results, pulled up some of the protists in NCBI and they had 
+the following error: "This record was removed as a result of standard genome annotation processing. Please see www.ncbi.nlm.nih.gov/genome/annotation_euk/process/ for more information."
+
+# Prior to 2023.01.15
+
+Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
+```{r}
+viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
+```
+
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
 (this is still the old rules)
@@
     keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 1
+    keep_score[input_seqs$Kaiju_Viral=="Viruses" & input_seqs$kaiju_match_ratio>=0.5] <- keep_score[input_seqs$Kaiju_Viral=="Viruses" & input_seqs$kaiju_match_ratio>=0.5] + 1
     #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
@@
   facet_grid(~true_virus, scales = "free")
+```
+
+```{r}
+viruses <- viruses_3to5kb
 ```
@@
 ```
+
+```{r}
+ggplot(viruses, aes(x=checkv_completeness, y=checkv_contamination, 
+                                  color=confusion_matrix_high_MCC, fill=confusion_matrix_high_MCC)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Contamination") +
+  xlab("Completeness") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~seqtype)
+```

### 2023-02-14 TestingPipeline/Scripts/testing_set_features_visualization.Rmd
Visualizing fragments analyses.

@@
 ```
+
+```{r}
+ggplot(viruses, aes(x=max_score, y=RNA)) +
+  geom_point() +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~max_score_group, scales = "free") +
+  xlab("VirSorter2 Max Score") +
+  ylab("VirSorter2 RNA Score")
+```
+
+```{r}
+viruses$hallmark_color <- "less than 2"
+viruses$hallmark_color[viruses$hallmark>=2] <- "greater than or equal to 2"
+```
+
+```{r}
+ggplot(viruses, aes(x=hallmark, color=hallmark_color, fill=hallmark_color)) +
+  geom_histogram() +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  scale_color_viridis_d(begin=0.2, end=0.8) +
+  scale_fill_viridis_d(begin=0.2, end=0.8, alpha=0.2) +
+  xlab("VS2 Hallmark Genes")
+```
+
+```{r}
+ggplot(viruses, aes(x=percent_unknown, y=checkv_length)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("Percent Unknown") +
+  ylab("Contig Length") +
+  geom_hline(yintercept=50000) +
+  geom_vline(xintercept=75)
+```
+
+```{r}
+ggplot(viruses, aes(x=percent_viral, y=viral)) +
+  geom_hex(bins = 30) +
+  scale_fill_continuous(type = "viridis", trans="log10") +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  xlab("Percent Viral - CheckV") +
+  ylab("Percent Viral - VirSorter2") +
+  geom_hline(yintercept=50) +
+  geom_vline(xintercept=50)
+```

### 2023-02-14 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
Visualizing fragments analyses.

--- 2023-02-14 TestingPipeline/Scripts/paper_figure_generation.Rmd
@@
 ---
-title: "Pipeline Performance Figure Generation"
+title: "Viral Sequence Sorting Tools - Rules on Fragments"
+author: Bridget Hegarty, James Riddell
+date: 07-22-2022
 output: html_notebook
 ---
-
-This R script creates all of the figures used to compare the performance across pipelines.
-
 This Rmarkdown file assesses the output of CheckV, DeepVirFinder, Kaiju,
@@
 
-Import the file that combines the results from each of the tools from running "combining_tool_output.Rmd":
-```{r}
-viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
-```
-
-```{r}
-viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
-```
-
-Function that allows for the comparing pieces of the pipeline
+Import the file that combines the results from each of the tools from running "combining_tool_output-fragments.Rmd":
+```{r}
+viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined-fragments.tsv")
+```
+
 ```{r getting_viral_set_1}
@@
 
-accounting for rules that have multiple tools in them
-```{r}
-accuracy_scores$numtools <- accuracy_scores$numrules
-
-for (i in 1:nrow(accuracy_scores)) {
-  toolcombo <- as.character(accuracy_scores$toolcombo[i])
-  if (substr(toolcombo, 1, 1)=="1") {
-    accuracy_scores$numtools[i] <- accuracy_scores$numtools[i] + 3
-  }
-
-  if (substr(toolcombo, 5, 5)=="1") {
-    accuracy_scores$numtools[i] <- accuracy_scores$numtools[i] + 4
-  }
-}
-
-accuracy_scores$numtools[accuracy_scores$numtools>6] <- 6
-```
-
-```{r}
-accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
-accuracy_scores$numtools <- as.factor(accuracy_scores$numtools)
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
-accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
-  pivot_longer(cols=c(precision, recall, MCC), 
-               names_to="performance_metric",
-               values_to="performance_metric_score")
-```
-
 ## Visualize how the precision, recall, and MCC changes across pipelines.
@@
 ```
-
-## differences based on genome size
-```{r}
-viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$confusion_matrix_high_MCC <- "true negative"
-viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC<1] <- "false negative"
-viruses$confusion_matrix_high_MCC[viruses$seqtype=="virus" & viruses$keep_score_high_MCC>=1] <- "true positive"
-viruses$confusion_matrix_high_MCC[viruses$seqtype!="virus" & viruses$keep_score_high_MCC>=1] <- "false positive"
-
-viruses$size_class <- "3-5kb"
-viruses$size_class[viruses$checkv_length>5000] <- "5-10kb"
-viruses$size_class[viruses$checkv_length>10000] <- ">10kb"
-
-confusion_by_taxa <- viruses %>% count(confusion_matrix_high_MCC, seqtype, size_class, Index)
-
-colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","size", "index", "count")
-
-confusion_vir_called <- confusion_by_taxa %>% filter(confusion_matrix=="true positive" | confusion_matrix=="false positive")
-
-type_count <- viruses %>% count(seqtype, size_class, Index)
-
-confusion_vir_called$per_viral <- 0
-
-for (i in c(1:nrow(confusion_vir_called))) {
-  confusion_vir_called$per_viral[i] <- confusion_vir_called$count[i]/type_count$n[type_count$seqtype==confusion_vir_called$seqtype[i] & 
-                                                                                    type_count$Index==confusion_vir_called$index[i] &
-                                                                                    type_count$size_class==confusion_vir_called$size[i]]*100
-}
-```
-
-```{r}
-confusion_vir_called_subset <- confusion_vir_called[confusion_vir_called$seqtype=="virus",]
-
-one.way <- aov(per_viral ~ size, data = confusion_vir_called_subset)
-
-summary(one.way)
-
-TukeyHSD(one.way)
-```
-
-```{r}
-confusion_vir_called <- confusion_vir_called %>% group_by(seqtype, size) %>%
-  summarise(mean=mean(per_viral), 
-            sd=sd(per_viral))
-
-confusion_vir_called$size <- factor(confusion_vir_called$size,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
-```
-
-```{r}
-ggplot(confusion_vir_called, aes(y=mean, x=size,
-                   fill=seqtype,
-                   color=seqtype)) +
-  geom_bar(stat="identity", position=position_dodge()) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
+comparing metric with and without tuning rules
+```{r}
+accuracy_scores_melt$tuning_inc <- "no"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1] <- "tv"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "tnv"
+accuracy_scores_melt$tuning_inc[substring(accuracy_scores_melt$toolcombo, 1, 1)==1 &
+                                  substring(accuracy_scores_melt$toolcombo, 3, 3)==1] <- "both"
+```
+
+```{r}
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score)) +
+  geom_boxplot() +
+  geom_point(aes(color=numrules, fill=numrules), alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
     legend.text=element_text(size=12),
@@
   ) +
-  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
-                 position=position_dodge(.9)) +
+  ylab("MCC") +
+  xlab("Number of Tools") +
   scale_fill_manual(name="",
@@
                      values = alpha(rev(pal(6)), 1)) +
-  xlab("Length") +
-  ylab("Sequences Called Viral (%)") 
-```
-```{r}
-one.way <- aov(mean ~ seqtype, data = confusion_vir_called)
-
-summary(one.way)
-
-TukeyHSD(one.way)
+  facet_wrap(~performance_metric)
+
+ggplot(accuracy_scores_melt, aes(x=tuning_inc, y=performance_metric_score,
+                                 color=numrules, fill=numrules)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("MCC") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=tuning_inc, y=performance_metric_score)) +
+  geom_boxplot() +
+  geom_boxplot(aes(color=numrules, fill=numrules)) +
+ # geom_point(aes(color=numrules, fill=numrules), alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Score") +
+  xlab("") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 0.7)) +
+  facet_wrap(~performance_metric)
+
 ```
@@
 ```{r}
-viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses_3to5kb <- viruses
+
+viruses_3to5kb$keep_score_high_all <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses_3to5kb$keep_score_high_recall <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = T,
                                               include_vibrant = F,
@@
                                               include_virsorter = T)
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+viruses_3to5kb$keep_score_high_precision <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses_3to5kb$keep_score_high_MCC <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses_3to5kb$keep_score_high_vs2 <- getting_viral_set_1(viruses_3to5kb, include_deepvirfinder = F,
                                               include_vibrant = F,
@@
                                               include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
+                                              include_tuning_not_viral = F,
                                               include_virsorter = F)
-viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$true_virus <- "not"
-viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
-
-viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_high"), size_class, true_virus) %>%
+
+viruses_3to5kb$true_virus <- "not"
+viruses_3to5kb$true_virus[viruses_3to5kb$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses_3to5kb %>% 
+  select(contains("keep_score_high"), true_virus) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
   mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, true_virus) %>%
+  group_by(rule_combination, viral_score, true_virus) %>%
   summarise(n = n())
-
-viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
-```
-
-```{r}
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+
 ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity", color="black") +
   theme_light() +
@@
   ) +
-  scale_fill_brewer(palette = "PuOr", ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
   xlab("") +
@@
 
-### Effect of combining tools
-```{r}
-viruses$keep_score_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
+## building up to all
+```{r}
+viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
                                               include_virsorter2 = F,
@@
 
-viruses$keep_score_dvf_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
+viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
                                               include_virsorter2 = F,
@@
                                               include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_dvf_vs_vb <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
                                               include_tuning_viral = F,
                                               include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_dvf_vs_vb_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
 
-viruses$keep_score_dvf_vs_vb_vs2_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
 
-viruses$keep_score_dvf_vs_vb_vs2_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
-
+viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+```
+
+```{r}
 viruses$true_virus <- "not"
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_dvf"), size_class, seqtype) %>%
+  select(contains("keep_score_vb"), true_virus) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
   mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  group_by(rule_combination, viral_score,true_virus) %>%
   summarise(n = n())
-
-viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
-```
-
-```{r}
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+
 ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity", color="black") +
   theme_light() +
@@
   ) +
-  scale_fill_brewer(palette = "PuOr", ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
   xlab("") +
   ylab("Number of Sequences") + 
-  scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
-                            "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) +
-  facet_grid(~seqtype, scales = "free")
+  scale_x_discrete(labels=c("VB", "VB+DVF", "VB+DVF+VS2", "VB+DVF+VS2+VS",
+                            "VB+DVF+VS2+VS+addition", "VB+DVF+VS2+VS+addition-removal")) +
+  facet_grid(~true_virus, scales = "free")
 ```
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_all_tv"), size_class, seqtype) %>%
+  select(contains("keep_score_all_tv"), seqtype) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
   mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
   summarise(n = n())
-
-viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
 ```
@@
                    fill=viral_score)) +
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity", color="black") +
   theme_light() +
@@
 
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity", color="black") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.5)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
- # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
- #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
-  facet_grid(size_class~seqtype, scales = "free")
-```
-
 ### Effect of tuning removal rules
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_all_ntv"), size_class, seqtype) %>%
+  select(contains("keep_score_all_ntv"), seqtype) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
 ```
-
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity", color="black") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 1)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
- # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
- #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
-  facet_grid(size_class~seqtype, scales = "free")
-```
-
-same idea but with a VS2 base
-```{r}
-viruses$keep_score_vs2_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = F)
-
-viruses$keep_score_vs2_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$true_virus <- "not"
-viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
-
-viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_vs2_ntv"), seqtype) %>%
-  pivot_longer(cols=contains("keep_score_vs2"), 
-               names_to="rule_combination",
-               values_to="viral_score") %>% 
-  mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, seqtype) %>%
-  summarise(n = n())
-```
-
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity", color="black") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  facet_grid(~seqtype, scales = "free")
-```

### 2023-02-20 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
added vs2 SOP v2
Compared further the keep2 category and analyzed the range of accuracy for the VS2 SOP given the manual curation is 0% or 100% accurate.

Fixed small bugs in fragments by changing the numrules to a factor data type so read as discrete not continuous

@@
 ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
-                                  color=numrules, fill=numrules)) +
+                                  color=as.factor(numrules), fill=numrules)) +
   geom_point(alpha=0.5) +

### 2023-03-23 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Updating figures with current rules.

@@
 
+comparing VS and VS2 viruses
+```{r}
+vs2_vs <- data.frame(vs2=viral_scores_two[,1],
+                     vs=viral_scores_two[,2],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,2],
+                     seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_vs$comb[vs2_vs$comb_score<1] <- "neither"
+vs2_vs$comb[vs2_vs$vs2==0.5 & vs2_vs$vs==0.5] <- "both_low"
+vs2_vs$comb[vs2_vs$vs2>=1 & vs2_vs$vs==0.5] <- "vs2_high_vs_low"
+vs2_vs$comb[vs2_vs$vs2>=0.5 & vs2_vs$vs>=1] <- "vs2_low_vs_high"
+vs2_vs$comb[vs2_vs$vs2>=1 & vs2_vs$vs==0] <- "vs2_only"
+vs2_vs$comb[vs2_vs$vs2==0 & vs2_vs$vs>=1] <- "vs_only"
+vs2_vs$comb[vs2_vs$vs2>=1 & vs2_vs$vs>=1] <- "both_high"
+
+#vs2_vb$true_virus <- "not"
+#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"
+
+vs2_vs_seqtype <- vs2_vs %>% 
+  group_by(comb, comb_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+table(vs2_vs$comb)
+```
+
+```{r}
+ggplot(vs2_vs_seqtype, aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
 comparing VS2 and VIBRANT viruses
@@
 vs2_vb_seqtype <- vs2_vb %>% 
-  group_by(comb, comb_score, seqtype, max_score_group) %>%
+  group_by(comb, comb_score, seqtype) %>%
   summarise(n = n())
@@
 table(vs2_vb$comb)
+```
+
+```{r}
+ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
 ```

### 2023-03-23 TestingPipeline/Scripts/examining_false_positives.Rmd
New visualizations for SI.
Visualizing fragments analyses.
Updating figures with current rules.

@@
 
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
 This section defines a viralness score "keep_score" based on the tool classifications. 
 A final keep_score above 1 indicates we will keep that sequence and call it viral.
-
-VIBRANT
-    Quality == "Complete Circular": +1
-    Quality == "High Quality Draft": +1
-    Quality == "Medium Quality Draft": +1
-    Quality == "Low Quality Draft" & provirus: +0.5
-
-Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +0.5
-    RNA >= 0.9: +1
-    lavidaviridae >= 0.9: +1
-    NCLDV >= 0.9: +1
-
-Virsorter
-    category ==  1,4: +1
-    category == 2,5: +0.5
-
-DeepVirFinder:
-    Score >= 0.7: +0.5
-
-Tuning - No Viral Signature:
-    Kaiju_viral = "cellular organisms": -0.5
-    If host_genes >50 and NOT provirus: -1 
-    If viral_genes == 0 and host_genes >= 1: -1
-    If 3*viral_genes <= host_genes and NOT provirus: -1
-    If length > 50,000 and hallmark <=1: -1
-    If length < 5000 and checkv completeness <= 75: -0.5
-
-Tuning - Viral Signature:
-    Kaiju_viral = "Viruses": +1
-    If %unknown >= 75 and length < 50000: +0.5
-    If %viral >= 50: +0.5
-    Hallmark > 2: +1
 
@@
   if (include_tuning_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
-    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
-    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
-    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4)
   }
@@
   if (include_tuning_not_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] <- keep_score[input_seqs$checkv_host_genes>50 & !input_seqs$provirus] - 1
-    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & !input_seqs$provirus] - 1 # consider accounting for RNA viruses
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
+    # tuning removal 
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5)
+
   }
@@
 ```
+
+# Just looking at the protists getting wrong
+
+```{r}
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
+```{r}
+ggplot(viruses, aes(y=kaiju_match_ratio, x=seqtype)) +
+  geom_boxplot() +
+  geom_point() +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") 
+```

### 2023-03-23 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Updating figures with current rules.

@@
 ```
+comparing each tool's score
+```{r}
+viruses$keep_score_ind_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_ind_"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```

### 2023-03-23 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating figures with current rules.

@@
 
+same idea but VS2 base
+```{r}
+viruses$keep_score_vs2_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_4_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_4_tv_2_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vs2_tv_"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(1, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange[1]), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
 ### Effect of tuning removal rules
@@
                                               include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
                                               include_tuning_not_viral = T,
@@
                                               include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
                                               include_tuning_not_viral = T,
@@
                                               include_tuning_viral = F,
-                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T,
                                               include_tuning_not_viral = T,
@@
                                               include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_0_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_all_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_all_tv_4_2_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = F)

### 2023-03-23 TestingPipeline/Scripts/comparing_to_VS2_SOP.Rmd

@@
 
-# VS2 SOP testing
+# VS2 SOP testing of version 2 and 3
 
@@
 
-```{r virsorter2_SOP}
-virsorter2_SOP <- function(input_seqs) {
+```{r virsorter2_SOP v2}
+virsorter2_SOP2 <- function(input_seqs) {
+  keep_score <- rep(0, nrow(input_seqs))
+  # Keep: viral_gene >0 OR score >=0.95 OR hallmark >2 OR (viral_gene =0 AND host_gene =0)
+  keep_score[(input_seqs$checkv_viral_genes > 0) |
+             (input_seqs$max_score >=0.95) |
+             (input_seqs$hallmark > 2) |
+             ((input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 0))] <- 1
+  # discard viral_gene =0 AND host_gene >1 OR (viral_gene =0 AND host_gene =1 AND length <10kb)
+  keep_score[(input_seqs$checkv_viral_genes == 0) & 
+             (input_seqs$checkv_host_genes > 1) |
+             ((input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length < 10000))] <- 0
+  # Manual check: (NOT in Keep1 OR Keep2) AND viral_gene =0 AND host_gene =1 AND length >=10kb
+  keep_score[(input_seqs$checkv_viral_genes == 0) & (input_seqs$checkv_host_genes == 1) & (input_seqs$checkv_length >= 10000)] <- 0.5
+  # implied by starting with zero matrix, unchanged will stay zero
+
+  return(keep_score)
+}
+
+```
+
+```{r virsorter2_SOP v3}
+virsorter2_SOP3 <- function(input_seqs) {
   keep_score <- rep(0, nrow(input_seqs))
@@
   # Keep2: viral_gene =0 AND (host_gene =0 OR score >=0.95 OR hallmark >2)
-  keep_score[(input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))] <- 1
+  keep_score[(input_seqs$checkv_viral_genes == 0) & ((input_seqs$checkv_host_genes == 0) | (input_seqs$max_score >=0.95) | (input_seqs$hallmark > 2))] <- 0.75
 
@@
 ```{r}
-viruses$keep_score_virsorter2_SOP <- virsorter2_SOP(viruses)
+viruses$keep_score_virsorter2_SOP3 <- virsorter2_SOP3(viruses)
+viruses$keep_score_virsorter2_SOP2 <- virsorter2_SOP2(viruses)
 ```
@@
 ```{r}
-viruses$confusion_matrix_vs2_sop <- "true negative"
-viruses$confusion_matrix_vs2_sop[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP<1] <- "false negative"
-viruses$confusion_matrix_vs2_sop[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP>=1] <- "true positive"
-viruses$confusion_matrix_vs2_sop[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP>=1] <- "false positive"
+viruses$confusion_matrix_vs2_sop2 <- "true negative"
+viruses$confusion_matrix_vs2_sop2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP2<1] <- "false negative"
+viruses$confusion_matrix_vs2_sop2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP2>=1] <- "true positive"
+viruses$confusion_matrix_vs2_sop2[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP2>=1] <- "false positive"
 
-viruses$accurate_vs2_SOP <- "false"
-viruses$accurate_vs2_SOP[grep("true", viruses$confusion_matrix_vs2_sop)] <- "true"
+viruses$accurate_vs2_SOP2 <- "false"
+viruses$accurate_vs2_SOP2[grep("true", viruses$confusion_matrix_vs2_sop2)] <- "true"
+```
+
+```{r}
+viruses$confusion_matrix_vs2_sop3 <- "true negative"
+viruses$confusion_matrix_vs2_sop3[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3<1] <- "false negative"
+viruses$confusion_matrix_vs2_sop3[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "true positive"
+viruses$confusion_matrix_vs2_sop3[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "false positive"
+
+viruses$accurate_vs2_SOP3 <- "false"
+viruses$accurate_vs2_SOP3[grep("true", viruses$confusion_matrix_vs2_sop3)] <- "true"
+```
+
+```{r}
+# investigate keep2
+keep2 = viruses[viruses$keep_score_virsorter2_SOP3 == 0.75,]
+nrow(keep2)
+```
+```{r}
+length(grep("virus", keep2$seqtype))
+```
+
+assume keep2 (0.75) and manual curation (0.5) has 100% accuracy to avoid having to check with DRAM v since it represents a small portion of the sequences
+
+```{r}
+viruses$confusion_matrix_vs2_keep2 <- "true negative"
+viruses$confusion_matrix_vs2_keep2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3<1] <- "false negative"
+viruses$confusion_matrix_vs2_keep2[viruses$seqtype=="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "true positive"
+viruses$confusion_matrix_vs2_keep2[viruses$seqtype!="virus" & viruses$keep_score_virsorter2_SOP3>=1] <- "false positive"
+viruses$confusion_matrix_vs2_keep2[viruses$keep_score_virsorter2_SOP3==0.75] <- "true positive"
+viruses$confusion_matrix_vs2_keep2[viruses$keep_score_virsorter2_SOP3==0.5] <- "true positive"
+
+viruses$accurate_vs2_keep2 <- "false"
+viruses$accurate_vs2_keep2[grep("true", viruses$confusion_matrix_vs2_keep2)] <- "true"
 ```
@@
 ```{r}
-length(grep("true", viruses$accurate_vs2_SOP))/nrow(viruses)
+length(grep("true", viruses$accurate_vs2_SOP2))/nrow(viruses)
+```
+```{r}
+length(grep("true", viruses$accurate_vs2_SOP3))/nrow(viruses)
 ```
@@
 ```{r}
-length(grep("true positive", viruses$confusion_matrix_vs2_sop))/length(grep("positive", viruses$confusion_matrix_vs2_sop))
+length(grep("true positive", viruses$confusion_matrix_vs2_sop2))/length(grep("positive", viruses$confusion_matrix_vs2_sop2))
+```
+```{r}
+length(grep("true positive", viruses$confusion_matrix_vs2_sop3))/length(grep("positive", viruses$confusion_matrix_vs2_sop3))
+```
+```{r}
+length(grep("true positive", viruses$confusion_matrix_vs2_keep2))/length(grep("positive", viruses$confusion_matrix_vs2_keep2))
 ```
@@
 ```{r}
-length(grep("true positive", viruses$confusion_matrix_vs2_sop))/(length(grep("true positive", viruses$confusion_matrix_vs2_prsop))+length(grep("false negative", viruses$confusion_matrix_vs2_sop)))
+length(grep("true positive", viruses$confusion_matrix_vs2_sop2))/(length(grep("true positive", viruses$confusion_matrix_vs2_sop2))+length(grep("false negative", viruses$confusion_matrix_vs2_sop2)))
 ```
+```{r}
+length(grep("true positive", viruses$confusion_matrix_vs2_sop3))/(length(grep("true positive", viruses$confusion_matrix_vs2_sop3))+length(grep("false negative", viruses$confusion_matrix_vs2_sop3)))
+```
+```{r}
+length(grep("true positive", viruses$confusion_matrix_vs2_keep2))/(length(grep("true positive", viruses$confusion_matrix_vs2_keep2))+length(grep("false negative", viruses$confusion_matrix_vs2_keep2)))
+```

### 2023-04-01 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set_dw.Rmd
Updating figures.

@@
 ```{r}
-viruses <- left_join(x=vs2_c, y=checkV, by = c("contig"))
+viruses <- left_join(x=checkV, y=vs2_c, by = c("contig"))
 viruses <- left_join(x=viruses, y=dvf_c, by = c("contig", "assembly"))

### 2023-04-01 TestingPipeline/Scripts/examining_false_positives-short_fragments.Rmd
Updating figures.

@@
                          include_tv_5=F) {
-    # tuning addition
-  if (include_tv_1) {
-    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- 0
-  }
+    # tuning removal
   if (include_tv_2) {
@@
   }
-  if (include_tv_4) {
-    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
-  }
 
@@
     current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  if (include_tv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
+  }
+
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- -3
   }

### 2023-04-01 TestingPipeline/Scripts/examining_false_positives.Rmd
Updating figures.

@@
 
-```{r getting_viral_set_1}
+```{r}
 getting_viral_set_1 <- function(input_seqs,
@@
 viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
+                                              include_vibrant = F,
                                               include_virsorter2 = T,

### 2023-04-01 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating figures.

@@
     current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
-  }
-  if (include_tv_4) {
-    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- 0
   }
@@
 ## Calculate the performance of each pipeline
-```{r}
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
-                      precision=rep(0, nrow(combos_list)*10),
-                       recall=rep(0, nrow(combos_list)*10),
-                       F1=rep(0, nrow(combos_list)*10),
-                       MCC=rep(0, nrow(combos_list)*10), 
-                      AUC=rep(0, nrow(combos_list)*10))
+quantifying variability between rule sets
+```{r}
+accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
+                      precision=rep(0, nrow(combos_list)*8),
+                       recall=rep(0, nrow(combos_list)*8),
+                       F1=rep(0, nrow(combos_list)*8),
+                       MCC=rep(0, nrow(combos_list)*8), 
+                      AUC=rep(0, nrow(combos_list)*8))
 
@@
                               build_score_list(viruses[viruses$Index==1,], combos_list))
-for (i in 2:10) {
+for (i in 2:8) {
   accuracy_scores <- rbind(accuracy_scores,
@@
 }
+```
+
+```{r}
+MCC_sd_quant <- cbind(accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1], as.data.frame(matrix(data=0, nrow=63, ncol=7)))
+
+for (i in (2:8)) {
+  temp <- accuracy_scores %>%
+    filter(testing_set_index<=i) %>%
+    select(MCC, toolcombo) %>%
+    group_by(toolcombo) %>%
+    summarise(sd = sd(MCC))
+
+  MCC_sd_quant[,i] <- temp$sd
+}
+
+colnames(MCC_sd_quant) <- c("testing_set_index", "1-2", "1-3", "1-4",
+                            "1-5", "1-6", "1-7", "1-8")
+
+MCC_sd_quant <- MCC_sd_quant %>% pivot_longer(cols=contains("-"))
+
+ggplot(MCC_sd_quant, aes(x=testing_set_index, y=value, 
+                                  color=name, fill=name)) +
+  geom_point(alpha=0.5)
+
+ggplot(MCC_sd_quant, aes(x=name, y=value)) +
+  geom_boxplot(alpha=0.5, color=pal(4)[1]) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 0),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Testing Sets") +
+  ylab("Standard Deviation") 
+```
+
+## Calculate the performance of each pipeline
+```{r}
+accuracy_scores <- accuracy_scores %>% filter(testing_set_index<=5)
 ```
@@
 ```
+
+### Visualize performance across all the rules
 
@@
 
-ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
+fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                   color=numrules, fill=numrules)) +
@@
   scale_fill_manual(name="Number of Rule Sets",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Rule Sets",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
+
+ggsave(
+  "../IntermediaryFiles/all_MCC_scatterplot.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in", "cm", "mm", "px"),
+  dpi = 300,
+  limitsize = TRUE
+)
 
@@
   scale_fill_manual(name="Number of Tools",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Tools",
-                     values = alpha(rev(pal(6)), 1))
-```
+                     values = alpha(rev(viridis(6)), 1))
+```
+
+making heatmap for the x-axis label
+```{r}
+viruses_1 <- accuracy_scores %>% filter(testing_set_index==1)
+
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2
+)
+
+ann_colors = list(
+  tnv = c("black", pal(6)[1]),
+  tv = c("black", pal(6)[3]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
+
+rownames(ann_col) <- combos_list$toolcombo
+
+viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==1)
+
+rownames(viral_scores_1) <- viral_scores_1$toolcombo
+
+viral_scores_1 <- viral_scores_1 %>%
+  select(c("MCC"))
+
+pheatmap::pheatmap(as.matrix(viral_scores_1),
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = NA,
+                  show_rownames = F,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows = F,
+                  cluster_cols = F,
+                  cellheight=3,
+                  filename = "../IntermediaryFiles/heatmap_MCC_labels.png"
+                  )
+```
+
 ```{r}
@@
   xlab("Number of Rule Sets") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 1)) +
   facet_wrap(~performance_metric)
@@
   scale_fill_manual(name="Number of Tools",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Tools",
-                     values = alpha(rev(pal(6)), 1)) +
+                     values = alpha(rev(viridis(6)), 1)) +
   facet_wrap(~performance_metric)
@@
   xlab("Number of Tools") +
-  scale_fill_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 1)) +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 1)) +
   facet_wrap(~performance_metric)
@@
 viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = F,
+                                              include_vibrant = T,
                                               include_virsorter2 = T,
@@
 viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_high"), size_class, true_virus) %>%
+  select(contains("keep_score_high"), true_virus) %>%
   pivot_longer(cols=contains("keep_score_"), 
@@
   mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, true_virus) %>%
+  group_by(rule_combination, viral_score, true_virus) %>%
   summarise(n = n())
 
-viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
-```
-
-```{r}
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
+
 ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
-  geom_bar(stat="identity") +
+  geom_bar(stat="identity", color="black") +
   theme_light() +
@@
   facet_grid(~true_virus, scales = "free")
+```
+
+```{r}
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_high"), max_score_group, true_virus) %>%
+  filter(true_virus=="virus") %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, max_score_group) %>%
+  summarise(n = n())
+
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~max_score_group, scales = "free") +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_high"), max_score_group, true_virus) %>%
+  filter(true_virus=="not") %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, max_score_group) %>%
+  summarise(n = n())
+
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~max_score_group, scales = "free") +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
 ```
@@
   mutate(viral_score=as.factor(round(viral_score))) %>%
-  group_by(rule_combination, viral_score, size_class, seqtype) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
   summarise(n = n())
 
-viruses_long_scores$size_class <- factor(viruses_long_scores$size_class,
-                                    levels = c("3-5kb", "5-10kb", ">10kb"))
-```
-
-```{r}
 pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
-pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
 
@@
     panel.border = element_blank(),
-    legend.position = "bottom"
+    legend.position = "bottom",
+    axis.text.x=element_text(size=8, angle = 45),
   ) +
@@
   ylab("Number of Sequences") + 
-  facet_grid(~seqtype, scales = "free")
+  facet_grid(~seqtype, scales = "free") +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
 ```
@@
     panel.border = element_blank(),
-    legend.position = "bottom"
+    legend.position = "bottom",
+    axis.text.x=element_text(size=8, angle = 45),
   ) +
   scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
   xlab("") +
   ylab("Number of Sequences") + 
-  facet_grid(~seqtype, scales = "free")
+  facet_grid(~seqtype, scales = "free") +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
 ```
@@
 ```
+
+comparing each tool's score
+```{r}
+viruses$keep_score_ind_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = T,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_ind_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_ind_"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+```
+Fig 4
+
+```{r}
+num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+
+for (i in 1:63) {
+  for (j in 1:63) {
+    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+
+    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
+  }
+}
+
+rownames(num_high_viruses) <- combos_list$toolcombo
+colnames(num_high_viruses) <- combos_list$toolcombo
+
+rownames(prop_high_viruses) <- combos_list$toolcombo
+colnames(prop_high_viruses) <- combos_list$toolcombo
+```
+
+```{r}
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    axis.text.x=element_text(size=8, angle = 45)
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free") +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 2)) +
+  scale_x_discrete(labels=rev(c("VirSorter2", "VirSorter", "VIBRANT",
+                            "tuning viral", "tuning not viral",
+                            "DeepVirFinder")))
+```
+
+# Comparing all of the tool methods - Fig 2A
+```{r}
+combos_list <- data.frame(toolcombo=rep(0, 64),
+                          tune_not_viral=rep(0, 64),
+                          DVF=rep(0, 64),
+                          tune_viral=rep(0, 64),
+                          VIBRANT=rep(0, 64),
+                          VS=rep(0, 64),
+                          VS2=rep(0, 64))
+p <- 1
+
+for (i in c(0,1)){
+  for (j in c(0,1)){
+    for (k in c(0,1)){
+      for (l in c(0,1)){
+        for (m in c(0,1)){
+          for (n in c(0,1)){
+            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
+            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
+                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
+            combos_list$tune_not_viral[p] <- i
+            combos_list$DVF[p] <- j
+            combos_list$tune_viral[p] <- k
+            combos_list$VIBRANT[p] <- l
+            combos_list$VS[p] <- m
+            combos_list$VS2[p] <- n
+            p <- p+1
+          }
+        }
+      }
+    }
+  }
+}
+
+combos_list <- combos_list[-1,]
+
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
+                          num_viruses=rep(0, nrow(combos_list)))
+
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning_viral = combos_list$tune_viral[i],
+                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+
+  if (max(viral_scores[,i])<=0) {
+    num_viruses$num_viruses[i] <- 0
+  }
+  else {
+    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
+  }
+  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
+
+  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
+}
+num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
+
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+viruses_1 <- viruses[viruses$Index=="1",]
+
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2
+)
+
+ann_row = data.frame(
+    score = viruses_1$seqtype
+)
+
+ann_colors = list(
+  tnv = c("black", pal(6)[1]),
+  tv = c("black", pal(6)[3]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6]),
+  score = c("archaea"=pal(6)[1], "bacteria"=pal(6)[2], "fungi"=pal(6)[3],
+              "plasmid"=pal(6)[4], "protist"=pal(6)[5], "virus"=pal(6)[6])
+)
+
+rownames(ann_col) <- combos_list$toolcombo
+rownames(ann_row) <- viruses_1$contig
+
+viral_scores_1 <- viral_scores[viruses$Index=="1",]
+
+colnames(viral_scores_1) <- combos_list$toolcombo
+rownames(viral_scores_1) <- viruses_1$contig
+
+pheatmap::pheatmap(t(viral_scores_1),
+                  annotation_col = ann_row,
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  #border_color = "black",
+                  color = c(rev(RColorBrewer::brewer.pal(3,"Oranges")),
+                            RColorBrewer::brewer.pal(5,"Purples") 
+                            ),
+                  show_rownames = F,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows = F,
+                  cluster_cols = F,
+                  cellheight=3,
+                  filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
+                  )
+```
+
+# comparing all rules against each other - Figure 5
+```{r}
+num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+
+for (i in 1:63) {
+  for (j in 1:63) {
+    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+
+    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
+  }
+}
+
+rownames(num_high_viruses) <- combos_list$toolcombo
+colnames(num_high_viruses) <- combos_list$toolcombo
+
+rownames(prop_high_viruses) <- combos_list$toolcombo
+colnames(prop_high_viruses) <- combos_list$toolcombo
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2,
+    num_vir = colSums(viral_scores>=1)+1
+)
+
+ann_colors = list(
+  num_vir=magma(100),
+  tnv = c("black", pal(6)[1]),
+  tv = c("black", pal(6)[3]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
+
+rownames(ann_col) <- combos_list$toolcombo
+
+pheatmap::pheatmap(prop_high_viruses,
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = "black",
+                  color = viridis(100),
+                  cutree_rows = 3,
+                  cutree_cols = 3,
+                  show_rownames = F,
+                  show_colnames = F,
+                  filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common.png",
+                  annotation_legend = F)
+```

### 2023-04-01 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
Updating figures with current rules.
Updating figures.

--- 2023-02-14 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
@@
 ```{r}
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
-                      precision=rep(0, nrow(combos_list)*10),
-                       recall=rep(0, nrow(combos_list)*10),
-                       F1=rep(0, nrow(combos_list)*10),
-                       MCC=rep(0, nrow(combos_list)*10), 
-                      AUC=rep(0, nrow(combos_list)*10))
+accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*5),
+                      precision=rep(0, nrow(combos_list)*5),
+                       recall=rep(0, nrow(combos_list)*5),
+                       F1=rep(0, nrow(combos_list)*5),
+                       MCC=rep(0, nrow(combos_list)*5), 
+                      AUC=rep(0, nrow(combos_list)*5))
 
@@
                               build_score_list(viruses[viruses$Index==1,], combos_list))
-for (i in 2:10) {
+for (i in 2:5) {
   accuracy_scores <- rbind(accuracy_scores,
@@
 ```{r}
+accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
+
+accuracy_scores_melt <- accuracy_scores %>% 
+  select(testing_set_index, precision, recall, MCC, numrules, toolcombo) %>%
+  pivot_longer(cols=c(precision, recall, MCC), 
+               names_to="performance_metric",
+               values_to="performance_metric_score")
+```
+
+```{r}
 ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
@@
 ```{r}
+accuracy_scores$numrules <- as.factor(accuracy_scores$numrules)
 one.way <- aov(MCC ~ numrules, data = accuracy_scores)

### 2023-04-02 TestingPipeline/Scripts/paper_figure_generation.Rmd
Fixing tv and tnv being swapped in one figure label.

@@
 ```{r}
-accuracy_scores <- accuracy_scores[order(accuracy_scores$MCC, decreasing=F),]
-accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+mean_MCC <- accuracy_scores %>% 
+  select(MCC, toolcombo) %>%
+  group_by(toolcombo) %>%
+  summarise(mean = mean(MCC)) %>%
+  arrange(mean)
+```
+
+```{r}
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_MCC$toolcombo)
 
@@
 ```{r}
-viruses_1 <- accuracy_scores %>% filter(testing_set_index==1)
-
 ann_col = data.frame(
-    tnv = combos_list$tune_not_viral,
-    tv = combos_list$tune_viral,
+    tnv = combos_list$tune_viral,
+    tv = combos_list$tune_not_viral,
     dvf = combos_list$DVF,
@@
 )
+rownames(ann_col) <- combos_list$toolcombo
 
 ann_colors = list(
+  tv = c("black", pal(6)[3]),
   tnv = c("black", pal(6)[1]),
-  tv = c("black", pal(6)[3]),
   dvf = c("black", pal(6)[2]),
@@
 
-rownames(ann_col) <- combos_list$toolcombo
-
-viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==1)
-
-rownames(viral_scores_1) <- viral_scores_1$toolcombo
-
-viral_scores_1 <- viral_scores_1 %>%
-  select(c("MCC"))
+viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
+#viral_scores_1 <- mean_MCC
+#rn <- viral_scores_1$toolcombo
+
+viral_scores_1 <- viral_scores_1 %>% select(c("MCC"))
+#viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
+rownames(viral_scores_1) <- rn
 
@@
                   border_color = NA,
-                  show_rownames = F,
+                  show_rownames = T,
                   show_colnames = F,
@@
                   )
-```
-
-```{r}
-mean_MCC <- accuracy_scores %>% 
-  select(MCC, toolcombo) %>%
-  group_by(toolcombo) %>%
-  summarise(mean = mean(MCC))
 ```

### 2023-04-06 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Updating SI figures.

@@
 viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
 ```
@@
 the chunks for the appropriate tools.
-
-```{r}
-viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
-```
 
@@
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
 ```

### 2023-04-06 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Updating SI figures.

@@
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.3)) +
+                     values = alpha(rev(viridis(6)), 0.3)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 0.7))
+                     values = alpha(rev(viridis(6)), 0.7))
 
@@
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.3)) +
+                     values = alpha(rev(viridis(6)), 0.3)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 0.7))
+                     values = alpha(rev(viridis(6)), 0.7))
 ```
@@
   scale_fill_manual(name="",
-                     values = alpha(pal(6), 0.3)) +
+                     values = alpha(viridis(6), 0.3)) +
   scale_color_manual(name="",
-                     values = alpha(pal(6), 0.7))
+                     values = alpha(viridis(6), 0.7))
 
@@
   scale_fill_manual(name="",
-                     values = alpha(pal(6), 0.3)) +
+                     values = alpha(viridis(6), 0.3)) +
   scale_color_manual(name="",
-                     values = alpha(pal(6), 0.7))
+                     values = alpha(viridis(6), 0.7))
 ```

### 2023-04-06 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating SI figures.

@@
                                                if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
-            combos_list$tune_not_viral[p] <- i
+            combos_list$tune_not_viral[p] <- k
             combos_list$DVF[p] <- j
-            combos_list$tune_viral[p] <- k
+            combos_list$tune_viral[p] <- i
             combos_list$VIBRANT[p] <- l
@@
   scale_fill_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
 ```
@@
   scale_fill_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
 
@@
 ann_col = data.frame(
-    tnv = combos_list$tune_viral,
-    tv = combos_list$tune_not_viral,
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
     dvf = combos_list$DVF,
@@
 ```
+
+# numbers in common between two tool combos
+```{r}
+combos_list_two <- combos_list
+combos_list_two$numtools <- str_count(combos_list_two$toolcombo, "1")
+combos_list_two <- combos_list_two[combos_list_two$numtools==1,]
+
+two_tool_pairs <- num_high_viruses[rownames(num_high_viruses) %in% combos_list_two$toolcombo,
+                                   colnames(num_high_viruses) %in% combos_list_two$toolcombo]
+
+colnames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+rownames(two_tool_pairs) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+```
+
+```{r}
+viral_scores_two <- viral_scores[,colnames(num_high_viruses) %in% combos_list_two$toolcombo]
+
+n <- 6
+viral_scores_two_comb <- matrix(data=NA, nrow=n, ncol=n)
+
+for (i in 1:n) {
+  for (j in 1:n) {
+    if (i==j) {
+      viral_scores_two_comb[i,j] <- sum(viral_scores_two[,i]>=1)
+    }
+    else if (i<j) {
+      viral_scores_two_comb[i,j] <- sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)
+    }
+    else if (i>j) {
+      viral_scores_two_comb[i,j] <- paste(round(sum((viral_scores_two[,i]+viral_scores_two[,j])>=1)/sum(viral_scores_two[,i]>=1 | viral_scores_two[,j]>=1)*100, digits=0), "%", sep="")
+    }
+  }
+}
+
+colnames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+rownames(viral_scores_two_comb) <- c("VS2", "VS", "VB", "tv", "DVF", "tnv")
+```
+
+```{r}
+viral_scores_two_comb
+```
+
+comparing VS2 and VIBRANT viruses - Figure S15
+```{r}
+vs2_vb <- data.frame(vs2=viral_scores_two[,1],
+                     vb=viral_scores_two[,3],
+                     comb_score=viral_scores_two[,1]+viral_scores_two[,3],
+                     seqtype=viruses$seqtype,
+                     max_score_group=viruses$max_score_group)
+
+vs2_vb$comb[vs2_vb$comb_score<1] <- "neither"
+vs2_vb$comb[vs2_vb$vs2==0.5 & vs2_vb$vb==0.5] <- "both_low"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0.5] <- "vs2_high_vb_low"
+vs2_vb$comb[vs2_vb$vs2>=0.5 & vs2_vb$vb==1] <- "vs2_low_vb_high"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb==0] <- "vs2_only"
+vs2_vb$comb[vs2_vb$vs2==0 & vs2_vb$vb>=1] <- "vb_only"
+vs2_vb$comb[vs2_vb$vs2>=1 & vs2_vb$vb>=1] <- "both_high"
+
+#vs2_vb$true_virus <- "not"
+#vs2_vb$true_virus[vs2_vb$seqtype=="virus"] <- "virus"
+
+vs2_vb_seqtype <- vs2_vb %>% 
+  group_by(comb, comb_score, seqtype) %>%
+  summarise(n = n())
+```
+
+```{r}
+table(vs2_vb$comb)
+```
+
+```{r}
+ggplot(vs2_vb_seqtype, aes(y=n, x=comb,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr", ) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
+# overlap in proportion of contigs for tool sets - Figure S16
+
+get number of tools and number of rules for each combination
+```{r}
+ann_col$numrules <- rowSums(ann_col[,1:6])
+ann_col$numtools <- ann_col$numrules
+
+for (i in 1:nrow(ann_col)) {
+  if (ann_col$tv[i]==1) {
+    ann_col$numtools[i] <- ann_col$numtools[i] + 3
+  }
+
+  if (ann_col$tnv[i]==1) {
+    ann_col$numtools[i] <- ann_col$numtools[i] + 5
+  }
+}
+
+ann_col$numtools[ann_col$numtools>6] <- 6
+
+ann_col$tool <- rownames(ann_col)
+```
+
+```{r}
+prop_high_viruses_long <- as.data.frame(prop_high_viruses)
+prop_high_viruses_long$tool <- combos_list$toolcombo
+
+prop_high_viruses_long <- as.data.frame(prop_high_viruses_long) %>% 
+  pivot_longer(!tool,
+               values_to="proportion",
+               names_to="tool2") %>%
+  inner_join(ann_col, by=c("tool"))
+
+```
+
+```{r}
+head(prop_high_viruses_long)
+```
+
+```{r}
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs2==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vs==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$vb==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.7))
+
+ggplot(prop_high_viruses_long[prop_high_viruses_long$dvf==1,], aes(x=as.factor(numrules), y=proportion, 
+                                  color=as.factor(numrules), fill=as.factor(numrules))) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Number of Rules") +
+  ylab("Proportion of Contigs in Common") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.7))
+```

### 2023-04-15 TestingPipeline/Scripts/paper_figure_generation.Rmd
TV vs TNV rule analysis.

@@
 
+## visualizing a select set of tools
+```{r}
+viruses$keep_score_high_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses$keep_score_high_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses$tv_vs_tv_ntv <- "not tv"
+viruses$tv_vs_tv_ntv[viruses$keep_score_high_tv>=1 & viruses$keep_score_high_tv_tnv>=1] <- "tnv_preserved"
+viruses$tv_vs_tv_ntv[viruses$keep_score_high_tv>=1 & viruses$keep_score_high_tv_tnv<1] <- "tnv lost"
+
+table(viruses$tv_vs_tv_ntv, viruses$true_virus)
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_high"), true_virus) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, true_virus) %>%
+  summarise(n = n())
+
+pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~true_virus, scales = "free")
+```
+
 # Comparing all of the tool methods - Fig 2A
@@
 ```
+
+# Comparing RefSeq vs VS2 database
+VS2 sequences used: (/scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/NCLDV/env/NCLDV_env.fna /scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/lavidaviridae/env/HQ_virophages.fna /scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/dsDNAphage/env/* /scratch/kwigg_root/kwigg/hegartyb/VSTE/non-refseq-genomes/dsDNAphage/provirus/VirSorterCuratedDataset_HQ_contigs.fna)
+
+```{r}
+refseq_vir <- read_tsv("../IntermediaryFiles/refseq_viral_genome_ids.txt",col_names = F)
+refseq_vir$X1 <- sub(">", "", refseq_vir$X1)
+refseq_vir <- separate(refseq_vir, col=X1, sep=" ", remove=T, into="id")
+refseq_vir$id <- sub("\\.", "_", refseq_vir$id)
+```
+
+```{r}
+table(refseq_vir$id %in% viruses$contig)
+```
+
+```{r}
+viruses_refseq_vir <- viruses[viruses$contig %in% refseq_vir$id,]
+viruses_not_refseq_vir <- viruses[!(viruses$contig %in% refseq_vir$id) & viruses$seqtype=="virus",]
+```
+
+```{r}
+table(viruses_refseq_vir$max_score_group)
+table(viruses_not_refseq_vir$max_score_group)
+```
+
+```{r}
+viruses_refseq_vir$keep_score_high_MCC <- getting_viral_set_1(viruses_refseq_vir, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses_not_refseq_vir$keep_score_high_MCC <- getting_viral_set_1(viruses_not_refseq_vir, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+sum(viruses_refseq_vir$keep_score_high_MCC>=1)/nrow(viruses_refseq_vir)
+```
+
+```{r}
+sum(viruses_not_refseq_vir$keep_score_high_MCC>=1)/nrow(viruses_not_refseq_vir)
+```

### 2023-04-15 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
TV vs TNV rule analysis.

@@
 viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined-fragments.tsv")
+```
+```{r}
+viruses$kaiju_match_ratio_color <- "less than 0.3"
+viruses$kaiju_match_ratio[viruses$kaiju_match_ratio>=0.3] <- "greater than or equal to 0.3"
+
+ggplot(viruses, aes(x=kaiju_match_ratio, color=kaiju_match_ratio_color, fill=kaiju_match_ratio_color)) +
+  geom_histogram() +
+  theme_bw() +
+  facet_wrap(~seqtype, scales = "free") +
+  #scale_color_viridis_d(begin=0.2, end=0.8) +
+  #scale_fill_viridis_d(begin=0.2, end=0.8, alpha=0.2) +
+  xlab("Kaiju Match Ratio")
 ```

### 2023-05-26 TestingPipeline/Scripts/comparing_methods_visualization.Rmd
Comparing 50:50 vs 90:10 not viral:viral splits.

@@
 ```{r}
-length(grep("true positive", viruses_refseq_vir$confusion_matrix_high_MCC))/nrow(viruses_refseq_vir)
+length(grep("true positive", viruses_refseq_vir$))/nrow(viruses_refseq_vir)
 ```
@@
 ```
+
+```{r}
+ggplot(vs2_tnv_seqtype, aes(y=n, x=seqtype,
+                   fill=as.factor(comb_score))) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_brewer(palette = "PuOr") +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(comb~max_score_group, scales = "free")
+```

### 2023-05-26 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Comparing 50:50 vs 90:10 not viral:viral splits.

@@
   summarise(n = n())
+
+viruses_long_scores$rule_type <- "viral sorting tool"
+viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_tv"] <- "tuning"
+viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_ntv"] <- "tuning"
+
 ```
@@
 pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
-pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
+
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Number of Sequences") + 
+  facet_grid(rule_type~seqtype, scales = "free") 
+```
+
+making a smaller test set of rules
+Function that allows for the comparing pieces of the pipeline
+```{r getting_viral_set_1}
+getting_viral_set_mod <- function(input_seqs,
+                                include_vibrant=FALSE, 
+                                include_virsorter2=FALSE,
+                                include_deepvirfinder=FALSE,
+                                include_tuning_viral=FALSE,
+                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
+                                include_tuning_not_viral=FALSE,
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
+                                include_virsorter=FALSE) {
+
+  keep_score <- rep(0, nrow(input_seqs))
+
+  if (include_vibrant) {
+    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
+  }
+
+  if (include_virsorter2) {
+    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
+  }
+
+  if (include_virsorter) {
+    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
+    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
+    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
+    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
+    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
+  }
+
+  if (include_deepvirfinder) {
+     # add if DVF calls viral
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+  }
+
+  if (include_tuning_viral) {
+    keep_score <- tuning_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=tv_1,
+                         include_tv_2=tv_2,
+                         include_tv_3=tv_3,
+                         include_tv_4=tv_4)
+  }
+
+  if (include_tuning_not_viral) {
+    # tuning removal 
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5)
+
+  }
+
+  return(keep_score)
+
+}
+
+tuning_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+  }
+  if (include_tv_2) {
+    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
+  }
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
+  if (include_tv_3) {
+    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
+  }
+  if (include_tv_4) {
+    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
+  }
+  return(current_score)
+}
+
+tuning_not_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=F) {
+    # tuning removal
+  if (include_tv_2) {
+    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
+  }
+  if (include_tv_3) {
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+  }
+
+  if (include_tv_5) {
+    # remove if DVF calls it not viral
+    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  if (include_tv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
+  }
+
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- -3
+  }
+
+  return(current_score)
+}
+```
+
+same idea but with a VS2 base
+```{r}
+viruses$keep_score_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=T, tv_4=T,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=T, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_1_2_3_4_tv_1_3_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=T, tv_4=T,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_all_tv_1_3_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=T, tv_4=T,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_1_2_3_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_1_3_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=T, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vs2"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_vs2"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
 
@@
   scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
   xlab("") +
@@
 ```
+
+```{r}
+viruses$keep_score_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_1_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_1_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_1_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_1_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_2_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_2_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=T, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_2_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=T,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_3_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_3_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=T,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_4_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=T,
+                                              include_virsorter = F)
+viruses$keep_score_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_ntv"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_ntv"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```
+
+```{r}
+viruses$keep_score_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=F, tv_2=T, tv_3=F, tv_4=F,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=F, tv_2=F, tv_3=T, tv_4=F,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=F, tv_2=F, tv_3=F, tv_4=T,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_1_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=T, tv_2=T, tv_3=F, tv_4=F,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_1_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=T, tv_2=F, tv_3=T, tv_4=F,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_1_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=T, tv_2=F, tv_3=F, tv_4=T,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_2_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=F, tv_2=T, tv_3=T, tv_4=F,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_2_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=F, tv_2=T, tv_3=F, tv_4=T,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_3_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             tv_1=F, tv_2=F, tv_3=T, tv_4=T,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+viruses$keep_score_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                             include_vibrant = F,
+                                             include_virsorter2 = F,
+                                             include_tuning_viral = T,
+                                             include_tuning_not_viral = F,
+                                             include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_tv"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_tv"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom"
+  ) +
+  scale_fill_manual(name="",
+                     values = alpha(c(pal_orange[1], pal_purple), 0.7)) +
+  xlab("") +
+  ylab("Number of Sequences") + 
+  facet_grid(~seqtype, scales = "free")
+```

### 2023-05-26 TestingPipeline/Scripts/paper_figure_generation_balanced_testing.Rmd
Comparing 50:50 vs 90:10 not viral:viral splits.

--- 2023-04-15 TestingPipeline/Scripts/paper_figure_generation.Rmd
@@
 viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
+```
+
+only 5 testing sets   
+```{r}
+viruses <- viruses %>% filter(Index<=5)
+```
+
+balancing testing set
+1000 viruses per testing set
+600 bacteria
+100 arachaea
+100 fungi
+100 plasmid
+100 protist
+```{r}
+viruses <- rbind(viruses %>% filter(seqtype=="archaea") %>%
+  group_by(Index) %>%
+  slice_sample(n=100),
+  viruses %>% filter(seqtype=="bacteria") %>%
+  group_by(Index) %>%
+  slice_sample(n=600),
+  viruses %>% filter(seqtype=="fungi") %>%
+  group_by(Index) %>%
+  slice_sample(n=100),
+  viruses %>% filter(seqtype=="plasmid") %>%
+  group_by(Index) %>%
+  slice_sample(n=100),
+  viruses %>% filter(seqtype=="protist") %>%
+  group_by(Index) %>%
+  slice_sample(n=100),
+  viruses %>% filter(seqtype=="virus") %>%
+  group_by(Index) %>%
+  slice_sample(n=1000)
+)
+
 ```
@@
 ```{r}
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
-                      precision=rep(0, nrow(combos_list)*8),
-                       recall=rep(0, nrow(combos_list)*8),
-                       F1=rep(0, nrow(combos_list)*8),
-                       MCC=rep(0, nrow(combos_list)*8), 
-                      AUC=rep(0, nrow(combos_list)*8))
+accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*5),
+                      precision=rep(0, nrow(combos_list)*5),
+                       recall=rep(0, nrow(combos_list)*5),
+                       F1=rep(0, nrow(combos_list)*5),
+                       MCC=rep(0, nrow(combos_list)*5), 
+                      AUC=rep(0, nrow(combos_list)*5))
 
@@
                               build_score_list(viruses[viruses$Index==1,], combos_list))
-for (i in 2:8) {
+for (i in 2:5) {
   accuracy_scores <- rbind(accuracy_scores,
@@
 }
+```
+
+```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
 ```
@@
 ```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-```
-
-```{r}
 accuracy_scores_melt <- accuracy_scores %>% 
@@
 
+getting set of tools with roughly equivalent precision
+```{r}
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 0"])
+
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 1 0 0"])
+
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 0 0 1"])
+
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 1"])
+```
+accounting for multiple t-tests with Bonferroni, then the following have the same precision:
+- vs2+tnv, vb+tnv, vb, and vs2
+
 recall
@@
 
+getting set of tools with roughly equivalent recall
+```{r}
+t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
+       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 1 1 1"])
+
+t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
+       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 0 1"])
+
+t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
+       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 1 0 1"])
+
+t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
+       accuracy_scores$recall[accuracy_scores$toolcombo=="1 0 0 0 1 1"])
+
+t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
+       accuracy_scores$recall[accuracy_scores$toolcombo=="1 0 0 1 1 1"])
+
+t.test(accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 0 0 1 1"], 
+       accuracy_scores$recall[accuracy_scores$toolcombo=="1 1 1 0 1 1"])
+```
+accounting for multiple t-tests with Bonferroni, then the following have the same recall:
+- vs2+vs+dvf+tv, vs2+vs+dvf+tv+vb, vs2+dvf+tv, vs2+vs+dvf+tv, vs2+dvf+tv+vb,
+
 precision vs recall
@@
 accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_MCC$toolcombo)
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
+                                  color=numrules, fill=numrules)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=20),
+  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("MCC") +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 1)) + 
+  ylim(c(0,1))
+```
+
+```{r}
 
@@
 ```
+```{r}
+max(accuracy_scores$MCC)
+```
+=78
+
+getting set of tools with roughly equivalent MCC
+```{r}
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="0 1 1 0 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 0 0 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 1 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 1 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 0 0 1"])
+```
+accounting for multiple t-tests with Bonferroni, then the following have the same MCC:
+- vs2+tnv, vs2+tnv+dvf, vs2, vs2+tnv+tv+vb, vs2+tnv+tv+vs
 
@@
 
+ggplot(accuracy_scores_melt[,], aes(x=numrules, y=performance_metric_score, 
+                                  color=numrules, fill=numrules)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Score") +
+  xlab("Number of Rule Sets") +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 1)) +
+  facet_wrap(~performance_metric)
+
 ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
@@
 
+```{r}
+viruses$keep_score_all_tv_only_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_all_tv_only_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_all_tv_only_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_all_tv_only_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_tv_only"))
+
+viruses_long_scores <- apply(viruses_long_scores, c(1,2), function(x) {if (x>0){x<-1} else{x<-0}})
+
+viruses_long_scores <- as.data.frame(cbind(viruses$true_virus, rowSums(viruses_long_scores)))
+
+table(viruses_long_scores$V1, viruses_long_scores$V2>1)
+```
+
 same idea but VS2 base
@@
                                               include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+ viruses$keep_score_vs2_tv_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
                                               include_virsorter = F)
@@
 pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
-pal_orange <- RColorBrewer::brewer.pal(1, "Oranges")
+pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
 
@@
   scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange[1]), pal_purple), 0.8)) +
+                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
   xlab("") +
@@
                                               include_tuning_viral = T,
-                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
                                               include_tuning_not_viral = T,
@@
                                               include_tuning_viral = T,
-                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
                                               include_tuning_not_viral = T,
@@
                                               include_tuning_viral = T,
-                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=T,
                                               include_tuning_not_viral = T,
@@
 ```{r}
+table(viruses$keep_score_all_ntv_0, viruses$keep_score_all_ntv_1)
+```
+
+```{r}
+table(viruses$keep_score_all_ntv_4, viruses$keep_score_all_ntv_5)
+```
+
+```{r}
 ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
@@
 ```{r}
-num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
-prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
-
-for (i in 1:63) {
-  for (j in 1:63) {
-    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
-
-    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
-    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
-  }
-}
-
-rownames(num_high_viruses) <- combos_list$toolcombo
-colnames(num_high_viruses) <- combos_list$toolcombo
-
-rownames(prop_high_viruses) <- combos_list$toolcombo
-colnames(prop_high_viruses) <- combos_list$toolcombo
-```
+table(viruses$true_virus, viruses$keep_score_ind_dvf)
+table(viruses$true_virus, viruses$keep_score_ind_vs2)
+table(viruses$true_virus, viruses$keep_score_ind_vs)
+table(viruses$true_virus, viruses$keep_score_ind_vb)
+```
+seems like false positive rate is way high for 0.5 across all
+
+```{r}
+table(viruses$true_virus, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2)
+```
+
+```{r}
+table(viruses$true_virus, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv)
+```
+
+```{r}
+table(viruses$keep_score_ind_vs2, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv)
+```
+with 0.5 rules: 1464 sequences predicted viral that weren't viral by vs2 on its own
+without them: 1234 sequences added
+
+```{r}
+table(viruses$keep_score_ind_vs2, viruses$keep_score_ind_dvf+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv,
+      viruses$true_virus)
+```
+with 0.5 rules: however, of these, only 432 (30%) are truly viruses
+without 0.5 rules: only 375 (30%) are truly viruses
+
+```{r}
+table(viruses$keep_score_ind_vs2, viruses$keep_score_ind_vb+viruses$keep_score_ind_vs2+viruses$keep_score_ind_ntv,
+      viruses$true_virus)
+```
+with 0.5 rules: however, of these, only 74 (7%) are truly viruses
+without 0.5 rules: only 87 (4%) are truly viruses
 
@@
                             "DeepVirFinder")))
+```
+
+```{r}
+num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
+
+for (i in 1:63) {
+  for (j in 1:63) {
+    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+
+    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
+    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
+  }
+}
+
+rownames(num_high_viruses) <- combos_list$toolcombo
+colnames(num_high_viruses) <- combos_list$toolcombo
+
+rownames(prop_high_viruses) <- combos_list$toolcombo
+colnames(prop_high_viruses) <- combos_list$toolcombo
 ```
@@
 pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+pal2 <- ggthemes::tableau_color_pal(palette="Tableau 20", type="regular")
 
@@
   vs2 = c("black", pal(6)[6]),
-  score = c("archaea"=pal(6)[1], "bacteria"=pal(6)[2], "fungi"=pal(6)[3],
-              "plasmid"=pal(6)[4], "protist"=pal(6)[5], "virus"=pal(6)[6])
+  score = c("archaea"=pal2(20)[13], "bacteria"=pal2(20)[9], "fungi"=pal2(20)[4],
+              "plasmid"=pal2(20)[19], "protist"=pal2(20)[5], "virus"=pal2(20)[17])
 )
@@
                   cellheight=3,
-                  filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
+                  filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores_balanced.png"
                   )

### 2023-05-26 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
Comparing 50:50 vs 90:10 not viral:viral splits.

@@
                                 include_tuning_not_viral=FALSE,
-                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
                                 include_virsorter=FALSE) {

### 2023-05-26 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated overarching heatmap.

@@
 ```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
 MCC_sd_quant <- cbind(accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1], as.data.frame(matrix(data=0, nrow=63, ncol=7)))
@@
 accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_MCC$toolcombo)
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
+                                  color=numrules, fill=numrules)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=20),
+  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("MCC") +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 1))
 
@@
   vs2 = c("black", pal(6)[6]),
-  score = c("archaea"=pal(6)[1], "bacteria"=pal(6)[2], "fungi"=pal(6)[3],
-              "plasmid"=pal(6)[4], "protist"=pal(6)[5], "virus"=pal(6)[6])
+  score = c("archaea"=pal2(20)[13], "bacteria"=pal2(20)[9], "fungi"=pal2(20)[20],
+              "plasmid"=pal2(20)[19], "protist"=pal2(20)[5], "virus"=pal2(20)[17])
 )
@@
 rownames(viral_scores_1) <- viruses_1$contig
+
+viral_scores_1 <- floor(viral_scores_1)
+
+seq_scores <- rowSums(viral_scores_1)
+
+seq_type <- table(viruses_1$seqtype)[c(6,2,1,4,5,3)]
+seq_type <- cumsum(seq_type)
+
+sub_seq_scores <- seq_scores[1:seq_type[1]]
+order_seq_scores_virus <- names(sub_seq_scores)[order(sub_seq_scores)]
+
+sub_seq_scores <- seq_scores[seq_type[1]:seq_type[2]]
+order_seq_scores_bacteria <- names(sub_seq_scores)[order(sub_seq_scores)]
+
+sub_seq_scores <- seq_scores[seq_type[2]:seq_type[3]]
+order_seq_scores_archaea <- names(sub_seq_scores)[order(sub_seq_scores)]
+
+sub_seq_scores <- seq_scores[seq_type[3]:seq_type[4]]
+order_seq_scores_plasmid <- names(sub_seq_scores)[order(sub_seq_scores)]
+
+sub_seq_scores <- seq_scores[seq_type[4]:seq_type[5]]
+order_seq_scores_protist <- names(sub_seq_scores)[order(sub_seq_scores)]
+
+sub_seq_scores <- seq_scores[seq_type[5]:seq_type[6]]
+order_seq_scores_fungi <- names(sub_seq_scores)[order(sub_seq_scores)]
+
+viral_scores_1 <- viral_scores_1[c(order_seq_scores_virus, 
+                                   order_seq_scores_bacteria,
+                                   order_seq_scores_archaea,
+                                   order_seq_scores_plasmid,
+                                   order_seq_scores_protist,
+                                   order_seq_scores_fungi),]
 
@@
                   #border_color = "black",
-                  color = c(rev(RColorBrewer::brewer.pal(3,"Oranges")),
-                            RColorBrewer::brewer.pal(5,"Purples") 
+                  color = c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
                             ),
@@
                   cellheight=3,
+                  gaps_col = c(seq_type[1], seq_type[2], seq_type[3], seq_type[4], seq_type[5]),
+                  legend_breaks=c(-3:5),
+                  legend_labels=c(-3:5),
                   filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"

### 2023-05-26 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Updated single rule rulesets.

@@
 
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
@@
   ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
   xlab("Ruleset") +
-  ylab("Number of Sequences") + 
+  ylab("Proportion") + 
   facet_grid(rule_type~seqtype, scales = "free")
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/individual_tool_rulesets.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
 ```

### 2023-05-26 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated MCC, precision, recall boxplot.

@@
 ```{r}
+accuracy_scores$ruletype <- "other"
+accuracy_scores$ruletype[accuracy_scores$toolcombo %in% c("0 0 1 0 0 1",
+                                                          "0 0 1 1 0 0",
+                                                          "0 0 0 1 0 0")] <- "high precision"
+accuracy_scores$ruletype[accuracy_scores$toolcombo %in% c("1 1 0 0 1 1",
+                                                          "1 1 0 1 1 1",
+                                                          "1 1 0 0 0 1",
+                                                          "1 1 0 1 0 1")] <- "high recall"
+accuracy_scores$ruletype[accuracy_scores$toolcombo %in% c("0 0 1 0 0 1",
+                                                          "0 1 1 0 0 1",
+                                                          "0 0 0 0 0 1",
+                                                          "1 0 1 1 0 1",
+                                                          "1 0 1 0 1 1",
+                                                          "1 0 1 0 0 1")] <- "high MCC"
+#accuracy_scores$ruletype[accuracy_scores$toolcombo %in% c("0 0 1 0 0 1")] <- "high MCC and high precision"
+accuracy_scores$ruletype[accuracy_scores$toolcombo %in% c("1 1 1 1 1 1")] <- "all"
+
+```
+
+```{r}
 accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo) %>%
+  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo, ruletype) %>%
   pivot_longer(cols=c(precision, recall, MCC), 
@@
                      values = alpha(rev(viridis(6)), 1))
-```
-
-```{r}
-mean_precision <- accuracy_scores %>% 
-  select(precision, toolcombo) %>%
-  group_by(toolcombo) %>%
-  summarise(mean = mean(precision))
-```
-
-getting set of tools with roughly equivalent precision
-```{r}
-t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 0"])
-
-t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 1 0 0"])
-
-t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 0 0 1"])
-
-t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 1"])
-```
-accounting for multiple t-tests with Bonferroni, then the following have the same precision:
-- vs2+tnv, vb+tnv, vb, and vs2
-
-recall
-```{r}
-accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
-accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
-
-ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_boxplot(alpha=0.5) +
   theme_light() +
@@
   xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Precision") +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
+```
+
+```{r}
+mean_precision <- accuracy_scores %>% 
+  select(precision, toolcombo) %>%
+  group_by(toolcombo) %>%
+  summarise(mean = mean(precision))
+```
+
+getting set of tools with roughly equivalent precision
+```{r}
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 0"])
+
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 1 0 0"])
+
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 0 0 0 1"])
+
+t.test(accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$precision[accuracy_scores$toolcombo=="0 0 1 1 0 1"])
+```
+accounting for multiple t-tests with Bonferroni, then the following have the same precision:
+- vs2+tnv, vb+tnv, vb, and vs2
+
+recall
+```{r}
+accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
+                                  color=numrules, fill=numrules)) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
   ylab("Recall") +
@@
                      values = alpha(rev(viridis(6)), 1))
+
+ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_boxplot(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Recall") +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
 ```
@@
 
-ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
+fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_boxplot(alpha=0.5) +
   theme_light() +
@@
     legend.text=element_text(size=12),
-    axis.title=element_text(size=20),
+    axis.title=element_text(size=16),
   ) +
@@
   ylab("MCC") +
-  scale_fill_manual(name="Number of Rule Sets",
-                     values = alpha(rev(viridis(6)), 0.5)) +
-  scale_color_manual(name="Number of Rule Sets",
-                     values = alpha(rev(viridis(6)), 1))
-
-fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
-                                  color=numrules, fill=numrules)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=20),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("MCC") +
-  scale_fill_manual(name="Number of Rule Sets",
-                     values = alpha(rev(viridis(6)), 0.5)) +
-  scale_color_manual(name="Number of Rule Sets",
-                     values = alpha(rev(viridis(6)), 1))
+  scale_fill_manual(name="Number of Rules",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="Number of Rules",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
+
+fig
 
@@
        accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 1 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"])
 
@@
                   filename = "../IntermediaryFiles/heatmap_MCC_labels.png"
+                  )
+```
+
+```{r}
+fig <- ggplot(accuracy_scores_melt, aes(x=numrules, y=performance_metric_score, 
+                                  color=numrules, fill=numrules)) +
+  geom_boxplot() +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Score") +
+  xlab("Number of Rule Sets") +
+  scale_fill_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 0.5)) +
+  scale_color_manual(name="Number of Rule Sets",
+                     values = alpha(rev(viridis(6)), 1)) +
+  facet_wrap(~performance_metric)
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/MCC_precision_recall_boxplots.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in", "cm", "mm", "px"),
+  dpi = 300,
+  limitsize = TRUE
 )

### 2023-05-26 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated tuning rules figures.

@@
   summarise(n = n())
-
-pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
-pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")
-
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+```
+
+```{r}
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
@@
     legend.position = "bottom",
-    axis.text.x=element_text(size=8, angle = 45),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  facet_grid(~seqtype, scales = "free") +
-  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
-```
-
-```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
-                   fill=viral_score)) +
-  geom_bar(stat="identity", color="black") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.5)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
- # scale_x_discrete(labels=c("DVF", "DVF+VS", "DVF+VS+VB", "DVF+VS+VB+VS2",
- #                          "DVF+VS+VB+VS2+addition", "DVF+VS+VB+VS2+addition-removal")) + 
-  facet_grid(size_class~seqtype, scales = "free")
-```
-
-```{r}
-viruses$keep_score_all_tv_only_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = T,
-                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_all_tv_only_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = T,
-                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_all_tv_only_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = T,
-                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_all_tv_only_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = F,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = T,
-                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$true_virus <- "not"
-viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
-
-viruses_long_scores <- viruses %>% 
-  select(contains("keep_score_all_tv_only"))
-
-viruses_long_scores <- apply(viruses_long_scores, c(1,2), function(x) {if (x>0){x<-1} else{x<-0}})
-
-viruses_long_scores <- as.data.frame(cbind(viruses$true_virus, rowSums(viruses_long_scores)))
-
-table(viruses_long_scores$V1, viruses_long_scores$V2>1)
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion") + 
+  facet_grid(~seqtype, scales = "free")
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/tuning_addition_rulesets.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
 ```
@@
 ```{r}
-pal_purple <- RColorBrewer::brewer.pal(4, "Purples")
-pal_orange <- RColorBrewer::brewer.pal(1, "Oranges")
-
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
@@
     panel.border = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange[1]), pal_purple), 0.8)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion") + 
   facet_grid(~seqtype, scales = "free")
+
+fig
 ```
@@
 ```{r}
-ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                    fill=viral_score)) +
@@
     legend.position = "bottom",
-    axis.text.x=element_text(size=8, angle = 45),
-  ) +
-  scale_fill_manual(name="",
-                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
-  xlab("") +
-  ylab("Number of Sequences") + 
-  facet_grid(~seqtype, scales = "free") +
-  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion") + 
+  facet_grid(~seqtype, scales = "free")
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/tuning_removal_rulesets.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
 ```
@@
                   filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common.png",
-                  annotation_legend = F)
+                  annotation_legend = T)
 ```

### 2023-05-26 EnvironmentalData/DrinkingWaterMetagenomes/identifying_viral_contigs_environmental_dw.Rmd
Effect of tuning rules on environmental data.

@@
 }
+```
+
+looking at tuning rules
+### Effect of tuning addition rules
+```{r}
+viruses$keep_score_all_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_tv")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion")
+
+fig
+```
+
+same idea but VS2 base
+```{r}
+viruses$keep_score_vs2_tv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_4_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_4_tv_2_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=T, tv_4=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+ viruses$keep_score_vs2_tv_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vs2_tv_")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion")
+
+fig
+```
+
+### Effect of tuning removal rules
+```{r}
+viruses$keep_score_all_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_ntv")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion")
+
+fig
+```
+
+### Effect of tuning removal rules -VS2
+```{r}
+viruses$keep_score_vs2_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_5 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+viruses$keep_score_vs2_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_vs2_ntv")) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score) %>%
+  summarise(n = n())
+```
+
+```{r}
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion")
+
+fig
 ```

### 2023-05-26 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated tuning figure.

@@
 
+### Effect of tuning rules - subset
+```{r}
+viruses$keep_score_all_ntv_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses$keep_score_all_ntv_neither <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              ntv_1=F, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_viral = T,
+                                              ntv_1=T, ntv_2=F, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=T, ntv_3=F, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_ntv_3 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              tv_1=T, tv_2=T, tv_3=F, tv_4=T,
+                                              include_tuning_viral = T,
+                                              ntv_1=F, ntv_2=F, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_0 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_1 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=T, tv_2=F, tv_3=F, tv_4=F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=T, tv_3=F, tv_4=F,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$keep_score_all_tv_4 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              tv_1=F, tv_2=F, tv_3=F, tv_4=T,
+                                              ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=F, ntv_5=F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+
+viruses$true_virus <- "not"
+viruses$true_virus[viruses$seqtype=="virus"] <- "virus"
+
+viruses_long_scores <- viruses %>% 
+  select(contains("keep_score_all_"), seqtype) %>%
+  pivot_longer(cols=contains("keep_score_"), 
+               names_to="rule_combination",
+               values_to="viral_score") %>% 
+  mutate(viral_score=as.factor(round(viral_score))) %>%
+  group_by(rule_combination, viral_score, seqtype) %>%
+  summarise(n = n())
+
+viruses_long_scores$tuning_type <- "tuning addition"
+viruses_long_scores$tuning_type[grepl("ntv", viruses_long_scores$rule_combination)] <- "tuning removal"
+viruses_long_scores$tuning_type[viruses_long_scores$rule_combination=="keep_score_all_ntv_all"] <- "all"
+viruses_long_scores$tuning_type[viruses_long_scores$rule_combination=="keep_score_all_ntv_neither"] <- "neither"
+
+fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
+                   fill=viral_score)) +
+  geom_bar(stat="identity", color="black") +
+  theme_light() +
+  coord_flip() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    legend.position = "bottom",
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  scale_fill_manual(name="Viral Score",
+                     values = alpha(c(rev(RColorBrewer::brewer.pal(9,"Oranges")[4:7]),
+                            RColorBrewer::brewer.pal(9,"Purples")[4:9] 
+                            ), 0.8)) +
+  xlab("Ruleset") +
+  ylab("Proportion") + 
+  facet_grid(tuning_type~seqtype, scales = "free")
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/tuning_rulesets.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 6,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
+
+```
+
 ### Effect of tuning addition rules
@@
   summarise(n = n())
-```
-
-```{r}
+
 fig <- ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
@@
 ggsave(
-  "../IntermediaryFiles/tuning_addition_rulesets.png",
+  "../IntermediaryFiles/tuning_addition_rulesets_all.png",
   plot = fig,
@@
 ggsave(
-  "../IntermediaryFiles/tuning_removal_rulesets.png",
+  "../IntermediaryFiles/tuning_removal_rulesets-all.png",
   plot = fig,

### 2023-05-27 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
updated x axis to match MCC graph
Also added a csv that has num_viruses without drinking water (being loaded from drinking water folder)

@@
 ```{r}
-viruses_dw <- read_csv("../DrinkingWaterMetagenomes/drinking_water_num_viruses.csv")
-```
-```{r}
-viruses_rest <- read_csv("/Users/hegartyb/Downloads/num_viruses_env.csv")
+env_path <- '/Users/riddellj/Documents/Research/VSTE/VSTE_Old_Pull/EnvironmentalData/'
+viruses_tara <- read_csv(paste(env_path, "tara_num_viruses.csv", sep=""))
+viruses_tara$environment <- "global_oceans"
+viruses_le <- read_csv(paste(env_path, "le_num_viruses.csv", sep=""))
+viruses_le$environment <- "Lake_Erie"
+viruses_lm <- read_csv(paste(env_path, "lake_michigan_num_viruses.csv", sep=""))
+viruses_lm$environment <- "Lake_Michigan"
+viruses_ww <- read_csv(paste(env_path, "wastewater_num_viruses.csv", sep=""))
+viruses_ww$environment <- "wastewater"
+viruses_dw <- read_csv(paste(env_path, "drinking_water_uknl_num_viruses.csv", sep=""))
+
+viruses_rest <- rbind(viruses_tara, viruses_le, viruses_lm, viruses_ww)
+write.csv(viruses_rest, 'env_rest_num_viruses.csv')
 ```
@@
 
-viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
-viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake_Erie"
-viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake_Michigan"
-viruses_rest$environment[viruses_rest$environment=="tara"] <- "global_oceans"
+# viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
+# viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake_Erie"
+# viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake_Michigan"
+# viruses_rest$environment[viruses_rest$environment=="tara"] <- "global_oceans"
 
@@
 viruses_all$best_coloring <- "others"
+# 
 viruses_all$best_coloring[viruses_all$toolcombo=="1 1 1 1 1 1"] <- "all"
-viruses_all$best_coloring[viruses_all$toolcombo=="0 1 1 1 1 1"] <- "high recall"
-viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 1 1 1"] <- "high MCC"
+# high recall
+viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 1 1"] <- "high recall"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 0 1 1"] <- "high recall"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 0 0 1"] <- "high recall"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 0 0 0 1 1"] <- "high recall"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 0 1"] <- "high recall"
+# VS2 Combo
 viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 0 1"] <- "VS2 combo"
-viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 0 0"] <- "high precision"
+# High Precision
+viruses_all$best_coloring[viruses_all$toolcombo=="0 0 0 0 0 1"] <- "high precision"
+viruses_all$best_coloring[viruses_all$toolcombo=="0 0 0 1 0 0"] <- "high precision"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 0 0 1 0 0"] <- "high precision"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 0 0 0 0 1"] <- "high precision"
+# High MCC
+viruses_all$best_coloring[viruses_all$toolcombo=="0 0 1 0 0 1"] <- "high MCC"
+viruses_all$best_coloring[viruses_all$toolcombo=="0 1 1 0 0 1"] <- "high MCC"
+viruses_all$best_coloring[viruses_all$toolcombo=="0 0 0 0 0 1"] <- "high MCC"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 1 0 1"] <- "high MCC"
+viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 1 1"] <- "high MCC"
 ```
 
+run commented-out lines after paper_figure_generation.Rmd if order changes
+
 ```{r}
+# dput output from paper_figure_generation of toolcombo_vector created with the following commands
+# toolcombo_MCC <- aggregate(accuracy_scores$MCC, list(accuracy_scores$toolcombo), mean)
+# toolcombo_MCC <- toolcombo_MCC[order(toolcombo_MCC$x),]
+# toolcombo_vector <- as.vector(toolcombo_MCC$Group.1)
+
+# hard-coded
+toolcombo_vector <- c("0 0 1 0 0 0", "0 1 0 0 0 0", "0 1 1 0 0 0", "0 1 0 0 1 0", 
+"0 0 0 0 1 0", "0 0 1 0 1 0", "0 1 1 0 1 0", "0 1 0 1 1 0", "0 1 0 1 0 0", 
+"0 0 0 1 1 0", "0 0 1 1 1 0", "0 1 1 1 1 0", "1 1 0 0 0 0", "0 1 1 1 0 0", 
+"0 0 0 1 0 0", "0 0 1 1 0 0", "1 1 0 1 1 0", "1 1 0 1 0 0", "1 1 0 0 1 0", 
+"1 1 0 1 1 1", "0 1 0 1 1 1", "1 1 0 0 1 1", "0 1 0 0 1 1", "1 1 0 1 0 1", 
+"0 1 0 1 0 1", "1 1 0 0 0 1", "0 0 0 1 1 1", "1 1 1 0 0 0", "1 0 0 1 1 0", 
+"1 0 0 1 1 1", "1 0 0 0 0 0", "0 0 0 0 1 1", "1 0 1 0 0 0", "1 0 0 0 1 0", 
+"1 1 1 1 0 0", "1 0 0 0 1 1", "0 1 1 1 1 1", "1 1 1 1 1 0", "0 0 1 1 1 1", 
+"0 1 0 0 0 1", "1 0 0 1 0 0", "0 0 0 1 0 1", "1 1 1 0 1 0", "1 0 1 1 1 0", 
+"0 1 1 0 1 1", "1 0 1 1 0 0", "0 0 1 0 1 1", "0 1 1 1 0 1", "1 1 1 1 1 1", 
+"1 0 0 1 0 1", "0 0 1 1 0 1", "1 0 1 0 1 0", "1 1 1 0 1 1", "1 0 0 0 0 1", 
+"1 1 1 1 0 1", "1 0 1 1 1 1", "1 1 1 0 0 1", "1 0 1 0 1 1", "1 0 1 1 0 1", 
+"0 0 0 0 0 1", "1 0 1 0 0 1", "0 1 1 0 0 1", "0 0 1 0 0 1")
+```
+
+```{r fig.align="center", fig.width = 5, fig.height = 2}
+viruses_all$toolcombo <- factor(viruses_all$toolcombo, levels=toolcombo_vector)
+
+tiff(filename='env_comparisons.tiff', units="in", width=15, height=6, res=300, compression = 'lzw')
 ggplot(viruses_all, aes(x=toolcombo, y=prop_vir, 
@@
     axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
+    axis.text.x=element_text(size=5, angle = 90),
     legend.text=element_text(size=12),
@@
   ) +
-  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Proportion of Contigs >3kb Viral") +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Prop of Contigs >3kb Viral") +
   scale_fill_manual(name="",
@@
   facet_wrap(~environment2)
+
+dev.off()
+```
+
+```{r}
+p.list = lapply(sort(unique(viruses_all$environment2)), function(i) {
+ggplot(viruses_all[viruses_all$environment2==i,], aes(x=toolcombo, y=prop_vir, 
+                                  color=best_coloring, fill=best_coloring)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=5, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Prop of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.3)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(viridis(6)), 0.7)) +
+  facet_wrap(~environment2)
+})
+p.list
 ```
@@
 ```{r}
-imp_toolcombs <- viruses_all %>% group_by(best_coloring, environment) %>%
+imp_toolcombs <- viruses_all %>% group_by(best_coloring, environment2) %>%
   summarize(median=median(prop_vir),
@@
   xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Proportion of Contigs >3kb Viral") +
+  ylab("Prop of Contigs >3kb Viral") +
   scale_fill_manual(name="",
@@
                      values = alpha(rev(viridis(6)), 0.7)) +
-  facet_wrap(~environment)
+  facet_wrap(~environment2)
 ```
@@
 ```
+
+# check how much tuning removal decreases viral ID by on average
+```{r}
+viruses_all <- viruses_all %>% 
+  mutate(
+    has_tnv = grepl('tnv', toolcombo2, fixed=TRUE),
+    has_tv = grepl('tv', toolcombo2, fixed=TRUE),
+  )
+has_both <- rep(0, nrow(viruses_all))
+has_both[viruses_all$has_tnv==TRUE] <- has_both[viruses_all$has_tnv==TRUE] + 0.5
+has_both[viruses_all$has_tv==TRUE] <- has_both[viruses_all$has_tv==TRUE] + 0.5
+viruses_all$has_both <- has_both
+
+metagenomes <- viruses_all %>% subset(type == 'metagenome')
+viromes <- viruses_all %>% subset(type == 'virome')
+
+# how much does the tuning removal decrease the proportion identified as viral?
+metaG_avg_decrease <- aggregate(metagenomes$prop_vir, list(metagenomes$has_tnv), mean)
+virome_avg_decrease <- aggregate(viromes$prop_vir, list(viromes$has_tnv), mean)
+metaG_avg_increase <- aggregate(metagenomes$prop_vir, list(metagenomes$has_tv), mean)
+virome_avg_increase <- aggregate(viromes$prop_vir, list(viromes$has_tv), mean)
+
+metaG_both <- aggregate(metagenomes$prop_vir, list(metagenomes$has_both), mean)
+virome_both <- aggregate(viromes$prop_vir, list(viromes$has_both), mean)
+```
+
+# get total number of viruses and assemblies
+```{r}
+viruses_vs2_combo <- viruses_all[viruses_all$toolcombo == "1 0 1 0 0 1",]
+viruses_vs2_combo$total_viruses <- viruses_vs2_combo$num_viruses / viruses_vs2_combo$prop_vir
+```

### 2023-05-27 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
Updated Figure 3A.

--- 2022-12-03 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
@@
 ```{r}
-viruses_dw$environment <- "drinking_water"
+viruses_dw$environment <- "drinking water"
 
 viruses_rest$environment[viruses_rest$environment=="waste_water"] <- "wastewater"
-viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake_Erie"
-viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake_Michigan"
-viruses_rest$environment[viruses_rest$environment=="tara"] <- "global_oceans"
+viruses_rest$environment[viruses_rest$environment=="lake_eerie"] <- "Lake Erie"
+viruses_rest$environment[viruses_rest$environment=="lake_michigan"] <- "Lake Michigan"
+viruses_rest$environment[viruses_rest$environment=="tara"] <- "global oceans"
 
@@
 viruses_all$environment2 <- viruses_all$environment
-viruses_all$environment2[viruses_all$assembly %in% c("49639", "49640", "49641", "49642")] <- "Lake_Erie_virome"
+viruses_all$environment2[viruses_all$assembly %in% c("49639", "49640", "49641", "49642")] <- "Lake Erie virome"
 viruses_all$environment2[viruses_all$assembly %in% c("49613", "49614", "49618", "49621",
-                                                     "49624", "49628", "49632", "49638")] <- "Lake_Erie_metagenome"
-viruses_all <- viruses_all[viruses_all$environment2!="Lake_Erie",]
+                                                     "49624", "49628", "49632", "49638")] <- "Lake Erie metagenome"
+viruses_all <- viruses_all[viruses_all$environment2!="Lake Erie",]
 ```
@@
 viruses_all$environment2 <- factor(viruses_all$environment2,
-                                    levels = c("drinking_water", "Lake_Erie_metagenome", 
-                                               "global_oceans",
-                                               "wastewater", "Lake_Erie_virome", 
-                                               "Lake_Michigan"))
+                                    levels = c("drinking water", "Lake Erie metagenome", 
+                                               "global oceans",
+                                               "wastewater", "Lake Erie virome", 
+                                               "Lake Michigan"))
 ```
@@
 ```{r}
-viruses_all$best_coloring <- "others"
-viruses_all$best_coloring[viruses_all$toolcombo=="1 1 1 1 1 1"] <- "all"
-viruses_all$best_coloring[viruses_all$toolcombo=="0 1 1 1 1 1"] <- "high recall"
-viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 1 1 1"] <- "high MCC"
-viruses_all$best_coloring[viruses_all$toolcombo=="1 0 1 0 0 1"] <- "VS2 combo"
-viruses_all$best_coloring[viruses_all$toolcombo=="1 1 0 1 0 0"] <- "high precision"
+viruses_all$ruletype <- "other"
+viruses_all$ruletype[viruses_all$toolcombo %in% c("0 0 1 0 0 1",
+                                                          "0 0 1 1 0 0",
+                                                          "0 0 0 1 0 0")] <- "high precision"
+viruses_all$ruletype[viruses_all$toolcombo %in% c("1 1 0 0 1 1",
+                                                          "1 1 0 1 1 1",
+                                                          "1 1 0 0 0 1",
+                                                          "1 1 0 1 0 1")] <- "high recall"
+viruses_all$ruletype[viruses_all$toolcombo %in% c("0 0 1 0 0 1",
+                                                          "0 1 1 0 0 1",
+                                                          "0 0 0 0 0 1",
+                                                          "1 0 1 1 0 1",
+                                                          "1 0 1 0 1 1",
+                                                          "1 0 1 0 0 1")] <- "high MCC"
+viruses_all$ruletype[viruses_all$toolcombo %in% c("1 1 1 1 1 1")] <- "all"
+```
+
+```{r}
+viruses_all$type <- "metagenome"
+viruses_all$type[viruses_all$environment2 %in% c("wastewater", "Lake Erie virome",
+                                                 "Lake Michigan")] <- "virome"
 ```
@@
 ggplot(viruses_all, aes(x=toolcombo, y=prop_vir, 
-                                  color=best_coloring, fill=best_coloring)) +
+                                  color=ruletype, fill=ruletype)) +
   geom_boxplot() +
   #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14, angle = 90),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black")
+  ) +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
+  facet_grid(~environment2, scales = "free")
+```
+
+```{r}
+viruses_all %>% 
+  select(prop_vir, toolcombo, environment2) %>%
+  filter(toolcombo=="1 0 0 1 0 0") %>%
+  filter(environment2=="wastewater") 
+  group_by(environment2) %>%
+  summarise(mean = mean(prop_vir))
+
+viruses_all %>% 
+  select(prop_vir, toolcombo, environment2) %>%
+  filter(toolcombo=="0 1 1 0 1 1") %>%
+  group_by(environment2) %>%
+  summarise(mean = mean(prop_vir))
+
+viruses_all %>% 
+  select(prop_vir, toolcombo, environment2) %>%
+  filter(toolcombo=="1 0 0 0 0 1") %>%
+  group_by(environment2) %>%
+  summarise(mean = mean(prop_vir))
+
+viruses_all %>% 
+  select(prop_vir, toolcombo, environment2) %>%
+  filter(toolcombo=="1 1 1 1 1 1") %>%
+  group_by(environment2) %>%
+  summarise(mean = mean(prop_vir))
+```
+
+```{r}
+fig <- ggplot(viruses_all, aes(x=ruletype, y=prop_vir, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=10),
+    axis.text.x=element_text(size=12, angle = 90, vjust = -0.1),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=12),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=12)
+  ) +
+  xlab("Environment") +
+  ylab("Proportion of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
+  facet_wrap(~environment2)
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/simplified_environmental.png",
+  plot = fig,
+  scale = 1,
+  width = 6.5,
+  height = 4,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
+```
+
+```{r}
+imp_toolcombs <- viruses_all %>% group_by(ruletype, environment2) %>%
+  summarize(median=median(prop_vir),
+            mean=mean(prop_vir),
+            sd=sd(prop_vir)
+            )
+```
+
+```{r}
+ggplot(imp_toolcombs, aes(x=ruletype, y=mean, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_point(size=3) +
   theme_light() +
@@
 ```
-
-```{r}
-viruses_all$type <- "metagenome"
-viruses_all$type[viruses_all$environment2 %in% c("wastewater", "Lake_Erie_virome",
-                                                 "Lake_Michigan")] <- "virome"
-```
-
-```{r}
-ggplot(viruses_all[viruses_all$best_coloring %in% c("high MCC", "high recall"),], aes(x=environment2, y=prop_vir, 
-                                  color=type, fill=type)) +
-  geom_boxplot() +
-  #geom_point() +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Environment") +
-  ylab("Proportion of Contigs >3kb Viral") +
-  scale_fill_manual(name="",
-                     values = alpha(magma(4)[2:3], 0.3)) +
-  scale_color_manual(name="",
-                     values = alpha(magma(4)[2:3], 0.7)) +
-  facet_wrap(~best_coloring)
-```
-
-```{r}
-imp_toolcombs <- viruses_all %>% group_by(best_coloring, environment) %>%
-  summarize(median=median(prop_vir),
-            mean=mean(prop_vir),
-            sd=sd(prop_vir)
-            )
-```
-
-```{r}
-ggplot(imp_toolcombs, aes(x=best_coloring, y=mean, 
-                                  color=best_coloring, fill=best_coloring)) +
-  geom_point(size=3) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Proportion of Contigs >3kb Viral") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(viridis(6)), 0.3)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(viridis(6)), 0.7)) +
-  facet_wrap(~environment)
-```
-
-```{r}
-imp_toolcombs_wide <- viruses_all %>% group_by(best_coloring, environment) %>%
-  summarize(median=median(prop_vir),
-            mean=mean(prop_vir),
-            sd=sd(prop_vir)
-            ) %>%
-  select(mean, environment, best_coloring) %>%
-  pivot_wider(names_from=environment, values_from=mean)
-```

### 2023-05-27 TestingPipeline/Scripts/individual_rule_set_visualizations.Rmd
Updated Figure 3A.

@@
 
-viruses_long_scores$rule_type <- "viral sorting tool"
+viruses_long_scores$rule_type <- "viral identification tool-based"
 viruses_long_scores$rule_type[viruses_long_scores$rule_combination=="keep_score_ind_tv"] <- "tuning"
@@
 ```{r}
+
+viruses_long_scores$seqtype <- factor(viruses_long_scores$seqtype, levels = c("virus", "bacteria", "archaea", "plasmid", "protist", "fungi"))
+
+#viruses_long_scores$rule_combination <- factor(viruses_long_scores$rule_combination, 
+#                                               levels = c(rev("keep_score_ind_vs2", "keep_score_ind_vs",
+#                                                              "keep_score_ind_vb", "keep_score_ind_dvf"),
+#                                                         rev("keep_score_ind_tv", "keep_score_ind_ntv")))
+
+viruses_long_scores$rule_type <- factor(viruses_long_scores$rule_type, levels = c("viral identification tool-based", "tuning"))
+
 pal_purple <- RColorBrewer::brewer.pal(4, "Purples")

### 2023-05-27 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated Figure 3A.

@@
 
-  auc <- round(auc(truepositive, keep_score),4)
-
-  performance <- c(precision, recall, F1, MCC, auc)
-  names(performance) <- c("precision", "recall", "F1", "MCC", "AUC")
+  prop_viral <- (TP+FP)/(TP+FP+TN+FN)
+
+  performance <- c(precision, recall, F1, MCC, prop_viral)
+  names(performance) <- c("precision", "recall", "F1", "MCC", "prop_viral")
 
@@
                        MCC=rep(0, nrow(combos)),
-                       AUC=rep(0, nrow(combos)))
+                       prop_viral=rep(0, nrow(combos)))
   for (i in 1:nrow(combos)) {
@@
                        MCC=rep(0, nrow(combos_list)*8), 
-                      AUC=rep(0, nrow(combos_list)*8))
+                      prop_viral=rep(0, nrow(combos_list)*8))
 
@@
 accuracy_scores_melt <- accuracy_scores %>% 
-  select(testing_set_index, precision, recall, MCC, numrules, numtools, toolcombo, ruletype) %>%
-  pivot_longer(cols=c(precision, recall, MCC), 
+  select(testing_set_index, precision, recall, MCC, prop_viral, numrules, numtools, toolcombo, ruletype) %>%
+  pivot_longer(cols=c(precision, recall, MCC, prop_viral), 
                names_to="performance_metric",
                values_to="performance_metric_score")
+```
+
+```{r}
+fig <- ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric=="prop_viral",], 
+              aes(x=ruletype, y=performance_metric_score, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=10),
+    axis.text.x=element_text(size=12, angle = 90, vjust = -0.1),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=12),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=12)
+  ) +
+  xlab("Environment") +
+  ylab("Proportion Called Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
+  geom_hline(yintercept = 0.1, linetype="dashed", color="dark grey")
+
+fig
+```
+
+```{r}
+ggsave(
+  "../IntermediaryFiles/simplified_performance_metricrule_types.png",
+  plot = fig,
+  scale = 1,
+  width = 6.5,
+  height = 4,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
 ```

### 2023-05-29 TestingPipeline/Scripts/paper_figure_generation.Rmd
Visualizing change by adding tnv.

@@
 )
+```
+
+visualizing change when add the tnv rule
+```{r}
+tnv_add_comp <- data.frame(test_set_index=rep(NA, ))
+```
+
+```{r}
+accuracy_scores_one_ts <- accuracy_scores[accuracy_scores$testing_set_index==1,]
+
+tnv <- rep(NA, 63)
+
+for (i in 1:nrow(accuracy_scores_one_ts)) {
+  tnv[i] <- as.numeric(substr(accuracy_scores_one_ts$toolcombo[i], 5, 5))
+}
+
+num_no_tnv <- 63-(sum(tnv))
+```
+
+```{r}
+no_tnv_prop <- data.frame(tool_combo=accuracy_scores_one_ts$toolcombo[grep("0", tnv)],
+                          prop_viral=rep(0, num_no_tnv),
+                          MCC=rep(0, num_no_tnv),
+                          precision=rep(0, num_no_tnv),
+                          recall=rep(0, num_no_tnv))
+
+tc_tnv <- "0"
+tc_no_tnv <- "0"
+prop_tnv <- 0
+j <- 1
+
+for (i in grep("0", tnv)) {
+  tc_no_tnv <- as.character(accuracy_scores_one_ts$toolcombo[i])
+  tc_tnv <- tc_no_tnv
+  substr(tc_tnv, 5, 5) <- "1"
+
+  prop_tnv <- accuracy_scores_one_ts$prop_viral[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$prop_viral[j] <-
+    abs(accuracy_scores_one_ts$prop_viral[i]-prop_tnv)/accuracy_scores_one_ts$prop_viral[i]*100
+
+  prop_tnv <- accuracy_scores_one_ts$MCC[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$MCC[j] <-
+    abs(accuracy_scores_one_ts$MCC[i]-prop_tnv)/accuracy_scores_one_ts$MCC[i]*100
+
+  prop_tnv <- accuracy_scores_one_ts$precision[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]*100
+  no_tnv_prop$precision[j] <-
+    abs(accuracy_scores_one_ts$precision[i]-prop_tnv)/accuracy_scores_one_ts$precision[i]
+
+  prop_tnv <- accuracy_scores_one_ts$recall[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$recall[j] <-
+    abs(accuracy_scores_one_ts$recall[i]-prop_tnv)/accuracy_scores_one_ts$recall[i]*100
+
+  j <- j+1
+}
+
+mean(no_tnv_prop$prop_viral)
+sd(no_tnv_prop$prop_viral)
+```
+
+```{r}
+no_tnv_prop_melt <- no_tnv_prop %>% 
+  select(precision, recall, MCC, prop_viral, tool_combo) %>%
+  pivot_longer(cols=c(precision, recall, MCC, prop_viral), 
+               names_to="performance_metric",
+               values_to="performance_metric_percent_diff")
+
+fig <- ggplot(no_tnv_prop_melt, 
+              aes(x=performance_metric, y=performance_metric_percent_diff)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=10),
+    axis.text.x=element_text(size=12),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=12),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=12)
+  ) +
+  xlab("Performance Metric") +
+  ylab("Percent Difference")
+
+fig
 ```

### 2023-05-29 TestingPipeline/Scripts/paper_figure_generation.Rmd
Visualizing change by adding tv.

--- 2023-05-27 TestingPipeline/Scripts/paper_figure_generation.Rmd
@@
 )
+```
+
+visualizing change when add the tnv rule
+```{r}
+accuracy_scores_one_ts <- accuracy_scores[accuracy_scores$testing_set_index==1,]
+
+tnv <- rep(NA, 63)
+
+for (i in 1:nrow(accuracy_scores_one_ts)) {
+  tnv[i] <- as.numeric(substr(accuracy_scores_one_ts$toolcombo[i], 5, 5))
+}
+
+num_no_tnv <- 63-(sum(tnv))
+```
+
+```{r}
+no_tnv_prop <- data.frame(tool_combo=accuracy_scores_one_ts$toolcombo[grep("0", tnv)],
+                          prop_viral=rep(0, num_no_tnv),
+                          MCC=rep(0, num_no_tnv),
+                          precision=rep(0, num_no_tnv),
+                          recall=rep(0, num_no_tnv))
+
+tc_tnv <- "0"
+tc_no_tnv <- "0"
+prop_tnv <- 0
+j <- 1
+
+for (i in grep("0", tnv)) {
+  tc_no_tnv <- as.character(accuracy_scores_one_ts$toolcombo[i])
+  tc_tnv <- tc_no_tnv
+  substr(tc_tnv, 5, 5) <- "1"
+
+  prop_tnv <- accuracy_scores_one_ts$prop_viral[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$prop_viral[j] <-
+    (prop_tnv-accuracy_scores_one_ts$prop_viral[i])/prop_tnv*100
+
+  prop_tnv <- accuracy_scores_one_ts$MCC[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$MCC[j] <-
+    (prop_tnv-accuracy_scores_one_ts$MCC[i])/prop_tnv*100
+
+  prop_tnv <- accuracy_scores_one_ts$precision[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$precision[j] <-
+    (prop_tnv-accuracy_scores_one_ts$precision[i])/prop_tnv*100
+
+  prop_tnv <- accuracy_scores_one_ts$recall[grep(tc_tnv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tnv_prop$recall[j] <-
+    (prop_tnv-accuracy_scores_one_ts$recall[i])/prop_tnv*100
+
+  j <- j+1
+}
+
+mean(no_tnv_prop$prop_viral)
+sd(no_tnv_prop$prop_viral)
+
+no_tnv_prop_melt <- no_tnv_prop %>% 
+  select(precision, recall, MCC, prop_viral, tool_combo) %>%
+  pivot_longer(cols=c(precision, recall, MCC, prop_viral), 
+               names_to="performance_metric",
+               values_to="performance_metric_percent_diff")
+
+```
+
+visualizing when add the tuning addition rule
+```{r}
+accuracy_scores_one_ts <- accuracy_scores[accuracy_scores$testing_set_index==1,]
+
+tv <- rep(NA, 63)
+
+for (i in 1:nrow(accuracy_scores_one_ts)) {
+  tv[i] <- as.numeric(substr(accuracy_scores_one_ts$toolcombo[i], 1, 1))
+}
+
+num_no_tv <- 63-(sum(tv))
+
+no_tv_prop <- data.frame(tool_combo=accuracy_scores_one_ts$toolcombo[grep("0", tv)],
+                          prop_viral=rep(0, num_no_tv),
+                          MCC=rep(0, num_no_tv),
+                          precision=rep(0, num_no_tv),
+                          recall=rep(0, num_no_tv))
+
+tc_tv <- "0"
+tc_no_tv <- "0"
+prop_tv <- 0
+j <- 1
+
+for (i in grep("0", tv)) {
+  tc_no_tv <- as.character(accuracy_scores_one_ts$toolcombo[i])
+  tc_tv <- tc_no_tv
+  substr(tc_tv, 1, 1) <- "1"
+
+  prop_tv <- accuracy_scores_one_ts$prop_viral[grep(tc_tv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tv_prop$prop_viral[j] <-
+    (prop_tv-accuracy_scores_one_ts$prop_viral[i])/prop_tv*100
+
+  prop_tv <- accuracy_scores_one_ts$MCC[grep(tc_tv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tv_prop$MCC[j] <-
+    (prop_tv-accuracy_scores_one_ts$MCC[i])/prop_tv*100
+
+  prop_tv <- accuracy_scores_one_ts$precision[grep(tc_tv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tv_prop$precision[j] <-
+    (prop_tv-accuracy_scores_one_ts$precision[i])/prop_tv*100
+
+  prop_tv <- accuracy_scores_one_ts$recall[grep(tc_tv,accuracy_scores_one_ts$toolcombo)][1]
+  no_tv_prop$recall[j] <-
+    abs(prop_tv-accuracy_scores_one_ts$recall[i])/prop_tv*100
+
+  j <- j+1
+}
+
+mean(no_tv_prop$prop_viral)
+sd(no_tv_prop$prop_viral)
+
+no_tv_prop_melt <- no_tv_prop %>% 
+  select(precision, recall, MCC, prop_viral, tool_combo) %>%
+  pivot_longer(cols=c(precision, recall, MCC, prop_viral), 
+               names_to="performance_metric",
+               values_to="performance_metric_percent_diff")
+```
+
+```{r}
+no_tv_and_ntv_prop_melt <- rbind(no_tv_prop_melt, no_tnv_prop_melt)
+
+no_tv_and_ntv_prop_melt$set <- c(rep("tuning addition", nrow(no_tv_prop_melt)),
+                                 rep("tuning removal", nrow(no_tnv_prop_melt)))
+
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+
+fig <- ggplot(no_tv_and_ntv_prop_melt, 
+              aes(x=performance_metric, y=performance_metric_percent_diff,
+                  color=set, fill=set)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=10),
+    axis.text.x=element_text(size=12),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=12),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=12)
+  ) +
+  xlab("Performance Metric") +
+  ylab("Percent Difference") +
+  geom_hline(yintercept = 1, color="grey") +
+  scale_fill_manual(name="",
+                     values = alpha(c(pal(6)[c(6,1)]), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(pal(6)[c(6,1)]), 0.6))
+
+fig
 ```

### 2023-05-29 EnvironmentalData/building_env_data_viral_features_set.Rmd
Updated environmental scripts

--- 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set.Rmd
@@
 
-calculate percent viral/host/unknown
+calculate percent viral/host/unknown/kaiju_match_ratio
 ```{r}
@@
 viruses$percent_unknown <- 100-(viruses$checkv_host_genes+viruses$checkv_viral_genes)/viruses$checkv_total_genes*100
+viruses$kaiju_match_ratio <- viruses$len/viruses$checkv_length
 ```

### 2023-05-29 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
Updated environmental scripts

@@
 ---
-title: "Building Environmental Viral Set and Visualization"
+title: "Building Metagenome Viral Set and Visualization Template"
 author: Bridget Hegarty, James Riddell
@@
 ---
-This Rmarkdown file takes the output from "building_env_data_viral_features_set.Rmd" and creates a set of viral contigs and visualizations.
+This Rmarkdown file takes outputs from viral identification tools and merges
+them into one file, "{dataset}_merged_viral_id_outputs.csv", which is the input
+for identifying_viral_metrics_environmental.Rmd.
 
@@
 
-TO DO:
-- add kaiju and DVF
-
 ```{r setup-library}
 library(ggplot2)
+library(stringr)
 library(plyr)
@@
 library(data.table)
+
 ```
@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "drinking_water" #replace with the dataset_name from "building_env_data_viral_features_set.Rmd"
+dataset_name <- "drinking_water_uknl"
 #name from building_env_data_viral_features_set.Rmd
-viruses_filename <- 'merged_viral_feature_outputs_nluk.csv'
-
-viruses <- read_tsv(viruses_filename)
+viruses_filename <- 'merged_viral_feature_outputs.tsv'
+# path of merged dataset
+# path = "/Users/riddellj/Documents/Research/VSTE/VSTE_Old_Pull/EnvironmentalData/DrinkingWaterMetagenomes/"
+path = "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/"
+setwd(path)
+viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+# write_csv(viruses[viruses$assembly %in% c("49639", "49640", "49641", "49642"),]paste(dataset_name, viruses_filename, 'virome', sep='_'))
 ```
@@
 
-TO: these rules are not written up correctly
-
 VIBRANT
+    Quality == "Complete Circular": +1
     Quality == "High Quality Draft": +1
@@
     Score >= 0.7: +0.5
+    Score >= 0.9: +0.5
 
-Kaiju:
-    Kaiju_viral = "cellular organisms": -1
+Tuning - Not Viral:
+    Kaiju_viral = "cellular organisms": -0.5
+    If host_genes >50 and NOT checkv_provirus: -1 
+    If viral_genes == 0 and host_genes >= 1: -1
+    If 3*viral_genes <= host_genes and NOT checkv_provirus: -1
+    If length > 50,000 and hallmark <=1: -1
+    If length < 5000 and checkv completeness <= 75: -0.5
+
+Tuning - Viral:
     Kaiju_viral = "Viruses": +1
-
-CheckV
-    viral_genes == 0 and host_genes >= 1: keep_score = 0
-    If 3*viral_genes <= host_genes: keep_score = 0
-    If length > 50,000 and hallmark == 0: keep_score = 0
-    If %unknown >= 75: +0.5
+    If %unknown >= 75 and length < 50000: +0.5
+    If %viral >= 50: +0.5
+    Hallmark > 2: +1
 
@@
 
+Function that allows for the comparing pieces of the pipeline
 ```{r getting_viral_set_1}
@@
   keep_score <- rep(0, nrow(input_seqs))
-
   if (include_vibrant) {
-    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
     keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
     keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
   }
@@
   if (include_virsorter2) {
-    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
-    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
+    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
   }
@@
   if (include_tuning_not_viral) {
-    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
-    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
-    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
-    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for RNA viruses
-    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
-    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 0.5 
-    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
+    # tuning removal 
+    keep_score <- tuning_not_viral(current_score=keep_score,
+                               ins=input_seqs,
+                         include_tv_1=ntv_1,
+                         include_tv_2=ntv_2,
+                         include_tv_3=ntv_3,
+                         include_tv_4=ntv_4,
+                         include_tv_5=ntv_5)
+
   }
@@
 }
+tuning_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T) {
+    # tuning addition
+  if (include_tv_1) {
+    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+  }
+  if (include_tv_2) {
+    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
+  }
+    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
+    #note: having kaiju removal rule didn't help
+  if (include_tv_3) {
+    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
+  }
+  if (include_tv_4) {
+    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
+  }
+  return(current_score)
+}
+tuning_not_viral <- function(current_score, ins, 
+                         include_tv_1=T,
+                         include_tv_2=T,
+                         include_tv_3=T,
+                         include_tv_4=T,
+                         include_tv_5=F) {
+    # tuning removal
+  if (include_tv_2) {
+    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
+  }
+  if (include_tv_3) {
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+  }
+
+  if (include_tv_5) {
+    # remove if DVF calls it not viral
+    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
+  }
+
+  if (include_tv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
+  }
+
+  if (include_tv_4) {
+    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- -3
+  }
+
+  return(current_score)
+}
 ```
+
+```{r}
+viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = T)
+viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+                                              include_vibrant = T,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = T)
+viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = T,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+```
+
+```{r}
+sum(viruses$keep_score_all>=1)
+sum(viruses$keep_score_high_recall>=1)
+sum(viruses$keep_score_high_precision>=1)
+sum(viruses$keep_score_high_MCC>=1)
+sum(viruses$keep_score_all>=0.5)
+sum(viruses$keep_score_high_recall>=0.5)
+sum(viruses$keep_score_high_precision>=0.5)
+sum(viruses$keep_score_high_MCC>=0.5)
+```
+
+```{r}
+write_csv(viruses[,c(1,43:45)], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
+```
+
+not having the rule, gives the same
 
@@
             combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
-            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
-                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
+            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
+                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
                                                if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
-            combos_list$tune_not_viral[p] <- i
+            combos_list$tune_not_viral[p] <- k
             combos_list$DVF[p] <- j
-            combos_list$tune_viral[p] <- k
+            combos_list$tune_viral[p] <- i
             combos_list$VIBRANT[p] <- l
@@
 }
-
-num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
+num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
 num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
@@
 ```{r}
+png(paste(dataset_name, 'virus_predicted.png', sep='_'))
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+# add this
 ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
@@
   ) +
-  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Num Viruses Predicted") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
+dev.off()
+```
 
-ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
-                                  color=numtools, fill=numtools)) +
+```{r}
+viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
+num_assemblies <- dim(table(viruses$assembly))
+assemblies <- names(table(viruses$assembly)) # add assemblies if not in viruses yet
+num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)*num_assemblies),
+                          toolcombo2=rep(0, nrow(combos_list)*num_assemblies),
+                          assembly=rep(0,nrow(combos_list)*num_assemblies),
+                          num_viruses=rep(0, nrow(combos_list)*num_assemblies),
+                          prop_vir=rep(0, nrow(combos_list)*num_assemblies)
+                          )
+n <- 1
+for (i in 1:nrow(combos_list)) {
+  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+                                            include_virsorter = combos_list$VS[i],
+                                            include_virsorter2 = combos_list$VS2[i],
+                                            include_tuning_viral = combos_list$tune_viral[i],
+                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
+                                            include_deepvirfinder = combos_list$DVF[i])
+  for (j in 1:num_assemblies) {
+    if (max(viral_scores[,i])<=0) {
+      num_viruses$num_viruses[n] <- 0
+    }
+    else {
+      num_viruses$num_viruses[n] <- sum(viral_scores[viruses$assembly==assemblies[j],i]>=1)
+      num_viruses$prop_vir[n] <- num_viruses$num_viruses[n]/table(viruses$assembly)[[j]]
+    }
+
+    num_viruses$toolcombo[n] <- combos_list$toolcombo[i]
+    num_viruses$toolcombo2[n] <- combos_list$toolcombo2[i]
+    num_viruses$assembly[n] <- assemblies[j]
+
+    n <- n+1
+  }
+}
+num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
+num_viruses <- num_viruses[order(num_viruses$prop_vir, decreasing=F),]
+num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
+num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
+num_viruses$numtools <- as.factor(num_viruses$numtools)
+```
+
+```{r}
+write_csv(num_viruses, paste(dataset_name, "num_viruses.csv", sep="_"))
+```
+
+```{r}
+png(paste(dataset_name, 'prop_viral_assembly.png', sep='_'))
+ggplot(num_viruses, aes(x=toolcombo, y=prop_vir, 
+                                  shape=numtools, color=assembly, fill=assembly)) +
   geom_point() +
@@
   ) +
-  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
-  ylab("Num Viruses Predicted")
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral")
+dev.off()
+```
+
+only plotting the average
+
+```{r}
+num_viruses$numtools <- as.numeric(num_viruses$numtools)
+median_prop_vir_by_tool <- num_viruses %>% 
+  select(toolcombo, prop_vir, numtools) %>%
+  group_by(toolcombo) %>%
+  summarize(median=median(prop_vir),
+         #sd=sd(prop_vir),
+         numtools=mean(numtools)) 
+median_prop_vir_by_tool$toolcombo <- factor(median_prop_vir_by_tool$toolcombo, 
+                                          levels = median_prop_vir_by_tool$toolcombo[order(median_prop_vir_by_tool$median, decreasing = F)])
+median_prop_vir_by_tool$numtools <- as.factor(median_prop_vir_by_tool$numtools)
 ```
@@
 ```{r}
-ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
-  geom_boxplot(aes(color=numtools)) +
-  geom_point(aes(color=numtools, fill=numtools)) +
+png(paste(dataset_name, 'prop_viral_average.png', sep='_'))
+ggplot(median_prop_vir_by_tool, aes(y=median, x=toolcombo,
+                                  color=numtools, fill=numtools, shape=numtools))  +
+  geom_hline(yintercept=0, color="grey") +
+  geom_point() +
+#  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
+#                 position=position_dodge(.9)) +
   theme_light() +
@@
   ) +
-  xlab("Number of Tools") +
-  ylab("Num Viruses Predicted")
-```
-
-```{r}
-viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
-viral_scores_nozeros <- viral_scores_nozeros + 1
-viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)
-
-colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
-```
-
-```{r}
-library(phyloseq)
-```
-
-```{r}
-tooldata <- num_viruses
-
-rownames(tooldata) <- tooldata$toolcombo2
-```
-
-```{r}
-physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
-                                     sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  geom_label(label=tooldata$toolcombo)
-
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", color="num_viruses") + 
-  geom_point(size = 3) +
-  theme_bw()
-```
-to do: try coloring above based on the F1 scores of the testing set on each combination
-
-```{r}
-bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
-clusters <- hclust(dist(bray_dist))
-plot(clusters)
-
-myclusters <- cutree(clusters, h=0.8)
-```
-
-```{r}
-names(myclusters[myclusters==1])
-names(myclusters[myclusters==2])
-names(myclusters[myclusters==3])
-names(myclusters[myclusters==4])
-names(myclusters[myclusters==5])
-
-myclusters_df <- tibble(combo=names(myclusters),
-                            cluster_index=myclusters)
-
-myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
-                                                            "Kaiju", "VIBRANT",
-                                                            "VirSorter", "VirSorter2"),
-                          sep=" ", remove = F)
-
-tool_count <- as.data.frame(rbind(table(myclusters_df$CheckV, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$DVF, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$Kaiju, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VIBRANT, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter, myclusters_df$cluster_index)[2,],
-                          table(myclusters_df$VirSorter2, myclusters_df$cluster_index)[2,])
-                    )
-
-tool_count$method <- c("CheckV", "DVF", "Kaiju", "VIBRANT", "VirSorter", "VirSorter2")
-
-tool_count <- melt(tool_count)
-
-colnames(tool_count) <- c("tool", "cluster_index", "tool_count")
-```
-
-```{r}
-pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
-
-ggplot(tool_count, aes(x=cluster_index, y=tool_count,
-                   fill=cluster_index,
-                   color=cluster_index)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral") +
   scale_fill_manual(name="",
@@
   scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1)) +
-  xlab("Cluster") +
-  ylab("Number of Times in Cluster") + 
-  facet_wrap(~tool, scales = "free")
+                     values = alpha(rev(pal(6)), 1))
+dev.off()
 ```
-
 ```{r}
-write_tsv(num_viruses, "dw_num_viruses_by_tool_combo.tsv")
-num_viruses <- read_tsv("dw_num_viruses_by_tool_combo.tsv")
-```
-
-```{r}
-accuracy_scores <- read_tsv("../../TestingPipeline/Scripts/20220927_accuracy_scores.tsv")
-```
-
-```{r}
-num_viruses_comp <- full_join(num_viruses, accuracy_scores[accuracy_scores$testing_set_index==1,], by=c("toolcombo", "numtools"))
-```
-
-```{r}
-num_viruses_comp$numtools <- as.factor(num_viruses_comp$numtools)
-```
-
-```{r}
-ggplot(num_viruses_comp, aes(x=num_viruses, y=F1, 
+png(paste(dataset_name, 'boxplot.png', sep='_'))
+ggplot(median_prop_vir_by_tool, aes(x=numtools, y=median, 
                                   color=numtools, fill=numtools)) +
+  geom_boxplot() +
   geom_point(alpha=0.5) +
@@
   ) +
-  ylab("F1") +
-  xlab("Num Viruses Predicted")
+  ylab("Proportion Viruses (%)") +
+  xlab("Number of Tools") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
+dev.off()
+```
 
-ggplot(num_viruses_comp, aes(x=num_viruses, y=precision, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
+by only a subset of the assemblies (UK/NL)
+
+```{r}
+num_viruses_uknl <- num_viruses
+num_viruses_uknl$residual <- "residual"
+num_viruses_uknl$residual[num_viruses_uknl$assembly=="Amsterdam_Leid" |
+                          num_viruses_uknl$assembly=="Amsterdam_Wee" |
+                          num_viruses_uknl$assembly=="Nieuwkoop_DHB" |
+                          num_viruses_uknl$assembly=="South_Holland_Dla" |
+                          num_viruses_uknl$assembly=="Spannenburg"
+                          ] <- "no residual"
+```
+
+```{r}
+num_viruses_uknl$numtools <- as.numeric(num_viruses_uknl$numtools)
+mean_prop_vir_by_tool <- num_viruses_uknl %>% 
+  select(toolcombo, prop_vir, numtools, residual) %>%
+  group_by(toolcombo, residual) %>%
+  summarize(prop_vir=mean(prop_vir),
+         #sd_per_vir=sd(prop_vir),
+         numtools=mean(numtools)) 
+mean_uknl <- mean_prop_vir_by_tool %>% 
+  group_by(toolcombo) %>%
+  summarize(mean=mean(prop_vir)) 
+mean_prop_vir_by_tool$toolcombo <- factor(mean_prop_vir_by_tool$toolcombo, 
+                                          levels = mean_uknl$toolcombo[order(mean_uknl$mean, decreasing = F)])
+mean_prop_vir_by_tool$numtools <- as.factor(mean_prop_vir_by_tool$numtools)
+```
+
+```{r}
+num_viruses_uknl$toolcombo <- factor(num_viruses_uknl$toolcombo, 
+                                          levels = levels(mean_prop_vir_by_tool$toolcombo))
+num_viruses_uknl$numtools <- as.factor(num_viruses_uknl$numtools)
+```
+
+```{r}
+png(paste(dataset_name, 'prop_viral_residual.png', sep='_'))
+ggplot(num_viruses_uknl) +
+  geom_hline(yintercept=0, color="grey") +
+  geom_point(aes(color=numtools, fill=numtools, y=prop_vir, x=toolcombo, shape=residual), alpha=0.5) +
+  geom_point(data=mean_prop_vir_by_tool, aes(x=toolcombo, 
+                                               y=prop_vir,
+                                             shape=residual),
+             color="black") +
   theme_light() +
@@
   ) +
-  ylab("Precision") +
-  xlab("Num Viruses Predicted")
-
-ggplot(num_viruses_comp, aes(x=num_viruses, y=recall, 
-                                  color=numtools, fill=numtools)) +
-  geom_point(alpha=0.5) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  ylab("Recall") +
-  xlab("Num Viruses Predicted")
+  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  ylab("Proportion of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(6)), 0.5)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(6)), 1))
+dev.off()
+```
+```{r}
+#t.test(num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="residual"],
+#       num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="no residual"]
+#       )
 ```
 
-# Visualizing how tools build on each other
+ordination: sample by tool combo
 
 ```{r}
-viruses$keep_score_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = F,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = F)
-
-viruses$keep_score_vb_dvf_vs2_vs <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = F,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = F,
-                                              include_virsorter = T)
-
-viruses$keep_score_vb_dvf_vs2_vs_tv_tnv <- getting_viral_set_1(viruses, include_deepvirfinder = T,
-                                              include_vibrant = T,
-                                              include_virsorter2 = T,
-                                              include_tuning_viral = T,
-                                              include_tuning_not_viral = T,
-                                              include_virsorter = T)
+library(phyloseq)
 ```
 
-# Considering how each method contributes to the final prediction
+make the sample by tool combo matrix
 ```{r}
-viruses_high <- viruses[viruses$keep_score_vb_dvf_vs2_vs_tv>=1,] 
-viruses_high_mod <- viruses_high %>% select(keep_score_vb,keep_score_vb_dvf, 
-                                            keep_score_vb_dvf_vs2, keep_score_vb_dvf_vs2_vs, 
-                                            keep_score_vb_dvf_vs2_vs_tv, keep_score_vb_dvf_vs2_vs_tv_tnv)
-#viruses_high_mod <- apply(viruses_high_mod, c(1,2), function(x) {if (x >= 1) {x <- 1} else {x <- 0}})
-viruses_high_mod <- as_tibble(viruses_high_mod)
-
+num_viruses_uknl_wide <- num_viruses_uknl %>% 
+  select(toolcombo, assembly, num_viruses) %>%
+  pivot_wider(names_from=assembly, values_from=num_viruses)
+num_viruses_uknl_wide <- as.data.frame(num_viruses_uknl_wide)
+rownames(num_viruses_uknl_wide) <- num_viruses_uknl_wide$toolcombo
+num_viruses_uknl_wide <- num_viruses_uknl_wide[,-1]
+num_viruses_uknl_wide <- num_viruses_uknl_wide[rowSums(num_viruses_uknl_wide)>0,]
 ```
@@
 ```{r}
-sm_m <- reshape2::melt(viruses_high_mod)
-colnames(sm_m) <- c("method", "viral_score")
+tooldata <- combos_list
+rownames(tooldata) <- tooldata$toolcombo
+tooldata <- tooldata[c(rownames(num_viruses_uknl_wide)),]
+tooldata$numtools <- str_count(tooldata$toolcombo, "1")
+tooldata$numtools <- as.factor(tooldata$numtools)
+tooldata$total_viruses <- rowSums(num_viruses_uknl_wide)
+```
+
+to reduce arch effect:
+```{r}
+num_viruses_uknl_wide <- num_viruses_uknl_wide/rowSums(num_viruses_uknl_wide)
 ```
@@
 ```{r}
-sm_m <- sm_m[sm_m$viral_score>0,]
-
-sm_m$score <- sm_m$viral_score
-
-sm_m$score[sm_m$viral_score==0.5] <- "0.5"
-sm_m$score[sm_m$viral_score>=1] <- "1"
-sm_m$score[sm_m$viral_score>=2] <- "2"
-sm_m$score[sm_m$viral_score>=3] <- "3"
-sm_m$score[sm_m$viral_score>=4] <- "4"
-sm_m$score[sm_m$viral_score>=5] <- "5"
-
-sm_m$score <- factor(sm_m$score, 
-                                       levels=c("0.5", "1", "2","3","4","5"))
+physeq_pooled <- phyloseq::phyloseq(phyloseq::otu_table(num_viruses_uknl_wide, taxa_are_rows = F),
+                                     phyloseq::sample_data(tooldata))
 ```
@@
 ```{r}
-ggplot(sm_m, aes(x=method, y=score,
-                   fill=score)) +
-  geom_bar(stat="identity") +
-  theme_light() +
-  coord_flip() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom"
-  ) +
-  scale_fill_manual(name = 'Number of Methods',
-                     values = alpha(c(viridis(6)), 1)) +
-  xlab("") +
-  ylab("Viral Score") +
-  coord_flip()
+ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
 ```
+
+```{r}
+png(paste(dataset_name, 'ordination.png', sep='_'))
+phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
+                          shape="numtools", 
+                          color="total_viruses") + 
+  geom_point(size = 3) +
+  theme_bw() +
+  scale_shape_manual(name="Num Tools",
+                     values = c(21,22,23,24,4,8)) +
+  scale_color_viridis_c()
+dev.off()
+```
+each point is tool combination
+
+Searching for novel viruses
+```{r}
+novel <- viruses[(viruses$keep_score_high_MCC >=1) & (viruses$percent_unknown > 75),]
+novel_kaiju_match_below_30 <- novel[novel$kaiju_match_ratio < 0.3,] %>% drop_na(kaiju_match_ratio)
+novel_kaiju_match_below_1 <- novel[novel$kaiju_match_ratio < 0.01,] %>% drop_na(kaiju_match_ratio)
+novel_by_type <- novel_kaiju_match_below_1 %>% group_by(max_score_group) %>% count()
+high_MCC_num_viruses <- sum(num_viruses[num_viruses$toolcombo=='0 0 1 0 0 1',]$num_viruses)
+```

### 2023-05-30 TestingPipeline/Scripts/paper_figure_generation_extended.Rmd
Difference of adding +0.5 rules.

--- 2023-05-27 TestingPipeline/Scripts/paper_figure_generation.Rmd
@@
 Function that allows for the comparing pieces of the pipeline
+
 ```{r getting_viral_set_1}
-getting_viral_set_1 <- function(input_seqs,
+getting_viral_set_2 <- function(input_seqs,
                                 include_vibrant=FALSE, 
@@
     keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
+    #keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
   }
@@
   if (include_virsorter2) {
-    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
-    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
+   # keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
+    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 1  
   }
@@
     keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
-    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
-    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
+    #keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
+    #keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
     keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
-    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
-    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
+    #keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
+    #keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
   }
@@
      # add if DVF calls viral
-    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
-    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
+    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 1
+   # keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
   }
@@
   for (i in 1:nrow(combos)) {
-    keep_score <- getting_viral_set_1(input_seqs, include_vibrant = combos$VIBRANT[i],
+    keep_score <- getting_viral_set_2(input_seqs, include_vibrant = combos$VIBRANT[i],
                                             include_virsorter = combos$VS[i],
@@
 ```{r}
-accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
-                      precision=rep(0, nrow(combos_list)*8),
-                       recall=rep(0, nrow(combos_list)*8),
-                       F1=rep(0, nrow(combos_list)*8),
-                       MCC=rep(0, nrow(combos_list)*8), 
-                      prop_viral=rep(0, nrow(combos_list)*8))
+accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*5),
+                      precision=rep(0, nrow(combos_list)*5),
+                       recall=rep(0, nrow(combos_list)*5),
+                       F1=rep(0, nrow(combos_list)*5),
+                       MCC=rep(0, nrow(combos_list)*5), 
+                      prop_viral=rep(0, nrow(combos_list)*5))
 
@@
 
-```{r}
-MCC_sd_quant <- cbind(accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1], as.data.frame(matrix(data=0, nrow=63, ncol=7)))
-
-for (i in (2:8)) {
-  temp <- accuracy_scores %>%
-    filter(testing_set_index<=i) %>%
-    select(MCC, toolcombo) %>%
-    group_by(toolcombo) %>%
-    summarise(sd = sd(MCC))
-
-  MCC_sd_quant[,i] <- temp$sd
-}
-
-colnames(MCC_sd_quant) <- c("testing_set_index", "1-2", "1-3", "1-4",
-                            "1-5", "1-6", "1-7", "1-8")
-
-MCC_sd_quant <- MCC_sd_quant %>% pivot_longer(cols=contains("-"))
-
-ggplot(MCC_sd_quant, aes(x=testing_set_index, y=value, 
-                                  color=name, fill=name)) +
-  geom_point(alpha=0.5)
-
-ggplot(MCC_sd_quant, aes(x=name, y=value)) +
-  geom_boxplot(alpha=0.5, color=pal(4)[1]) +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 0),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Testing Sets") +
-  ylab("Standard Deviation") 
-```
-
 ## Calculate the performance of each pipeline
-```{r}
-accuracy_scores <- accuracy_scores %>% filter(testing_set_index<=5)
-```
 
@@
 
-```{r}
-ggsave(
-  "../IntermediaryFiles/simplified_performance_metricrule_types.png",
-  plot = fig,
-  scale = 1,
-  width = 6.5,
-  height = 4,
-  units = c("in"),
-  dpi = 300,
-  limitsize = TRUE
-)
-```
-
 ### Visualize performance across all the rules
@@
 
-ggsave(
-  "../IntermediaryFiles/all_MCC_scatterplot.png",
-  plot = fig,
-  scale = 1,
-  width = 8,
-  height = 4,
-  units = c("in", "cm", "mm", "px"),
-  dpi = 300,
-  limitsize = TRUE
-)
-
 ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
@@
 ```
+
 ```{r}
@@
 TukeyHSD(one.way)
+```
+
+#comparing based on 0.5 rules or not
+
+```{r}
+#accuracy_scores_melt_no_0.5 <- accuracy_scores_melt
+```
+
+```{r}
+percent_diff <- data.frame(toolcombo=accuracy_scores_melt$toolcombo,
+                           percent_diff=(accuracy_scores_melt$performance_metric_score -
+                   accuracy_scores_melt_no_0.5$performance_metric_score)/
+  accuracy_scores_melt$performance_metric_score,
+                          performance_metric=accuracy_scores_melt$performance_metric)
+)
+
+fig <- ggplot(percent_diff, 
+              aes(x=performance_metric, y=percent_diff)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=10),
+    axis.text.x=element_text(size=12),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=12),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=12)
+  ) +
+  xlab("Performance Metric") +
+  ylab("Percent Difference") +
+  geom_hline(yintercept = 0, color="grey")
+
+fig
+```
+
+```{r}
+t.test(accuracy_scores_melt$performance_metric_score[accuracy_scores_melt$performance_metric=="MCC"],
+       accuracy_scores_melt_no_0.5$performance_metric_score[accuracy_scores_melt_no_0.5$performance_metric=="MCC"])
+```
+
+```{r}
+t.test(accuracy_scores_melt$performance_metric_score[accuracy_scores_melt$performance_metric=="precision"],
+       accuracy_scores_melt_no_0.5$performance_metric_score[accuracy_scores_melt_no_0.5$performance_metric=="precision"])
+```
+
+```{r}
+t.test(accuracy_scores_melt$performance_metric_score[accuracy_scores_melt$performance_metric=="recall"],
+       accuracy_scores_melt_no_0.5$performance_metric_score[accuracy_scores_melt_no_0.5$performance_metric=="recall"])
+```
+
+```{r}
+t.test(accuracy_scores_melt$performance_metric_score[accuracy_scores_melt$performance_metric=="prop_viral"],
+       accuracy_scores_melt_no_0.5$performance_metric_score[accuracy_scores_melt_no_0.5$performance_metric=="prop_viral"])
 ```

### 2023-05-31 TestingPipeline/Scripts/paper_figure_generation.Rmd
Some minor figure updates.

@@
 
-# Comparing all of the tool methods - Fig 2A
+# Comparing all of the tool methods - Fig 2A (also needed for 5)
 ```{r}
@@
                   filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common.png",
-                  annotation_legend = T)
+                  annotation_legend = F)
 ```

### 2023-05-31 TestingPipeline/Scripts/paper_figure_generation.Rmd
Changing location of row annotations in giant heatmap.

@@
     vs = combos_list$VS,
-    vs2 = combos_list$VS2,
-    num_vir = colSums(viral_scores>=1)+1
+    vs2 = combos_list$VS2
 )
+
+ann_row = data.frame(
+    num_vir = colSums(viral_scores>=1)+1)
 
@@
 rownames(ann_col) <- combos_list$toolcombo
+rownames(ann_row) <- combos_list$toolcombo
 
 pheatmap::pheatmap(prop_high_viruses,
-                  annotation_row = ann_col,
+                  annotation_row = ann_row,
+                  annotation_col=ann_col,
                   annotation_colors = ann_colors,

### 2023-05-31 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
Updating proportion of viral sequences Fig 8.

--- 2023-05-27 EnvironmentalData/CrossEnvironmentComparisons/comparing_rule_sets_across_environments.Rmd
@@
 viruses_dw <- read_csv(paste(env_path, "drinking_water_uknl_num_viruses.csv", sep=""))
+#viruses_dw <- read_csv("../DrinkingWaterMetagenomes/drinking_water_num_viruses.csv")
 
@@
 write.csv(viruses_rest, 'env_rest_num_viruses.csv')
+#viruses_rest <- read.csv('env_rest_num_viruses.csv')
+
 ```
@@
 viruses_all <- rbind(viruses_dw, viruses_rest)
+#viruses_all <- rbind(viruses_dw, viruses_rest[-1])
+
 ```
@@
 ```
+
+# simplified version comparing rulesets based on their accuracy type
+
+```{r}
+viruses_all$ruletype <- viruses_all$best_coloring
+viruses_all$ruletype[viruses_all$best_coloring=="VS2 combo"] <- "high MCC"
+
+fig <- ggplot(viruses_all, aes(x=ruletype, y=prop_vir, 
+                                  color=ruletype, fill=ruletype)) +
+  geom_boxplot() +
+  #geom_point() +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=10),
+    axis.text.x=element_text(size=12, angle = 90, vjust = -0.1),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=12),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=12)
+  ) +
+  xlab("Environment") +
+  ylab("Proportion of Contigs >3kb Viral") +
+  scale_fill_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.4)) +
+  scale_color_manual(name="",
+                     values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6)) +
+  facet_wrap(~environment2)
+fig
+
+ggsave(
+  "simplified_environmental.png",
+  plot = fig,
+  scale = 1,
+  width = 6.5,
+  height = 4,
+  units = c("in"),
+  dpi = 300,
+  limitsize = TRUE
+)
+```

### 2023-05-31 TestingPipeline/Scripts/paper_figure_generation.Rmd
Fixing prop_in_common figure.

@@
 colnames(prop_high_viruses) <- combos_list$toolcombo
+
+combos_list$ruletype <- "other"
+combos_list$ruletype[combos_list$toolcombo %in% c("0 0 1 0 0 1",
+                                          "0 0 1 1 0 0",
+                                          "0 0 0 1 0 0")] <- "high precision"
+combos_list$ruletype[combos_list$toolcombo %in% c("1 1 0 0 1 1",
+                                          "1 1 0 1 1 1",
+                                          "1 1 0 0 0 1",
+                                          "1 1 0 1 0 1")] <- "high recall"
+combos_list$ruletype[combos_list$toolcombo %in% c("0 0 1 0 0 1",
+                                          "0 1 1 0 0 1",
+                                          "0 0 0 0 0 1",
+                                          "1 0 1 1 0 1",
+                                          "1 0 1 0 1 1",
+                                          "1 0 1 0 0 1")] <- "high MCC"
+#combos_lists$ruletype[accuracy_scores$toolcombo %in% c("0 0 1 0 0 1")] <- "high MCC and high precision"
+combos_list$ruletype[combos_list$toolcombo %in% c("1 1 1 1 1 1")] <- "all"
+
+combos_list$ruletype <- as.factor(combos_list$ruletype)
 ```
@@
     vs = combos_list$VS,
-    vs2 = combos_list$VS2
+    vs2 = combos_list$VS2,
+    ruleset_type = combos_list$ruletype
 )
@@
   vs = c("black", pal(6)[5]),
-  vs2 = c("black", pal(6)[6])
+  vs2 = c("black", pal(6)[6]),
+  ruleset_type = c("all" = magma(7)[3],
+                   "high MCC" = magma(7)[4],
+                   "high precision" = magma(7)[5],
+                   "high recall" = magma(7)[6], 
+                   "other" = "grey")
 )

### 2023-06-16 TestingPipeline/Scripts/paper_figure_generation.Rmd
Minor updates to MCC figures for short fragments.

@@
 ```{r}
-fig <- ggplot(accuracy_scores_melt, aes(x=numrules, y=performance_metric_score, 
+fig <- ggplot(accuracy_scores_melt[accuracy_scores_melt$toolcombo!="0 0 1 0 0 0" & accuracy_scores_melt$performance_metric!="prop_viral",], aes(x=numrules, y=performance_metric_score, 
                                   color=numrules, fill=numrules)) +
@@
     panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
+    #axis.ticks.y = element_blank(),
     legend.position = "bottom",
@@
                      values = alpha(rev(viridis(6)), 1)) +
-  facet_wrap(~performance_metric)
+  facet_wrap(~performance_metric) +
+  geom_hline(yintercept = 0, color="grey")
 

### 2023-06-28 TestingPipeline/Scripts/paper_figure_generation.Rmd
Figure 5 with three breaks.

@@
                   annotation_legend = T)
+
+pheatmap::pheatmap(prop_high_viruses,
+                  annotation_row = ann_row,
+                  annotation_col=ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = "black",
+                  color = c(rep(viridis(3)[1], 5), rep(viridis(3)[2], 4), viridis(3)[3]),
+                  cutree_rows = 3,
+                  cutree_cols = 3,
+                  show_rownames = F,
+                  show_colnames = F,
+                  filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common_3_breaks.png",
+                  annotation_legend = T)
 ```

### 2023-06-28 TestingPipeline/Scripts/examining_false_positives.Rmd
New figure for proportion vs number of VOGs.

@@
 
-```{r}
+```{r getting_viral_set_1}
 getting_viral_set_1 <- function(input_seqs,
@@
 ```{r}
+ggplot(annotations_vb_fp_vog_count, 
+       aes(x=num_vog, y=prop_vog, 
+           color=seqtype, fill=seqtype)) +
+  geom_smooth(formula = y ~ log(x), se=F) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Proportion Genes VOGs") +
+  xlab("Number of VOGs") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(3)), 0.2)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(3)), 1)) +
+  facet_wrap(~seqtype, scales="free_x") +
+  ylim(c(0,1))
+
+ggplot(annotations_vb_fp_vog_count, 
+       aes(x=num_vog, y=prop_vog, 
+           color=seqtype, fill=seqtype)) +
+  geom_smooth(method="lm", se=F) +
+  geom_point(alpha=0.5) +
+  theme_light() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.border = element_blank(),
+    axis.ticks.y = element_blank(),
+    legend.position = "bottom",
+    axis.text.y=element_text(size=14),
+    axis.text.x=element_text(size=14),
+    legend.text=element_text(size=12),
+    axis.title=element_text(size=16),
+  ) +
+  ylab("Proportion Genes VOGs") +
+  xlab("Number of VOGs") +
+  scale_fill_manual(name="",
+                     values = alpha(rev(pal(3)), 0.2)) +
+  scale_color_manual(name="",
+                     values = alpha(rev(pal(3)), 1)) +
+  facet_wrap(~seqtype, scales="free_x") +
+  ylim(c(0,1))
+
 ggplot(annotations_vb_fp_vog_count, 
@@
   scale_fill_manual(name="",
-                     values = alpha(rev(pal(3)), 0.5)) +
+                     values = alpha(rev(pal(3)), 0.2)) +
   scale_color_manual(name="",
-                     values = alpha(rev(pal(3)), 1))
+                     values = alpha(rev(pal(3)), 1)) +
+  facet_wrap(~seqtype, scales="free_x") +
+  ylim(c(0,1))
+```
+
+```{r}
+
+cor.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="bacteria"], annotations_vb_fp_vog_count$num_vog[annotations_vb_fp_vog_count$seqtype=="bacteria"])
+
+cor.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="virus"], annotations_vb_fp_vog_count$num_vog[annotations_vb_fp_vog_count$seqtype=="virus"])
+
+cor.test(annotations_vb_fp_vog_count$prop_vog[annotations_vb_fp_vog_count$seqtype=="plasmid"], annotations_vb_fp_vog_count$num_vog[annotations_vb_fp_vog_count$seqtype=="plasmid"])
 ```

### 2023-06-29 TestingPipeline/Scripts/paper_figure_generation.Rmd
Some updates to flowchart.

@@
                   annotation_legend = T)
+
+pheatmap::pheatmap(prop_high_viruses,
+                  annotation_row = ann_row,
+                  annotation_col=ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = "black",
+                  color = c(viridis(4)[1], rep(viridis(4)[2], 4), rep(viridis(4)[3], 4), viridis(4)[4]),
+                  cutree_rows = 3,
+                  cutree_cols = 3,
+                  show_rownames = F,
+                  show_colnames = F,
+                  filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common_4_breaks.png",
+                  annotation_legend = T)
+
+pheatmap::pheatmap(prop_high_viruses,
+                  annotation_row = ann_row,
+                  annotation_col=ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = "black",
+                  color = viridis(5),
+                  cutree_rows = 3,
+                  cutree_cols = 3,
+                  show_rownames = F,
+                  show_colnames = F,
+                  filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common_5_breaks.png",
+                  annotation_legend = T)
 ```
@@
 ```
+
+```{r}
+viruses_from_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = T,
+                                              include_tuning_not_viral = F,
+                                              include_virsorter = F)
+
+table(viruses_from_tv, viruses$seqtype)
+```
+
+```{r}
+viruses_from_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+                                              include_vibrant = F,
+                                              include_virsorter2 = F,
+                                              include_tuning_viral = F,
+                                              include_tuning_not_viral = T,
+                                              include_virsorter = F)
+
+table(viruses_from_ntv, viruses$seqtype)
+```

### 2023-07-06 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd

--- 2023-04-15 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
@@
 ```{r}
+pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
+```
+
+```{r}
 accuracy_scores <- accuracy_scores[order(accuracy_scores$precision, decreasing=F),]
 accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+accuracy_scores$numrules <- factor(accuracy_scores$numrules, levels = unique(accuracy_scores$numrules))
 
@@
   ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
   ylab("Precision") +
   scale_fill_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
 ```
@@
   ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
   ylab("Recall") +
   scale_fill_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Rules",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
 ```
@@
   ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
+  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
   ylab("MCC") +
   scale_fill_manual(name="Number of Rule Sets",
-                     values = alpha(rev(pal(6)), 0.5)) +
+                     values = alpha(rev(viridis(6)), 0.5)) +
   scale_color_manual(name="Number of Rule Sets",
-                     values = alpha(rev(pal(6)), 1))
+                     values = alpha(rev(viridis(6)), 1))
 

### 2023-08-02 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updated figure S1.

@@
 
-colnames(MCC_sd_quant) <- c("testing_set_index", "1-2", "1-3", "1-4",
-                            "1-5", "1-6", "1-7", "1-8")
-
-MCC_sd_quant <- MCC_sd_quant %>% pivot_longer(cols=contains("-"))
-
-ggplot(MCC_sd_quant, aes(x=testing_set_index, y=value, 
-                                  color=name, fill=name)) +
-  geom_point(alpha=0.5)
+colnames(MCC_sd_quant) <- c("testing_set_index", "2", "3", "4",
+                            "5", "6", "7", "8")
+
+MCC_sd_quant <- MCC_sd_quant %>% pivot_longer(cols=!contains("testing_set_index"))
 
@@
   ) +
-  xlab("Testing Sets") +
+  xlab("Number of Testing Sets Included") +
   ylab("Standard Deviation") 
+```
+```{r}
+t.test(MCC_sd_quant$value[MCC_sd_quant$name==2],MCC_sd_quant$value[MCC_sd_quant$name==3])
+t.test(MCC_sd_quant$value[MCC_sd_quant$name==3],MCC_sd_quant$value[MCC_sd_quant$name==4])
+t.test(MCC_sd_quant$value[MCC_sd_quant$name==4],MCC_sd_quant$value[MCC_sd_quant$name==5])
+t.test(MCC_sd_quant$value[MCC_sd_quant$name==5],MCC_sd_quant$value[MCC_sd_quant$name==6])
+t.test(MCC_sd_quant$value[MCC_sd_quant$name==6],MCC_sd_quant$value[MCC_sd_quant$name==7])
+t.test(MCC_sd_quant$value[MCC_sd_quant$name==7],MCC_sd_quant$value[MCC_sd_quant$name==8])
 ```

### 2023-08-02 TestingPipeline/Scripts/paper_figure_generation_balanced_testing.Rmd
Updating the 50:50 split piece.

@@
 
-ggsave(
-  "../IntermediaryFiles/all_MCC_scatterplot.png",
-  plot = fig,
-  scale = 1,
-  width = 8,
-  height = 4,
-  units = c("in", "cm", "mm", "px"),
-  dpi = 300,
-  limitsize = TRUE
-)
+fig
 
@@
 ```
-=78
+=91
 
@@
 ```{r}
-t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$MCC[accuracy_scores$toolcombo=="0 1 1 0 0 1"])
-
-t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"])
-
-t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 0 0 0 1"])
-
-t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 1 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 0 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
        accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 1 0 1"])
 
-t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 1 1"])
-
-t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="0 0 1 0 0 1"], 
-       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 0 0 1"])
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 0 1 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 1 1 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 0 0 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 1 0 1"])
+
+t.test(accuracy_scores$MCC[accuracy_scores$toolcombo=="1 0 1 0 0 1"], 
+       accuracy_scores$MCC[accuracy_scores$toolcombo=="1 1 1 1 1 1"])
 ```
@@
                   cellheight=3,
-                  filename = "../IntermediaryFiles/heatmap_MCC_labels.png"
+                  filename = "../IntermediaryFiles/heatmap_MCC_labels_50-50.png"
                   )

### 2023-08-18 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updates to MCC, recall, precision boxplots.

@@
 ```{r}
-fig <- ggplot(accuracy_scores_melt[accuracy_scores_melt$toolcombo!="0 0 1 0 0 0" & accuracy_scores_melt$performance_metric!="prop_viral",], aes(x=numrules, y=performance_metric_score, 
-                                  color=numrules, fill=numrules)) +
-  geom_boxplot() +
-  geom_point(alpha=0.5) +
+fig <- ggplot(accuracy_scores_melt[accuracy_scores_melt$toolcombo!="0 0 1 0 0 0" & accuracy_scores_melt$performance_metric!="prop_viral",], aes(x=numrules, y=performance_metric_score)) +
+  geom_boxplot(alpha=0.1, fill="black") +
+  geom_point(alpha=0.2, fill="black", color="black") +
   theme_light() +
@@
     axis.title=element_text(size=16),
+    strip.background = element_rect(fill="white", color="grey"),
+    strip.text = element_text(color="black", size=14)
   ) +
@@
   xlab("Number of Rule Sets") +
-  scale_fill_manual(name="Number of Rule Sets",
-                     values = alpha(rev(viridis(6)), 0.5)) +
-  scale_color_manual(name="Number of Rule Sets",
-                     values = alpha(rev(viridis(6)), 1)) +
   facet_wrap(~performance_metric) +
@@
 ggsave(
-  "../IntermediaryFiles/MCC_precision_recall_boxplots.png",
+  "../IntermediaryFiles/MCC_precision_recall_boxplots_grey.png",
   plot = fig,

### 2023-08-29 build_features_table.Rmd
User-friendly RMarkdown files for viral sorting
The first file, build_features_table.Rmd will combine viral ID outputs.
The second file, identify_viruses.Rmd, will take the build_features_table.Rmd output and classify which contigs to call viral based on the outputs.

--- 2022-10-26 EnvironmentalData/DrinkingWaterMetagenomes/building_env_data_viral_features_set.Rmd
@@
 ---
-title: "Building Environmental Data Viral Features Set"
-author: Bridget Hegarty, James Riddell
-date: 10-26-2022
-output: html_notebook
+title: "viral_contig_identification_user"
+author: "James Riddell"
+date: "8/25/2023"
+output: html_document
 ---
+
 This Rmarkdown file takes outputs from viral identification tools and merges
@@
 
-    General filenames for each tool:
+    General output filenames used from each tool:
         Checkv: quality_summary.tsv
@@
 3) Check each chunk and ensure all columns are accounted for. For example,
-if you don't have "assembly" and called it something else, please either
+if you don't have an "assembly" column and called it something else, please either
 remove it from this script or modify your input files.
@@
 kj_path <- "/Users/hegartyb/Documents/BlueSkyPostdoc/ViralSortingToolsEvaluation/Temp/kaiju_outputs_merged.tsv"
-# KB cutoff
-KB_CUTOFF <- 3000
+# length cutoff (unit: basepairs)
+BP_CUTOFF <- 3000
 
@@
 
-# All Viral Contigs
+Here we will read in all the viral ID output files to 1) standardize their
+column names, and 2) standardize the contig names. The contig name will act
+as the key to join all of the outputs into one dataframe, so make sure it is
+the same across tool outputs.
+
+If you are not using all of the tool outputs, there is no need to run the
+code chunks for tools you are not using.
 
@@
 
-```{r}
-# Quick glimpse of cellular vs viral sequence classification by Kaiju by assembly
-table(kj_c$assembly, kj_c$Kaiju_Viral)
-```
+Now we will merge all of the outputs together into one dataframe, 'viruses'.
+If you are not using all of the tools, then there is no need to run the lines
+below for the tools you are not using.
 
@@
 ```{r}
-viruses <- full_join(x=vs2_c, y=checkV, by = c("uniq_contig", "assembly"))
+viruses <- full_join(x=checkV, y=vs2_c, by = c("uniq_contig", "assembly"))
 viruses <- full_join(x=viruses, y=dvf_c, by = c("uniq_contig", "assembly"))
@@
 
-Remove contigs not greater than the basepair length cutoff (3000)
+Remove contigs not greater than the basepair length cutoff. If you prefer
+faster code, you can also move this step to the beginning with checkV, and
+perform left joins instead of full joins to the checkV dataframe, assuming
+all contigs are represented in the CheckV dataframe.
 
 ```{r}
-viruses <- viruses %>% filter(checkv_length > KB_CUTOFF)
+viruses <- viruses %>% filter(checkv_length > BP_CUTOFF)
 ```
@@
 Use the next code chunk to check if any contigs are missing from checkv.
-If yes, then there is a mix up in contig names since checkv should contain
-all contig names even if they are identified as non-viral.
+If there are any rows present, these are located in other viral ID outputs,
+but not CheckV. Please go back and change the contig names so that they match
+the pattern in CheckV, or check your CheckV output to make sure it contains
+as many rows as headers in your fasta file (with the bp cutoff).
 ```{r}
@@
 
-calculate percent viral/host/unknown
+Calculate percent viral/host/unknown/kaiju_match_ratio columns
+These will be used in the classification step.
 ```{r}

### 2023-08-29 identify_viruses.Rmd
User-friendly RMarkdown files for viral sorting
The first file, build_features_table.Rmd will combine viral ID outputs.
The second file, identify_viruses.Rmd, will take the build_features_table.Rmd output and classify which contigs to call viral based on the outputs.

--- 2023-05-29 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
@@
 ---
-title: "Building Metagenome Viral Set and Visualization Template"
-author: Bridget Hegarty, James Riddell
-date: 10-26-2022
-output: html_notebook
+title: "identify_viruses"
+author: "James Riddell"
+date: "8/25/2023"
+output: html_document
 ---
+
 This Rmarkdown file takes outputs from viral identification tools and merges
@@
 # path of merged dataset
-# path = "/Users/riddellj/Documents/Research/VSTE/VSTE_Old_Pull/EnvironmentalData/DrinkingWaterMetagenomes/"
 path = "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/"
 setwd(path)
+
+# adjust to read file in.
 viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
-# write_csv(viruses[viruses$assembly %in% c("49639", "49640", "49641", "49642"),]paste(dataset_name, viruses_filename, 'virome', sep='_'))
 ```
-
-TODO: ADD ENVIRONMENTAL METRICS LIKE SHANNON, ALPHA DIVERSITY.
 
@@
 A final keep_score above 1 indicates we will keep that sequence and call it viral.
+
+Rule
+  subrule
+  ...
+  subrule
 
@@
 
+Virsorter
+    category ==  1 or 4: +1
+    category == 2, 3, 5, or 6: +0.5
+
 Virsorter2
-    Viral >= 50: +0.5
-    Viral >= 0.95: +0.5
-    RNA >= 0.9: +1
-    lavidaviridae >= 0.9: +1
-    NCLDV >= 0.9: +1
-
-Virsorter
-    category ==  1,4: +1
-    category == 2,5: +0.5
+    max_score >= 0.5: +0.5
+    max_score >= 0.95: +0.5 (additive to the previous subrule)
 
 DeepVirFinder:
-    Score >= 0.7: +0.5
-    Score >= 0.9: +0.5
+    Score >= 0.7 & p-value <= 0.05 & checkv_length < 20kb: +0.5
+    Score >= 0.9 & p-value <= 0.05 & checkv_length < 20kb: +0.5 (additive to the previous subrule)
+
+Tuning Addition:
+    Kaiju_viral == "Viruses" & kaiju_match_ratio >= 0.3: +1
+    If %unknown >= 75 and checkv_length < 50000: +0.5
+    If checkv_%viral >= 50 or vs2_%viral >= 50%: +0.5
+    Hallmark >= 3: +1
 
 Tuning - Not Viral:
-    Kaiju_viral = "cellular organisms": -0.5
-    If host_genes >50 and NOT checkv_provirus: -1 
-    If viral_genes == 0 and host_genes >= 1: -1
-    If 3*viral_genes <= host_genes and NOT checkv_provirus: -1
+    viral_genes == 0 & host_genes >= 1: -1
+    If host_genes >50 and checkv_provirus == False: -1 
+    If 3*viral_genes <= host_genes and  checkv_provirus == False: -1
     If length > 50,000 and hallmark <=1: -1
-    If length < 5000 and checkv completeness <= 75: -0.5
-
-Tuning - Viral:
-    Kaiju_viral = "Viruses": +1
-    If %unknown >= 75 and length < 50000: +0.5
-    If %viral >= 50: +0.5
-    Hallmark > 2: +1
 
@@
 Function that allows for the comparing pieces of the pipeline
-```{r getting_viral_set_1}
-getting_viral_set_1 <- function(input_seqs,
+```{r compute_viral_score}
+
+#' The function compute_viral_score takes in a dataframe of sequences with viral
+#' identification tool outputs and assigns a viralness score, "keep_score", based
+#' on the classifications of each viral identification tool. The function returns
+#' a list of keep_scores, one per observation of the input_seqs dataframe.
+
+#' This function is used by assigning another column of the input_seqs dataframe
+#' to the output of the function, associating a keep_score to each sequence.
+
+#' Sequences with a keep_score >= 1 are considered viral. Refer to the manuscript
+#' Figure 7 for which tool combinations to use based on the sample type. Refer
+#' to the manuscript Figures 3-6 for performance of each combination.
+
+#' parameters:
+#'  input_seqs: dataframe with input sequence names and viral ID tool outputs
+#'              from "build_features_table.Rmd"
+#'  include_*:  boolean variables if the specified rule is to be used in
+#'              computing the keep_score.
+#'  tv_*:       sub-rules of tuning addition that can be individually turned on 
+#'              or off.
+#'  tnv_*:      sub-rules of tuning removal that can be individually turned on 
+#'              or off.
+compute_viral_score <- function(input_seqs,
                                 include_vibrant=FALSE, 
+                                include_virsorter=FALSE,
                                 include_virsorter2=FALSE,
@@
                                 include_tuning_not_viral=FALSE,
-                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
-                                include_virsorter=FALSE) {
+                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T) {
 
@@
   if (include_deepvirfinder) {
-     # add if DVF calls viral
     keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
@@
                                ins=input_seqs,
-                         include_tv_1=ntv_1,
-                         include_tv_2=ntv_2,
-                         include_tv_3=ntv_3,
-                         include_tv_4=ntv_4,
-                         include_tv_5=ntv_5)
+                         include_tnv_1=ntv_1,
+                         include_tnv_2=ntv_2,
+                         include_tnv_3=ntv_3,
+                         include_tnv_4=ntv_4)
 
@@
   if (include_tv_1) {
-    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
+    current_score[ins$hallmark>=3] <- current_score[ins$hallmark>=3] + 1
   }
@@
   }
-    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
-    #note: having kaiju removal rule didn't help
   if (include_tv_3) {
@@
 tuning_not_viral <- function(current_score, ins, 
-                         include_tv_1=T,
-                         include_tv_2=T,
-                         include_tv_3=T,
-                         include_tv_4=T,
-                         include_tv_5=F) {
+                         include_tnv_1=T,
+                         include_tnv_2=T,
+                         include_tnv_3=T,
+                         include_tnv_4=T) {
     # tuning removal
-  if (include_tv_2) {
+  if (include_tnv_1) {
+    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
+  }
+
+  if (include_tnv_2) {
     current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
   }
-  if (include_tv_3) {
+  if (include_tnv_3) {
     current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
@@
 
-  if (include_tv_5) {
-    # remove if DVF calls it not viral
-    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
-  }
-
-  if (include_tv_1) {
-    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
-  }
-
-  if (include_tv_4) {
+  if (include_tnv_4) {
     current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- -3
@@
 
+The next code chunk is where you can create your own combination of tools,
+or use one of the presets provided. Simply change each include_* rule to True
+to include the rule, and False to exclude it.
+
 ```{r}
-viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_all <- compute_viral_score(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
                                               include_virsorter = T)
-viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
+viruses$keep_score_high_recall <- compute_viral_score(viruses, include_deepvirfinder = T,
                                               include_vibrant = T,
@@
                                               include_virsorter = T)
-viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+viruses$keep_score_high_MCC <- compute_viral_score(viruses, include_deepvirfinder = F,
                                               include_vibrant = F,
@@
                                               include_virsorter = F)
-viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
+viruses$keep_score_high_precision <- compute_viral_score(viruses, include_deepvirfinder = F,
                                               include_vibrant = F,
@@
 sum(viruses$keep_score_high_MCC>=1)
-sum(viruses$keep_score_all>=0.5)
-sum(viruses$keep_score_high_recall>=0.5)
-sum(viruses$keep_score_high_precision>=0.5)
-sum(viruses$keep_score_high_MCC>=0.5)
 ```
 
+Write sequence names and viral scores to a file.
 ```{r}
-write_csv(viruses[,c(1,43:45)], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
+write_csv(viruses[,c('contig','provirus','keep_score_all', 'keep_score_high_MCC', 'keep_score_high_precision', 'keep_score_high_recall')], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
 ```
 
-not having the rule, gives the same
+Write a list of sequences that are considered virus to file. Change high MCC
+to whichever combination you choose to filter by.
 
-Create list of contigs considered viral
 ```{r}
-#c_list <- data.frame(viruses_high$checkv_uniq_contig)
-#write_tsv(c_list, paste(dataset_name, "viruses_high_uniq_contig_list.tsv", sep='_'))
+virus_contig_list <- viruses[viruses$keep_score_high_MCC >= 1, c('contig', 'keep_score_high_MCC')]
+
+write_csv(virus_contig_list, paste(dataset_name, 'virus_contig_list.csv', sep='_'))
 ```
 
-Export keep_score_high and list of contigs considered viral by the combined tool.
+Most will be done with using this tool at this stage. You can take the virus
+contig list and use it to filter your fasta file for only contigs classified as
+viral by whichever tool combination chosen above.
 
-```{r}
-#write_tsv(viruses_high, paste(dataset_name, "viruses_high.tsv", sep='_'))
-```
+Below are code chunks for comparing the number of viruses identified by different tool combinations.
+These are not necessary to run, but beware that number of contigs called virus DOES NOT correlate to
+total number of viruses, and may just be including more false positives. It is always recommended
+to check contigs by looking at functional annotations (from DRAMv, for example) to confirm if 
+the sequence is viral or not.
 
@@
 for (i in 1:nrow(combos_list)) {
-  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
+  viral_scores[,i] <- compute_viral_score(viruses, include_vibrant = combos_list$VIBRANT[i],
                                             include_virsorter = combos_list$VS[i],
@@
 ```
+
+This chunk will generate a plot of all tool combinations and how many viruses
+were identified from the sample from each type.
 
@@
 ```
-
-by only a subset of the assemblies (UK/NL)
-
-```{r}
-num_viruses_uknl <- num_viruses
-num_viruses_uknl$residual <- "residual"
-num_viruses_uknl$residual[num_viruses_uknl$assembly=="Amsterdam_Leid" |
-                          num_viruses_uknl$assembly=="Amsterdam_Wee" |
-                          num_viruses_uknl$assembly=="Nieuwkoop_DHB" |
-                          num_viruses_uknl$assembly=="South_Holland_Dla" |
-                          num_viruses_uknl$assembly=="Spannenburg"
-                          ] <- "no residual"
-```
-
-```{r}
-num_viruses_uknl$numtools <- as.numeric(num_viruses_uknl$numtools)
-mean_prop_vir_by_tool <- num_viruses_uknl %>% 
-  select(toolcombo, prop_vir, numtools, residual) %>%
-  group_by(toolcombo, residual) %>%
-  summarize(prop_vir=mean(prop_vir),
-         #sd_per_vir=sd(prop_vir),
-         numtools=mean(numtools)) 
-mean_uknl <- mean_prop_vir_by_tool %>% 
-  group_by(toolcombo) %>%
-  summarize(mean=mean(prop_vir)) 
-mean_prop_vir_by_tool$toolcombo <- factor(mean_prop_vir_by_tool$toolcombo, 
-                                          levels = mean_uknl$toolcombo[order(mean_uknl$mean, decreasing = F)])
-mean_prop_vir_by_tool$numtools <- as.factor(mean_prop_vir_by_tool$numtools)
-```
-
-```{r}
-num_viruses_uknl$toolcombo <- factor(num_viruses_uknl$toolcombo, 
-                                          levels = levels(mean_prop_vir_by_tool$toolcombo))
-num_viruses_uknl$numtools <- as.factor(num_viruses_uknl$numtools)
-```
-
-```{r}
-png(paste(dataset_name, 'prop_viral_residual.png', sep='_'))
-ggplot(num_viruses_uknl) +
-  geom_hline(yintercept=0, color="grey") +
-  geom_point(aes(color=numtools, fill=numtools, y=prop_vir, x=toolcombo, shape=residual), alpha=0.5) +
-  geom_point(data=mean_prop_vir_by_tool, aes(x=toolcombo, 
-                                               y=prop_vir,
-                                             shape=residual),
-             color="black") +
-  theme_light() +
-  theme(
-    panel.grid.major.y = element_blank(),
-    panel.border = element_blank(),
-    axis.ticks.y = element_blank(),
-    legend.position = "bottom",
-    axis.text.y=element_text(size=14),
-    axis.text.x=element_text(size=14, angle = 90),
-    legend.text=element_text(size=12),
-    axis.title=element_text(size=16),
-  ) +
-  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
-  ylab("Proportion of Contigs >3kb Viral") +
-  scale_fill_manual(name="",
-                     values = alpha(rev(pal(6)), 0.5)) +
-  scale_color_manual(name="",
-                     values = alpha(rev(pal(6)), 1))
-dev.off()
-```
-```{r}
-#t.test(num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="residual"],
-#       num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="no residual"]
-#       )
-```
-
-ordination: sample by tool combo
-
-```{r}
-library(phyloseq)
-```
-
-make the sample by tool combo matrix
-```{r}
-num_viruses_uknl_wide <- num_viruses_uknl %>% 
-  select(toolcombo, assembly, num_viruses) %>%
-  pivot_wider(names_from=assembly, values_from=num_viruses)
-num_viruses_uknl_wide <- as.data.frame(num_viruses_uknl_wide)
-rownames(num_viruses_uknl_wide) <- num_viruses_uknl_wide$toolcombo
-num_viruses_uknl_wide <- num_viruses_uknl_wide[,-1]
-num_viruses_uknl_wide <- num_viruses_uknl_wide[rowSums(num_viruses_uknl_wide)>0,]
-```
-
-```{r}
-tooldata <- combos_list
-rownames(tooldata) <- tooldata$toolcombo
-tooldata <- tooldata[c(rownames(num_viruses_uknl_wide)),]
-tooldata$numtools <- str_count(tooldata$toolcombo, "1")
-tooldata$numtools <- as.factor(tooldata$numtools)
-tooldata$total_viruses <- rowSums(num_viruses_uknl_wide)
-```
-
-to reduce arch effect:
-```{r}
-num_viruses_uknl_wide <- num_viruses_uknl_wide/rowSums(num_viruses_uknl_wide)
-```
-
-```{r}
-physeq_pooled <- phyloseq::phyloseq(phyloseq::otu_table(num_viruses_uknl_wide, taxa_are_rows = F),
-                                     phyloseq::sample_data(tooldata))
-```
-
-```{r}
-ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
-```
-
-```{r}
-png(paste(dataset_name, 'ordination.png', sep='_'))
-phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
-                          shape="numtools", 
-                          color="total_viruses") + 
-  geom_point(size = 3) +
-  theme_bw() +
-  scale_shape_manual(name="Num Tools",
-                     values = c(21,22,23,24,4,8)) +
-  scale_color_viridis_c()
-dev.off()
-```
-each point is tool combination
-
-Searching for novel viruses
-```{r}
-novel <- viruses[(viruses$keep_score_high_MCC >=1) & (viruses$percent_unknown > 75),]
-novel_kaiju_match_below_30 <- novel[novel$kaiju_match_ratio < 0.3,] %>% drop_na(kaiju_match_ratio)
-novel_kaiju_match_below_1 <- novel[novel$kaiju_match_ratio < 0.01,] %>% drop_na(kaiju_match_ratio)
-novel_by_type <- novel_kaiju_match_below_1 %>% group_by(max_score_group) %>% count()
-high_MCC_num_viruses <- sum(num_viruses[num_viruses$toolcombo=='0 0 1 0 0 1',]$num_viruses)
-```

### 2023-08-29 build_features_table.Rmd
Fixed file-to-file inconsistencies
Thanks to James Tan for finding and reporting on these bugs, as well as providing solutions.

@@
 ```{r inputs}
+setwd('/Users/riddellj/Documents/GitHub/VSTE/RunFilesForUsers')
+
 # dataset name for file organization and outputs for metrics file
@@
 # checkV
-checkV_path <- "DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
+checkV_path <- "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/DrinkingWaterToolOutput/merged_checkV_allcontigs.tsv"
 # VIBRANT
-vibrant_path <- "DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
+vibrant_path <- "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/DrinkingWaterToolOutput/vibrant_merged_contig_list_all.txt"
 # DeepVirFinder
-dvf_path <- "DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
+dvf_path <- "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/DrinkingWaterToolOutput/dvf_outputs_merged.tsv"
 # Virsorter
-vs_path <- "DrinkingWaterToolOutput/vs_merged_all.txt"
+vs_path <- "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/DrinkingWaterToolOutput/vs_merged_all.txt"
 # Virsorter2
-vs2_path <- "DrinkingWaterToolOutput/vs2_merged_file.txt"
+vs2_path <- "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/DrinkingWaterToolOutput/vs2_merged_file.txt"
 # Kaiju
-kj_path <- "/Users/hegartyb/Documents/BlueSkyPostdoc/ViralSortingToolsEvaluation/Temp/kaiju_outputs_merged.tsv"
+kj_path <- "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/DrinkingWaterToolOutput/kaiju_outputs_merged.tsv"
 # length cutoff (unit: basepairs)
@@
         )
+checkV$checkv_provirus <- checkV$checkv_provirus == 'Yes'
 checkV$method = 'checkv'
@@
 vb_c$method <- "vibrant"
-vb_c$vibrant_prophage <- "No"
-vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- "Yes"
+vb_c$vibrant_prophage <- FALSE
+vb_c$vibrant_prophage[grep("_fragment_", vb_c$contig)] <- TRUE
 vb_c$contig <- gsub("_fragment_.*", "", vb_c$contig)
@@
 ```{r}
-v_missing <- viruses[is.na(viruses$checkv_uniq_contig),]
+v_missing <- viruses[is.na(viruses$checkv_quality),]
 ```
@@
 viruses$hallmark[is.na(viruses$hallmark)] <- 0
+viruses$max_score[is.na(viruses$max_score)] <- 0
 

### 2023-08-29 identify_viruses.Rmd
Fixed file-to-file inconsistencies
Thanks to James Tan for finding and reporting on these bugs, as well as providing solutions.

@@
 # dataset name for file organization and outputs for metrics file
-dataset_name <- "drinking_water_uknl"
+dataset_name <- "drinking_water"
 #name from building_env_data_viral_features_set.Rmd
-viruses_filename <- 'merged_viral_feature_outputs.tsv'
+viruses_filename <- 'merged_viral_feature_outputs.csv'
 # path of merged dataset
-path = "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/"
+path = "/Users/riddellj/Documents/GitHub/VSTE/RunFilesForUsers/"
 setwd(path)
@@
 # adjust to read file in.
-viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
+viruses <- read_csv(paste(dataset_name, viruses_filename, sep='_'))
 ```
@@
     keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
-    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
+    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$vibrant_prophage] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$vibrant_prophage] + 0.5
   }
@@
   if (include_tnv_1) {
-    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
+    current_score[(ins$checkv_host_genes>50) & !ins$checkv_provirus] <- -3
   }
@@
   if (include_tnv_3) {
-    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
+    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$checkv_provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$checkv_provirus] - 1
   }
@@
 ```{r}
-write_csv(viruses[,c('contig','provirus','keep_score_all', 'keep_score_high_MCC', 'keep_score_high_precision', 'keep_score_high_recall')], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
+write_csv(viruses[,c('contig','checkv_provirus','keep_score_all', 'keep_score_high_MCC', 'keep_score_high_precision', 'keep_score_high_recall')], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
 ```

### 2023-12-13 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating SI2 and 5 to match figure in the main body of the text.

@@
 ```{r}
-accuracy_scores <- accuracy_scores[order(accuracy_scores$precision, decreasing=F),]
-accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+mean_precision <- accuracy_scores %>% 
+  select(precision, toolcombo) %>%
+  group_by(toolcombo) %>%
+  dplyr::summarise(mean = mean(precision)) %>%
+  arrange(mean)
+```
+
+```{r}
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_precision$toolcombo)
+
+#accuracy_scores <- accuracy_scores[order(accuracy_scores$precision, decreasing=F),]
+#accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
 
@@
 
-ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
+fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=precision, 
                                   color=ruletype, fill=ruletype)) +
@@
                      values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/all_precision_scatterplot.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in", "cm", "mm", "px"),
+  dpi = 300,
+  limitsize = TRUE
+)
+```
+
+```{r}
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2
+)
+rownames(ann_col) <- combos_list$toolcombo
+
+ann_colors = list(
+  tv = c("black", pal(6)[3]),
+  tnv = c("black", pal(6)[1]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
+
+#viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
+viral_scores_1 <- mean_precision
+rn <- viral_scores_1$toolcombo
+
+viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
+viral_scores_1 <- as.matrix(viral_scores_1)
+#viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
+rownames(viral_scores_1) <- rn
+
+pheatmap::pheatmap(viral_scores_1,
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = NA,
+                  show_rownames = T,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows = F,
+                  cluster_cols = F,
+                  cellheight=3,
+                  filename = "../IntermediaryFiles/heatmap_precision_labels_S2A.png"
+                  )
 ```
@@
 ```{r}
-accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
-accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+mean_recall <- accuracy_scores %>% 
+  select(recall, toolcombo) %>%
+  group_by(toolcombo) %>%
+  dplyr::summarise(mean = mean(recall)) %>%
+  arrange(mean)
+```
+
+```{r}
+#accuracy_scores <- accuracy_scores[order(accuracy_scores$recall, decreasing=F),]
+#accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = unique(accuracy_scores$toolcombo))
+
+accuracy_scores$toolcombo <- factor(accuracy_scores$toolcombo, levels = mean_recall$toolcombo)
 
@@
 
-ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
+fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=recall, 
                                   color=ruletype, fill=ruletype)) +
@@
                      values = alpha(c(rev(magma(7)[3:6]), "grey"), 0.6))
+
+fig
+
+ggsave(
+  "../IntermediaryFiles/all_recall_scatterplot.png",
+  plot = fig,
+  scale = 1,
+  width = 8,
+  height = 4,
+  units = c("in", "cm", "mm", "px"),
+  dpi = 300,
+  limitsize = TRUE
+)
+```
+
+making heatmap for the x-axis label - Figure S2B
+```{r}
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2
+)
+rownames(ann_col) <- combos_list$toolcombo
+
+ann_colors = list(
+  tv = c("black", pal(6)[3]),
+  tnv = c("black", pal(6)[1]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
+
+#viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
+viral_scores_1 <- mean_recall
+rn <- viral_scores_1$toolcombo
+
+#viral_scores_1 <- viral_scores_1 %>% select(c("recall"))
+viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
+rownames(viral_scores_1) <- rn
+
+pheatmap::pheatmap(as.matrix(viral_scores_1),
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = NA,
+                  show_rownames = T,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows = F,
+                  cluster_cols = F,
+                  cellheight=3,
+                  filename = "../IntermediaryFiles/heatmap_recall_labels_S2B.png"
+                  )
 ```
@@
 ```
+
 MCC
@@
   group_by(toolcombo) %>%
-  summarise(mean = mean(MCC)) %>%
+  dplyr::summarise(mean = mean(MCC)) %>%
   arrange(mean)

### 2023-12-13 TestingPipeline/Scripts/visualizing_rules_on_short_fragments.Rmd
Updating SI2 and 5 to match figure in the main body of the text.

@@
 ```
+
+making heatmap for the x-axis label - Figure S5
+```{r}
+ann_col = data.frame(
+    tnv = combos_list$tune_not_viral,
+    tv = combos_list$tune_viral,
+    dvf = combos_list$DVF,
+    vb = combos_list$VIBRANT,
+    vs = combos_list$VS,
+    vs2 = combos_list$VS2
+)
+rownames(ann_col) <- combos_list$toolcombo
+
+ann_colors = list(
+  tv = c("black", pal(6)[3]),
+  tnv = c("black", pal(6)[1]),
+  dvf = c("black", pal(6)[2]),
+  vb = c("black", pal(6)[4]),
+  vs = c("black", pal(6)[5]),
+  vs2 = c("black", pal(6)[6])
+)
+
+viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==5)
+#viral_scores_1 <- mean_MCC
+rn <- viral_scores_1$toolcombo
+
+viral_scores_1 <- viral_scores_1 %>% select(c("MCC"))
+#viral_scores_1 <- viral_scores_1 %>% select(c("mean"))
+rownames(viral_scores_1) <- rn
+
+pheatmap::pheatmap(as.matrix(viral_scores_1),
+                  annotation_row = ann_col,
+                  annotation_colors = ann_colors,
+                  border_color = NA,
+                  show_rownames = T,
+                  show_colnames = F,
+                  annotation_legend = F,
+                  cluster_rows = F,
+                  cluster_cols = F,
+                  cellheight=3,
+                  filename = "../IntermediaryFiles/heatmap_MCC_labels_S5.png"
+                  )
+```
+
 ```{r}

### 2023-12-14 01-build-feature-table.R
initial commit of user-friendly version
This is my first attempt at more user-friendly code based on feedback from a few people who tried to use our testing scripts for their own virus sorting.

@@ -1,0 +1,385 @@
+#!/usr/bin/env Rscript
+
+#############################################################
+## Takes outputs from viral identification tools and merges
+## them into one file, "merged_viral_id_outputs.csv", which is the input
+## for identify_viruses.R
+##
+## NOTE: This script is a blueprint for cleaning up the feature table. Your table may
+## be different from ours and will require modifying the script for the data cleaning process
+## I have tried my bes to make it as user friendly as possible, but will require some modifications
+## depending on how your files are named and which parameters and versions 
+## were used for each viral ID output, since these can influence the filenames.
+## These changes to the data wrangling process can be made in the 2 - load data section of this script.
+##
+## Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
+## Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.
+#############################################################
+
+
+#############################################################
+# Please read the below documentation to ensure your input
+# files are in the correct format before merging.
+#############################################################
+
+# 1) For this script's inputs, the user must merge each viral identification
+# tool output with the following specified filenames into one tab-separated file.
+# Replace the filenames with their paths if needed into the variables in the
+# 'inputs' chunk.
+# 
+#     General output filenames used from each tool:
+#         Checkv: quality_summary.tsv
+#         VIBRANT: VIBRANT_genome_quality_${assembly}.tsv
+#         Virsorter: VIRSorter_global-phage-signal.csv
+#         Virsorter2: final-viral-score.tsv
+#         DeepVirFinder: ${assembly}.fasta_gt2500bp_dvfpred.txt
+#         Kaiju: ${assembly}.kaiju.names.out
+# 
+# 2) Make sure all fasta headers (assembly, contig) are consistent within and
+# across tools. Chunks for each tool do contain some lines for cleaning up these features, but due to their variability it will be the user's responsibility to
+# make sure they match across tools.
+# 
+# 3) Check each chunk and ensure all columns are accounted for. For example,
[…]

### 2023-12-14 02-identify-viruses.R

@@ -1,0 +1,247 @@
+
+
+
+
+
+
+##############################
+# 1 - Load libraries
+##############################
+
+library(ggplot2)
+library(stringr)
+library(plyr)
+library(reshape2)
+library(viridis)
+library(tidyr)
+library(dplyr)
+library(readr)
+library(data.table)
+
+##############################
+# 2 - Set parameters
+##############################
+
+option_list = list(
+  make_option(c("-i", "--infile"), type="character", default="./merged_viral_feature_outputs.csv", 
+              help="input path/filename from 01-build-feature-table.R [default = %default]", metavar="character"),
+  
+  make_option(c("-a", "--virsorter2"), type="character", default=FALSE, 
+              help="specify if <tool_name> was included in <infile> generation [default = %default]", metavar="character"),
+  
+  make_option(c("-b", "--virsorter"), type="character", default=FALSE, 
+              help="specify if <tool_name> was included in <infile> generation [default = %default]", metavar="character"),
+  
+  make_option(c("-c", "--vibrant"), type="character", default=FALSE, 
+              help="specify if <tool_name> was included in <infile> generation [default = %default]", metavar="character"),
+  
+  make_option(c("-d", "--deepvirfinder"), type="character", default=FALSE, 
+              help="specify if <tool_name> was included in <infile> generation [default = %default]", metavar="character"),
+  
+  make_option(c("-t", "--tuning_viral"), type="character", default=FALSE, 
+              help="specify if <tuning_viral> is wished to be used. Requires virsorter2, checkv, and kaiju. [default = %default]", metavar="character"),
+  
+  make_option(c("-n", "--tuning_not_viral"), type="character", default=FALSE, 
+              help="specify if <tuning_viral> is wished to be used. Requires CheckV. [default = %default]", metavar="character")
+); 
+
+opt_parser = OptionParser(option_list=option_list);
+opt = parse_args(opt_parser);
+
+
+
+##############################
+# 2 - compute viral score
+##############################
+
+# This section defines a viralness score "keep_score" based on the tool classifications. 
+# A final keep_score above 1 indicates we will keep that sequence and call it viral.
+# 
+# Rule
+# subrule
+# ...
+# subrule
+# 
+# VIBRANT
+# Quality == "Complete Circular": +1
+# Quality == "High Quality Draft": +1
+# Quality == "Medium Quality Draft": +1
[…]
